### 2022-02-16 draft/draft.Rmd
adding draft

---
title: Manifold Clustering for Latent Structure Block Models

# to produce blinded version set to 1
blinded: 0

authors: 
- name: John Koo
  affiliation: Department of YYY, University of XXX

keywords:
- block models, community detection, coordinate descent, latent structure models, manifold clustering, random dot product graph

abstract: |
  The text of your abstract. 200 or fewer words.

bibliography: bibliography.bib
output: rticles::asa_article
urlcolor: blue
header-includes:
- \usepackage{float}
- \usepackage{mathtools}
- \usepackage{natbib}
- \usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
- \setcitestyle{numbers,square,comma}
- \usepackage{verbatim}
- \usepackage{amsthm}
- \usepackage{comment}
---

\newcommand{\diag}{\mathrm{diag}}
\newcommand{\tr}{\mathrm{Tr}}
\newcommand{\blockdiag}{\mathrm{blockdiag}}
\newcommand{\indep}{\stackrel{\mathrm{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\mathrm{iid}}{\sim}}
\newcommand{\Bernoulli}{\mathrm{Bernoulli}}
\newcommand{\Betadist}{\mathrm{Beta}}
\newcommand{\BG}{\mathrm{BernoulliGraph}}
\newcommand{\Uniform}{\mathrm{Uniform}}
\newcommand{\PABM}{\mathrm{PABM}}
\newcommand{\RDPG}{\mathrm{RDPG}}
\newcommand{\GRDPG}{\mathrm{GRDPG}}
\newcommand{\Multinomial}{\mathrm{Multinomial}}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\theoremstyle{remark}
\newtheorem{remark}{Remark}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\as}{\stackrel{\mathrm{a.s.}}{\to}}

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      # eval = FALSE,
                      comment = NA, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.pos = 'H', 
                      fig.align = 'center', 
                      fig.height = 3, 
                      fig.width = 5, 
                      fig.dpi = 300)

options(xtable.comment = FALSE, 
        xtable.table.placement = 'H')
```

```{r}
import::from(magrittr, `%>%`, `%<>%`)
library(ggplot2)
source('../functions.R')
source('~/dev/pabm-grdpg/functions.R')
source('https://mtrosset.pages.iu.edu/Courses/675/graph.r')

theme_set(theme_bw())

doMC::registerDoMC(8)
```

# Introduction

# Latent Structure Block Models

All block models are latent structure block models.

# Methods

Here, we provide two algorithms for LSBM community detection.

## Nonintersecting Manifolds

### Preliminary Theory

#### Distributions of differences of order statistics

Let $D_i = X_{(i+1)} - X_{(i)}$. Then if $\max_i D_i < \delta$, we have sufficient separation of points in $\mathcal{M}_1$. 
Then it is sufficient to quantify $P(\max_i D_i > \delta)$ as a function of $n$ and $\delta$ and show that this converges to zero as $n$ grows to $\infty$.

We denote $f(x)$ as the density of each $X_i$, $g_i(x)$ as the density of $X_{(i)}$, 
$g_{ij}(x, y)$ as the joint density of $X_{(i)}, X_{(j)}$, and $h_i(d)$ as the density of $D_i$ (with corresponding capital letters for the cumulative distribution functions).

The following are taken as given[^wikipedia]:

1. $g_i(x) = \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x))^{n-i} f(x)$.
2. $g_{ij}(x, y) = \frac{n!}{(i-1)! (j-i-1)! (n-j)!} (F(x))^{i-1} (F(y) - F(x))^{j-i-1} (1 - F(y))^{n-j} f(x) f(y)$.
3. By convolution, $h_i(d) = \int_0^{1} g_{i, i+1} (x, x + d) \dd x$.

[^wikipedia]: https://en.wikipedia.org/wiki/Order_statistic

\begin{lemma}[The probability density function of $D_i$]
\begin{equation}
\label{eq:pdf}
h_i(d) = \int_0^{1-d} \frac{n!}{(i-1)! (n-i-1)!} (F(x))^{i-1} (1 - F(x+d))^{n-i-1} f(x) f(x+d) \dd x 
\end{equation}
\end{lemma}

\begin{proof}
This is just a direct consequence of 2 and 3 under the given statements. 
We also note that because the support of $X_i$ is $[0, 1]$, 
the integral only needs to be evaluated from $0$ to $1 - d$ 
because of the $f(x+d)$ and $1 - F(x+d)$ terms.
\end{proof}

\begin{lemma}[The cumulative distribution function of $D_i$]

\begin{equation}
\label{eq:cdf}
P(D_i < \delta) = H_i(\delta) = 1 - \int_0^{1-\delta} \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \dd x
\end{equation}
\end{lemma}

\begin{proof}
$$
\begin{aligned}
H_i(\delta) & = \int_x^{x+\delta} h_i(d) \dd d \\
& = \int_x^{x+\delta} \int_0^{1} \frac{n!}{(i-1)! (n-i-1)!} ((F(x))^{i-1} (1 - F(x+d))^{n-i-1} f(x) f(x+d) \dd x \dd d \\
& = \int_0^{1} \frac{n!}{(i-1)! (n-i-1)!} (F(x))^{i-1} f(x) \int_x^{x+\delta} (1 - F(x+d))^{n-i-1} f(x+d) \dd d \dd x \\
& = \int_0^{1} \frac{n!}{(i-1)! (n-i-1)!} (F(x))^{i-1} f(x) \int_{F(x)}^{F(x+\delta)} (1 - u)^{n-i-1} \dd u \dd x \\
& = \int_0^{1} \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} f(x) ((1 - F(x))^{n-i} - (1 - F(x + \delta))^{n-i}) \dd x \\ 
& = \int_0^1 g_i(x) \dd x - \int_0^1 \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \dd x \\
& = 1 - \int_0^1 \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \dd x \\
\end{aligned}
$$

Because of the $x + \delta$ term, we can't actually evaluate this integral all the way up to $1$, and so we are left with 
$$
\begin{aligned}
& = 1 - \int_0^{1 - \delta} \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \dd x.
\end{aligned}
$$
\end{proof}

#### Uniform case

\begin{lemma}[Differences between order statistics of a uniform distribution]
If $X_1, ..., X_n \iid \Uniform(0, 1)$, then each $D_i \sim \Betadist(1, n)$.
\end{lemma}

\begin{proof}
We begin with Eq. (\ref{eq:pdf}), plugging in $f(x) = 1$ and $F(x) = x$:

$$
h_i(d) = \int_0^{1-d} \frac{n!}{(i-1)! (n-i-1)!} x^{i-1} (1-x-d)^{n-i-1} \dd x
$$

Then we proceed with integration by parts, setting 
$u = x^{i-1} \implies du = (i-1) x^{i-2}$ and 
$dv = (1-x-d)^{n-i-1} dx \implies v = -\frac{1}{n-i} (1-x-d)^{n-i-1}$. 
Note that $u v |_0^{1-d} = 0$ in this case. This yields

$$
= \frac{n!}{(i-1)! (n-i-1)!} \int \frac{i-1}{n-i} x^{i-2} (1-x-d)^{n-i} \dd x
$$

Then applying integration by parts again until the $x^p$ term disappears, we get:

$$
\begin{aligned}
& = \frac{n!}{(i-1)! (n-i-1)!} \frac{(i-1)!}{(n-i) \cdots (n-2)} \int_0^{1-d} (1-x-d)^{n-2} \dd x \\
& = -\frac{n (n-1)}{n-1} (1-x-d)^{n-1} \Big|_0^{1-d} \\
& = n (1 - d)^{n-1}
\end{aligned}
$$

This the density function for $\Betadist(1, n)$, completing the proof.
\end{proof}

\begin{theorem}
Let $X_1, ..., X_n \iid \Uniform(0, 1)$. 
Then for any $\epsilon$ and $\delta > 0$, 
there exists an $N = O \big(\frac{-\log \epsilon}{\delta} \big)$ such that 
$P(\max_i X_{(i+1)} - X_{(i)} < \delta) \geq 1 - \epsilon$ when $n > N$.
\end{theorem}

\begin{proof}[Proof (sketch)]
Since $X_{(i+1)} - X_{(i)} = D_i \sim \Betadist(1, n)$, 
$P(X_{(i+1)} - X_{(i)} < \delta) = 1 - (1 - \delta)^n $. This yields

$$
\begin{aligned}
P(\max_i D_i < \delta) & \geq (P(D_i < \delta))^{n-1} \\
& = (1 - (1 - \delta)^n)^{n-1} \\
& \approx e^{-n \exp(-n \delta)}.
\end{aligned}
$$

In the limit $n \to \infty$, this goes to 1.
\end{proof}

#### General case for one-dimensional manifolds

\begin{theorem}
\label{thm:generalized}
Let $X_1, ..., X_n \iid F$ with support $[0, 1]$, and suppose $f(x)$ is continuous and $f(x) \geq a > 0$ everywhere on the support. 
Let $D_i = X_{(i+1)} - X_{(i)}$. 
Then for any $\epsilon > 0$, there exists $N > 0$ such that $P(\max_i D_i < \delta) \geq 1 - \epsilon$ when $n > N$.
\end{theorem}

\begin{proof}[Proof (sketch)]
We start with Eq. (\ref{eq:cdf}):

$$P(D_i \leq \delta) = 1 - \int_0^{1-\delta} \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \dd x.$$

Making the approximation $F(x+\delta) \approx F(x) + \delta f(x)$ 
and bounding $f(x) \geq a$, we get:

$$P(D_i \leq \delta) \geq 1 - \int_0^{1-\delta} \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x) - a \delta)^{n-i} f(x) \dd x.$$

Then making the substitution $u = F(x) \implies du = f(x) \dd x$, we obtain

$$1 - \int_0^{F(1-\delta)} \frac{n!}{(n-i)! (i-1)!} u^{i-1} (1 - u - a \delta)^{n-i} \dd u$$

Evaluating the integral yields

$$P(D_i < \delta) = 1 - (1 - a \delta)^n + (1 - F(1-\delta) - a \delta)^n.$$

Then as before,

$$
\begin{aligned}
P(\max_i D_i < \delta) & = P(\text{all } D_i < \delta) \\
& = 1 - P(\text{some } D_i > \delta) \\
& \geq 1 - \sum_i^{n-1} P(D_i > \delta) \\
& = 1 - (n - 1) (1 - a \delta)^n + (n - 1) (1 - F(1 - \delta) - a \delta)^n
\end{aligned}
$$

This converges to $1$ in the limit $n \to \infty$.

We can also approximate $F(1 - \delta) \approx 1 - a \delta$, which yields 
$1 - (n - 1) (1 - a \delta)^n$. Setting this $\geq 1 - \epsilon$ ...
\end{proof}

#### Extension to multidimensional manifolds

Here, we extend the results on the line to the unit hypercube. The following theorem is a direct consequence of Lemma 2 of \citet{trosset2020rehabilitating}.

\begin{theorem}
\label{thm:multidim}
Let $X_1, ..., X_n \iid F$ with support $[0, 1]^r$, and $f(x) \geq a > 0$ everywhere on the support. 
Define $E_n$ as the event that an $\eta$-neighborhood graph constructed from the sample is connected. 
Then for any $\epsilon > 0$, there exists $N = O \bigg( \frac{\log \epsilon \eta^r / r^{r/2}}{\log (1 - \frac{a \eta^r}{r^{r / 2}})} \bigg)$ such that $P(E_n) > 1 - \epsilon$ when $n \geq N$.
\end{theorem}

\begin{proof}[Proof (sketch)]
Divide the hypercube $[0, 1]^r$ into a grid of sub-hypercubes of side length at most $\eta / \sqrt{r}$. 
$E_n$ is satisfied if each sub-hypercube contains at least one $X_i$ from the sample.

$$
\begin{aligned}
P(E_n) & = 1 - P(\text{some cells don't contain } X_i) \\
& \geq 1 - \sum_k^{\lceil \sqrt{r} / \eta \rceil^r} \prod_i^n P(X_i \text{ is not in the } k^{th} \text{ hypercube}) \\
& \geq 1 - \lceil \sqrt{r} / \eta \rceil^r (1 - a \eta^r / r^{r/2})^n
\end{aligned}
$$

Setting this $\geq 1 - \epsilon$ and solving for $n$ yields the desired rate. 
\end{proof}

## Intersecting Manifolds

\begin{algorithm}[h]
  \DontPrintSemicolon
  \SetAlgoLined
  \KwData{Adjacency matrix $A$, number of communities $K$, embedding dimensions $p$, $q$, stopping criterion $\epsilon$}
  \KwResult{Community assignments $1, ..., K$, curves $g_1, ..., g_K$}
    Compute $X$, the ASE of $A$ using the $p$ most positive and $q$ most negative eigenvalues and their corresponding eigenvectors.\;
    Initialize community labels $z_1, ..., z_n$.\;
    \Repeat {$\sum_k \sum_{i \in C_k} \|x_i - g_k(t_i)\|^2 < \epsilon$} {
      \For {$k = 1, ..., K$} {
        Define $X_k$ as the rows of $X$ for which $z_i = k$.\;
        Fit curve $g_k$ and positions $t_{k_i}$ to $X_k$ by minimizing $\sum_{k_i} \|x_{k_i} - g_k(t_{k_i})\|^2$.\;
      }
      \For {$k = 1, ..., K$} {
        Assign $z_i \leftarrow \arg\min_\ell \|x_i - g_\ell(t_i)\|^2$.\
      }
    }
  \caption{$K$-Curves Clustering.}
\end{algorithm}

\begin{theorem}
\label{k-curves-clustering}
Let each $g_k$ be smooth. 
Then $K$-curves clustering converges to a stationary point of the objective, 
$\sum_k \sum_{i \in C_k} \|x_i - g_k(t_i)\|^2$.
\end{theorem}

\begin{proof}
$K$-curves clustering is a batch coordinate descent algorithm. 
Thus, in order to show that it converges to a stationary point, it is sufficient to show that each descent step decreases the objective function. 
\end{proof}

# Examples

\begin{example}
Here, $K = 2$ with $g_1(t) = \begin{bmatrix} t^2 & 2 t (1 - t) \end{bmatrix}^\top$ and $g_2(t) = \begin{bmatrix} 2 t (1 - t) & (1 - t) ^ 2 \end{bmatrix}^\top$. We draw $n_1 = n_2 = 2^8$ points uniformly from each curve.

```{r, fig.height = 2, fig.cap = 'Latent positions, labeled by curve/community.'}
f1 <- function(t) {
  x1 <- t ^ 2
  x2 <- 2 * t - 2 * t ^ 2
  return(cbind(x1, x2))
}

f2 <- function(t) {
  x1 <- 2 * t - 2 * t ^ 2
  x2 <- 1 - 2 * t + t ^ 2
  return(cbind(x1, x2))
}

set.seed(314159)

n1 <- 2 ** 8
n2 <- n1
n <- n1 + n2
z <- c(rep(1, n1), rep(2, n2))

a <- 1
b <- 1
t1 <- rbeta(n1, a, b)
t2 <- rbeta(n2, a, b)
T <- rbind(t1, t2)

X1 <- f1(t1)
X2 <- f2(t2)
X <- rbind(X1, X2)

ggplot() + 
  geom_point(aes(x = X[, 1], y = X[, 2], colour = factor(z)), size = .5) + 
  coord_fixed() + 
  labs(x = expression(x[1]), y = expression(x[2]), colour = NULL) + 
  theme_bw() + 
  theme(legend.position = 'none')

P <- X %*% t(X)
diag(P) <- 0
A <- draw.graph(P)
Xhat <- embedding(A, 2, 0)
```

We draw $A \sim \RDPG(X)$ and obtain the following ASE:

```{r, fig.height = 3, fig.cap = 'ASE of an RDPG drawn from the latent positions, labeled by curve/community.'}
ggplot() + 
  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2], colour = factor(z)), size = .5) + 
  coord_fixed() + 
  labs(x = expression(x[1]), y = expression(x[2]), colour = NULL) + 
  theme_bw() + 
  theme(legend.position = 'none')
```

We then try applying $K$-curves clustering to this graph. 
The first three are with random initial labels, forcing the intercept to be zero. 
The fourth initializes the labels randomly but allows the intercept to be nonzero. 
The fifth initializes the labels by spectral clustering with the normalized Laplacian, again forcing the intercept to be zero. 
The sixth also initializes via spectral clustering but allows the intercept to be nonzero.

```{r, cache = TRUE}
random.starts <- 4
eps <- 1e-3
out <- lapply(seq(random.starts), function(i) {
  set.seed(i)
  intercept <- (i / random.starts > .5 + eps)
  manifold.clustering(Xhat, 2, parallel = TRUE, intercept = intercept, 
                      maxit = 50,
                      verbose = FALSE)
})
names(out) <- paste('random', seq(random.starts))
names(out) <- sapply(seq(random.starts), function(i) {
  if (i / random.starts < .5 + eps) {
    paste('random', i, '\n(w/o intercept)')
  } else {
    paste('random', i, '\n(w/ intercept)')
  }
})

# out[['random\n(w/ intercept)']] <- manifold.clustering(Xhat, 2,
#                                           parallel = TRUE,
#                                           intercept = TRUE, 
#                                           maxit = 50, 
#                                           initialization = 'random',
#                                           verbose = FALSE)

out[['spectral clust\n(w/o intercept)']] <- manifold.clustering(
  Xhat, 2, 
  parallel = TRUE, 
  intercept = FALSE, 
  maxit = 50, 
  initialization = 'spectral',
  verbose = FALSE)

# out[['ground truth']] <- manifold.clustering(Xhat, 2,
#                                             parallel = TRUE,
#                                             intercept = FALSE,
#                                             maxit = 50,
#                                             initialization = z,
#                                             verbose = FALSE)

out[['spectral clust\n(w/ intercept)']] <- manifold.clustering(
  Xhat, 2,
  parallel = TRUE,
  intercept = TRUE,
  maxit = 50,
  initialization = 'spectral',
  verbose = FALSE)
```

```{r, fig.cap = 'Clustering loss vs. iteration for each run of K-curve clustering.', fig.height = 2, fig.width = 6}
# max.iter <- max(sapply(out, function(x) x$niter))
max.loss <- max(sapply(out, function(x) max(x$loss[-1])))
# plot(out[[1]]$loss, xlim = c(0, max.iter), ylim = c(0, max.loss),
#      ylab = 'clustering loss', xlab = 'iteration')
# # lines(out[[1]]$loss)
# 
# for (i in seq_along(out)) {
#   points(out[[i]]$loss, col = i)
#   lines(out[[i]]$loss, col = i)
# }
# legend('topright', 
#        title = 'initialization',
#        cex = .75, 
#        bty = 'n',
#        legend = names(out),
#        col = seq_along(out),
#        pch = 1)

loss.df <- plyr::ldply(seq_along(out), function(i) {
  dplyr::data_frame(loss = out[[i]]$loss,
                    initialization = factor(names(out),
                                            levels = names(out))[i]) %>%
    dplyr::mutate(iteration = seq(dplyr::n()))
})

ggplot(loss.df) + 
  geom_point(aes(x = iteration, y = loss, colour = initialization)) + 
  geom_line(aes(x = iteration, y = loss, colour = initialization)) + 
  # ylim(0, max.loss) + 
  labs(y = 'clustering loss')
```

```{r, fig.width = 6, fig.height = 5, fig.cap = 'ASE labeled by estimated community labels for each initialization strategy.'}
plots.list <- lapply(seq_along(out), function(i) {
  zhat.manifold <- out[[i]]$z

  t. <- seq(0, 1, .01)
  Y1 <- bezier.curve(t., out[[i]]$p[[1]])
  Y2 <- bezier.curve(t., out[[i]]$p[[2]])

  # Y1 <- out[[i]]$X[[1]][order(out[[i]]$t[[1]]), ]
  # Y2 <- out[[i]]$X[[2]][order(out[[i]]$t[[2]]), ]

  ggplot() + 
    geom_point(aes(x = Xhat[, 1], y = Xhat[, 2], 
                   colour = factor(zhat.manifold)),
               size = .5) + 
    labs(x = expression(x[1]), y = expression(x[2]), colour = NULL,
         title = names(out)[i]) + 
    theme_bw() + 
    # coord_fixed() +
    theme(legend.position = 'none') +
    geom_path(aes(x = Y1[, 1], y = Y1[, 2])) +
    geom_path(aes(x = Y2[, 1], y = Y2[, 2])) + 
    xlim(min(Xhat[, 1]), max(Xhat[, 1])) + ylim(min(Xhat[, 2]), max(Xhat[, 2]))
})

gridExtra::grid.arrange(grobs = plots.list, ncol = 3)
```

\end{example}

# Discussion

### 2022-02-16 draft/draft.tex
adding draft

--- 2022-02-16 draft/draft.Rmd
@@
----
-title: Manifold Clustering for Latent Structure Block Models
+% !TeX program = pdfLaTeX
+\documentclass[12pt]{article}
+\usepackage{amsmath}
+\usepackage{graphicx,psfrag,epsf}
+\usepackage{enumerate}
+\usepackage{natbib}
+\usepackage{textcomp}
+\usepackage[hyphens]{url} % not crucial - just used below for the URL
+\usepackage{hyperref}
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 
-# to produce blinded version set to 1
-blinded: 0
+%\pdfminorversion=4
+% NOTE: To produce blinded version, replace "0" with "1" below.
+\newcommand{\blind}{0}
 
-authors: 
-- name: John Koo
-  affiliation: Department of YYY, University of XXX
+% DON'T change margins - should be 1 inch all around.
+\addtolength{\oddsidemargin}{-.5in}%
+\addtolength{\evensidemargin}{-.5in}%
+\addtolength{\textwidth}{1in}%
+\addtolength{\textheight}{1.3in}%
+\addtolength{\topmargin}{-.8in}%
 
-keywords:
-- block models, community detection, coordinate descent, latent structure models, manifold clustering, random dot product graph
+%% load any required packages here
 
-abstract: |
+% Pandoc citation processing
+
+\usepackage{float}
+\usepackage{mathtools}
+\usepackage{natbib}
+\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
+\setcitestyle{numbers,square,comma}
+\usepackage{verbatim}
+\usepackage{amsthm}
+\usepackage{comment}
+
+\begin{document}
+
+\def\spacingset#1{\renewcommand{\baselinestretch}%
+{#1}\small\normalsize} \spacingset{1}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\if0\blind
+{
+  \title{\bf Manifold Clustering for Latent Structure Block Models}
+
+  \author{
+        John Koo \\
+    Department of YYY, University of XXX\\
+      }
+  \maketitle
+} \fi
+
+\if1\blind
+{
+  \bigskip
+  \bigskip
+  \bigskip
+  \begin{center}
+    {\LARGE\bf Manifold Clustering for Latent Structure Block Models}
+  \end{center}
+  \medskip
+} \fi
+
+\bigskip
+\begin{abstract}
 The text of your abstract. 200 or fewer words.
+\end{abstract}
 
-bibliography: bibliography.bib
-output: rticles::asa_article
-urlcolor: blue
-header-includes:
-- \usepackage{float}
-- \usepackage{mathtools}
-- \usepackage{natbib}
-- \usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
-- \setcitestyle{numbers,square,comma}
-- \usepackage{verbatim}
-- \usepackage{amsthm}
-- \usepackage{comment}
----
+\noindent%
+{\it Keywords:} block models, community detection, coordinate descent,
+latent structure models, manifold clustering, random dot product graph
+\vfill
+
+\newpage
+\spacingset{1.45} % DON'T change the spacing!
 
@@
 
-```{r setup, include = FALSE}
-knitr::opts_chunk$set(echo = FALSE, 
-                      # eval = FALSE,
-                      comment = NA, 
-                      warning = FALSE, 
-                      message = FALSE, 
-                      fig.pos = 'H', 
-                      fig.align = 'center', 
-                      fig.height = 3, 
-                      fig.width = 5, 
-                      fig.dpi = 300)
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
-options(xtable.comment = FALSE, 
-        xtable.table.placement = 'H')
-```
-
-```{r}
-import::from(magrittr, `%>%`, `%<>%`)
-library(ggplot2)
-source('../functions.R')
-source('~/dev/pabm-grdpg/functions.R')
-source('https://mtrosset.pages.iu.edu/Courses/675/graph.r')
-
-theme_set(theme_bw())
-
-doMC::registerDoMC(8)
-```
-
-# Introduction
-
-# Latent Structure Block Models
+\hypertarget{latent-structure-block-models}{%
+\section{Latent Structure Block
+Models}\label{latent-structure-block-models}}
 
@@
 
-# Methods
+\hypertarget{methods}{%
+\section{Methods}\label{methods}}
 
@@
 
-## Nonintersecting Manifolds
+\hypertarget{nonintersecting-manifolds}{%
+\subsection{Nonintersecting Manifolds}\label{nonintersecting-manifolds}}
 
-### Preliminary Theory
+\hypertarget{preliminary-theory}{%
+\subsubsection{Preliminary Theory}\label{preliminary-theory}}
 
-#### Distributions of differences of order statistics
+\hypertarget{distributions-of-differences-of-order-statistics}{%
+\paragraph{Distributions of differences of order
+statistics}\label{distributions-of-differences-of-order-statistics}}
 
-Let $D_i = X_{(i+1)} - X_{(i)}$. Then if $\max_i D_i < \delta$, we have sufficient separation of points in $\mathcal{M}_1$. 
-Then it is sufficient to quantify $P(\max_i D_i > \delta)$ as a function of $n$ and $\delta$ and show that this converges to zero as $n$ grows to $\infty$.
+Let \(D_i = X_{(i+1)} - X_{(i)}\). Then if \(\max_i D_i < \delta\), we
+have sufficient separation of points in \(\mathcal{M}_1\). Then it is
+sufficient to quantify \(P(\max_i D_i > \delta)\) as a function of \(n\)
+and \(\delta\) and show that this converges to zero as \(n\) grows to
+\(\infty\).
 
-We denote $f(x)$ as the density of each $X_i$, $g_i(x)$ as the density of $X_{(i)}$, 
-$g_{ij}(x, y)$ as the joint density of $X_{(i)}, X_{(j)}$, and $h_i(d)$ as the density of $D_i$ (with corresponding capital letters for the cumulative distribution functions).
+We denote \(f(x)\) as the density of each \(X_i\), \(g_i(x)\) as the
+density of \(X_{(i)}\), \(g_{ij}(x, y)\) as the joint density of
+\(X_{(i)}, X_{(j)}\), and \(h_i(d)\) as the density of \(D_i\) (with
+corresponding capital letters for the cumulative distribution
+functions).
 
-The following are taken as given[^wikipedia]:
+The following are taken as given\footnote{\url{https://en.wikipedia.org/wiki/Order_statistic}}:
 
-1. $g_i(x) = \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x))^{n-i} f(x)$.
-2. $g_{ij}(x, y) = \frac{n!}{(i-1)! (j-i-1)! (n-j)!} (F(x))^{i-1} (F(y) - F(x))^{j-i-1} (1 - F(y))^{n-j} f(x) f(y)$.
-3. By convolution, $h_i(d) = \int_0^{1} g_{i, i+1} (x, x + d) \dd x$.
-
-[^wikipedia]: https://en.wikipedia.org/wiki/Order_statistic
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  \(g_i(x) = \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x))^{n-i} f(x)\).
+\item
+  \(g_{ij}(x, y) = \frac{n!}{(i-1)! (j-i-1)! (n-j)!} (F(x))^{i-1} (F(y) - F(x))^{j-i-1} (1 - F(y))^{n-j} f(x) f(y)\).
+\item
+  By convolution,
+  \(h_i(d) = \int_0^{1} g_{i, i+1} (x, x + d) \mathrm{d}x\).
+\end{enumerate}
 
@@
 \label{eq:pdf}
-h_i(d) = \int_0^{1-d} \frac{n!}{(i-1)! (n-i-1)!} (F(x))^{i-1} (1 - F(x+d))^{n-i-1} f(x) f(x+d) \dd x 
+h_i(d) = \int_0^{1-d} \frac{n!}{(i-1)! (n-i-1)!} (F(x))^{i-1} (1 - F(x+d))^{n-i-1} f(x) f(x+d) \mathrm{d}x 
 \end{equation}
@@
 \label{eq:cdf}
-P(D_i < \delta) = H_i(\delta) = 1 - \int_0^{1-\delta} \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \dd x
+P(D_i < \delta) = H_i(\delta) = 1 - \int_0^{1-\delta} \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \mathrm{d}x
 \end{equation}
@@
 \begin{aligned}
-H_i(\delta) & = \int_x^{x+\delta} h_i(d) \dd d \\
-& = \int_x^{x+\delta} \int_0^{1} \frac{n!}{(i-1)! (n-i-1)!} ((F(x))^{i-1} (1 - F(x+d))^{n-i-1} f(x) f(x+d) \dd x \dd d \\
-& = \int_0^{1} \frac{n!}{(i-1)! (n-i-1)!} (F(x))^{i-1} f(x) \int_x^{x+\delta} (1 - F(x+d))^{n-i-1} f(x+d) \dd d \dd x \\
-& = \int_0^{1} \frac{n!}{(i-1)! (n-i-1)!} (F(x))^{i-1} f(x) \int_{F(x)}^{F(x+\delta)} (1 - u)^{n-i-1} \dd u \dd x \\
-& = \int_0^{1} \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} f(x) ((1 - F(x))^{n-i} - (1 - F(x + \delta))^{n-i}) \dd x \\ 
-& = \int_0^1 g_i(x) \dd x - \int_0^1 \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \dd x \\
-& = 1 - \int_0^1 \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \dd x \\
+H_i(\delta) & = \int_x^{x+\delta} h_i(d) \mathrm{d}d \\
+& = \int_x^{x+\delta} \int_0^{1} \frac{n!}{(i-1)! (n-i-1)!} ((F(x))^{i-1} (1 - F(x+d))^{n-i-1} f(x) f(x+d) \mathrm{d}x \mathrm{d}d \\
+& = \int_0^{1} \frac{n!}{(i-1)! (n-i-1)!} (F(x))^{i-1} f(x) \int_x^{x+\delta} (1 - F(x+d))^{n-i-1} f(x+d) \mathrm{d}d \mathrm{d}x \\
+& = \int_0^{1} \frac{n!}{(i-1)! (n-i-1)!} (F(x))^{i-1} f(x) \int_{F(x)}^{F(x+\delta)} (1 - u)^{n-i-1} \mathrm{d}u \mathrm{d}x \\
+& = \int_0^{1} \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} f(x) ((1 - F(x))^{n-i} - (1 - F(x + \delta))^{n-i}) \mathrm{d}x \\ 
+& = \int_0^1 g_i(x) \mathrm{d}x - \int_0^1 \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \mathrm{d}x \\
+& = 1 - \int_0^1 \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \mathrm{d}x \\
 \end{aligned}
@@
 \begin{aligned}
-& = 1 - \int_0^{1 - \delta} \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \dd x.
+& = 1 - \int_0^{1 - \delta} \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \mathrm{d}x.
 \end{aligned}
@@
 
-#### Uniform case
+\hypertarget{uniform-case}{%
+\paragraph{Uniform case}\label{uniform-case}}
 
 \begin{lemma}[Differences between order statistics of a uniform distribution]
-If $X_1, ..., X_n \iid \Uniform(0, 1)$, then each $D_i \sim \Betadist(1, n)$.
+If $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}\mathrm{Uniform}(0, 1)$, then each $D_i \sim \mathrm{Beta}(1, n)$.
 \end{lemma}
@@
 $$
-h_i(d) = \int_0^{1-d} \frac{n!}{(i-1)! (n-i-1)!} x^{i-1} (1-x-d)^{n-i-1} \dd x
+h_i(d) = \int_0^{1-d} \frac{n!}{(i-1)! (n-i-1)!} x^{i-1} (1-x-d)^{n-i-1} \mathrm{d}x
 $$
@@
 $$
-= \frac{n!}{(i-1)! (n-i-1)!} \int \frac{i-1}{n-i} x^{i-2} (1-x-d)^{n-i} \dd x
+= \frac{n!}{(i-1)! (n-i-1)!} \int \frac{i-1}{n-i} x^{i-2} (1-x-d)^{n-i} \mathrm{d}x
 $$
@@
 \begin{aligned}
-& = \frac{n!}{(i-1)! (n-i-1)!} \frac{(i-1)!}{(n-i) \cdots (n-2)} \int_0^{1-d} (1-x-d)^{n-2} \dd x \\
+& = \frac{n!}{(i-1)! (n-i-1)!} \frac{(i-1)!}{(n-i) \cdots (n-2)} \int_0^{1-d} (1-x-d)^{n-2} \mathrm{d}x \\
 & = -\frac{n (n-1)}{n-1} (1-x-d)^{n-1} \Big|_0^{1-d} \\
@@
 
-This the density function for $\Betadist(1, n)$, completing the proof.
+This the density function for $\mathrm{Beta}(1, n)$, completing the proof.
 \end{proof}
@@
 \begin{theorem}
-Let $X_1, ..., X_n \iid \Uniform(0, 1)$. 
+Let $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}\mathrm{Uniform}(0, 1)$. 
 Then for any $\epsilon$ and $\delta > 0$, 
@@
 \begin{proof}[Proof (sketch)]
-Since $X_{(i+1)} - X_{(i)} = D_i \sim \Betadist(1, n)$, 
+Since $X_{(i+1)} - X_{(i)} = D_i \sim \mathrm{Beta}(1, n)$, 
 $P(X_{(i+1)} - X_{(i)} < \delta) = 1 - (1 - \delta)^n $. This yields
@@
 
-#### General case for one-dimensional manifolds
+\hypertarget{general-case-for-one-dimensional-manifolds}{%
+\paragraph{General case for one-dimensional
+manifolds}\label{general-case-for-one-dimensional-manifolds}}
 
@@
 \label{thm:generalized}
-Let $X_1, ..., X_n \iid F$ with support $[0, 1]$, and suppose $f(x)$ is continuous and $f(x) \geq a > 0$ everywhere on the support. 
+Let $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}F$ with support $[0, 1]$, and suppose $f(x)$ is continuous and $f(x) \geq a > 0$ everywhere on the support. 
 Let $D_i = X_{(i+1)} - X_{(i)}$. 
@@
 
-$$P(D_i \leq \delta) = 1 - \int_0^{1-\delta} \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \dd x.$$
+$$P(D_i \leq \delta) = 1 - \int_0^{1-\delta} \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \mathrm{d}x.$$
 
@@
 
-$$P(D_i \leq \delta) \geq 1 - \int_0^{1-\delta} \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x) - a \delta)^{n-i} f(x) \dd x.$$
+$$P(D_i \leq \delta) \geq 1 - \int_0^{1-\delta} \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x) - a \delta)^{n-i} f(x) \mathrm{d}x.$$
 
-Then making the substitution $u = F(x) \implies du = f(x) \dd x$, we obtain
+Then making the substitution $u = F(x) \implies du = f(x) \mathrm{d}x$, we obtain
 
-$$1 - \int_0^{F(1-\delta)} \frac{n!}{(n-i)! (i-1)!} u^{i-1} (1 - u - a \delta)^{n-i} \dd u$$
+$$1 - \int_0^{F(1-\delta)} \frac{n!}{(n-i)! (i-1)!} u^{i-1} (1 - u - a \delta)^{n-i} \mathrm{d}u$$
 
@@
 
-#### Extension to multidimensional manifolds
+\hypertarget{extension-to-multidimensional-manifolds}{%
+\paragraph{Extension to multidimensional
+manifolds}\label{extension-to-multidimensional-manifolds}}
 
-Here, we extend the results on the line to the unit hypercube. The following theorem is a direct consequence of Lemma 2 of \citet{trosset2020rehabilitating}.
+Here, we extend the results on the line to the unit hypercube. The
+following theorem is a direct consequence of Lemma 2 of
+\citet{trosset2020rehabilitating}.
 
@@
 \label{thm:multidim}
-Let $X_1, ..., X_n \iid F$ with support $[0, 1]^r$, and $f(x) \geq a > 0$ everywhere on the support. 
+Let $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}F$ with support $[0, 1]^r$, and $f(x) \geq a > 0$ everywhere on the support. 
 Define $E_n$ as the event that an $\eta$-neighborhood graph constructed from the sample is connected. 
@@
 
-## Intersecting Manifolds
+\hypertarget{intersecting-manifolds}{%
+\subsection{Intersecting Manifolds}\label{intersecting-manifolds}}
 
@@
 
-# Examples
+\hypertarget{examples}{%
+\section{Examples}\label{examples}}
 
@@
 
-```{r, fig.height = 2, fig.cap = 'Latent positions, labeled by curve/community.'}
-f1 <- function(t) {
-  x1 <- t ^ 2
-  x2 <- 2 * t - 2 * t ^ 2
-  return(cbind(x1, x2))
+\begin{figure}[H]
+
+{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-2-1}
+
 }
 
-f2 <- function(t) {
-  x1 <- 2 * t - 2 * t ^ 2
-  x2 <- 1 - 2 * t + t ^ 2
-  return(cbind(x1, x2))
+\caption{Latent positions, labeled by curve/community.}\label{fig:unnamed-chunk-2}
+\end{figure}
+
+We draw $A \sim \mathrm{RDPG}(X)$ and obtain the following ASE:
+
+\begin{figure}[H]
+
+{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-3-1}
+
 }
 
-set.seed(314159)
-
-n1 <- 2 ** 8
-n2 <- n1
-n <- n1 + n2
-z <- c(rep(1, n1), rep(2, n2))
-
-a <- 1
-b <- 1
-t1 <- rbeta(n1, a, b)
-t2 <- rbeta(n2, a, b)
-T <- rbind(t1, t2)
-
-X1 <- f1(t1)
-X2 <- f2(t2)
-X <- rbind(X1, X2)
-
-ggplot() + 
-  geom_point(aes(x = X[, 1], y = X[, 2], colour = factor(z)), size = .5) + 
-  coord_fixed() + 
-  labs(x = expression(x[1]), y = expression(x[2]), colour = NULL) + 
-  theme_bw() + 
-  theme(legend.position = 'none')
-
-P <- X %*% t(X)
-diag(P) <- 0
-A <- draw.graph(P)
-Xhat <- embedding(A, 2, 0)
-```
-
-We draw $A \sim \RDPG(X)$ and obtain the following ASE:
-
-```{r, fig.height = 3, fig.cap = 'ASE of an RDPG drawn from the latent positions, labeled by curve/community.'}
-ggplot() + 
-  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2], colour = factor(z)), size = .5) + 
-  coord_fixed() + 
-  labs(x = expression(x[1]), y = expression(x[2]), colour = NULL) + 
-  theme_bw() + 
-  theme(legend.position = 'none')
-```
+\caption{ASE of an RDPG drawn from the latent positions, labeled by curve/community.}\label{fig:unnamed-chunk-3}
+\end{figure}
 
@@
 
-```{r, cache = TRUE}
-random.starts <- 4
-eps <- 1e-3
-out <- lapply(seq(random.starts), function(i) {
-  set.seed(i)
-  intercept <- (i / random.starts > .5 + eps)
-  manifold.clustering(Xhat, 2, parallel = TRUE, intercept = intercept, 
-                      maxit = 50,
-                      verbose = FALSE)
-})
-names(out) <- paste('random', seq(random.starts))
-names(out) <- sapply(seq(random.starts), function(i) {
-  if (i / random.starts < .5 + eps) {
-    paste('random', i, '\n(w/o intercept)')
-  } else {
-    paste('random', i, '\n(w/ intercept)')
+\begin{figure}[H]
+
+{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-5-1}
+
 }
-})
 
-# out[['random\n(w/ intercept)']] <- manifold.clustering(Xhat, 2,
-#                                           parallel = TRUE,
-#                                           intercept = TRUE, 
-#                                           maxit = 50, 
-#                                           initialization = 'random',
-#                                           verbose = FALSE)
+\caption{Clustering loss vs. iteration for each run of K-curve clustering.}\label{fig:unnamed-chunk-5}
+\end{figure}
 
-out[['spectral clust\n(w/o intercept)']] <- manifold.clustering(
-  Xhat, 2, 
-  parallel = TRUE, 
-  intercept = FALSE, 
-  maxit = 50, 
-  initialization = 'spectral',
-  verbose = FALSE)
+\begin{figure}[H]
 
-# out[['ground truth']] <- manifold.clustering(Xhat, 2,
-#                                             parallel = TRUE,
-#                                             intercept = FALSE,
-#                                             maxit = 50,
-#                                             initialization = z,
-#                                             verbose = FALSE)
+{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-6-1}
 
-out[['spectral clust\n(w/ intercept)']] <- manifold.clustering(
-  Xhat, 2,
-  parallel = TRUE,
-  intercept = TRUE,
-  maxit = 50,
-  initialization = 'spectral',
-  verbose = FALSE)
-```
+}
 
-```{r, fig.cap = 'Clustering loss vs. iteration for each run of K-curve clustering.', fig.height = 2, fig.width = 6}
-# max.iter <- max(sapply(out, function(x) x$niter))
-max.loss <- max(sapply(out, function(x) max(x$loss[-1])))
-# plot(out[[1]]$loss, xlim = c(0, max.iter), ylim = c(0, max.loss),
-#      ylab = 'clustering loss', xlab = 'iteration')
-# # lines(out[[1]]$loss)
-# 
-# for (i in seq_along(out)) {
-#   points(out[[i]]$loss, col = i)
-#   lines(out[[i]]$loss, col = i)
-# }
-# legend('topright', 
-#        title = 'initialization',
-#        cex = .75, 
-#        bty = 'n',
-#        legend = names(out),
-#        col = seq_along(out),
-#        pch = 1)
-
-loss.df <- plyr::ldply(seq_along(out), function(i) {
-  dplyr::data_frame(loss = out[[i]]$loss,
-                    initialization = factor(names(out),
-                                            levels = names(out))[i]) %>%
-    dplyr::mutate(iteration = seq(dplyr::n()))
-})
-
-ggplot(loss.df) + 
-  geom_point(aes(x = iteration, y = loss, colour = initialization)) + 
-  geom_line(aes(x = iteration, y = loss, colour = initialization)) + 
-  # ylim(0, max.loss) + 
-  labs(y = 'clustering loss')
-```
-
-```{r, fig.width = 6, fig.height = 5, fig.cap = 'ASE labeled by estimated community labels for each initialization strategy.'}
-plots.list <- lapply(seq_along(out), function(i) {
-  zhat.manifold <- out[[i]]$z
-
-  t. <- seq(0, 1, .01)
-  Y1 <- bezier.curve(t., out[[i]]$p[[1]])
-  Y2 <- bezier.curve(t., out[[i]]$p[[2]])
-
-  # Y1 <- out[[i]]$X[[1]][order(out[[i]]$t[[1]]), ]
-  # Y2 <- out[[i]]$X[[2]][order(out[[i]]$t[[2]]), ]
-
-  ggplot() + 
-    geom_point(aes(x = Xhat[, 1], y = Xhat[, 2], 
-                   colour = factor(zhat.manifold)),
-               size = .5) + 
-    labs(x = expression(x[1]), y = expression(x[2]), colour = NULL,
-         title = names(out)[i]) + 
-    theme_bw() + 
-    # coord_fixed() +
-    theme(legend.position = 'none') +
-    geom_path(aes(x = Y1[, 1], y = Y1[, 2])) +
-    geom_path(aes(x = Y2[, 1], y = Y2[, 2])) + 
-    xlim(min(Xhat[, 1]), max(Xhat[, 1])) + ylim(min(Xhat[, 2]), max(Xhat[, 2]))
-})
-
-gridExtra::grid.arrange(grobs = plots.list, ncol = 3)
-```
+\caption{ASE labeled by estimated community labels for each initialization strategy.}\label{fig:unnamed-chunk-6}
+\end{figure}
 
@@
 
-# Discussion
+\bibliographystyle{agsm}
+\bibliography{bibliography.bib}
+
+\end{document}

### 2022-02-16 functions.R

@@ -1,0 +1,456 @@
+laplacian.eigenmap <- function(W, d = 2, normalized = TRUE) {
+  n <- nrow(W)
+  if (normalized) {
+    L <- normalized.laplacian(W)
+  } else {
+    L <- graph.laplacian(W)
+  }
+  L.eigen <- eigen(L, symmetric = TRUE)
+  sweep(L.eigen$vectors[, seq(n - 1, n - d)], 2, 
+        sqrt(L.eigen$values[seq(n - 1, n - d)]), `/`)
+}
+
+graph.laplacian <- function(W) {
+  return(diag(colSums(W)) - W)
+}
+
+spectral.clustering <- function(W, K = 2, d = K, normalized = TRUE) {
+  Y <- laplacian.eigenmap(W, d, normalized)
+  kmeans(Y, K)$cluster
+}
+
+estimate.one.t.bezier <- function(x, p, 
+                                  min.t = 0, max.t = 2, 
+                                  intercept = TRUE, 
+                                  tol = 1e-9) {
+  polynomial <- function(t, p) {
+    if (intercept) {
+      r <- nrow(p) - 1
+      T <- sapply(seq(0, r), function(s) {
+        choose(r, s) * (1 - t) ^ (r - s) * t ^ s
+      })
+    } else {
+      r <- nrow(p)
+      T <- sapply(seq(r), function(s) {
+        choose(r, s) * (1 - t) ^ (r - s) * t ^ s
+      })
+    }
+    as.numeric(T %*% p)
+  }
+  l <- function(t) log(as.numeric(crossprod(polynomial(t, p) - x)))
+  minima <- optimize(l, c(min.t, max.t), tol = tol)$minimum
+  # if (abs(minima) == max.t) warning(minima)
+  # if (length(minima) > 1) warning(minima)
+  return(max(minima))
+}
+
+estimate.t.bezier <- function(X, p, t.prev, 
+                              intercept = TRUE, 
+                              min.t = 0, max.t = 2, 
+                              tol = 1e-9,
+                              parallel = FALSE) {
+  if (missing(t.prev)) {
+    t.hat <- plyr::laply(seq_len(nrow(X)), function(i) {
+      estimate.one.t.bezier(X[i, ], p, 
+                            min.t = min.t, max.t = max.t, 
+                            tol = tol, intercept = intercept)
+    }, .parallel = parallel)
+  } else {
[…]

### 2022-02-23 draft/draft.Rmd
update to draft

@@
 library(ggplot2)
+library(gganimate)
 source('../functions.R')
@@
 
-doMC::registerDoMC(8)
+doMC::registerDoMC(parallel::detectCores())
 ```
@@
 ```{r, cache = TRUE}
-random.starts <- 4
+maxit <- 50
+animate <- TRUE
+animation.dir <- '~/dev/manifold-block-models/draft'
+random.starts <- 2
 eps <- 1e-3
@@
   intercept <- (i / random.starts > .5 + eps)
-  manifold.clustering(Xhat, 2, parallel = TRUE, intercept = intercept, 
-                      maxit = 50,
-                      verbose = FALSE)
+  manifold.clustering(Xhat, 2, 
+                      parallel = TRUE, 
+                      intercept = intercept, 
+                      maxit = maxit,
+                      eps = 1e-3,
+                      verbose = FALSE,
+                      animate = animate,
+                      animation.dir = animation.dir,
+                      animation.title = i)
 })
@@
 
-# out[['random\n(w/ intercept)']] <- manifold.clustering(Xhat, 2,
-#                                           parallel = TRUE,
-#                                           intercept = TRUE, 
-#                                           maxit = 50, 
-#                                           initialization = 'random',
-#                                           verbose = FALSE)
-
 out[['spectral clust\n(w/o intercept)']] <- manifold.clustering(
@@
   intercept = FALSE, 
-  maxit = 50, 
+  maxit = maxit, 
+  eps = 1e-3,
   initialization = 'spectral',
-  verbose = FALSE)
-
-# out[['ground truth']] <- manifold.clustering(Xhat, 2,
-#                                             parallel = TRUE,
-#                                             intercept = FALSE,
-#                                             maxit = 50,
-#                                             initialization = z,
-#                                             verbose = FALSE)
+  verbose = FALSE,
+  animate = animate,
+  animation.dir = animation.dir,
+  animation.title = 'spec-clust-no-intercept')
 
@@
   intercept = TRUE,
-  maxit = 50,
+  maxit = maxit,
+  eps = 1e-3,
   initialization = 'spectral',
-  verbose = FALSE)
+  verbose = FALSE,
+  animate = animate,
+  animation.dir = animation.dir,
+  animation.title = 'spec-clust-w-intercept')
+
+out[['ground truth\n(w/o intercept)']] <- manifold.clustering(
+  Xhat, 2,
+  parallel = TRUE,
+  intercept = FALSE,
+  maxit = maxit,
+  eps = 1e-3,
+  initialization = z,
+  verbose = FALSE,
+  animate = animate,
+  animation.dir = animation.dir,
+  animation.title = 'groundtruth-no-intercept')
+
+out[['ground truth\n(w/ intercept)']] <- manifold.clustering(
+  Xhat, 2,
+  parallel = TRUE,
+  intercept = TRUE,
+  maxit = maxit,
+  eps = 1e-3,
+  initialization = z,
+  verbose = FALSE,
+  animate = animate,
+  animation.dir = animation.dir,
+  animation.title = 'groundtruth-w-intercept')
 ```
@@
 loss.df <- plyr::ldply(seq_along(out), function(i) {
-  dplyr::data_frame(loss = out[[i]]$loss,
+  dplyr::tibble(loss = out[[i]]$loss,
                 initialization = factor(names(out),
@@
   # ylim(0, max.loss) + 
+  # scale_y_log10() + 
   labs(y = 'clustering loss')
@@
 
+\begin{example}
+
+\end{example}
+
+# Simulation Study
+
 # Discussion

### 2022-02-23 draft/draft.Rmd
another animation

@@
 maxit <- 50
-animate <- TRUE
+# animate <- TRUE
 animation.dir <- '~/dev/manifold-block-models/draft'

### 2022-02-28 functions.R
fixed animations

@@ -22,7 +22,8 @@ spectral.clustering <- function(W, K = 2, d = K, normalized = TRUE) {
 estimate.one.t.bezier <- function(x, p, 
                                   min.t = 0, max.t = 2, 
                                   intercept = TRUE, 
-                                  tol = 1e-9) {
+                                  tol = 1e-9,
+                                  method = 'polyroot') {
   polynomial <- function(t, p) {
     if (intercept) {
       r <- nrow(p) - 1
@@ -38,7 +39,34 @@ estimate.one.t.bezier <- function(x, p,
     as.numeric(T %*% p)
   }
   l <- function(t) log(as.numeric(crossprod(polynomial(t, p) - x)))
-  minima <- optimize(l, c(min.t, max.t), tol = tol)$minimum
+  if (method == 'optimize') {
+    minima <- optimize(l, c(min.t, max.t), tol = tol)$minimum
+  } else if (method == 'optim') {
+    minima <- optim((min.t + max.t) / 2, l, method = 'SANN')$par
+  } else if (method == 'polyroot') {
+    if (!intercept) {
+      p <- rbind(0, p)
+    }
+    p0 <- p[1, ]
+    p1 <- p[2, ]
+    p2 <- p[3, ]
+    b0 <- sum((x - p0) * (p1 - p0))
+    b1 <- sum((x - p0) * (p0 - 2 * p1 + p2) - 2 * (p1 - p0) ^ 2)
+    b2 <- -3 * sum((p1 - p0) * (p0 - 2 * p1 + p2))
+    b3 <- -sum((p0 - 2 * p1 + p2) ^ 2)
+    roots <- polyroot(c(b0, b1, b2, b3))
+    roots <- Re(roots[abs(Im(roots)) < tol])
+    if (length(roots) > 1) {
+      minima <- roots[which.min(sapply(roots, l))]
+    } else if (length(roots) == 1) {
+      minima <- roots
+    } else if (length(roots) == 0) {
+      warning('failed to find a real root')
+      minima <- optimize(l, c(min.t, max.t), tol = tol)$minimum
+    }
+  } else {
+    stop('invalid method for finding t')
+  }
   # if (abs(minima) == max.t) warning(minima)
   # if (length(minima) > 1) warning(minima)
   return(max(minima))
@@ -294,9 +322,12 @@ manifold.clustering <- function(X, K = 2,
                                 maxit = 50,
                                 k.isomap = as.integer(sqrt(nrow(X)) / K),
[…]

### 2022-03-02 draft/draft.Rmd
updated draft

--- 2022-02-23 draft/draft.Rmd
@@
 maxit <- 50
-animate <- TRUE
+animate <- FALSE
 animation.dir <- '~/dev/manifold-block-models/draft'
@@
 
-\begin{example}
+\begin{example}[Macaque visuotactile brain areas and connections]
+
+```{r}
+data(macaque, package = 'igraphdata')
+A <- as.matrix(igraph::as_adjacency_matrix(macaque, type = 'both'))
+A <- sign(A + t(A))
+z <- as.numeric(factor(igraph::vertex_attr(macaque)$shape))
+
+Xhat <- embedding(A, 2, 0)
+plot(Xhat, col = z, asp = 1)
+```
+
+```{r cache = TRUE}
+set.seed(123)
+z.spectral <- spectral.clustering(A)
+macaque.out <- manifold.clustering(Xhat, 2, 
+                                   initialization = z.spectral, 
+                                   intercept = TRUE, 
+                                   parallel = TRUE, 
+                                   verbose = FALSE, 
+                                   animate = animate, 
+                                   curve.init = 'random',
+                                   animation.title = 'macaque')
+```
+
+```{r}
+t.1 <- seq(min(macaque.out$t[[1]]), max(macaque.out$t[[1]]), .01)
+t.2 <- seq(min(macaque.out$t[[2]]), max(macaque.out$t[[2]]), .01)
+Y1 <- bezier.curve(t.1, macaque.out$p[[1]])
+Y2 <- bezier.curve(t.2, macaque.out$p[[2]])
+
+ggplot() + 
+  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2], 
+                 colour = factor(macaque.out$z)),
+             size = .5) + 
+  labs(x = expression(x[1]), y = expression(x[2]), colour = NULL) + 
+  theme_bw() + 
+  coord_fixed() +
+  theme(legend.position = 'none') +
+  geom_path(aes(x = Y1[, 1], y = Y1[, 2])) +
+  geom_path(aes(x = Y2[, 1], y = Y2[, 2])) + 
+  xlim(min(Xhat[, 1]), max(Xhat[, 1])) + ylim(min(Xhat[, 2]), max(Xhat[, 2]))
+```
+
+\end{example}
+
+\begin{example}[Non-intersecting curves]
+
+```{r}
+n1 <- 2 ** 7
+n2 <- 2 ** 7
+n <- n1 + n2
+z <- c(rep(1, n1), rep(2, n2))
+u1 <- runif(n1)
+u2 <- runif(n2)
+x1 <- cos(pi / 3 * u1)
+y1 <- sin(pi / 3 * u1)
+x2 <- 1 + cos(pi / 3 * u2 + pi)
+y2 <- 1 + sin(pi / 3 * u2 + pi)
+data.matrix <- cbind(c(x1, x2), c(y1, y2))
+
+plot(data.matrix, col = z * 2, asp = 1, xlab = NA, ylab = NA,
+     main = 'Latent configuration')
+```
+
+```{r}
+P <- data.matrix %*% diag(c(1, 1)) %*% t(data.matrix)
+A <- draw.graph(P)
+Xhat <- embedding(A, 2, 0)
+plot(Xhat, col = z * 2, asp = 1, main = 'ASE')
+```
+
+```{r cache = TRUE}
+z.init <- hclust(dist(Xhat), method = 'single') %>% 
+  cutree(2)
+nonintersect.out <- manifold.clustering(Xhat, 2, 
+                                        initialization = z.init,
+                                        intercept = TRUE, 
+                                        animate = animate,
+                                        animation.title = 'nonintersect')
+```
+
+```{r}
+t.1 <- seq(min(nonintersect.out$t[[1]]), max(nonintersect.out$t[[1]]), .01)
+t.2 <- seq(min(nonintersect.out$t[[2]]), max(nonintersect.out$t[[2]]), .01)
+Y1 <- bezier.curve(t.1, nonintersect.out$p[[1]])
+Y2 <- bezier.curve(t.2, nonintersect.out$p[[2]])
+
+ggplot() + 
+  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2], 
+                 colour = factor(nonintersect.out$z)),
+             size = .5) + 
+  labs(x = expression(x[1]), y = expression(x[2]), colour = NULL) + 
+  theme_bw() + 
+  coord_fixed() +
+  theme(legend.position = 'none') +
+  geom_path(aes(x = Y1[, 1], y = Y1[, 2])) +
+  geom_path(aes(x = Y2[, 1], y = Y2[, 2])) + 
+  xlim(min(Xhat[, 1]), max(Xhat[, 1])) + ylim(min(Xhat[, 2]), max(Xhat[, 2]))
+```
 

### 2022-03-02 draft/draft.tex
update to draft
updated draft

@@
 
+\begin{example}[Macaque visuotactile brain areas and connections]
+
+\begin{center}\includegraphics{draft_files/figure-latex/unnamed-chunk-7-1} \end{center}
+
+\begin{center}\includegraphics{draft_files/figure-latex/unnamed-chunk-9-1} \end{center}
+
+\end{example}
+
+\begin{example}[Non-intersecting curves]
+
+\begin{center}\includegraphics{draft_files/figure-latex/unnamed-chunk-10-1} \end{center}
+
+\begin{center}\includegraphics{draft_files/figure-latex/unnamed-chunk-11-1} \end{center}
+
+\begin{center}\includegraphics{draft_files/figure-latex/unnamed-chunk-13-1} \end{center}
+
+\end{example}
+
+\hypertarget{simulation-study}{%
+\section{Simulation Study}\label{simulation-study}}
+
 \bibliographystyle{agsm}

### 2022-03-02 functions.R

@@ -167,12 +167,12 @@ estimate.bezier.curve.2 <- function(X,
         p <- matrix(rnorm(d * r), nrow = r, ncol = d)
       }
       t.hat <- estimate.t.bezier(X, p, intercept = intercept, parallel = parallel)
-      if (!intercept) {
-        if (sum((c(0, 0) %*% p) ^ 2) > sum((c(1, 1) %*% p) ^ 2)) {
-        # if (sum(X[which.min(t.hat), ] ^ 2) > sum(X[which.max(t.hat), ] ^ 2)) {
-          t.hat <- 1 - t.hat
-        }
-      }
+      # if (!intercept) {
+      #   if (sum((c(0, 0) %*% p) ^ 2) > sum((c(1, 1) %*% p) ^ 2)) {
+      #   # if (sum(X[which.min(t.hat), ] ^ 2) > sum(X[which.max(t.hat), ] ^ 2)) {
+      #     t.hat <- 1 - t.hat
+      #   }
+      # }
       
     } else if (initialization == 'isomap') {
       isomap.out <- estimate.bezier.curve.isomap(X, k.isomap, weights, 
@@ -188,9 +188,6 @@ estimate.bezier.curve.2 <- function(X,
     # if (!intercept) {
     #   if (sum(X[which.min(t.hat), ] ^ 2) > sum(X[which.max(t.hat), ] ^ 2)) {
     #     t.hat <- 1 - t.hat
-    #     T <- construct.bezier.model.matrix(t.hat, r, intercept = intercept)
-    #     # p <- solve(t(T) %*% W %*% T, t(T) %*% W %*% X)
-    #     # t.hat <- estimate.t.bezier(X, p, intercept = intercept, parallel = parallel)
     #   }
     # }
   }
@@ -256,6 +253,8 @@ estimate.bezier.curve.2 <- function(X,
     }
   }
   
+  # T <- construct.bezier.model.matrix(t.hat, r, intercept = intercept)
+  # p <- solve(t(T) %*% W %*% T, t(T) %*% W %*% X)
   X.hat <- T %*% p
   
   return(list(p = p, 
@@ -289,12 +288,14 @@ estimate.bezier.curve.isomap <- function(X, k = as.integer(sqrt(nrow(X))),
   D <- knn.graph(X, k)
   
   t.hat <- as.vector(cmdscale(sqrt(D), 1))
+  
[…]

### 2022-03-05 draft/draft.Rmd
fixed a few bugs

@@
 out <- lapply(seq(random.starts), function(i) {
+  intercept <- (i / random.starts > .5 + eps)
   set.seed(i)
-  intercept <- (i / random.starts > .5 + eps)
   manifold.clustering(Xhat, 2, 
@@
 
-  t. <- seq(0, 1, .01)
-  Y1 <- bezier.curve(t., out[[i]]$p[[1]])
-  Y2 <- bezier.curve(t., out[[i]]$p[[2]])
+  t.1 <- seq(min(out[[i]]$t[[1]]), max(out[[i]]$t[[1]]), .01)
+  t.2 <- seq(min(out[[i]]$t[[2]]), max(out[[i]]$t[[2]]), .01)
+  Y1 <- bezier.curve(t.1, out[[i]]$p[[1]])
+  Y2 <- bezier.curve(t.2, out[[i]]$p[[2]])
 
@@
 ```{r cache = TRUE}
-set.seed(123)
+# set.seed(123)
 z.spectral <- spectral.clustering(A)
 macaque.out <- manifold.clustering(Xhat, 2, 
-                                   initialization = z.spectral, 
-                                   intercept = TRUE, 
+                                   # initialization = 'random',
+                                   # initialization = z.spectral, 
+                                   initialization = z,
+                                   intercept = FALSE, 
                                    parallel = TRUE, 
@@
                                    animate = animate, 
-                                   curve.init = 'random',
+                                   curve.init = 'x',
                                    animation.title = 'macaque')
@@
 ```{r}
-t.1 <- seq(min(macaque.out$t[[1]]), max(macaque.out$t[[1]]), .01)
-t.2 <- seq(min(macaque.out$t[[2]]), max(macaque.out$t[[2]]), .01)
+t.1 <- seq(min(c(0, macaque.out$t[[1]])), max(macaque.out$t[[1]]), .01)
+t.2 <- seq(min(c(0, macaque.out$t[[2]])), max(macaque.out$t[[2]]), .01)
 Y1 <- bezier.curve(t.1, macaque.out$p[[1]])
@@
   geom_path(aes(x = Y1[, 1], y = Y1[, 2])) +
-  geom_path(aes(x = Y2[, 1], y = Y2[, 2])) + 
-  xlim(min(Xhat[, 1]), max(Xhat[, 1])) + ylim(min(Xhat[, 2]), max(Xhat[, 2]))
+  geom_path(aes(x = Y2[, 1], y = Y2[, 2]))
 ```
@@
 
-plot(data.matrix, col = z * 2, asp = 1, xlab = NA, ylab = NA,
-     main = 'Latent configuration')
+plot(data.matrix, col = z * 2, asp = 1, xlab = NA, ylab = NA)
 ```
@@
 Xhat <- embedding(A, 2, 0)
-plot(Xhat, col = z * 2, asp = 1, main = 'ASE')
+# plot(Xhat, col = z * 2, asp = 1)
 ```

### 2022-03-05 functions.R

@@ -20,11 +20,11 @@ spectral.clustering <- function(W, K = 2, d = K, normalized = TRUE) {
 }
 
 estimate.one.t.bezier <- function(x, p, 
-                                  min.t = 0, max.t = 2, 
+                                  min.t = 0, max.t = 1, 
                                   intercept = TRUE, 
                                   tol = 1e-9,
                                   method = 'polyroot') {
-  polynomial <- function(t, p) {
+  polynomial <- function(t, p, intercept) {
     if (intercept) {
       r <- nrow(p) - 1
       T <- sapply(seq(0, r), function(s) {
@@ -38,11 +38,19 @@ estimate.one.t.bezier <- function(x, p,
     }
     as.numeric(T %*% p)
   }
-  l <- function(t) log(as.numeric(crossprod(polynomial(t, p) - x)))
+  l <- function(t, intercept) {
+    log(as.numeric(crossprod(polynomial(t, p, intercept) - x)))
+  }
   if (method == 'optimize') {
-    minima <- optimize(l, c(min.t, max.t), tol = tol)$minimum
+    minima <- optimize(l, 
+                       c(min.t, max.t),
+                       intercept = intercept, 
+                       tol = tol)$minimum
   } else if (method == 'optim') {
-    minima <- optim((min.t + max.t) / 2, l, method = 'SANN')$par
+    minima <- optim(par = (min.t + max.t) / 2, 
+                    fn = l, 
+                    intercept = intercept, 
+                    method = 'SANN')$par
   } else if (method == 'polyroot') {
     if (!intercept) {
       p <- rbind(0, p)
@@ -56,13 +64,18 @@ estimate.one.t.bezier <- function(x, p,
     b3 <- -sum((p0 - 2 * p1 + p2) ^ 2)
     roots <- polyroot(c(b0, b1, b2, b3))
     roots <- Re(roots[abs(Im(roots)) < tol])
+    roots <- roots[roots >= min.t]
+    roots <- roots[roots <= max.t]
+    roots <- c(roots, min.t, max.t)
     if (length(roots) > 1) {
-      minima <- roots[which.min(sapply(roots, l))]
+      minima <- roots[which.min(sapply(roots, l, intercept = TRUE))]
     } else if (length(roots) == 1) {
       minima <- roots
     } else if (length(roots) == 0) {
       warning('failed to find a real root')
-      minima <- optimize(l, c(min.t, max.t), tol = tol)$minimum
[…]

### 2022-03-08 draft/draft.Rmd
fixed a few bugs

@@
 
+```{r}
+animate <- FALSE
+```
+
 ```{r, cache = TRUE}
 maxit <- 50
-animate <- FALSE
 animation.dir <- '~/dev/manifold-block-models/draft'
 random.starts <- 2
-eps <- 1e-3
+eps <- 1e-6
+normalize <- TRUE
+
 out <- lapply(seq(random.starts), function(i) {
@@
                       maxit = maxit,
+                      normalize = normalize, 
                       eps = 1e-3,
@@
                       animation.dir = animation.dir,
-                      animation.title = i)
+                      animation.title = paste('random', i, ifelse(intercept,
+                                                                  'w-intercept',
+                                                                  'no-intercept'),
+                                              sep = '-'))
 })
@@
   intercept = FALSE, 
+  normalize = normalize, 
   maxit = maxit, 
@@
   intercept = TRUE,
+  normalize = normalize, 
   maxit = maxit,
@@
   intercept = FALSE,
+  normalize = normalize, 
   maxit = maxit,
@@
 ```{r cache = TRUE}
-# set.seed(123)
+set.seed(123)
 z.spectral <- spectral.clustering(A)
@@
                                    # initialization = 'random',
-                                   # initialization = z.spectral, 
-                                   initialization = z,
+                                   initialization = z.spectral,
+                                   # initialization = z,
                                    intercept = FALSE, 
@@
   geom_path(aes(x = Y2[, 1], y = Y2[, 2]))
+```
+
+```{r}
+plot(macaque.out$loss, type = 'l')
 ```
@@
 nonintersect.out <- manifold.clustering(Xhat, 2, 
-                                        initialization = z.init,
+                                        # initialization = z.init,
+                                        initialization = 'random', 
                                         intercept = TRUE, 

### 2022-03-08 functions.R

@@ -163,6 +163,8 @@ estimate.bezier.curve.2 <- function(X,
                                     normalize.method = 'umvue',
                                     k.isomap = as.integer(sqrt(nrow(X))),
                                     intercept = TRUE,
+                                    min.t = 0, 
+                                    max.t = 1,
                                     parallel = TRUE) {
   n <- nrow(X)
   d <- ncol(X)
@@ -221,7 +223,7 @@ estimate.bezier.curve.2 <- function(X,
   niter <- 0
   while (TRUE) {
     mse.prev <- mse
-
+    
     # if (!intercept) {
     #   if (sum((c(0, 0) %*% p) ^ 2) > sum((c(1, 1) %*% p) ^ 2)) {
     #   # if (sum(X[which.min(t.hat), ] ^ 2) > sum(X[which.max(t.hat), ] ^ 2)) {
@@ -248,6 +250,8 @@ estimate.bezier.curve.2 <- function(X,
     
     t.hat <- estimate.t.bezier(X, p, t.hat, 
                                intercept = intercept, 
+                               min.t = min.t,
+                               max.t = max.t, 
                                parallel = parallel)
     
     mse <- bezier.mse(X, t.hat, p, intercept = intercept)
@@ -337,6 +341,9 @@ manifold.clustering <- function(X, K = 2,
                                 maxit = 50,
                                 k.isomap = as.integer(sqrt(nrow(X)) / K),
                                 curve.init = 'isomap',
+                                min.t = 0, 
+                                max.t = 1, 
+                                normalize = FALSE, 
                                 eps = 1e-3,
                                 parallel = TRUE,
                                 verbose = FALSE,
@@ -344,6 +351,7 @@ manifold.clustering <- function(X, K = 2,
                                 animation.dir = '.',
                                 animation.title = 'plot') {
   n <- nrow(X)
+  r <- 2
   
   if (length(initialization) == n) {
     z.hat <- initialization
@@ -370,59 +378,57 @@ manifold.clustering <- function(X, K = 2,
                                          Y = numeric(0),
                                          iter = numeric(0),
                                          type = character(0))
+    curves.df <- data.frame(X = numeric(0),
+                            Y = numeric(0),
+                            Z = integer(0),
+                            iter = numeric(0),
+                            type = character(0))
+    
+    t.seq <- seq(min.t, max.t, length.out = 100)
   }
   
   while (TRUE) {
     z.hat.prev <- z.hat
[…]

### 2022-03-09 draft/draft.Rmd
still working on code for higher order polynomials

@@
   Xhat, 2, 
+  A = A, 
   parallel = TRUE, 
@@
   Xhat, 2,
+  A = A, 
   parallel = TRUE,
@@
 plots.list <- lapply(seq_along(out), function(i) {
-  zhat.manifold <- out[[i]]$z
-
-  t.1 <- seq(min(out[[i]]$t[[1]]), max(out[[i]]$t[[1]]), .01)
-  t.2 <- seq(min(out[[i]]$t[[2]]), max(out[[i]]$t[[2]]), .01)
-  Y1 <- bezier.curve(t.1, out[[i]]$p[[1]])
-  Y2 <- bezier.curve(t.2, out[[i]]$p[[2]])
-
-  # Y1 <- out[[i]]$X[[1]][order(out[[i]]$t[[1]]), ]
-  # Y2 <- out[[i]]$X[[2]][order(out[[i]]$t[[2]]), ]
-
-  ggplot() + 
-    geom_point(aes(x = Xhat[, 1], y = Xhat[, 2], 
-                   colour = factor(zhat.manifold)),
-               size = .5) + 
-    labs(x = expression(x[1]), y = expression(x[2]), colour = NULL,
-         title = names(out)[i]) + 
-    theme_bw() + 
-    # coord_fixed() +
-    theme(legend.position = 'none') +
-    geom_path(aes(x = Y1[, 1], y = Y1[, 2])) +
-    geom_path(aes(x = Y2[, 1], y = Y2[, 2])) + 
-    xlim(min(Xhat[, 1]), max(Xhat[, 1])) + ylim(min(Xhat[, 2]), max(Xhat[, 2]))
+  plot.estimated.curves(Xhat, out[[i]])
 })
@@
 \begin{example}[Macaque visuotactile brain areas and connections]
+
+\cite{https://doi.org/10.1111/j.1460-9568.2006.04678.x}
 
@@
 ```{r}
-t.1 <- seq(min(c(0, macaque.out$t[[1]])), max(macaque.out$t[[1]]), .01)
-t.2 <- seq(min(c(0, macaque.out$t[[2]])), max(macaque.out$t[[2]]), .01)
-Y1 <- bezier.curve(t.1, macaque.out$p[[1]])
-Y2 <- bezier.curve(t.2, macaque.out$p[[2]])
-
-ggplot() + 
-  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2], 
-                 colour = factor(macaque.out$z)),
-             size = .5) + 
-  labs(x = expression(x[1]), y = expression(x[2]), colour = NULL) + 
-  theme_bw() + 
-  coord_fixed() +
-  theme(legend.position = 'none') +
-  geom_path(aes(x = Y1[, 1], y = Y1[, 2])) +
-  geom_path(aes(x = Y2[, 1], y = Y2[, 2]))
+plot.estimated.curves(Xhat, macaque.out)
 ```
@@
 nonintersect.out <- manifold.clustering(Xhat, 2, 
-                                        # initialization = z.init,
-                                        initialization = 'random', 
+                                        initialization = z.init,
+                                        # initialization = 'random', 
                                         intercept = TRUE, 
@@
 ```{r}
-t.1 <- seq(min(nonintersect.out$t[[1]]), max(nonintersect.out$t[[1]]), .01)
-t.2 <- seq(min(nonintersect.out$t[[2]]), max(nonintersect.out$t[[2]]), .01)
-Y1 <- bezier.curve(t.1, nonintersect.out$p[[1]])
-Y2 <- bezier.curve(t.2, nonintersect.out$p[[2]])
-
-ggplot() + 
-  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2], 
-                 colour = factor(nonintersect.out$z)),
-             size = .5) + 
-  labs(x = expression(x[1]), y = expression(x[2]), colour = NULL) + 
-  theme_bw() + 
-  coord_fixed() +
-  theme(legend.position = 'none') +
-  geom_path(aes(x = Y1[, 1], y = Y1[, 2])) +
-  geom_path(aes(x = Y2[, 1], y = Y2[, 2]))
+plot.estimated.curves(Xhat, nonintersect.out)
 ```

### 2022-03-09 draft/draft.tex
fixed a few bugs
still working on code for higher order polynomials

@@
 
-{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-5-1}
+{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-6-1}
 
@@
 
-\caption{Clustering loss vs. iteration for each run of K-curve clustering.}\label{fig:unnamed-chunk-5}
+\caption{Clustering loss vs. iteration for each run of K-curve clustering.}\label{fig:unnamed-chunk-6}
 \end{figure}
@@
 
-{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-6-1}
+{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-7-1}
 
@@
 
-\caption{ASE labeled by estimated community labels for each initialization strategy.}\label{fig:unnamed-chunk-6}
+\caption{ASE labeled by estimated community labels for each initialization strategy.}\label{fig:unnamed-chunk-7}
 \end{figure}
@@
 
-\begin{center}\includegraphics{draft_files/figure-latex/unnamed-chunk-7-1} \end{center}
+\cite{https://doi.org/10.1111/j.1460-9568.2006.04678.x}
 
-\begin{center}\includegraphics{draft_files/figure-latex/unnamed-chunk-9-1} \end{center}
+\begin{center}\includegraphics{draft_files/figure-latex/unnamed-chunk-8-1} \end{center}
+
+\begin{center}\includegraphics{draft_files/figure-latex/unnamed-chunk-10-1} \end{center}
+
+\begin{center}\includegraphics{draft_files/figure-latex/unnamed-chunk-11-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{draft_files/figure-latex/unnamed-chunk-10-1} \end{center}
+\begin{center}\includegraphics{draft_files/figure-latex/unnamed-chunk-12-1} \end{center}
 
-\begin{center}\includegraphics{draft_files/figure-latex/unnamed-chunk-11-1} \end{center}
-
-\begin{center}\includegraphics{draft_files/figure-latex/unnamed-chunk-13-1} \end{center}
+\begin{center}\includegraphics{draft_files/figure-latex/unnamed-chunk-15-1} \end{center}
 

### 2022-03-09 functions.R

@@ -73,7 +73,6 @@ estimate.one.t.bezier <- function(x, p,
       minima <- roots
     } else if (length(roots) == 0) {
       warning('failed to find a real root')
-      # minima <- optimize(l, c(min.t, max.t), intercept = FALSE, tol = tol)$minimum
       minima <- c(min.t, max.t)
       minima <- minima[which.min(sapply(minima, l, intercept = TRUE))]
     }
@@ -83,6 +82,48 @@ estimate.one.t.bezier <- function(x, p,
   # if (abs(minima) == max.t) warning(minima)
   # if (length(minima) > 1) warning(minima)
   return(max(minima))
+}
+
+bezier.coefs.to.d.polynomial.coefs <- function(p, x, intercept = TRUE) {
+  if (!intercept) {
+    p <- rbind(0, p)
+  }
+  
+  degree <- nrow(p) - 1
+  
+  c. <- sapply(seq(0, degree), function(R) {
+    if (R == 0) {
+      x - p[1, ]
+    } else {
+      sapply(seq(0, R), function(r) {
+        (-1) ^ (R - r) * choose(R, r) * p[r + 1, ]
+      }) %>% 
+        rowSums()
+    }
+  }) %>% 
+    t() %>% 
+    unname()
+  
+  c.orig <- sapply(seq(0, degree), function(r) {
+    c.[r + 1, ] * choose(degree, r) * (-1) ^ (degree - r - 1)
+  }) %>% 
+    t()
+  c.deriv <- sapply(seq(0, degree - 1), function(r) {
+    c.[r + 2, ] * choose(degree - 1, r) * (-1) ^ (degree - r - 1)
+  }) %>% 
+    t()
+  
+  cross <- c.orig %*% t(c.deriv) %>% 
+    rbind(0) %>% 
+    cbind(0) %>% 
+    cbind(0)
+  
+  sapply(seq(0, degree + 1), function(i) {
+    sapply(seq(0, i), function(j) {
+      cross[i - j + 1, j + 1] * (-1) ^ (j * i)
+    }) %>% 
+      sum()
+  })
 }
 
 estimate.t.bezier <- function(X, p, t.prev, 
@@ -157,7 +198,7 @@ estimate.bezier.curve.2 <- function(X,
                                     init.params, 
                                     weights,
                                     eps = 1e-3, 
-                                    maxit = 200,
+                                    maxit = 100,
                                     initialization = 'isomap', 
                                     normalize = FALSE, 
[…]

### 2022-03-16 draft/draft.Rmd
minor updates

@@
 
+\begin{algorithm}[h]
+\DontPrintSemicolon
+\SetAlgoLined
+\KwData{Adjacency matrix $A$, number of communities $K$, embedding dimensions $p$ and $q$.}
+\KwResult{Community assignments $1, ..., K$.}
+Compute $X$, the ASE of $A$ using the $p$ most positive and $q$ most negative eigenvalues and their corresponding eigenvectors.\;
+Apply single linkage clustering with $K$ communities on the ASE.\;
+\caption{ASE clustering for nonintersecting communities.}
+\end{algorithm}
+
 ## Intersecting Manifolds
@@
 Initialize community labels $z_1, ..., z_n$.\;
-\Repeat {$\sum_k \sum_{i \in C_k} \|x_i - g_k(t_i)\|^2 < \epsilon$} {
+\Repeat {the change in $\sum_k \sum_{i \in C_k} \|x_i - g_k(t_i)\|^2$ is less than $\epsilon$} {
 \For {$k = 1, ..., K$} {
@@
 }
-\caption{$K$-Curves Clustering.}
+\caption{$K$-curves clustering.}
 \end{algorithm}
@@
 \end{proof}
+
+\begin{algorithm}[h]
+\DontPrintSemicolon
+\SetAlgoLined
+\KwData{Adjacency matrix $A$, number of communities $K$, embedding dimensions $p$, $q$, stopping criterion $\epsilon$, $m_k \leq n_k$ known community assignments for each community}
+\KwResult{Community assignments $1, ..., K$, curves $g_1, ..., g_K$}
+Compute $X$, the ASE of $A$ using the $p$ most positive and $q$ most negative eigenvalues and their corresponding eigenvectors.\;
+Fit curves $g_1, ..., g_K$ using each of the $m_1, ..., m_K$ points with known community labels by minimizing $\sum_{j=1}^{m_i} \|x_j - g_k(t_j)\|^2$.\;
+Assign labels $z_1, ..., z_n$ to each $x_1, ..., x_n$ by minimizing $\|x_i - g_k(t_i)\|^2$ for $k$, holding the initial known labels constant.\; 
+\Repeat {the change in $\sum_k \sum_{i \in C_k} \|x_i - g_k(t_i)\|^2$ is less than $\epsilon$} {
+\For {$k = 1, ..., K$} {
+Define $X_k$ as the rows of $X$ for which $z_i = k$.\;
+Fit curve $g_k$ and positions $t_{k_i}$ to $X_k$ by minimizing $\sum_{k_i} \|x_{k_i} - g_k(t_{k_i})\|^2$.\;
+}
+\For {$k = 1, ..., K$} {
+Assign $z_i \leftarrow \arg\min_\ell \|x_i - g_\ell(t_i)\|^2$, holding the known initial labels constant.\
+}
+}
+\caption{Semi-supervised $K$-curves clustering.}
+\end{algorithm}
+
+\begin{theorem}
+Let each $g_k$ be smooth. 
+Then as $n \to \infty$, the proportion of misclustered vertices approaches $0$ with probability $1$.
+\end{theorem}
 

### 2022-03-16 draft/draft.tex
minor updates

@@
 
+\begin{algorithm}[h]
+\DontPrintSemicolon
+\SetAlgoLined
+\KwData{Adjacency matrix $A$, number of communities $K$, embedding dimensions $p$ and $q$.}
+\KwResult{Community assignments $1, ..., K$.}
+Compute $X$, the ASE of $A$ using the $p$ most positive and $q$ most negative eigenvalues and their corresponding eigenvectors.\;
+Apply single linkage clustering with $K$ communities on the ASE.\;
+\caption{ASE clustering for nonintersecting communities.}
+\end{algorithm}
+
 \hypertarget{intersecting-manifolds}{%

### 2022-03-16 functions.R

@@ -38,7 +38,7 @@ estimate.one.t.bezier <- function(x, p,
     }
     as.numeric(T %*% p)
   }
-  l <- function(t, intercept) {
+  l <- function(t, intercept = TRUE) {
     log(as.numeric(crossprod(polynomial(t, p, intercept) - x)))
   }
   if (method == 'optimize') {
@@ -55,14 +55,18 @@ estimate.one.t.bezier <- function(x, p,
     if (!intercept) {
       p <- rbind(0, p)
     }
-    p0 <- p[1, ]
-    p1 <- p[2, ]
-    p2 <- p[3, ]
-    b0 <- sum((x - p0) * (p1 - p0))
-    b1 <- sum((x - p0) * (p0 - 2 * p1 + p2) - 2 * (p1 - p0) ^ 2)
-    b2 <- -3 * sum((p1 - p0) * (p0 - 2 * p1 + p2))
-    b3 <- -sum((p0 - 2 * p1 + p2) ^ 2)
-    roots <- polyroot(c(b0, b1, b2, b3))
+    
+    # p0 <- p[1, ]
+    # p1 <- p[2, ]
+    # p2 <- p[3, ]
+    # b0 <- sum((x - p0) * (p1 - p0))
+    # b1 <- sum((x - p0) * (p0 - 2 * p1 + p2) - 2 * (p1 - p0) ^ 2)
+    # b2 <- -3 * sum((p1 - p0) * (p0 - 2 * p1 + p2))
+    # b3 <- -sum((p0 - 2 * p1 + p2) ^ 2)
+    # b <- c(b0, b1, b2, b3)
+    
+    b <- bezier.coefs.to.d.polynomial.coefs(p, x, TRUE)
+    roots <- polyroot(b)
     roots <- Re(roots[abs(Im(roots)) < tol])
     roots <- roots[roots >= min.t]
     roots <- roots[roots <= max.t]
@@ -93,10 +97,10 @@ bezier.coefs.to.d.polynomial.coefs <- function(p, x, intercept = TRUE) {
   
   c. <- sapply(seq(0, degree), function(R) {
     if (R == 0) {
-      x - p[1, ]
+      p[1, ] - x
     } else {
       sapply(seq(0, R), function(r) {
-        (-1) ^ (R - r) * choose(R, r) * p[r + 1, ]
+        (-1) ^ r * choose(R, r) * p[r + 1, ]
       }) %>% 
         rowSums()
     }
[…]

### 2022-04-13 draft/draft.Rmd
updated draft

@@
 - \usepackage{comment}
+- \usepackage{amsfonts}
 ---
@@
 \newcommand{\as}{\stackrel{\mathrm{a.s.}}{\to}}
+\newcommand{\ER}{\text{Erd\"{o}s-R\'{e}nyi}}
 
@@
 
+We define a *Bernoulli graph* as a random graph model for which edge probabilities are contained in an edge probability matrix $P \in [0, 1]^{n \times n}$, and an edge occurs between vertices $i$ and $j$ with probability $P_{ij}$.
+Common random graph models then impose structure on $P$, based on various assumptions about the way in which the data are generated, or to allow $P$ to be estimated. 
+One example is the $\ER$ model, in which all edge probabilities are fixed, i.e., $P_{ij} = p$.
+
+One common analysis task for graph and network data is community detection, which assumes that each vertex of a graph has a hidden community label.
+The goal of the analysis is then to retrieve these labels.
+In order to perform this analysis as a statistical inference task is to define a probability model with inherent community structure.
+We call such models *block models*:
+First, each vertex is assigned a label $z_1, ..., z_n \in \{1, 2, ..., K\}$ where $K \ll n$. 
+Then each edge probability $P_{ij}$ is said to depend on the labels $z_i$ and $z_j$, possibly along with some other parameters. 
+For example, the stochastic block model (SBM) sets a fixed edge probability for each pair of communities, i.e., $P_{ij} = \omega_{z_i, z_j}$. 
+The degree-corrected block model (DCBM) assigns an additional parameter $\theta_i$ to each vertex by which edge probabilities are scaled, i.e., $P_{ij} = \theta_i \theta_j \omega_{z_i, z_j}$. 
+The popularity adjusted block model (PABM) assigns $K$ parameters to each vertex $\lambda_{i1}, \lambda_{i2}, ..., \lambda_{iK}$ that describe that vertex's affinity toward each community; 
+the edge probability between vertices $i$ and $j$ is then defined as the product of vertex $i$'s affinity toward vertex $j$'s community and vertex $j$'s affinity toward vertex $i$'s community, i.e., $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$.
+
+The three block model types, as well as the $\ER$ model, impose structure on $P$, including on the rank of $P$. 
+$P$ has rank 1 for the $\ER$ model, rank $K$ for the SBM and DCBM, and rank $K^2$ for the PABM. 
+This provides the intuition behind another family of Bernoulli graphs called the *random dot product graph* (RDPG) and *generalized random dot product graph* (GRDPG). 
+In the RDPG, each vertex has a corresponding latent vector in $d$-dimensional Euclidean space, where $d$ is the rank of $P$ and $P$ is positive semidefinite.
+Then the edge probability between each pair of vertices is defined as the inner product between the corresponding latent vectors, i.e., $P_{ij} = x_i^\top x_j$.
+If the latent vectors are collected in a data matrix $X = \bigl[ x_1 \mid \cdots \mid x_n \bigr]^\top$, then the edge probability matrix for the RDPG is $P = X X^\top$.
+Similarly, the edge probability between each pair of vertices for the GRDPG is defined as the indefinite inner product between the corresponding latent vectors, i.e., $P_{ij} = x_i^\top I_{p,q} x_j$, where $I_{p,q} = \blockdiag(I_p, -I_q)$ and $p + q = d$.
+Then the edge probability matrix for the GRDPG is $P = X I_{p,q} X^\top$.
+This allows for a model similar to the RDPG for non-positive semidefinite $P$.
+While the RDPG and GRDPG do not necessarily have community structure, it has been shown that block models are specific cases of the RDPG or GRDPG in which latent vectors are organized by community.
+This includes the SBM, in which communities correspond to point masses, DCBM, in which communities correspond to line segments, and PABM, in which communities correspond to orthogonal subspaces.
+In this work, we extend this idea to communities organized into more general latent structures.
+In particular, we assume that each community corresponds to a manifold in the latent space.
+
 # Latent Structure Block Models
 
-All block models are latent structure block models.
+All block models are latent structure models.
+
+Insert something about ASE and its consistency here.
 
@@
 
-Here, we provide two algorithms for LSBM community detection.
+Here, we provide two algorithms for MBM community detection.
+
+First, we consider the case in which communities correspond to manifolds in the latent space that do not intersect and are separated by some finite distance.
+In this scenario, we use the convergence of the ASE to show that single linkage clustering on the latent space produces a clustering such that the total number of misclustered vertices goes to zero, with high probability.
+
+Next, we consider the case in which communities correspond to one-dimensional manifolds in the latent space and may or may not intersect.
+In this scenario, we propose an alternating coordinate descent algorithm that alternates between estimating the structure of the manifolds and the community labels, which we call $K$-curves clustering.
+We again use the convergence of the ASE to show that under certain conditions, $K$-curves clustering produces a clustering such that the proportion of misclustered vertices goes to zero, with high probability.
 
 ## Nonintersecting Manifolds
+
+In this section, we consider the following scenario: Suppose that each community is represented by a manifold $\mathcal{M}_k$, $k \in \{1, ..., K\}$ in the latent space of a RDPG or GDRPG, and the manifolds do not intersect each other.
+Define $\delta = \min\limits_{k, \ell} \min\limits_{x \in \mathcal{M}_k, y \in \mathcal{M}_\ell} \|x - y\|$, the minimum distance between two manifolds. 
+We assume that $\delta > 0$.
+
+First, we consider the case in which each $\mathcal{M}_k \subset \mathbb{R}^d$ is a one-dimensional manifold, which can be characterized by a function $g_k : [0, 1] \to \mathcal{M}_k$. 
+Let $F$ be a probability distribution with support $[0, 1]$. 
+Then we define a mixture model as follows:
+
+1. Draw $t_1, ..., t_n \iid F$.
+2. Draw $z_1, ..., z_n \iid \Multinomial(\alpha_1, ..., \alpha_K)$, the community labels.
+3. Let each $x_i = g_{z_i}(t_i)$ be the latent vector for vertex $v_i$.  
+Assume $\min_{k, \ell} \min_{t, s} \|g_k(t) - g_\ell(s)\| = \delta > 0$. 
+4. Draw $A \sim \RDPG(X)$ or $A \sim \GRDPG_{p,q}(X)$.
 
@@
 
+In this section, we again consider the setting for the RDPG or GRDPG in which each community lies on a manifold in the latent space. 
+However, this time, we do not assume that the manifolds are nonintersecting.
+
 \begin{algorithm}[h]
@@
 
+$K$-curves clustering assumes that the functional form of $g_k$ is known. 
+The choice of $g_k$ affects the difficulty of the algorithm. 
+As a balance between flexibility and ease of estimation, we consider the case where each $g_k$ is a Bezier polynomial of degree $R$ with coefficients $p_k$. 
+Then we have $g_k(t) = g(t; p_k) = \sum_{r=0}^R p_k^{(r)} \binom{R}{r} (1-t)^{R-r} t^r$.
+
+Given $\{t_i\}$ and $\{z_i\}$, it is straightforward to obtain $\hat{p}_k = \arg\min_p \sum_{k_i} \|x_{k_i} - g_k (t_{k_i}; p)\|^2$
+$$\hat{p}_k = (T_k^\top T_k)^{-1} T_k^\top X_k,$$
+where $T_k$ is an $n_k \times (R+1)$ matrix with rows $\begin{bmatrix} (1 - t_{k_i})^R & (1 - t_{k_i})^{R-1} t_{k_i} & \cdots & (1 - t_{k_i}) t_{k_i}^{R-1} & t_{k_i}^R \end{bmatrix}$.
+Estimation of $\{t_i\}$ given $\{z_i\}$ and $\{p_k\}$ is more difficult. 
+Each $t_i$ can be estimated separately: 
+\begin{equation} \label{eq:min-t}
+\hat{t}_i = \arg\min_t \|x_i - g(t; p_{z_i})\|^2. 
+\end{equation}
+This is equivalent to solving $0 = (x_i - g(t; p_{z_i}))^\top (\dot{g}(t; p_{z_i}))$. 
+Setting $c^{(s)} = \sum_{r=0}^s (-1)^{s-r} \binom{R}{r} p^{(r)}_{z_i}$ for $s \neq 0$ and $c^{(0)} = p^{(0)}_{z_i} - x_i$, let $c = \begin{bmatrix} c^{(0)} & \cdots & c^{(R)} \end{bmatrix}^\top$.
+Then solving Eq.
+\ref{eq:min-t} is equivalent to finding the real roots of a polynomial with coefficients that are the sums of the reverse diagonals of $C D^\top$, where $C_{ij} = c_{ij} (-1)^i \binom{R}{i}$ and $D_{ij} = c_{i-1,j} (-1)^{i-1} \binom{R-1}{i-1}$.
+
 \begin{algorithm}[h]
@@
 \begin{theorem}
-Let each $g_k$ be smooth. 
-Then as $n \to \infty$, the proportion of misclustered vertices approaches $0$ with probability $1$.
+Let each $g(\cdot; p_k)$ be a nonintersecting Bezier polynomial of order $R$, 
+and a GRDPG is drawn from vectors that lie on the curves. 
+Suppose we observe the true labels of $m_k$ vertices from each community, and each $m_k > R + 1$.
+Suppose further that latent vectors $x_j = g(t_i; p_{z_j})$ that correspond to vertices with observed labels are such that
+Then as $n \to \infty$, the proportion of misclustered vertices from $K$-curves clustering approaches $0$ with probability $1$.
 \end{theorem}

### 2022-04-13 draft/draft.tex
updated draft

@@
 \usepackage{comment}
+\usepackage{amsfonts}
 
@@
 \section{Introduction}\label{introduction}}
+
+We define a \emph{Bernoulli graph} as a random graph model for which
+edge probabilities are contained in an edge probability matrix
+\(P \in [0, 1]^{n \times n}\), and an edge occurs between vertices \(i\)
+and \(j\) with probability \(P_{ij}\). Common random graph models then
+impose structure on \(P\), based on various assumptions about the way in
+which the data are generated, or to allow \(P\) to be estimated. One
+example is the \(\text{Erd\"{o}s-R\'{e}nyi}\) model, in which all edge
+probabilities are fixed, i.e., \(P_{ij} = p\).
+
+One common analysis task for graph and network data is community
+detection, which assumes that each vertex of a graph has a hidden
+community label. The goal of the analysis is then to retrieve these
+labels. In order to perform this analysis as a statistical inference
+task is to define a probability model with inherent community structure.
+We call such models \emph{block models}: First, each vertex is assigned
+a label \(z_1, ..., z_n \in \{1, 2, ..., K\}\) where \(K \ll n\). Then
+each edge probability \(P_{ij}\) is said to depend on the labels \(z_i\)
+and \(z_j\), possibly along with some other parameters. For example, the
+stochastic block model (SBM) sets a fixed edge probability for each pair
+of communities, i.e., \(P_{ij} = \omega_{z_i, z_j}\). The
+degree-corrected block model (DCBM) assigns an additional parameter
+\(\theta_i\) to each vertex by which edge probabilities are scaled,
+i.e., \(P_{ij} = \theta_i \theta_j \omega_{z_i, z_j}\). The popularity
+adjusted block model (PABM) assigns \(K\) parameters to each vertex
+\(\lambda_{i1}, \lambda_{i2}, ..., \lambda_{iK}\) that describe that
+vertex's affinity toward each community; the edge probability between
+vertices \(i\) and \(j\) is then defined as the product of vertex
+\(i\)'s affinity toward vertex \(j\)'s community and vertex \(j\)'s
+affinity toward vertex \(i\)'s community, i.e.,
+\(P_{ij} = \lambda_{i z_j} \lambda_{j z_i}\).
+
+The three block model types, as well as the
+\(\text{Erd\"{o}s-R\'{e}nyi}\) model, impose structure on \(P\),
+including on the rank of \(P\). \(P\) has rank 1 for the
+\(\text{Erd\"{o}s-R\'{e}nyi}\) model, rank \(K\) for the SBM and DCBM,
+and rank \(K^2\) for the PABM. This provides the intuition behind
+another family of Bernoulli graphs called the \emph{random dot product
+graph} (RDPG) and \emph{generalized random dot product graph} (GRDPG).
+In the RDPG, each vertex has a corresponding latent vector in
+\(d\)-dimensional Euclidean space, where \(d\) is the rank of \(P\) and
+\(P\) is positive semidefinite. Then the edge probability between each
+pair of vertices is defined as the inner product between the
+corresponding latent vectors, i.e., \(P_{ij} = x_i^\top x_j\). If the
+latent vectors are collected in a data matrix
+\(X = \bigl[ x_1 \mid \cdots \mid x_n \bigr]^\top\), then the edge
+probability matrix for the RDPG is \(P = X X^\top\). Similarly, the edge
+probability between each pair of vertices for the GRDPG is defined as
+the indefinite inner product between the corresponding latent vectors,
+i.e., \(P_{ij} = x_i^\top I_{p,q} x_j\), where
+\(I_{p,q} = \mathrm{blockdiag}(I_p, -I_q)\) and \(p + q = d\). Then the
+edge probability matrix for the GRDPG is \(P = X I_{p,q} X^\top\). This
+allows for a model similar to the RDPG for non-positive semidefinite
+\(P\). While the RDPG and GRDPG do not necessarily have community
+structure, it has been shown that block models are specific cases of the
+RDPG or GRDPG in which latent vectors are organized by community. This
+includes the SBM, in which communities correspond to point masses, DCBM,
+in which communities correspond to line segments, and PABM, in which
+communities correspond to orthogonal subspaces. In this work, we extend
+this idea to communities organized into more general latent structures.
+In particular, we assume that each community corresponds to a manifold
+in the latent space.
 
@@
 
-All block models are latent structure block models.
+All block models are latent structure models.
+
+Insert something about ASE and its consistency here.
 
@@
 
-Here, we provide two algorithms for LSBM community detection.
+Here, we provide two algorithms for MBM community detection.
+
+First, we consider the case in which communities correspond to manifolds
+in the latent space that do not intersect and are separated by some
+finite distance. In this scenario, we use the convergence of the ASE to
+show that single linkage clustering on the latent space produces a
+clustering such that the total number of misclustered vertices goes to
+zero, with high probability.
+
+Next, we consider the case in which communities correspond to
+one-dimensional manifolds in the latent space and may or may not
+intersect. In this scenario, we propose an alternating coordinate
+descent algorithm that alternates between estimating the structure of
+the manifolds and the community labels, which we call \(K\)-curves
+clustering. We again use the convergence of the ASE to show that under
+certain conditions, \(K\)-curves clustering produces a clustering such
+that the proportion of misclustered vertices goes to zero, with high
+probability.
 
@@
 \subsection{Nonintersecting Manifolds}\label{nonintersecting-manifolds}}
+
+In this section, we consider the following scenario: Suppose that each
+community is represented by a manifold \(\mathcal{M}_k\),
+\(k \in \{1, ..., K\}\) in the latent space of a RDPG or GDRPG, and the
+manifolds do not intersect each other. Define
+\(\delta = \min\limits_{k, \ell} \min\limits_{x \in \mathcal{M}_k, y \in \mathcal{M}_\ell} \|x - y\|\),
+the minimum distance between two manifolds. We assume that
+\(\delta > 0\).
+
+First, we consider the case in which each
+\(\mathcal{M}_k \subset \mathbb{R}^d\) is a one-dimensional manifold,
+which can be characterized by a function
+\(g_k : [0, 1] \to \mathcal{M}_k\). Let \(F\) be a probability
+distribution with support \([0, 1]\). Then we define a mixture model as
+follows:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Draw \(t_1, ..., t_n \stackrel{\mathrm{iid}}{\sim}F\).
+\item
+  Draw
+  \(z_1, ..., z_n \stackrel{\mathrm{iid}}{\sim}\mathrm{Multinomial}(\alpha_1, ..., \alpha_K)\),
+  the community labels.
+\item
+  Let each \(x_i = g_{z_i}(t_i)\) be the latent vector for vertex
+  \(v_i\).\\
+  Assume
+  \(\min_{k, \ell} \min_{t, s} \|g_k(t) - g_\ell(s)\| = \delta > 0\).
+\item
+  Draw \(A \sim \mathrm{RDPG}(X)\) or
+  \(A \sim \mathrm{GRDPG}_{p,q}(X)\).
+\end{enumerate}
 
@@
 
+In this section, we again consider the setting for the RDPG or GRDPG in
+which each community lies on a manifold in the latent space. However,
+this time, we do not assume that the manifolds are nonintersecting.
+
 \begin{algorithm}[h]
@@
 
+\(K\)-curves clustering assumes that the functional form of \(g_k\) is
+known. The choice of \(g_k\) affects the difficulty of the algorithm. As
+a balance between flexibility and ease of estimation, we consider the
+case where each \(g_k\) is a Bezier polynomial of degree \(R\) with
+coefficients \(p_k\). Then we have
+\(g_k(t) = g(t; p_k) = \sum_{r=0}^R p_k^{(r)} \binom{R}{r} (1-t)^{R-r} t^r\).
+
+Given \(\{t_i\}\) and \(\{z_i\}\), it is straightforward to obtain
+\(\hat{p}_k = \arg\min_p \sum_{k_i} \|x_{k_i} - g_k (t_{k_i}; p)\|^2\)
+\[\hat{p}_k = (T_k^\top T_k)^{-1} T_k^\top X_k,\] where \(T_k\) is an
+\(n_k \times (R+1)\) matrix with rows
+\(\begin{bmatrix} (1 - t_{k_i})^R & (1 - t_{k_i})^{R-1} t_{k_i} & \cdots & (1 - t_{k_i}) t_{k_i}^{R-1} & t_{k_i}^R \end{bmatrix}\).
+Estimation of \(\{t_i\}\) given \(\{z_i\}\) and \(\{p_k\}\) is more
+difficult. Each \(t_i\) can be estimated separately:
+\begin{equation} \label{eq:min-t}
+\hat{t}_i = \arg\min_t \|x_i - g(t; p_{z_i})\|^2. 
+\end{equation} This is equivalent to solving
+\(0 = (x_i - g(t; p_{z_i}))^\top (\dot{g}(t; p_{z_i}))\). Setting
+\(c^{(s)} = \sum_{r=0}^s (-1)^{s-r} \binom{R}{r} p^{(r)}_{z_i}\) for
+\(s \neq 0\) and \(c^{(0)} = p^{(0)}_{z_i} - x_i\), let
+\(c = \begin{bmatrix} c^{(0)} & \cdots & c^{(R)} \end{bmatrix}^\top\).
+Then solving Eq. \ref{eq:min-t} is equivalent to finding the real roots
+of a polynomial with coefficients that are the sums of the reverse
+diagonals of \(C D^\top\), where \(C_{ij} = c_{ij} (-1)^i \binom{R}{i}\)
+and \(D_{ij} = c_{i-1,j} (-1)^{i-1} \binom{R-1}{i-1}\).
+
 \begin{algorithm}[h]

### 2022-04-13 functions.R

@@ -396,7 +396,29 @@ manifold.clustering <- function(X, K = 2,
   n <- nrow(X)
   
   if (length(initialization) == n) {
-    z.hat <- initialization
+    if (all(!is.na(initialization))) {
+      z.hat <- initialization
+    } else {
+      z.hat.partial <- initialization[!is.na(initialization)]
+      X.partial <- X[!is.na(initialization), ]
+      clustering.partial <- manifold.clustering(
+        X = X.partial, 
+        K = K,
+        degree = degree,
+        initialization = z.hat.partial,
+        intercept = intercept,
+        maxit = maxit,
+        k.isomap = k.isomap * length(z.hat.partial) / n,
+        curve.init = curve.init,
+        min.t = min.t, 
+        max.t = max.t, 
+        normalize = normalize,
+        eps = eps,
+        parallel = paralle,
+        verbose = verbose, 
+        animate = FALSE)
+      z.hat <- clustering.partial$z
+    }
   } else if (initialization == 'spectral') {
     if (missing(A)) {
       stop('adjacency matrix required for spectral initialization')

### 2022-04-20 draft/draft.Rmd
still working on manuscript

@@
 - \usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
-- \setcitestyle{numbers,square,comma}
 - \usepackage{verbatim}
@@
 
+\begin{definition}[Manifold block model]
+Let $p, q \geq 0$, $d = p + q \geq 1$, $K \geq 1$, and $n \geq 1$ be integers.
+If there are manifolds $\mathcal{M}_1, ..., \mathcal{M}_K \subset \mathcal{X}$ for $\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p,q} y \in [0, 1] \}$, and 
+$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}$ such that each $x_i \in \mathcal{M}_k$ for some $k \in [K]$, then $A \sim \GRDPG_{p,q}(X; \rho_n)$ is a \emph{manifold block model}.
+\end{definition}
+
+In practice, we define each manifold $\mathcal{M}_k$ by a function $f_k : [0, 1]^r \to \mathcal{X}$, where $1 \geq r < d$ is the dimensionality of manifold $\mathcal{M}_k$.
+
 Insert something about ASE and its consistency here.
+
+\begin{example}
+Let $p = 2$, $q = 0$, $K = 2$, and $r = 1$. 
+Define two manifolds by $f_1(t) = \begin{bmatrix} \cos(\frac{\pi}{3} t) & \sin(\frac{\pi}{3} t) \end{bmatrix}^\top$ and $f_2(t) = \begin{bmatrix} 1 - \cos(\frac{\pi}{3} t) & 1 - \sin(\frac{\pi}{3} t) \end{bmatrix}^\top$.
+Draw $t_1, ..., t_n \sim \Uniform(0, 1)$ and $z_1, ..., z_n \Multinomial(\frac{1}{2}, \frac{1}{2})$, and compute latent vectors $x_i = f_{z_i}(t_i)$, which are collected in data matrix $\begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$.
+Finally, let $A \sim \RDPG(X)$.
+
+```{r fig1, fig.cap = 'Manifold block model described in Example 1.
+The latent configuration is on the left, and a random dot product graph drawn from the latent configuration is on the middle, and the ASE is on the right.', fig.height = 2.5, fig.width = 7}
+set.seed(123456)
+n1 <- 2 ** 5
+n2 <- 2 ** 5
+n <- n1 + n2
+z <- c(rep(1, n1), rep(2, n2))
+u1 <- runif(n1)
+u2 <- runif(n2)
+x1 <- cos(pi / 3 * u1)
+y1 <- sin(pi / 3 * u1)
+x2 <- 1 + cos(pi / 3 * u2 + pi)
+y2 <- 1 + sin(pi / 3 * u2 + pi)
+data.matrix <- cbind(c(x1, x2), c(y1, y2))
+A <- draw.graph(data.matrix %*% t(data.matrix))
+Xhat <- embedding(A, 2, 0)
+
+par(mfrow = c(1, 3))
+# par(fig = c(0, .7, 0, 1))
+plot(data.matrix, col = z * 2, asp = 1,
+     xlab = NA, ylab = NA, xaxt = 'n', yaxt = 'n',
+     xlim = c(-.1, 1.1), ylim = c(0, 1))
+# ggplot() +
+#   geom_point(aes(x = data.matrix[, 1], y = data.matrix[, 2], colour = factor(z))) +
+#   labs(x = NULL, y = NULL) +
+#   scale_colour_brewer(palette = 'Set1') +
+#   theme(legend.position = 'none') +
+#   coord_fixed()
+# par(fig = c(.3, 1, 0, 1))
+qgraph::qgraph(A, groups = factor(z), legend = FALSE)
+plot(Xhat, col = z * 2, asp = 1, 
+     xlab = NA, ylab = NA, xaxt = 'n', yaxt = 'n')
+```
+\end{example}
 
@@
 Here, we provide two algorithms for MBM community detection. 
-
 First, we consider the case in which communities correspond to manifolds in the latent space that do not intersect and are separated by some finite distance.
@@
 
-### Preliminary Theory
-
-#### Distributions of differences of order statistics
-
-Let $D_i = X_{(i+1)} - X_{(i)}$. Then if $\max_i D_i < \delta$, we have sufficient separation of points in $\mathcal{M}_1$. 
-Then it is sufficient to quantify $P(\max_i D_i > \delta)$ as a function of $n$ and $\delta$ and show that this converges to zero as $n$ grows to $\infty$.
-
-We denote $f(x)$ as the density of each $X_i$, $g_i(x)$ as the density of $X_{(i)}$, 
-$g_{ij}(x, y)$ as the joint density of $X_{(i)}, X_{(j)}$, and $h_i(d)$ as the density of $D_i$ (with corresponding capital letters for the cumulative distribution functions).
-
-The following are taken as given[^wikipedia]:
-
-1. $g_i(x) = \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x))^{n-i} f(x)$.
-2. $g_{ij}(x, y) = \frac{n!}{(i-1)! (j-i-1)! (n-j)!} (F(x))^{i-1} (F(y) - F(x))^{j-i-1} (1 - F(y))^{n-j} f(x) f(y)$.
-3. By convolution, $h_i(d) = \int_0^{1} g_{i, i+1} (x, x + d) \dd x$.
-
-[^wikipedia]: https://en.wikipedia.org/wiki/Order_statistic
-
-\begin{lemma}[The probability density function of $D_i$]
-\begin{equation}
-\label{eq:pdf}
-h_i(d) = \int_0^{1-d} \frac{n!}{(i-1)! (n-i-1)!} (F(x))^{i-1} (1 - F(x+d))^{n-i-1} f(x) f(x+d) \dd x 
-\end{equation}
-\end{lemma}
-
-\begin{proof}
-This is just a direct consequence of 2 and 3 under the given statements. 
-We also note that because the support of $X_i$ is $[0, 1]$, 
-the integral only needs to be evaluated from $0$ to $1 - d$ 
-because of the $f(x+d)$ and $1 - F(x+d)$ terms.
-\end{proof}
-
-\begin{lemma}[The cumulative distribution function of $D_i$]
-
-\begin{equation}
-\label{eq:cdf}
-P(D_i < \delta) = H_i(\delta) = 1 - \int_0^{1-\delta} \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \dd x
-\end{equation}
-\end{lemma}
-
-\begin{proof}
-$$
-\begin{aligned}
-H_i(\delta) & = \int_x^{x+\delta} h_i(d) \dd d \\
-& = \int_x^{x+\delta} \int_0^{1} \frac{n!}{(i-1)! (n-i-1)!} ((F(x))^{i-1} (1 - F(x+d))^{n-i-1} f(x) f(x+d) \dd x \dd d \\
-& = \int_0^{1} \frac{n!}{(i-1)! (n-i-1)!} (F(x))^{i-1} f(x) \int_x^{x+\delta} (1 - F(x+d))^{n-i-1} f(x+d) \dd d \dd x \\
-& = \int_0^{1} \frac{n!}{(i-1)! (n-i-1)!} (F(x))^{i-1} f(x) \int_{F(x)}^{F(x+\delta)} (1 - u)^{n-i-1} \dd u \dd x \\
-& = \int_0^{1} \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} f(x) ((1 - F(x))^{n-i} - (1 - F(x + \delta))^{n-i}) \dd x \\ 
-& = \int_0^1 g_i(x) \dd x - \int_0^1 \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \dd x \\
-& = 1 - \int_0^1 \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \dd x \\
-\end{aligned}
-$$
-
-Because of the $x + \delta$ term, we can't actually evaluate this integral all the way up to $1$, and so we are left with 
-$$
-\begin{aligned}
-& = 1 - \int_0^{1 - \delta} \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \dd x.
-\end{aligned}
-$$
-\end{proof}
-
-#### Uniform case
-
-\begin{lemma}[Differences between order statistics of a uniform distribution]
-If $X_1, ..., X_n \iid \Uniform(0, 1)$, then each $D_i \sim \Betadist(1, n)$.
-\end{lemma}
-
-\begin{proof}
-We begin with Eq. (\ref{eq:pdf}), plugging in $f(x) = 1$ and $F(x) = x$:
-
-$$
-h_i(d) = \int_0^{1-d} \frac{n!}{(i-1)! (n-i-1)!} x^{i-1} (1-x-d)^{n-i-1} \dd x
-$$
-
-Then we proceed with integration by parts, setting 
-$u = x^{i-1} \implies du = (i-1) x^{i-2}$ and 
-$dv = (1-x-d)^{n-i-1} dx \implies v = -\frac{1}{n-i} (1-x-d)^{n-i-1}$. 
-Note that $u v |_0^{1-d} = 0$ in this case. This yields
-
-$$
-= \frac{n!}{(i-1)! (n-i-1)!} \int \frac{i-1}{n-i} x^{i-2} (1-x-d)^{n-i} \dd x
-$$
-
-Then applying integration by parts again until the $x^p$ term disappears, we get:
-
-$$
-\begin{aligned}
-& = \frac{n!}{(i-1)! (n-i-1)!} \frac{(i-1)!}{(n-i) \cdots (n-2)} \int_0^{1-d} (1-x-d)^{n-2} \dd x \\
-& = -\frac{n (n-1)}{n-1} (1-x-d)^{n-1} \Big|_0^{1-d} \\
-& = n (1 - d)^{n-1}
-\end{aligned}
-$$
-
-This the density function for $\Betadist(1, n)$, completing the proof.
-\end{proof}
-
-\begin{theorem}
-Let $X_1, ..., X_n \iid \Uniform(0, 1)$. 
-Then for any $\epsilon$ and $\delta > 0$, 
-there exists an $N = O \big(\frac{-\log \epsilon}{\delta} \big)$ such that 
-$P(\max_i X_{(i+1)} - X_{(i)} < \delta) \geq 1 - \epsilon$ when $n > N$.
+\begin{theorem}[Community detection for one-dimensional nonintersecting manifolds without noise]
+\label{nonintersect-no-noise}
+Let $x_1, ..., x_n$ be points sampled on $K$ one-dimensional manifolds $\mathcal{M}_1, ..., \mathcal{M}_K$, defined by arc-length parameterizations $g_1, ..., g_K$. 
+Suppose $\delta = \min_{k, \ell} \min_{t, s} \| g_k(t) - g_\ell(s) \| > 0$. 
+Let $A_n$ be the event such that $\max_k \max_i \|x_{(i+1)}^{(k)} - x_{(i)}^{(k)}\| > \delta$, where $x_{(i)}^{(k)}$ is the $i^{th}$ order statistic on the $k^{th}$ manifold. 
+Then for any $\epsilon > 0$, there exists an $N \in \mathbb{N}$ such that when $n > N$, $P(A_n) < \epsilon$. 
 \end{theorem}
 
-\begin{proof}[Proof (sketch)]
-Since $X_{(i+1)} - X_{(i)} = D_i \sim \Betadist(1, n)$, 
-$P(X_{(i+1)} - X_{(i)} < \delta) = 1 - (1 - \delta)^n $. This yields
-
-$$
-\begin{aligned}
-P(\max_i D_i < \delta) & \geq (P(D_i < \delta))^{n-1} \\
-& = (1 - (1 - \delta)^n)^{n-1} \\
-& \approx e^{-n \exp(-n \delta)}.
-\end{aligned}
-$$
-
-In the limit $n \to \infty$, this goes to 1.
-\end{proof}
-
-#### General case for one-dimensional manifolds
-
-\begin{theorem}
-\label{thm:generalized}
-Let $X_1, ..., X_n \iid F$ with support $[0, 1]$, and suppose $f(x)$ is continuous and $f(x) \geq a > 0$ everywhere on the support. 
-Let $D_i = X_{(i+1)} - X_{(i)}$. 
-Then for any $\epsilon > 0$, there exists $N > 0$ such that $P(\max_i D_i < \delta) \geq 1 - \epsilon$ when $n > N$.
+\begin{theorem}[Community detection for one-dimensional nonintersecting manifolds with noise]
+\label{nonintersect-with-noise}
+Suppose the same setup as in Theorem \ref{nonintersect-no-noise}. 
+Define $y_i = x_i + e_i$ and $B_n$ as the event such that $\|y_{(i+1)}^{(k)} - y_{(i)}^{(k)}\| > \|y_{(i)}^{(k)} - y_{(j)}^{(\ell)}\|$ for some $i$, $j$, $k$, and $\ell$. 
+Then if $\max_i \|e_i\| < \delta / 3$, for any $\epsilon > 0$, there exists an $N \in \mathbb{N}$ such that when $n > N$, $P(B_n) < \epsilon$. 
 \end{theorem}
 
-\begin{proof}[Proof (sketch)]
-We start with Eq. (\ref{eq:cdf}):
-
-$$P(D_i \leq \delta) = 1 - \int_0^{1-\delta} \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \dd x.$$
-
-Making the approximation $F(x+\delta) \approx F(x) + \delta f(x)$ 
-and bounding $f(x) \geq a$, we get:
-
-$$P(D_i \leq \delta) \geq 1 - \int_0^{1-\delta} \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x) - a \delta)^{n-i} f(x) \dd x.$$
-
-Then making the substitution $u = F(x) \implies du = f(x) \dd x$, we obtain
-
-$$1 - \int_0^{F(1-\delta)} \frac{n!}{(n-i)! (i-1)!} u^{i-1} (1 - u - a \delta)^{n-i} \dd u$$
-
-Evaluating the integral yields
-
-$$P(D_i < \delta) = 1 - (1 - a \delta)^n + (1 - F(1-\delta) - a \delta)^n.$$
-
-Then as before,
-
-$$
-\begin{aligned}
-P(\max_i D_i < \delta) & = P(\text{all } D_i < \delta) \\
-& = 1 - P(\text{some } D_i > \delta) \\
-& \geq 1 - \sum_i^{n-1} P(D_i > \delta) \\
-& = 1 - (n - 1) (1 - a \delta)^n + (n - 1) (1 - F(1 - \delta) - a \delta)^n
-\end{aligned}
-$$
-
-This converges to $1$ in the limit $n \to \infty$.
-
-We can also approximate $F(1 - \delta) \approx 1 - a \delta$, which yields 
-$1 - (n - 1) (1 - a \delta)^n$. Setting this $\geq 1 - \epsilon$ ...
-\end{proof}
-
-#### Extension to multidimensional manifolds
-
-Here, we extend the results on the line to the unit hypercube. The following theorem is a direct consequence of Lemma 2 of \citet{trosset2020rehabilitating}.
-
-\begin{theorem}
-\label{thm:multidim}
-Let $X_1, ..., X_n \iid F$ with support $[0, 1]^r$, and $f(x) \geq a > 0$ everywhere on the support. 
-Define $E_n$ as the event that an $\eta$-neighborhood graph constructed from the sample is connected. 
-Then for any $\epsilon > 0$, there exists $N = O \bigg( \frac{\log \epsilon \eta^r / r^{r/2}}{\log (1 - \frac{a \eta^r}{r^{r / 2}})} \bigg)$ such that $P(E_n) > 1 - \epsilon$ when $n \geq N$.
+\begin{theorem}[Community detection for one-dimensional nonintersecting manifold block models]
+\label{nonintersect-mbm}
+Suppose we have the same setup as in Theorem \ref{nonintersect-no-noise}. 
+Let $p, q \geq 0$ such that $p + q = d$ where $d$ is the dimension of the range of each $g_k$. 
+Suppose we draw $A \sim \GRDPG_{p,q}(X; \rho_n)$ for some sparsity parameter $\rho_n \in (0, 1]$, and let $\hat{X}$ be the ASE of $A$. 
+Then for any $\epsilon > 0$, there exists an $N \in \mathbb{N}$ such that when $n > N$, single linkage clustering produces zero community detection error with probability 1. 
 \end{theorem}
-
-\begin{proof}[Proof (sketch)]
-Divide the hypercube $[0, 1]^r$ into a grid of sub-hypercubes of side length at most $\eta / \sqrt{r}$. 
-$E_n$ is satisfied if each sub-hypercube contains at least one $X_i$ from the sample.
-
-$$
-\begin{aligned}
-P(E_n) & = 1 - P(\text{some cells don't contain } X_i) \\
-& \geq 1 - \sum_k^{\lceil \sqrt{r} / \eta \rceil^r} \prod_i^n P(X_i \text{ is not in the } k^{th} \text{ hypercube}) \\
-& \geq 1 - \lceil \sqrt{r} / \eta \rceil^r (1 - a \eta^r / r^{r/2})^n
-\end{aligned}
-$$
-
-Setting this $\geq 1 - \epsilon$ and solving for $n$ yields the desired rate. 
-\end{proof}
 
@@
 \KwData{Adjacency matrix $A$, number of communities $K$, embedding dimensions $p$ and $q$.}
-\KwResult{Community assignments $1, ..., K$.}
-Compute $X$, the ASE of $A$ using the $p$ most positive and $q$ most negative eigenvalues and their corresponding eigenvectors.\;
-Apply single linkage clustering with $K$ communities on the ASE.\;
+\KwResult{Community assignments $z_1, ..., z_n \in \{1, ..., K\}$.}
+Compute $\hat{X}$, the ASE of $A$ using the $p$ most positive and $q$ most negative eigenvalues and their corresponding eigenvectors.\;
+Apply single linkage clustering with $K$ communities on $\hat{X}$\;
 \caption{ASE clustering for nonintersecting communities.}
@@
 
-\begin{example}[Macaque visuotactile brain areas and connections]
-
-\cite{https://doi.org/10.1111/j.1460-9568.2006.04678.x}
+\begin{example}[Macaque visuotactile brain areas and connections \citep{https://doi.org/10.1111/j.1460-9568.2006.04678.x}]
 
@@
 # Discussion
+
+\appendix
+
+\section{Proofs of Theorems}
+
+To prove theorems \ref{nonintersect-no-noise}, \ref{nonintersect-with-noise}, and \ref{nonintersect-mbm}, we first need to introduce some theory for order statistics. 
+Let $D_i = X_{(i+1)} - X_{(i)}$. Then if $\max_i D_i < \delta$, we have sufficient separation of points in $\mathcal{M}_1$. 
+Then it is sufficient to quantify $P(\max_i D_i > \delta)$ as a function of $n$ and $\delta$ and show that this converges to zero as $n$ grows to $\infty$.
+
+We denote $f(x)$ as the density of each $X_i$, $g_i(x)$ as the density of $X_{(i)}$, 
+$g_{ij}(x, y)$ as the joint density of $X_{(i)}, X_{(j)}$, and $h_i(d)$ as the density of $D_i$ (with corresponding capital letters for the cumulative distribution functions).
+
+The following are taken as given[^wikipedia]:
+
+1. $g_i(x) = \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x))^{n-i} f(x)$.
+2. $g_{ij}(x, y) = \frac{n!}{(i-1)! (j-i-1)! (n-j)!} (F(x))^{i-1} (F(y) - F(x))^{j-i-1} (1 - F(y))^{n-j} f(x) f(y)$.
+3. By convolution, $h_i(d) = \int_0^{1} g_{i, i+1} (x, x + d) \dd x$.
+
+[^wikipedia]: https://en.wikipedia.org/wiki/Order_statistic
+
+\begin{lemma}[The probability density function of $D_i$]
+\begin{equation}
+\label{eq:pdf}
+h_i(d) = \int_0^{1-d} \frac{n!}{(i-1)! (n-i-1)!} (F(x))^{i-1} (1 - F(x+d))^{n-i-1} f(x) f(x+d) \dd x 
+\end{equation}
+\end{lemma}
+
+\begin{proof}
+This is just a direct consequence of 2 and 3 under the given statements. 
+We also note that because the support of $X_i$ is $[0, 1]$, 
+the integral only needs to be evaluated from $0$ to $1 - d$ 
+because of the $f(x+d)$ and $1 - F(x+d)$ terms.
+\end{proof}
+
+\begin{lemma}[The cumulative distribution function of $D_i$]
+
+\begin{equation}
+\label{eq:cdf}
+P(D_i < \delta) = H_i(\delta) = 1 - \int_0^{1-\delta} \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \dd x
+\end{equation}
+\end{lemma}
+
+\begin{proof}
+$$
+\begin{aligned}
+H_i(\delta) & = \int_x^{x+\delta} h_i(d) \dd d \\
+& = \int_x^{x+\delta} \int_0^{1} \frac{n!}{(i-1)! (n-i-1)!} ((F(x))^{i-1} (1 - F(x+d))^{n-i-1} f(x) f(x+d) \dd x \dd d \\
+& = \int_0^{1} \frac{n!}{(i-1)! (n-i-1)!} (F(x))^{i-1} f(x) \int_x^{x+\delta} (1 - F(x+d))^{n-i-1} f(x+d) \dd d \dd x \\
+& = \int_0^{1} \frac{n!}{(i-1)! (n-i-1)!} (F(x))^{i-1} f(x) \int_{F(x)}^{F(x+\delta)} (1 - u)^{n-i-1} \dd u \dd x \\
+& = \int_0^{1} \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} f(x) ((1 - F(x))^{n-i} - (1 - F(x + \delta))^{n-i}) \dd x \\ 
+& = \int_0^1 g_i(x) \dd x - \int_0^1 \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \dd x \\
+& = 1 - \int_0^1 \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \dd x \\
+\end{aligned}
+$$
+
+Because of the $x + \delta$ term, we can't actually evaluate this integral all the way up to $1$, and so we are left with 
+$$
+\begin{aligned}
+& = 1 - \int_0^{1 - \delta} \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \dd x.
+\end{aligned}
+$$
+\end{proof}
+
+We now focus on the case where points are sampled uniformly on the unit interval to build up to more general distributions.
+
+\begin{lemma}[Differences between order statistics of a uniform distribution]
+If $X_1, ..., X_n \iid \Uniform(0, 1)$, then each $D_i \sim \Betadist(1, n)$.
+\end{lemma}
+
+\begin{proof}
+We begin with Eq. (\ref{eq:pdf}), plugging in $f(x) = 1$ and $F(x) = x$:
+
+$$
+h_i(d) = \int_0^{1-d} \frac{n!}{(i-1)! (n-i-1)!} x^{i-1} (1-x-d)^{n-i-1} \dd x
+$$
+
+Then we proceed with integration by parts, setting 
+$u = x^{i-1} \implies du = (i-1) x^{i-2}$ and 
+$dv = (1-x-d)^{n-i-1} dx \implies v = -\frac{1}{n-i} (1-x-d)^{n-i-1}$. 
+Note that $u v |_0^{1-d} = 0$ in this case. This yields
+
+$$
+= \frac{n!}{(i-1)! (n-i-1)!} \int \frac{i-1}{n-i} x^{i-2} (1-x-d)^{n-i} \dd x
+$$
+
+Then applying integration by parts again until the $x^p$ term disappears, we get:
+
+$$
+\begin{aligned}
+& = \frac{n!}{(i-1)! (n-i-1)!} \frac{(i-1)!}{(n-i) \cdots (n-2)} \int_0^{1-d} (1-x-d)^{n-2} \dd x \\
+& = -\frac{n (n-1)}{n-1} (1-x-d)^{n-1} \Big|_0^{1-d} \\
+& = n (1 - d)^{n-1}
+\end{aligned}
+$$
+
+This the density function for $\Betadist(1, n)$, completing the proof.
+\end{proof}
+
+\begin{lemma}
+Let $X_1, ..., X_n \iid \Uniform(0, 1)$. 
+Then for any $\epsilon$ and $\delta > 0$, 
+there exists an $N = O \big(\frac{-\log \epsilon}{\delta} \big)$ such that 
+$P(\max_i X_{(i+1)} - X_{(i)} < \delta) \geq 1 - \epsilon$ when $n > N$.
+\end{lemma}
+
+\begin{proof}[Proof (sketch)]
+Since $X_{(i+1)} - X_{(i)} = D_i \sim \Betadist(1, n)$, 
+$P(X_{(i+1)} - X_{(i)} < \delta) = 1 - (1 - \delta)^n $. This yields
+
+$$
+\begin{aligned}
+P(\max_i D_i < \delta) & \geq (P(D_i < \delta))^{n-1} \\
+& = (1 - (1 - \delta)^n)^{n-1} \\
+& \approx e^{-n \exp(-n \delta)}.
+\end{aligned}
+$$
+
+In the limit $n \to \infty$, this goes to 1.
+\end{proof}
+
+Now we extend this lemma to general distributions on the unit interval.
+
+\begin{lemma}
+Let $X_1, ..., X_n \iid F$ with support $[0, 1]$, and suppose $f(x)$ is continuous and $f(x) \geq a > 0$ everywhere on the support. 
+Let $D_i = X_{(i+1)} - X_{(i)}$. 
+Then for any $\epsilon > 0$, there exists $N > 0$ such that $P(\max_i D_i < \delta) \geq 1 - \epsilon$ when $n > N$.
+\end{lemma}
+
+\begin{proof}[Proof (sketch)]
+We start with Eq. (\ref{eq:cdf}):
+
+$$P(D_i \leq \delta) = 1 - \int_0^{1-\delta} \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \dd x.$$
+
+Making the approximation $F(x+\delta) \approx F(x) + \delta f(x)$ 
+and bounding $f(x) \geq a$, we get:
+
+$$P(D_i \leq \delta) \geq 1 - \int_0^{1-\delta} \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x) - a \delta)^{n-i} f(x) \dd x.$$
+
+Then making the substitution $u = F(x) \implies du = f(x) \dd x$, we obtain
+
+$$1 - \int_0^{F(1-\delta)} \frac{n!}{(n-i)! (i-1)!} u^{i-1} (1 - u - a \delta)^{n-i} \dd u$$
+
+Evaluating the integral yields
+
+$$P(D_i < \delta) = 1 - (1 - a \delta)^n + (1 - F(1-\delta) - a \delta)^n.$$
+
+Then as before,
+
+$$
+\begin{aligned}
+P(\max_i D_i < \delta) & = P(\text{all } D_i < \delta) \\
+& = 1 - P(\text{some } D_i > \delta) \\
+& \geq 1 - \sum_i^{n-1} P(D_i > \delta) \\
+& = 1 - (n - 1) (1 - a \delta)^n + (n - 1) (1 - F(1 - \delta) - a \delta)^n
+\end{aligned}
+$$
+
+This converges to $1$ in the limit $n \to \infty$.
+\end{proof}
+
+Here, we extend the results on the line to the unit hypercube. The following theorem is a direct consequence of Lemma 2 of \citet{trosset2020rehabilitating}.
+
+\begin{lemma}
+\label{thm:multidim}
+Let $X_1, ..., X_n \iid F$ with support $[0, 1]^r$, and $f(x) \geq a > 0$ everywhere on the support. 
+Define $E_n$ as the event that an $\eta$-neighborhood graph constructed from the sample is connected. 
+Then for any $\epsilon > 0$, there exists $N = O \bigg( \frac{\log \epsilon \eta^r / r^{r/2}}{\log (1 - \frac{a \eta^r}{r^{r / 2}})} \bigg)$ such that $P(E_n) > 1 - \epsilon$ when $n \geq N$.
+\end{lemma}
+
+\begin{proof}[Proof (sketch)]
+Divide the hypercube $[0, 1]^r$ into a grid of sub-hypercubes of side length at most $\eta / \sqrt{r}$. 
+$E_n$ is satisfied if each sub-hypercube contains at least one $X_i$ from the sample.
+
+$$
+\begin{aligned}
+P(E_n) & = 1 - P(\text{some cells don't contain } X_i) \\
+& \geq 1 - \sum_k^{\lceil \sqrt{r} / \eta \rceil^r} \prod_i^n P(X_i \text{ is not in the } k^{th} \text{ hypercube}) \\
+& \geq 1 - \lceil \sqrt{r} / \eta \rceil^r (1 - a \eta^r / r^{r/2})^n
+\end{aligned}
+$$
+
+Setting this $\geq 1 - \epsilon$ and solving for $n$ yields the desired rate. 
+\end{proof}

### 2022-04-20 draft/draft.tex
still working on manuscript

@@
 \usepackage[linesnumbered,ruled,vlined]{algorithm2e}
-\setcitestyle{numbers,square,comma}
 \usepackage{verbatim}
@@
 
+\begin{definition}[Manifold block model]
+Let $p, q \geq 0$, $d = p + q \geq 1$, $K \geq 1$, and $n \geq 1$ be integers.
+If there are manifolds $\mathcal{M}_1, ..., \mathcal{M}_K \subset \mathcal{X}$ for $\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p,q} y \in [0, 1] \}$, and 
+$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}$ such that each $x_i \in \mathcal{M}_k$ for some $k \in [K]$, then $A \sim \mathrm{GRDPG}_{p,q}(X; \rho_n)$ is a \emph{manifold block model}.
+\end{definition}
+
+In practice, we define each manifold \(\mathcal{M}_k\) by a function
+\(f_k : [0, 1]^r \to \mathcal{X}\), where \(1 \geq r < d\) is the
+dimensionality of manifold \(\mathcal{M}_k\).
+
 Insert something about ASE and its consistency here.
+
+\begin{example}
+Let $p = 2$, $q = 0$, $K = 2$, and $r = 1$. 
+Define two manifolds by $f_1(t) = \begin{bmatrix} \cos(\frac{\pi}{3} t) & \sin(\frac{\pi}{3} t) \end{bmatrix}^\top$ and $f_2(t) = \begin{bmatrix} 1 - \cos(\frac{\pi}{3} t) & 1 - \sin(\frac{\pi}{3} t) \end{bmatrix}^\top$.
+Draw $t_1, ..., t_n \sim \mathrm{Uniform}(0, 1)$ and $z_1, ..., z_n \mathrm{Multinomial}(\frac{1}{2}, \frac{1}{2})$, and compute latent vectors $x_i = f_{z_i}(t_i)$, which are collected in data matrix $\begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$.
+Finally, let $A \sim \mathrm{RDPG}(X)$.
+
+\begin{figure}[H]
+
+{\centering \includegraphics{draft_files/figure-latex/fig1-1}
+
+}
+
+\caption{Manifold block model described in Example 1.
+The latent configuration is on the left, and a random dot product graph drawn from the latent configuration is on the middle, and the ASE is on the right.}\label{fig:fig1}
+\end{figure}
+\end{example}
 
@@
 
-Here, we provide two algorithms for MBM community detection.
-
-First, we consider the case in which communities correspond to manifolds
-in the latent space that do not intersect and are separated by some
-finite distance. In this scenario, we use the convergence of the ASE to
-show that single linkage clustering on the latent space produces a
-clustering such that the total number of misclustered vertices goes to
-zero, with high probability.
+Here, we provide two algorithms for MBM community detection. First, we
+consider the case in which communities correspond to manifolds in the
+latent space that do not intersect and are separated by some finite
+distance. In this scenario, we use the convergence of the ASE to show
+that single linkage clustering on the latent space produces a clustering
+such that the total number of misclustered vertices goes to zero, with
+high probability.
 
@@
 
-\hypertarget{preliminary-theory}{%
-\subsubsection{Preliminary Theory}\label{preliminary-theory}}
-
-\hypertarget{distributions-of-differences-of-order-statistics}{%
-\paragraph{Distributions of differences of order
-statistics}\label{distributions-of-differences-of-order-statistics}}
-
-Let \(D_i = X_{(i+1)} - X_{(i)}\). Then if \(\max_i D_i < \delta\), we
-have sufficient separation of points in \(\mathcal{M}_1\). Then it is
-sufficient to quantify \(P(\max_i D_i > \delta)\) as a function of \(n\)
-and \(\delta\) and show that this converges to zero as \(n\) grows to
-\(\infty\).
-
-We denote \(f(x)\) as the density of each \(X_i\), \(g_i(x)\) as the
-density of \(X_{(i)}\), \(g_{ij}(x, y)\) as the joint density of
-\(X_{(i)}, X_{(j)}\), and \(h_i(d)\) as the density of \(D_i\) (with
-corresponding capital letters for the cumulative distribution
-functions).
-
-The following are taken as given\footnote{\url{https://en.wikipedia.org/wiki/Order_statistic}}:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  \(g_i(x) = \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x))^{n-i} f(x)\).
-\item
-  \(g_{ij}(x, y) = \frac{n!}{(i-1)! (j-i-1)! (n-j)!} (F(x))^{i-1} (F(y) - F(x))^{j-i-1} (1 - F(y))^{n-j} f(x) f(y)\).
-\item
-  By convolution,
-  \(h_i(d) = \int_0^{1} g_{i, i+1} (x, x + d) \mathrm{d}x\).
-\end{enumerate}
-
-\begin{lemma}[The probability density function of $D_i$]
-\begin{equation}
-\label{eq:pdf}
-h_i(d) = \int_0^{1-d} \frac{n!}{(i-1)! (n-i-1)!} (F(x))^{i-1} (1 - F(x+d))^{n-i-1} f(x) f(x+d) \mathrm{d}x 
-\end{equation}
-\end{lemma}
-
-\begin{proof}
-This is just a direct consequence of 2 and 3 under the given statements. 
-We also note that because the support of $X_i$ is $[0, 1]$, 
-the integral only needs to be evaluated from $0$ to $1 - d$ 
-because of the $f(x+d)$ and $1 - F(x+d)$ terms.
-\end{proof}
-
-\begin{lemma}[The cumulative distribution function of $D_i$]
-
-\begin{equation}
-\label{eq:cdf}
-P(D_i < \delta) = H_i(\delta) = 1 - \int_0^{1-\delta} \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \mathrm{d}x
-\end{equation}
-\end{lemma}
-
-\begin{proof}
-$$
-\begin{aligned}
-H_i(\delta) & = \int_x^{x+\delta} h_i(d) \mathrm{d}d \\
-& = \int_x^{x+\delta} \int_0^{1} \frac{n!}{(i-1)! (n-i-1)!} ((F(x))^{i-1} (1 - F(x+d))^{n-i-1} f(x) f(x+d) \mathrm{d}x \mathrm{d}d \\
-& = \int_0^{1} \frac{n!}{(i-1)! (n-i-1)!} (F(x))^{i-1} f(x) \int_x^{x+\delta} (1 - F(x+d))^{n-i-1} f(x+d) \mathrm{d}d \mathrm{d}x \\
-& = \int_0^{1} \frac{n!}{(i-1)! (n-i-1)!} (F(x))^{i-1} f(x) \int_{F(x)}^{F(x+\delta)} (1 - u)^{n-i-1} \mathrm{d}u \mathrm{d}x \\
-& = \int_0^{1} \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} f(x) ((1 - F(x))^{n-i} - (1 - F(x + \delta))^{n-i}) \mathrm{d}x \\ 
-& = \int_0^1 g_i(x) \mathrm{d}x - \int_0^1 \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \mathrm{d}x \\
-& = 1 - \int_0^1 \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \mathrm{d}x \\
-\end{aligned}
-$$
-
-Because of the $x + \delta$ term, we can't actually evaluate this integral all the way up to $1$, and so we are left with 
-$$
-\begin{aligned}
-& = 1 - \int_0^{1 - \delta} \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \mathrm{d}x.
-\end{aligned}
-$$
-\end{proof}
-
-\hypertarget{uniform-case}{%
-\paragraph{Uniform case}\label{uniform-case}}
-
-\begin{lemma}[Differences between order statistics of a uniform distribution]
-If $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}\mathrm{Uniform}(0, 1)$, then each $D_i \sim \mathrm{Beta}(1, n)$.
-\end{lemma}
-
-\begin{proof}
-We begin with Eq. (\ref{eq:pdf}), plugging in $f(x) = 1$ and $F(x) = x$:
-
-$$
-h_i(d) = \int_0^{1-d} \frac{n!}{(i-1)! (n-i-1)!} x^{i-1} (1-x-d)^{n-i-1} \mathrm{d}x
-$$
-
-Then we proceed with integration by parts, setting 
-$u = x^{i-1} \implies du = (i-1) x^{i-2}$ and 
-$dv = (1-x-d)^{n-i-1} dx \implies v = -\frac{1}{n-i} (1-x-d)^{n-i-1}$. 
-Note that $u v |_0^{1-d} = 0$ in this case. This yields
-
-$$
-= \frac{n!}{(i-1)! (n-i-1)!} \int \frac{i-1}{n-i} x^{i-2} (1-x-d)^{n-i} \mathrm{d}x
-$$
-
-Then applying integration by parts again until the $x^p$ term disappears, we get:
-
-$$
-\begin{aligned}
-& = \frac{n!}{(i-1)! (n-i-1)!} \frac{(i-1)!}{(n-i) \cdots (n-2)} \int_0^{1-d} (1-x-d)^{n-2} \mathrm{d}x \\
-& = -\frac{n (n-1)}{n-1} (1-x-d)^{n-1} \Big|_0^{1-d} \\
-& = n (1 - d)^{n-1}
-\end{aligned}
-$$
-
-This the density function for $\mathrm{Beta}(1, n)$, completing the proof.
-\end{proof}
-
-\begin{theorem}
-Let $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}\mathrm{Uniform}(0, 1)$. 
-Then for any $\epsilon$ and $\delta > 0$, 
-there exists an $N = O \big(\frac{-\log \epsilon}{\delta} \big)$ such that 
-$P(\max_i X_{(i+1)} - X_{(i)} < \delta) \geq 1 - \epsilon$ when $n > N$.
+\begin{theorem}[Community detection for one-dimensional nonintersecting manifolds without noise]
+\label{nonintersect-no-noise}
+Let $x_1, ..., x_n$ be points sampled on $K$ one-dimensional manifolds $\mathcal{M}_1, ..., \mathcal{M}_K$, defined by arc-length parameterizations $g_1, ..., g_K$. 
+Suppose $\delta = \min_{k, \ell} \min_{t, s} \| g_k(t) - g_\ell(s) \| > 0$. 
+Let $A_n$ be the event such that $\max_k \max_i \|x_{(i+1)}^{(k)} - x_{(i)}^{(k)}\| > \delta$, where $x_{(i)}^{(k)}$ is the $i^{th}$ order statistic on the $k^{th}$ manifold. 
+Then for any $\epsilon > 0$, there exists an $N \in \mathbb{N}$ such that when $n > N$, $P(A_n) < \epsilon$. 
 \end{theorem}
 
-\begin{proof}[Proof (sketch)]
-Since $X_{(i+1)} - X_{(i)} = D_i \sim \mathrm{Beta}(1, n)$, 
-$P(X_{(i+1)} - X_{(i)} < \delta) = 1 - (1 - \delta)^n $. This yields
-
-$$
-\begin{aligned}
-P(\max_i D_i < \delta) & \geq (P(D_i < \delta))^{n-1} \\
-& = (1 - (1 - \delta)^n)^{n-1} \\
-& \approx e^{-n \exp(-n \delta)}.
-\end{aligned}
-$$
-
-In the limit $n \to \infty$, this goes to 1.
-\end{proof}
-
-\hypertarget{general-case-for-one-dimensional-manifolds}{%
-\paragraph{General case for one-dimensional
-manifolds}\label{general-case-for-one-dimensional-manifolds}}
-
-\begin{theorem}
-\label{thm:generalized}
-Let $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}F$ with support $[0, 1]$, and suppose $f(x)$ is continuous and $f(x) \geq a > 0$ everywhere on the support. 
-Let $D_i = X_{(i+1)} - X_{(i)}$. 
-Then for any $\epsilon > 0$, there exists $N > 0$ such that $P(\max_i D_i < \delta) \geq 1 - \epsilon$ when $n > N$.
+\begin{theorem}[Community detection for one-dimensional nonintersecting manifolds with noise]
+\label{nonintersect-with-noise}
+Suppose the same setup as in Theorem \ref{nonintersect-no-noise}. 
+Define $y_i = x_i + e_i$ and $B_n$ as the event such that $\|y_{(i+1)}^{(k)} - y_{(i)}^{(k)}\| > \|y_{(i)}^{(k)} - y_{(j)}^{(\ell)}\|$ for some $i$, $j$, $k$, and $\ell$. 
+Then if $\max_i \|e_i\| < \delta / 3$, for any $\epsilon > 0$, there exists an $N \in \mathbb{N}$ such that when $n > N$, $P(B_n) < \epsilon$. 
 \end{theorem}
 
-\begin{proof}[Proof (sketch)]
-We start with Eq. (\ref{eq:cdf}):
-
-$$P(D_i \leq \delta) = 1 - \int_0^{1-\delta} \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \mathrm{d}x.$$
-
-Making the approximation $F(x+\delta) \approx F(x) + \delta f(x)$ 
-and bounding $f(x) \geq a$, we get:
-
-$$P(D_i \leq \delta) \geq 1 - \int_0^{1-\delta} \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x) - a \delta)^{n-i} f(x) \mathrm{d}x.$$
-
-Then making the substitution $u = F(x) \implies du = f(x) \mathrm{d}x$, we obtain
-
-$$1 - \int_0^{F(1-\delta)} \frac{n!}{(n-i)! (i-1)!} u^{i-1} (1 - u - a \delta)^{n-i} \mathrm{d}u$$
-
-Evaluating the integral yields
-
-$$P(D_i < \delta) = 1 - (1 - a \delta)^n + (1 - F(1-\delta) - a \delta)^n.$$
-
-Then as before,
-
-$$
-\begin{aligned}
-P(\max_i D_i < \delta) & = P(\text{all } D_i < \delta) \\
-& = 1 - P(\text{some } D_i > \delta) \\
-& \geq 1 - \sum_i^{n-1} P(D_i > \delta) \\
-& = 1 - (n - 1) (1 - a \delta)^n + (n - 1) (1 - F(1 - \delta) - a \delta)^n
-\end{aligned}
-$$
-
-This converges to $1$ in the limit $n \to \infty$.
-
-We can also approximate $F(1 - \delta) \approx 1 - a \delta$, which yields 
-$1 - (n - 1) (1 - a \delta)^n$. Setting this $\geq 1 - \epsilon$ ...
-\end{proof}
-
-\hypertarget{extension-to-multidimensional-manifolds}{%
-\paragraph{Extension to multidimensional
-manifolds}\label{extension-to-multidimensional-manifolds}}
-
-Here, we extend the results on the line to the unit hypercube. The
-following theorem is a direct consequence of Lemma 2 of
-\citet{trosset2020rehabilitating}.
-
-\begin{theorem}
-\label{thm:multidim}
-Let $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}F$ with support $[0, 1]^r$, and $f(x) \geq a > 0$ everywhere on the support. 
-Define $E_n$ as the event that an $\eta$-neighborhood graph constructed from the sample is connected. 
-Then for any $\epsilon > 0$, there exists $N = O \bigg( \frac{\log \epsilon \eta^r / r^{r/2}}{\log (1 - \frac{a \eta^r}{r^{r / 2}})} \bigg)$ such that $P(E_n) > 1 - \epsilon$ when $n \geq N$.
+\begin{theorem}[Community detection for one-dimensional nonintersecting manifold block models]
+\label{nonintersect-mbm}
+Suppose we have the same setup as in Theorem \ref{nonintersect-no-noise}. 
+Let $p, q \geq 0$ such that $p + q = d$ where $d$ is the dimension of the range of each $g_k$. 
+Suppose we draw $A \sim \mathrm{GRDPG}_{p,q}(X; \rho_n)$ for some sparsity parameter $\rho_n \in (0, 1]$, and let $\hat{X}$ be the ASE of $A$. 
+Then for any $\epsilon > 0$, there exists an $N \in \mathbb{N}$ such that when $n > N$, single linkage clustering produces zero community detection error with probability 1. 
 \end{theorem}
-
-\begin{proof}[Proof (sketch)]
-Divide the hypercube $[0, 1]^r$ into a grid of sub-hypercubes of side length at most $\eta / \sqrt{r}$. 
-$E_n$ is satisfied if each sub-hypercube contains at least one $X_i$ from the sample.
-
-$$
-\begin{aligned}
-P(E_n) & = 1 - P(\text{some cells don't contain } X_i) \\
-& \geq 1 - \sum_k^{\lceil \sqrt{r} / \eta \rceil^r} \prod_i^n P(X_i \text{ is not in the } k^{th} \text{ hypercube}) \\
-& \geq 1 - \lceil \sqrt{r} / \eta \rceil^r (1 - a \eta^r / r^{r/2})^n
-\end{aligned}
-$$
-
-Setting this $\geq 1 - \epsilon$ and solving for $n$ yields the desired rate. 
-\end{proof}
 
@@
 
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
+
+\appendix
+
+\section{Proofs of Theorems}
+
+To prove theorems \ref{nonintersect-no-noise},
+\ref{nonintersect-with-noise}, and \ref{nonintersect-mbm}, we first need
+to introduce some theory for order statistics. Let
+\(D_i = X_{(i+1)} - X_{(i)}\). Then if \(\max_i D_i < \delta\), we have
+sufficient separation of points in \(\mathcal{M}_1\). Then it is
+sufficient to quantify \(P(\max_i D_i > \delta)\) as a function of \(n\)
+and \(\delta\) and show that this converges to zero as \(n\) grows to
+\(\infty\).
+
+We denote \(f(x)\) as the density of each \(X_i\), \(g_i(x)\) as the
+density of \(X_{(i)}\), \(g_{ij}(x, y)\) as the joint density of
+\(X_{(i)}, X_{(j)}\), and \(h_i(d)\) as the density of \(D_i\) (with
+corresponding capital letters for the cumulative distribution
+functions).
+
+The following are taken as given\footnote{\url{https://en.wikipedia.org/wiki/Order_statistic}}:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  \(g_i(x) = \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x))^{n-i} f(x)\).
+\item
+  \(g_{ij}(x, y) = \frac{n!}{(i-1)! (j-i-1)! (n-j)!} (F(x))^{i-1} (F(y) - F(x))^{j-i-1} (1 - F(y))^{n-j} f(x) f(y)\).
+\item
+  By convolution,
+  \(h_i(d) = \int_0^{1} g_{i, i+1} (x, x + d) \mathrm{d}x\).
+\end{enumerate}
+
+\begin{lemma}[The probability density function of $D_i$]
+\begin{equation}
+\label{eq:pdf}
+h_i(d) = \int_0^{1-d} \frac{n!}{(i-1)! (n-i-1)!} (F(x))^{i-1} (1 - F(x+d))^{n-i-1} f(x) f(x+d) \mathrm{d}x 
+\end{equation}
+\end{lemma}
+
+\begin{proof}
+This is just a direct consequence of 2 and 3 under the given statements. 
+We also note that because the support of $X_i$ is $[0, 1]$, 
+the integral only needs to be evaluated from $0$ to $1 - d$ 
+because of the $f(x+d)$ and $1 - F(x+d)$ terms.
+\end{proof}
+
+\begin{lemma}[The cumulative distribution function of $D_i$]
+
+\begin{equation}
+\label{eq:cdf}
+P(D_i < \delta) = H_i(\delta) = 1 - \int_0^{1-\delta} \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \mathrm{d}x
+\end{equation}
+\end{lemma}
+
+\begin{proof}
+$$
+\begin{aligned}
+H_i(\delta) & = \int_x^{x+\delta} h_i(d) \mathrm{d}d \\
+& = \int_x^{x+\delta} \int_0^{1} \frac{n!}{(i-1)! (n-i-1)!} ((F(x))^{i-1} (1 - F(x+d))^{n-i-1} f(x) f(x+d) \mathrm{d}x \mathrm{d}d \\
+& = \int_0^{1} \frac{n!}{(i-1)! (n-i-1)!} (F(x))^{i-1} f(x) \int_x^{x+\delta} (1 - F(x+d))^{n-i-1} f(x+d) \mathrm{d}d \mathrm{d}x \\
+& = \int_0^{1} \frac{n!}{(i-1)! (n-i-1)!} (F(x))^{i-1} f(x) \int_{F(x)}^{F(x+\delta)} (1 - u)^{n-i-1} \mathrm{d}u \mathrm{d}x \\
+& = \int_0^{1} \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} f(x) ((1 - F(x))^{n-i} - (1 - F(x + \delta))^{n-i}) \mathrm{d}x \\ 
+& = \int_0^1 g_i(x) \mathrm{d}x - \int_0^1 \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \mathrm{d}x \\
+& = 1 - \int_0^1 \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \mathrm{d}x \\
+\end{aligned}
+$$
+
+Because of the $x + \delta$ term, we can't actually evaluate this integral all the way up to $1$, and so we are left with 
+$$
+\begin{aligned}
+& = 1 - \int_0^{1 - \delta} \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \mathrm{d}x.
+\end{aligned}
+$$
+\end{proof}
+
+We now focus on the case where points are sampled uniformly on the unit
+interval to build up to more general distributions.
+
+\begin{lemma}[Differences between order statistics of a uniform distribution]
+If $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}\mathrm{Uniform}(0, 1)$, then each $D_i \sim \mathrm{Beta}(1, n)$.
+\end{lemma}
+
+\begin{proof}
+We begin with Eq. (\ref{eq:pdf}), plugging in $f(x) = 1$ and $F(x) = x$:
+
+$$
+h_i(d) = \int_0^{1-d} \frac{n!}{(i-1)! (n-i-1)!} x^{i-1} (1-x-d)^{n-i-1} \mathrm{d}x
+$$
+
+Then we proceed with integration by parts, setting 
+$u = x^{i-1} \implies du = (i-1) x^{i-2}$ and 
+$dv = (1-x-d)^{n-i-1} dx \implies v = -\frac{1}{n-i} (1-x-d)^{n-i-1}$. 
+Note that $u v |_0^{1-d} = 0$ in this case. This yields
+
+$$
+= \frac{n!}{(i-1)! (n-i-1)!} \int \frac{i-1}{n-i} x^{i-2} (1-x-d)^{n-i} \mathrm{d}x
+$$
+
+Then applying integration by parts again until the $x^p$ term disappears, we get:
+
+$$
+\begin{aligned}
+& = \frac{n!}{(i-1)! (n-i-1)!} \frac{(i-1)!}{(n-i) \cdots (n-2)} \int_0^{1-d} (1-x-d)^{n-2} \mathrm{d}x \\
+& = -\frac{n (n-1)}{n-1} (1-x-d)^{n-1} \Big|_0^{1-d} \\
+& = n (1 - d)^{n-1}
+\end{aligned}
+$$
+
+This the density function for $\mathrm{Beta}(1, n)$, completing the proof.
+\end{proof}
+
+\begin{lemma}
+Let $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}\mathrm{Uniform}(0, 1)$. 
+Then for any $\epsilon$ and $\delta > 0$, 
+there exists an $N = O \big(\frac{-\log \epsilon}{\delta} \big)$ such that 
+$P(\max_i X_{(i+1)} - X_{(i)} < \delta) \geq 1 - \epsilon$ when $n > N$.
+\end{lemma}
+
+\begin{proof}[Proof (sketch)]
+Since $X_{(i+1)} - X_{(i)} = D_i \sim \mathrm{Beta}(1, n)$, 
+$P(X_{(i+1)} - X_{(i)} < \delta) = 1 - (1 - \delta)^n $. This yields
+
+$$
+\begin{aligned}
+P(\max_i D_i < \delta) & \geq (P(D_i < \delta))^{n-1} \\
+& = (1 - (1 - \delta)^n)^{n-1} \\
+& \approx e^{-n \exp(-n \delta)}.
+\end{aligned}
+$$
+
+In the limit $n \to \infty$, this goes to 1.
+\end{proof}
+
+Now we extend this lemma to general distributions on the unit interval.
+
+\begin{lemma}
+Let $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}F$ with support $[0, 1]$, and suppose $f(x)$ is continuous and $f(x) \geq a > 0$ everywhere on the support. 
+Let $D_i = X_{(i+1)} - X_{(i)}$. 
+Then for any $\epsilon > 0$, there exists $N > 0$ such that $P(\max_i D_i < \delta) \geq 1 - \epsilon$ when $n > N$.
+\end{lemma}
+
+\begin{proof}[Proof (sketch)]
+We start with Eq. (\ref{eq:cdf}):
+
+$$P(D_i \leq \delta) = 1 - \int_0^{1-\delta} \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \mathrm{d}x.$$
+
+Making the approximation $F(x+\delta) \approx F(x) + \delta f(x)$ 
+and bounding $f(x) \geq a$, we get:
+
+$$P(D_i \leq \delta) \geq 1 - \int_0^{1-\delta} \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x) - a \delta)^{n-i} f(x) \mathrm{d}x.$$
+
+Then making the substitution $u = F(x) \implies du = f(x) \mathrm{d}x$, we obtain
+
+$$1 - \int_0^{F(1-\delta)} \frac{n!}{(n-i)! (i-1)!} u^{i-1} (1 - u - a \delta)^{n-i} \mathrm{d}u$$
+
+Evaluating the integral yields
+
+$$P(D_i < \delta) = 1 - (1 - a \delta)^n + (1 - F(1-\delta) - a \delta)^n.$$
+
+Then as before,
+
+$$
+\begin{aligned}
+P(\max_i D_i < \delta) & = P(\text{all } D_i < \delta) \\
+& = 1 - P(\text{some } D_i > \delta) \\
+& \geq 1 - \sum_i^{n-1} P(D_i > \delta) \\
+& = 1 - (n - 1) (1 - a \delta)^n + (n - 1) (1 - F(1 - \delta) - a \delta)^n
+\end{aligned}
+$$
+
+This converges to $1$ in the limit $n \to \infty$.
+\end{proof}
+
+Here, we extend the results on the line to the unit hypercube. The
+following theorem is a direct consequence of Lemma 2 of
+\citet{trosset2020rehabilitating}.
+
+\begin{lemma}
+\label{thm:multidim}
+Let $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}F$ with support $[0, 1]^r$, and $f(x) \geq a > 0$ everywhere on the support. 
+Define $E_n$ as the event that an $\eta$-neighborhood graph constructed from the sample is connected. 
+Then for any $\epsilon > 0$, there exists $N = O \bigg( \frac{\log \epsilon \eta^r / r^{r/2}}{\log (1 - \frac{a \eta^r}{r^{r / 2}})} \bigg)$ such that $P(E_n) > 1 - \epsilon$ when $n \geq N$.
+\end{lemma}
+
+\begin{proof}[Proof (sketch)]
+Divide the hypercube $[0, 1]^r$ into a grid of sub-hypercubes of side length at most $\eta / \sqrt{r}$. 
+$E_n$ is satisfied if each sub-hypercube contains at least one $X_i$ from the sample.
+
+$$
+\begin{aligned}
+P(E_n) & = 1 - P(\text{some cells don't contain } X_i) \\
+& \geq 1 - \sum_k^{\lceil \sqrt{r} / \eta \rceil^r} \prod_i^n P(X_i \text{ is not in the } k^{th} \text{ hypercube}) \\
+& \geq 1 - \lceil \sqrt{r} / \eta \rceil^r (1 - a \eta^r / r^{r/2})^n
+\end{aligned}
+$$
+
+Setting this $\geq 1 - \epsilon$ and solving for $n$ yields the desired rate. 
+\end{proof}
+
 \bibliographystyle{agsm}

### 2022-04-22 draft/draft.Rmd
added proof of theorem 1

@@
 ---
-title: Manifold Clustering for Latent Structure Block Models
+title: Manifold Clustering in the Generalized Random Dot Product Graph
 
@@
 If there are manifolds $\mathcal{M}_1, ..., \mathcal{M}_K \subset \mathcal{X}$ for $\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p,q} y \in [0, 1] \}$, and 
-$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}$ such that each $x_i \in \mathcal{M}_k$ for some $k \in [K]$, then $A \sim \GRDPG_{p,q}(X; \rho_n)$ is a \emph{manifold block model}.
+$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$ such that each $x_i \in \mathcal{M}_k$ for some $k \in [K]$, then $A \sim \GRDPG_{p,q}(X; \rho_n)$ is a \emph{manifold block model}.
 \end{definition} 
-
-In practice, we define each manifold $\mathcal{M}_k$ by a function $f_k : [0, 1]^r \to \mathcal{X}$, where $1 \geq r < d$ is the dimensionality of manifold $\mathcal{M}_k$.
+In practice, we define each manifold $\mathcal{M}_k$ by a continuous function $g_k : [0, 1]^r \to \mathcal{X}$, where $1 \leq r < d$ is the dimensionality of manifold $\mathcal{M}_k$.
 
@@
 Define two manifolds by $f_1(t) = \begin{bmatrix} \cos(\frac{\pi}{3} t) & \sin(\frac{\pi}{3} t) \end{bmatrix}^\top$ and $f_2(t) = \begin{bmatrix} 1 - \cos(\frac{\pi}{3} t) & 1 - \sin(\frac{\pi}{3} t) \end{bmatrix}^\top$.
-Draw $t_1, ..., t_n \sim \Uniform(0, 1)$ and $z_1, ..., z_n \Multinomial(\frac{1}{2}, \frac{1}{2})$, and compute latent vectors $x_i = f_{z_i}(t_i)$, which are collected in data matrix $\begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$.
+Draw $t_1, ..., t_n \sim \Uniform(0, 1)$ and $z_1, ..., z_n \iid \Multinomial(\frac{1}{2}, \frac{1}{2})$, and compute latent vectors $x_i = f_{z_i}(t_i)$, which are collected in data matrix $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$.
 Finally, let $A \sim \RDPG(X)$.
@@
 plot(data.matrix, col = z * 2, asp = 1,
-     xlab = NA, ylab = NA, xaxt = 'n', yaxt = 'n',
+     xlab = NA, ylab = NA, 
+     # xaxt = 'n', yaxt = 'n',
      xlim = c(-.1, 1.1), ylim = c(0, 1))
@@
 plot(Xhat, col = z * 2, asp = 1, 
-     xlab = NA, ylab = NA, xaxt = 'n', yaxt = 'n')
+     # xaxt = 'n', yaxt = 'n', 
+     xlab = NA, ylab = NA)
 ```
@@
 To prove theorems \ref{nonintersect-no-noise}, \ref{nonintersect-with-noise}, and \ref{nonintersect-mbm}, we first need to introduce some theory for order statistics. 
-Let $D_i = X_{(i+1)} - X_{(i)}$. Then if $\max_i D_i < \delta$, we have sufficient separation of points in $\mathcal{M}_1$. 
-Then it is sufficient to quantify $P(\max_i D_i > \delta)$ as a function of $n$ and $\delta$ and show that this converges to zero as $n$ grows to $\infty$.
+Let $D_i^{(k)} = X_{(i+1)}^{(k)} - X_{(i)}^{(k)}$, where $X_{(i)}^{(k)}$ is the $i^{th}$ order statistic of univariate sample $X_1^{(k)}, ..., X_n^{(k)} \iid F^{(k)}$ and $F^{(k)}$ is a continuous distribution on manifold $\mathcal{M}_k$.
+If $\max_i D_i^{(k)} < \delta$, then there is sufficient separation of points between each manifold. 
+Then it is sufficient to quantify $P(\max_i D_i^{(k)} > \delta)$ for each $k$ as a function of $n$ and $\delta$ and show that this converges to zero as $n$ grows to $\infty$.
 
-We denote $f(x)$ as the density of each $X_i$, $g_i(x)$ as the density of $X_{(i)}$, 
+For ease of notation, we drop the superscript denoting community/manifold membership until lemma 6. 
+Denote $f(x)$ as the density of each $F$, $g_i(x)$ as the density of $X_{(i)}$, 
 $g_{ij}(x, y)$ as the joint density of $X_{(i)}, X_{(j)}$, and $h_i(d)$ as the density of $D_i$ (with corresponding capital letters for the cumulative distribution functions).
@@
 \begin{lemma}[The probability density function of $D_i$]
+Suppose $X_1, ..., X_n \iid F$ and the support of $F$ is the unit interval. 
+Let $D_i = X_{(i+1)} - X_{(i)}$. 
+Then the density of each $D_i$ is:
+
 \begin{equation}
@@
 \begin{proof}
-This is just a direct consequence of 2 and 3 under the given statements. 
+This is a direct consequence of 2 and 3. 
 We also note that because the support of $X_i$ is $[0, 1]$, 
@@
 \begin{lemma}[The cumulative distribution function of $D_i$]
+Suppose $X_1, ..., X_n \iid F$ and the support of $F$ is the unit interval. 
+Let $D_i = X_{(i+1)} - X_{(i)}$. 
+Then the distribution function of each $D_i$ is:
 
@@
 
-Because of the $x + \delta$ term, we can't actually evaluate this integral all the way up to $1$, and so we are left with 
+Because of the $x + \delta$ term and the fact that the support is $[0, 1]$, the integrand is zero above $1 - \delta$, so we are left with
+
 $$
@@
 $$
-h_i(d) = \int_0^{1-d} \frac{n!}{(i-1)! (n-i-1)!} x^{i-1} (1-x-d)^{n-i-1} \dd x
+h_i(d) = \int_0^{1-d} \frac{n!}{(i-1)! (n-i-1)!} x^{i-1} (1-x-d)^{n-i-1} \dd x.
 $$
@@
 $$
-= \frac{n!}{(i-1)! (n-i-1)!} \int \frac{i-1}{n-i} x^{i-2} (1-x-d)^{n-i} \dd x
+= \frac{n!}{(i-1)! (n-i-1)!} \int \frac{i-1}{n-i} x^{i-2} (1-x-d)^{n-i} \dd x.
 $$
@@
 & = -\frac{n (n-1)}{n-1} (1-x-d)^{n-1} \Big|_0^{1-d} \\
-& = n (1 - d)^{n-1}
+& = n (1 - d)^{n-1}.
 \end{aligned}
@@
 
-\begin{proof}[Proof (sketch)]
+\begin{proof}
 Since $X_{(i+1)} - X_{(i)} = D_i \sim \Betadist(1, n)$, 
@@
 \begin{lemma}
+\label{lem:lim-d}
 Let $X_1, ..., X_n \iid F$ with support $[0, 1]$, and suppose $f(x)$ is continuous and $f(x) \geq a > 0$ everywhere on the support. 
@@
 
-\begin{proof}[Proof (sketch)]
+\begin{proof}
 We start with Eq. (\ref{eq:cdf}):
@@
 
-Here, we extend the results on the line to the unit hypercube. The following theorem is a direct consequence of Lemma 2 of \citet{trosset2020rehabilitating}.
+\begin{proof}[Proof of Theorem \ref{nonintersect-no-noise}]
+For this theorem to be true, there are two additional conditions: 
+First, each $n_k \to \infty$ as $n \to \infty$, or the sample size on each manifold grows with $n$, and second, the number of manifolds $K$ remains fixed. 
+Then this is a direct consequence of Lemma \ref{lem:lim-d}. 
+Here, we denote $m = \min_k n_k$ as the sample size of the manifold with the fewest points.
+
+$$
+\begin{aligned}
+P(\max_{k,i} \|X_{(i+1)}^{(k)} - X_{(i)}^{(k)}\| < \delta) & = P(\forall k, i ~ D_i^{(k)} < \delta) \\
+& \geq \prod_{k=1}^K P(D_i^{(k)} < \delta) \\
+& \geq (1 - (m-1) (1 - a \delta)^{m} + (m-1) (1 - F(1-\delta) - a \delta)^{m})^K
+\end{aligned}
+$$
+Since as $n \to \infty$, $m \to \infty$ as well, and $K$ is fixed, this approaches $(1)^K = 1$. 
+\end{proof}
+
+Here, we extend the results on the line to the unit hypercube. The following theorem is based on Lemma 2 of \citet{trosset2020rehabilitating}.
 
@@
 \end{proof}
+
+\section{Details on Fitting Bezier Curves with Noise}

### 2022-04-22 draft/draft.tex
added proof of theorem 1

@@
 {
-  \title{\bf Manifold Clustering for Latent Structure Block Models}
+  \title{\bf Manifold Clustering in the Generalized Random Dot Product
+Graph}
 
@@
   \begin{center}
-    {\LARGE\bf Manifold Clustering for Latent Structure Block Models}
+    {\LARGE\bf Manifold Clustering in the Generalized Random Dot Product
+Graph}
   \end{center}
@@
 If there are manifolds $\mathcal{M}_1, ..., \mathcal{M}_K \subset \mathcal{X}$ for $\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p,q} y \in [0, 1] \}$, and 
-$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}$ such that each $x_i \in \mathcal{M}_k$ for some $k \in [K]$, then $A \sim \mathrm{GRDPG}_{p,q}(X; \rho_n)$ is a \emph{manifold block model}.
+$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$ such that each $x_i \in \mathcal{M}_k$ for some $k \in [K]$, then $A \sim \mathrm{GRDPG}_{p,q}(X; \rho_n)$ is a \emph{manifold block model}.
 \end{definition}
 
-In practice, we define each manifold \(\mathcal{M}_k\) by a function
-\(f_k : [0, 1]^r \to \mathcal{X}\), where \(1 \geq r < d\) is the
-dimensionality of manifold \(\mathcal{M}_k\).
+In practice, we define each manifold \(\mathcal{M}_k\) by a continuous
+function \(g_k : [0, 1]^r \to \mathcal{X}\), where \(1 \leq r < d\) is
+the dimensionality of manifold \(\mathcal{M}_k\).
 
@@
 Define two manifolds by $f_1(t) = \begin{bmatrix} \cos(\frac{\pi}{3} t) & \sin(\frac{\pi}{3} t) \end{bmatrix}^\top$ and $f_2(t) = \begin{bmatrix} 1 - \cos(\frac{\pi}{3} t) & 1 - \sin(\frac{\pi}{3} t) \end{bmatrix}^\top$.
-Draw $t_1, ..., t_n \sim \mathrm{Uniform}(0, 1)$ and $z_1, ..., z_n \mathrm{Multinomial}(\frac{1}{2}, \frac{1}{2})$, and compute latent vectors $x_i = f_{z_i}(t_i)$, which are collected in data matrix $\begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$.
+Draw $t_1, ..., t_n \sim \mathrm{Uniform}(0, 1)$ and $z_1, ..., z_n \stackrel{\mathrm{iid}}{\sim}\mathrm{Multinomial}(\frac{1}{2}, \frac{1}{2})$, and compute latent vectors $x_i = f_{z_i}(t_i)$, which are collected in data matrix $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$.
 Finally, let $A \sim \mathrm{RDPG}(X)$.
@@
 to introduce some theory for order statistics. Let
-\(D_i = X_{(i+1)} - X_{(i)}\). Then if \(\max_i D_i < \delta\), we have
-sufficient separation of points in \(\mathcal{M}_1\). Then it is
-sufficient to quantify \(P(\max_i D_i > \delta)\) as a function of \(n\)
+\(D_i^{(k)} = X_{(i+1)}^{(k)} - X_{(i)}^{(k)}\), where \(X_{(i)}^{(k)}\)
+is the \(i^{th}\) order statistic of univariate sample
+\(X_1^{(k)}, ..., X_n^{(k)} \stackrel{\mathrm{iid}}{\sim}F^{(k)}\) and
+\(F^{(k)}\) is a continuous distribution on manifold \(\mathcal{M}_k\).
+If \(\max_i D_i^{(k)} < \delta\), then there is sufficient separation of
+points between each manifold. Then it is sufficient to quantify
+\(P(\max_i D_i^{(k)} > \delta)\) for each \(k\) as a function of \(n\)
 and \(\delta\) and show that this converges to zero as \(n\) grows to
@@
 
-We denote \(f(x)\) as the density of each \(X_i\), \(g_i(x)\) as the
-density of \(X_{(i)}\), \(g_{ij}(x, y)\) as the joint density of
-\(X_{(i)}, X_{(j)}\), and \(h_i(d)\) as the density of \(D_i\) (with
-corresponding capital letters for the cumulative distribution
-functions).
+For ease of notation, we drop the superscript denoting
+community/manifold membership until lemma 6. Denote \(f(x)\) as the
+density of each \(F\), \(g_i(x)\) as the density of \(X_{(i)}\),
+\(g_{ij}(x, y)\) as the joint density of \(X_{(i)}, X_{(j)}\), and
+\(h_i(d)\) as the density of \(D_i\) (with corresponding capital letters
+for the cumulative distribution functions).
 
@@
 \begin{lemma}[The probability density function of $D_i$]
+Suppose $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}F$ and the support of $F$ is the unit interval. 
+Let $D_i = X_{(i+1)} - X_{(i)}$. 
+Then the density of each $D_i$ is:
+
 \begin{equation}
@@
 \begin{lemma}[The cumulative distribution function of $D_i$]
+Suppose $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}F$ and the support of $F$ is the unit interval. 
+Let $D_i = X_{(i+1)} - X_{(i)}$. 
+Then the distribution function of each $D_i$ is:
 
@@
 $$
-h_i(d) = \int_0^{1-d} \frac{n!}{(i-1)! (n-i-1)!} x^{i-1} (1-x-d)^{n-i-1} \mathrm{d}x
+h_i(d) = \int_0^{1-d} \frac{n!}{(i-1)! (n-i-1)!} x^{i-1} (1-x-d)^{n-i-1} \mathrm{d}x.
 $$
@@
 $$
-= \frac{n!}{(i-1)! (n-i-1)!} \int \frac{i-1}{n-i} x^{i-2} (1-x-d)^{n-i} \mathrm{d}x
+= \frac{n!}{(i-1)! (n-i-1)!} \int \frac{i-1}{n-i} x^{i-2} (1-x-d)^{n-i} \mathrm{d}x.
 $$
@@
 
-\begin{proof}[Proof (sketch)]
+\begin{proof}
 Since $X_{(i+1)} - X_{(i)} = D_i \sim \mathrm{Beta}(1, n)$, 
@@
 \begin{lemma}
+\label{lem:lim-d}
 Let $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}F$ with support $[0, 1]$, and suppose $f(x)$ is continuous and $f(x) \geq a > 0$ everywhere on the support. 
@@
 
+\begin{proof}[Proof of Theorem \ref{nonintersect-no-noise}]
+For this theorem to be true, there are two additional conditions: 
+First, each $n_k \to \infty$ as $n \to \infty$, or the sample size on each manifold grows with $n$, and second, the number of manifolds $K$ remains fixed. 
+Then this is a direct consequence of Lemma \ref{lem:lim-d}. 
+Here, we denote $m = \min_k n_k$ as the sample size of the manifold with the fewest points.
+
+$$
+\begin{aligned}
+P(\max_{k,i} \|X_{(i+1)}^{(k)} - X_{(i)}^{(k)}\| < \delta) & = P(\forall k, i ~ D_i^{(k)} < \delta) \\
+& \geq \prod_{k=1}^K P(D_i^{(k)} < \delta) \\
+& \geq (1 - (m-1) (1 - a \delta)^{m} + (m-1) (1 - F(1-\delta) - a \delta)^{m})^K
+\end{aligned}
+$$
+Since as $n \to \infty$, $m \to \infty$ as well, and $K$ is fixed, this approaches $(1)^K = 1$. 
+\end{proof}
+
 Here, we extend the results on the line to the unit hypercube. The
-following theorem is a direct consequence of Lemma 2 of
+following theorem is based on Lemma 2 of
 \citet{trosset2020rehabilitating}.
@@
 
+\section{Details on Fitting Bezier Curves with Noise}
+
 \bibliographystyle{agsm}

### 2022-04-24 draft/draft.Rmd
still cleaning up proofs of theorems 2 and 3

@@
 \label{nonintersect-with-noise}
-Suppose the same setup as in Theorem \ref{nonintersect-no-noise}. 
+Suppose the same setup as in theorem \ref{nonintersect-no-noise}. 
 Define $y_i = x_i + e_i$ and $B_n$ as the event such that $\|y_{(i+1)}^{(k)} - y_{(i)}^{(k)}\| > \|y_{(i)}^{(k)} - y_{(j)}^{(\ell)}\|$ for some $i$, $j$, $k$, and $\ell$. 
@@
 \label{nonintersect-mbm}
-Suppose we have the same setup as in Theorem \ref{nonintersect-no-noise}. 
+Suppose we have the same setup as in theorem \ref{nonintersect-no-noise}. 
 Let $p, q \geq 0$ such that $p + q = d$ where $d$ is the dimension of the range of each $g_k$. 
@@
 
-This converges to $1$ in the limit $n \to \infty$.
+Since $(1 - a \delta) < 1$ and $(1 - F(1-\delta) - a \delta) < 1$, this converges to $1$ in the limit $n \to \infty$. 
+Then for any $\epsilon \in (0, 1)$, we can always solve for an $n$ such that $1 - (n-1) (1 - a \delta)^n (n-1) (1 - F(1 - \delta) - a \delta)^n \geq 1 - \epsilon$ holds. 
 \end{proof}
 
-\begin{proof}[Proof of Theorem \ref{nonintersect-no-noise}]
-For this theorem to be true, there are two additional conditions: 
-First, each $n_k \to \infty$ as $n \to \infty$, or the sample size on each manifold grows with $n$, and second, the number of manifolds $K$ remains fixed. 
+\begin{proof}[Proof of theorem \ref{nonintersect-no-noise}]
+This theorem requires two additional conditions: 
+First, each $n_k \to \infty$ as $n \to \infty$, and second, the number of manifolds $K$ remains fixed. 
+These are satisified by the model assumptions. 
 Then this is a direct consequence of Lemma \ref{lem:lim-d}. 
-Here, we denote $m = \min_k n_k$ as the sample size of the manifold with the fewest points.
+Denoting $m = \min_k n_k$ as the sample size of the manifold with the fewest points,
 
@@
 
-Here, we extend the results on the line to the unit hypercube. The following theorem is based on Lemma 2 of \citet{trosset2020rehabilitating}.
+\begin{proof}[Proof of theorem \ref{nonintersect-with-noise}]
+\end{proof}
+
+\begin{proof}[Proof of theorem \ref{nonintersect-mbm}]
+\end{proof}
+
+Here, we extend the results on the line to the unit hypercube. The following lemma is based on lemma 2 of \citet{trosset2020rehabilitating}.
 

### 2022-04-24 draft/draft.tex
still cleaning up proofs of theorems 2 and 3

@@
 
+\begin{proof}[Proof of theorem \ref{nonintersect-with-noise}]
+\end{proof}
+
+\begin{proof}[Proof of theorem \ref{nonintersect-mbm}]
+\end{proof}
+
 Here, we extend the results on the line to the unit hypercube. The
-following theorem is based on Lemma 2 of
+following lemma is based on lemma 2 of
 \citet{trosset2020rehabilitating}.

### 2022-04-27 draft/draft.Rmd
still working ...

@@
 \end{definition}
-In practice, we define each manifold $\mathcal{M}_k$ by a continuous function $g_k : [0, 1]^r \to \mathcal{X}$, where $1 \leq r < d$ is the dimensionality of manifold $\mathcal{M}_k$.
 
-Insert something about ASE and its consistency here.
+In practice, we often define each manifold $\mathcal{M}_k$ by a continuous function $g_k : [0, 1]^r \to \mathcal{X}$, where $1 \leq r < d$ is the dimensionality of manifold $\mathcal{M}_k$. 
+To complete the probability model, we define a probability distribution $F$ with support $[0, 1]^r$ and draw the community memberships from a multinomial distribution.
+
+The full mixture model can be described as follows:
+
+1. Draw labels $z_1, ..., z_n \iid \Multinomial(\alpha_1, ..., \alpha_K)$. 
+2. Draw latent vectors by first drawing each $t_1, ..., t_n \iid F$ and then computing each $x_i = g_{z_i}(t_i)$. 
+3. Let $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$, 
+and draw $A \sim \RDPG(X)$ or $A \sim \GRDPG_{p,q}(X)$.
 
@@
 Let $p = 2$, $q = 0$, $K = 2$, and $r = 1$. 
-Define two manifolds by $f_1(t) = \begin{bmatrix} \cos(\frac{\pi}{3} t) & \sin(\frac{\pi}{3} t) \end{bmatrix}^\top$ and $f_2(t) = \begin{bmatrix} 1 - \cos(\frac{\pi}{3} t) & 1 - \sin(\frac{\pi}{3} t) \end{bmatrix}^\top$.
-Draw $t_1, ..., t_n \sim \Uniform(0, 1)$ and $z_1, ..., z_n \iid \Multinomial(\frac{1}{2}, \frac{1}{2})$, and compute latent vectors $x_i = f_{z_i}(t_i)$, which are collected in data matrix $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$.
+Define two one-dimensional manifolds by $f_1(t) = \begin{bmatrix} \cos(\frac{\pi}{3} t) & \sin(\frac{\pi}{3} t) \end{bmatrix}^\top$ and $f_2(t) = \begin{bmatrix} 1 - \cos(\frac{\pi}{3} t) & 1 - \sin(\frac{\pi}{3} t) \end{bmatrix}^\top$.
+Draw $t_1, ..., t_n \iid \Uniform(0, 1)$ and $z_1, ..., z_n \iid \Multinomial(\frac{1}{2}, \frac{1}{2})$, and compute latent vectors $x_i = f_{z_i}(t_i)$, which are collected in data matrix $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$.
 Finally, let $A \sim \RDPG(X)$.
+Fig \ref{fig:example1} shows the latent configuration drawn from this latent distribution, a random dot product graph drawn from the latent configuration, and the adjacency spectral embedding of the graph.
 
-```{r fig1, fig.cap = 'Manifold block model described in Example 1.
+```{r example1, fig.cap = 'Manifold block model described in Example 1.
 The latent configuration is on the left, and a random dot product graph drawn from the latent configuration is on the middle, and the ASE is on the right.', fig.height = 2.5, fig.width = 7}
@@
 ## Nonintersecting Manifolds
+\label{section:nonintersecting}
 
-In this section, we consider the following scenario: Suppose that each community is represented by a manifold $\mathcal{M}_k$, $k \in \{1, ..., K\}$ in the latent space of a RDPG or GDRPG, and the manifolds do not intersect each other.
-Define $\delta = \min\limits_{k, \ell} \min\limits_{x \in \mathcal{M}_k, y \in \mathcal{M}_\ell} \|x - y\|$, the minimum distance between two manifolds. 
-We assume that $\delta > 0$.
+In this section, we consider the following scenario: 
+Suppose that each community is represented by a manifold $\mathcal{M}_k$, $k \in \{1, ..., K\}$ in the latent space of a RDPG or GDRPG, and the manifolds do not intersect each other. 
+Define $\delta = \min\limits_{k \neq \ell} \min\limits_{x \in \mathcal{M}_k, y \in \mathcal{M}_\ell} \|x - y\|$, the minimum distance between two manifolds. 
+We assume that $\delta > 0$, i.e., the manifolds do not intersect.
 
-First, we consider the case in which each $\mathcal{M}_k \subset \mathbb{R}^d$ is a one-dimensional manifold, which can be characterized by a function $g_k : [0, 1] \to \mathcal{M}_k$. 
-Let $F$ be a probability distribution with support $[0, 1]$. 
-Then we define a mixture model as follows:
-
-1. Draw $t_1, ..., t_n \iid F$.
-2. Draw $z_1, ..., z_n \iid \Multinomial(\alpha_1, ..., \alpha_K)$, the community labels.
-3. Let each $x_i = g_{z_i}(t_i)$ be the latent vector for vertex $v_i$.  
-Assume $\min_{k, \ell} \min_{t, s} \|g_k(t) - g_\ell(s)\| = \delta > 0$. 
-4. Draw $A \sim \RDPG(X)$ or $A \sim \GRDPG_{p,q}(X)$.
-
-\begin{theorem}[Community detection for one-dimensional nonintersecting manifolds without noise]
-\label{nonintersect-no-noise}
-Let $x_1, ..., x_n$ be points sampled on $K$ one-dimensional manifolds $\mathcal{M}_1, ..., \mathcal{M}_K$, defined by arc-length parameterizations $g_1, ..., g_K$. 
-Suppose $\delta = \min_{k, \ell} \min_{t, s} \| g_k(t) - g_\ell(s) \| > 0$. 
-Let $A_n$ be the event such that $\max_k \max_i \|x_{(i+1)}^{(k)} - x_{(i)}^{(k)}\| > \delta$, where $x_{(i)}^{(k)}$ is the $i^{th}$ order statistic on the $k^{th}$ manifold. 
-Then for any $\epsilon > 0$, there exists an $N \in \mathbb{N}$ such that when $n > N$, $P(A_n) < \epsilon$. 
-\end{theorem}
-
-\begin{theorem}[Community detection for one-dimensional nonintersecting manifolds with noise]
-\label{nonintersect-with-noise}
-Suppose the same setup as in theorem \ref{nonintersect-no-noise}. 
-Define $y_i = x_i + e_i$ and $B_n$ as the event such that $\|y_{(i+1)}^{(k)} - y_{(i)}^{(k)}\| > \|y_{(i)}^{(k)} - y_{(j)}^{(\ell)}\|$ for some $i$, $j$, $k$, and $\ell$. 
-Then if $\max_i \|e_i\| < \delta / 3$, for any $\epsilon > 0$, there exists an $N \in \mathbb{N}$ such that when $n > N$, $P(B_n) < \epsilon$. 
-\end{theorem}
-
-\begin{theorem}[Community detection for one-dimensional nonintersecting manifold block models]
-\label{nonintersect-mbm}
-Suppose we have the same setup as in theorem \ref{nonintersect-no-noise}. 
-Let $p, q \geq 0$ such that $p + q = d$ where $d$ is the dimension of the range of each $g_k$. 
-Suppose we draw $A \sim \GRDPG_{p,q}(X; \rho_n)$ for some sparsity parameter $\rho_n \in (0, 1]$, and let $\hat{X}$ be the ASE of $A$. 
-Then for any $\epsilon > 0$, there exists an $N \in \mathbb{N}$ such that when $n > N$, single linkage clustering produces zero community detection error with probability 1. 
-\end{theorem}
+In the noiseless setting, if the subsample on each manifold is sufficiently dense, it is possible to construct for each manifold an $\eta_k$-neighborhood graph for each manifold for some $\eta_k > 0$ such that the graph is connected.
+Then if $\max_k \eta_k = \eta < \delta$, an $\eta$-neighborhood graph for the entire sample will consist of $K$ disconnected subgraphs that map onto each manifold. 
+Equivalently, we can apply single-linkage clustering. 
+The remainder of this section explores under which conditions these criteria are met for the latent configuration, in which latent vectors lie exactly on manifolds, as well as the ASE, which introduces noise.
 
@@
 
+Consider the specific case in which each $\mathcal{M}_k \subset \mathbb{R}^d$ is a one-dimensional manifold. 
+Let $F_k$ be a probability distribution with support $\mathcal{M}_k$. 
+Then we define a mixture model as follows:
+
+1. Draw labels $z_1, ..., z_n \iid \Multinomial(\alpha_1, ..., \alpha_K)$. 
+2. Draw latent vectors each as $x_i \indep F_{z_i}$ for distributions $F_1, ..., F_K$ with respective supports $\mathcal{M}_1, ..., \mathcal{M}_K$. 
+3. Let $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$, 
+and draw $A \sim \RDPG(X)$ or $A \sim \GRDPG_{p,q}(X)$.
+
+Note that here, we redefine the model to ignore $g_1, ..., g_K$, the parameterizations of each manifold. 
+Instead, we sample points directly on the manifolds themselves. 
+We will return to the parameterizations in Section \ref{section:intersecting}.
+
+In this setting, although each manifold is in $d$-dimensional space, since the manifolds are one-dimensional, it is possible to consider an ordering of points sampled on the manifold. 
+In the original latent space in which there is no noise, the condition for which community detection is possible via $\eta$-neighborhood graph clustering or single linkage clustering is each pair of adjacent order statistics on each manifold being less than $\delta$ apart.
+
+\begin{theorem}[Community detection for one-dimensional nonintersecting manifolds without noise]
+\label{nonintersect-no-noise}
+Let $x_1, ..., x_n$ be points sampled on $K$ one-dimensional manifolds $\mathcal{M}_1, ..., \mathcal{M}_K$. 
+Suppose $\delta = \min_{k \neq \ell} \min_{x_i \in \mathcal{M}_k, x_j \in \mathcal{M}_\ell} \| x_i - x_j \| > 0$. 
+Let $A_n$ be the event such that $\max_k \max_i \|x_{(i+1)}^{(k)} - x_{(i)}^{(k)}\| < \delta$, where $x_{(i)}^{(k)}$ is the $i^{th}$ order statistic on the $k^{th}$ manifold. 
+Then for any $\epsilon > 0$, there exists an $N \in \mathbb{N}$ such that when $n > N$, $P(A_n) > 1 - \epsilon$. 
+\end{theorem}
+
+Sampling a graph from this latent configuration and constructing an ASE from the graph introduces error relative to the original latent vectors. 
+If the errors are small, which is guaranteed by the consistency property of the ASE for sufficiently large $n$, then  sufficient separation between embedding points belonging to different manifolds is still guaranteed.
+
+\begin{theorem}[Community detection for one-dimensional nonintersecting manifolds with noise]
+\label{nonintersect-with-noise}
+Suppose the same setup as in theorem \ref{nonintersect-no-noise}. 
+Define $y_i = x_i + e_i$ and $B_n$ as the event such that $\|y_{(i+1)}^{(k)} - y_{(i)}^{(k)}\| < \|y_{(i)}^{(k)} - y_{(j)}^{(\ell)}\|$ for any $k \neq \ell$. 
+Here, $y_{(i)}^{(k)}$ is the $i^{th}$ order statistic on the $k^{th}$ manifold plus its corresponding noise. 
+Then if $\max_i \|e_i\| < \delta / 3$, for any $\epsilon > 0$, there exists an $N \in \mathbb{N}$ such that when $n > N$, $P(B_n) > 1 - \epsilon$. 
+\end{theorem}
+
+\begin{theorem}[Community detection for one-dimensional nonintersecting manifold block models]
+\label{nonintersect-mbm}
+Suppose we have the same setup as in theorem \ref{nonintersect-no-noise}. 
+Let $p, q \geq 0$ such that $p + q = d$ where $d$ is the dimension of the range of each $g_k$. 
+Suppose we draw $A \sim \GRDPG_{p,q}(X; \rho_n)$ for some sparsity parameter $\rho_n \in (0, 1]$, and let $\hat{X}$ be the ASE of $A$. 
+Then for any $\epsilon > 0$, there exists an $N \in \mathbb{N}$ such that when $n > N$, single linkage clustering produces zero community detection error with high probability. 
+\end{theorem}
+
 ## Intersecting Manifolds
+\label{section:intersecting}
 
@@
 However, this time, we do not assume that the manifolds are nonintersecting. 
+We also restrict this case to one-dimensional manifolds which are each described by $g_k : [0, 1] \to \mathcal{X}$. 
+Then we define a mixture model as follows:
+
+1. Draw $t_1, ..., t_n \iid F$ for probability distribution $F$ with support $[0, 1]$.
+2. Draw $z_1, ..., z_n \iid \Multinomial(\alpha_1, ..., \alpha_K)$, the community labels.
+3. Let each $x_i = g_{z_i}(t_i)$ be the latent vector for vertex $v_i$, and collect the latent vectors into matrix $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$.
+4. Draw $A \sim \RDPG(X)$ or $A \sim \GRDPG_{p,q}(X)$.
 
@@
 First, each $n_k \to \infty$ as $n \to \infty$, and second, the number of manifolds $K$ remains fixed. 
-These are satisified by the model assumptions. 
+These are satisfied by the model assumptions. 
 Then this is a direct consequence of Lemma \ref{lem:lim-d}. 
@@
 $$
-Since as $n \to \infty$, $m \to \infty$ as well, and $K$ is fixed, this approaches $(1)^K = 1$. 
+Since as $n \to \infty$, $m \to \infty$ as well, and $K$ is fixed, this approaches $(1)^K = 1$. Then setting this $\geq 1 - \epsilon$, it is always possible to solve this for $m$. 
 \end{proof}
@@
 \begin{proof}[Proof of theorem \ref{nonintersect-with-noise}]
+Let $\max_i \|e_i\| \leq \xi$. 
 \end{proof}

### 2022-04-27 draft/draft.tex
still working ...

@@
 
-In practice, we define each manifold \(\mathcal{M}_k\) by a continuous
-function \(g_k : [0, 1]^r \to \mathcal{X}\), where \(1 \leq r < d\) is
-the dimensionality of manifold \(\mathcal{M}_k\).
+In practice, we often define each manifold \(\mathcal{M}_k\) by a
+continuous function \(g_k : [0, 1]^r \to \mathcal{X}\), where
+\(1 \leq r < d\) is the dimensionality of manifold \(\mathcal{M}_k\). To
+complete the probability model, we define a probability distribution
+\(F\) with support \([0, 1]^r\) and draw the community memberships from
+a multinomial distribution.
 
-Insert something about ASE and its consistency here.
+The full mixture model can be described as follows:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Draw labels
+  \(z_1, ..., z_n \stackrel{\mathrm{iid}}{\sim}\mathrm{Multinomial}(\alpha_1, ..., \alpha_K)\).
+\item
+  Draw latent vectors by first drawing each
+  \(t_1, ..., t_n \stackrel{\mathrm{iid}}{\sim}F\) and then computing
+  each \(x_i = g_{z_i}(t_i)\).
+\item
+  Let \(X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top\), and
+  draw \(A \sim \mathrm{RDPG}(X)\) or
+  \(A \sim \mathrm{GRDPG}_{p,q}(X)\).
+\end{enumerate}
 
@@
 Let $p = 2$, $q = 0$, $K = 2$, and $r = 1$. 
-Define two manifolds by $f_1(t) = \begin{bmatrix} \cos(\frac{\pi}{3} t) & \sin(\frac{\pi}{3} t) \end{bmatrix}^\top$ and $f_2(t) = \begin{bmatrix} 1 - \cos(\frac{\pi}{3} t) & 1 - \sin(\frac{\pi}{3} t) \end{bmatrix}^\top$.
-Draw $t_1, ..., t_n \sim \mathrm{Uniform}(0, 1)$ and $z_1, ..., z_n \stackrel{\mathrm{iid}}{\sim}\mathrm{Multinomial}(\frac{1}{2}, \frac{1}{2})$, and compute latent vectors $x_i = f_{z_i}(t_i)$, which are collected in data matrix $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$.
+Define two one-dimensional manifolds by $f_1(t) = \begin{bmatrix} \cos(\frac{\pi}{3} t) & \sin(\frac{\pi}{3} t) \end{bmatrix}^\top$ and $f_2(t) = \begin{bmatrix} 1 - \cos(\frac{\pi}{3} t) & 1 - \sin(\frac{\pi}{3} t) \end{bmatrix}^\top$.
+Draw $t_1, ..., t_n \stackrel{\mathrm{iid}}{\sim}\mathrm{Uniform}(0, 1)$ and $z_1, ..., z_n \stackrel{\mathrm{iid}}{\sim}\mathrm{Multinomial}(\frac{1}{2}, \frac{1}{2})$, and compute latent vectors $x_i = f_{z_i}(t_i)$, which are collected in data matrix $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$.
 Finally, let $A \sim \mathrm{RDPG}(X)$.
+Fig \ref{fig:example1} shows the latent configuration drawn from this latent distribution, a random dot product graph drawn from the latent configuration, and the adjacency spectral embedding of the graph.
 
@@
 
-{\centering \includegraphics{draft_files/figure-latex/fig1-1}
+{\centering \includegraphics{draft_files/figure-latex/example1-1}
 
@@
 \caption{Manifold block model described in Example 1.
-The latent configuration is on the left, and a random dot product graph drawn from the latent configuration is on the middle, and the ASE is on the right.}\label{fig:fig1}
+The latent configuration is on the left, and a random dot product graph drawn from the latent configuration is on the middle, and the ASE is on the right.}\label{fig:example1}
 \end{figure}
@@
 
+\label{section:nonintersecting}
+
 In this section, we consider the following scenario: Suppose that each
@@
 manifolds do not intersect each other. Define
-\(\delta = \min\limits_{k, \ell} \min\limits_{x \in \mathcal{M}_k, y \in \mathcal{M}_\ell} \|x - y\|\),
+\(\delta = \min\limits_{k \neq \ell} \min\limits_{x \in \mathcal{M}_k, y \in \mathcal{M}_\ell} \|x - y\|\),
 the minimum distance between two manifolds. We assume that
-\(\delta > 0\).
+\(\delta > 0\), i.e., the manifolds do not intersect.
 
-First, we consider the case in which each
-\(\mathcal{M}_k \subset \mathbb{R}^d\) is a one-dimensional manifold,
-which can be characterized by a function
-\(g_k : [0, 1] \to \mathcal{M}_k\). Let \(F\) be a probability
-distribution with support \([0, 1]\). Then we define a mixture model as
-follows:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Draw \(t_1, ..., t_n \stackrel{\mathrm{iid}}{\sim}F\).
-\item
-  Draw
-  \(z_1, ..., z_n \stackrel{\mathrm{iid}}{\sim}\mathrm{Multinomial}(\alpha_1, ..., \alpha_K)\),
-  the community labels.
-\item
-  Let each \(x_i = g_{z_i}(t_i)\) be the latent vector for vertex
-  \(v_i\).\\
-  Assume
-  \(\min_{k, \ell} \min_{t, s} \|g_k(t) - g_\ell(s)\| = \delta > 0\).
-\item
-  Draw \(A \sim \mathrm{RDPG}(X)\) or
-  \(A \sim \mathrm{GRDPG}_{p,q}(X)\).
-\end{enumerate}
-
-\begin{theorem}[Community detection for one-dimensional nonintersecting manifolds without noise]
-\label{nonintersect-no-noise}
-Let $x_1, ..., x_n$ be points sampled on $K$ one-dimensional manifolds $\mathcal{M}_1, ..., \mathcal{M}_K$, defined by arc-length parameterizations $g_1, ..., g_K$. 
-Suppose $\delta = \min_{k, \ell} \min_{t, s} \| g_k(t) - g_\ell(s) \| > 0$. 
-Let $A_n$ be the event such that $\max_k \max_i \|x_{(i+1)}^{(k)} - x_{(i)}^{(k)}\| > \delta$, where $x_{(i)}^{(k)}$ is the $i^{th}$ order statistic on the $k^{th}$ manifold. 
-Then for any $\epsilon > 0$, there exists an $N \in \mathbb{N}$ such that when $n > N$, $P(A_n) < \epsilon$. 
-\end{theorem}
-
-\begin{theorem}[Community detection for one-dimensional nonintersecting manifolds with noise]
-\label{nonintersect-with-noise}
-Suppose the same setup as in theorem \ref{nonintersect-no-noise}. 
-Define $y_i = x_i + e_i$ and $B_n$ as the event such that $\|y_{(i+1)}^{(k)} - y_{(i)}^{(k)}\| > \|y_{(i)}^{(k)} - y_{(j)}^{(\ell)}\|$ for some $i$, $j$, $k$, and $\ell$. 
-Then if $\max_i \|e_i\| < \delta / 3$, for any $\epsilon > 0$, there exists an $N \in \mathbb{N}$ such that when $n > N$, $P(B_n) < \epsilon$. 
-\end{theorem}
-
-\begin{theorem}[Community detection for one-dimensional nonintersecting manifold block models]
-\label{nonintersect-mbm}
-Suppose we have the same setup as in theorem \ref{nonintersect-no-noise}. 
-Let $p, q \geq 0$ such that $p + q = d$ where $d$ is the dimension of the range of each $g_k$. 
-Suppose we draw $A \sim \mathrm{GRDPG}_{p,q}(X; \rho_n)$ for some sparsity parameter $\rho_n \in (0, 1]$, and let $\hat{X}$ be the ASE of $A$. 
-Then for any $\epsilon > 0$, there exists an $N \in \mathbb{N}$ such that when $n > N$, single linkage clustering produces zero community detection error with probability 1. 
-\end{theorem}
+In the noiseless setting, if the subsample on each manifold is
+sufficiently dense, it is possible to construct for each manifold an
+\(\eta_k\)-neighborhood graph for each manifold for some \(\eta_k > 0\)
+such that the graph is connected. Then if
+\(\max_k \eta_k = \eta < \delta\), an \(\eta\)-neighborhood graph for
+the entire sample will consist of \(K\) disconnected subgraphs that map
+onto each manifold. Equivalently, we can apply single-linkage
+clustering. The remainder of this section explores under which
+conditions these criteria are met for the latent configuration, in which
+latent vectors lie exactly on manifolds, as well as the ASE, which
+introduces noise.
 
@@
 
+Consider the specific case in which each
+\(\mathcal{M}_k \subset \mathbb{R}^d\) is a one-dimensional manifold.
+Let \(F_k\) be a probability distribution with support
+\(\mathcal{M}_k\). Then we define a mixture model as follows:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Draw labels
+  \(z_1, ..., z_n \stackrel{\mathrm{iid}}{\sim}\mathrm{Multinomial}(\alpha_1, ..., \alpha_K)\).
+\item
+  Draw latent vectors each as
+  \(x_i \stackrel{\mathrm{ind}}{\sim}F_{z_i}\) for distributions
+  \(F_1, ..., F_K\) with respective supports
+  \(\mathcal{M}_1, ..., \mathcal{M}_K\).
+\item
+  Let \(X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top\), and
+  draw \(A \sim \mathrm{RDPG}(X)\) or
+  \(A \sim \mathrm{GRDPG}_{p,q}(X)\).
+\end{enumerate}
+
+Note that here, we redefine the model to ignore \(g_1, ..., g_K\), the
+parameterizations of each manifold. Instead, we sample points directly
+on the manifolds themselves. We will return to the parameterizations in
+Section \ref{section:intersecting}.
+
+In this setting, although each manifold is in \(d\)-dimensional space,
+since the manifolds are one-dimensional, it is possible to consider an
+ordering of points sampled on the manifold. In the original latent space
+in which there is no noise, the condition for which community detection
+is possible via \(\eta\)-neighborhood graph clustering or single linkage
+clustering is each pair of adjacent order statistics on each manifold
+being less than \(\delta\) apart.
+
+\begin{theorem}[Community detection for one-dimensional nonintersecting manifolds without noise]
+\label{nonintersect-no-noise}
+Let $x_1, ..., x_n$ be points sampled on $K$ one-dimensional manifolds $\mathcal{M}_1, ..., \mathcal{M}_K$. 
+Suppose $\delta = \min_{k \neq \ell} \min_{x_i \in \mathcal{M}_k, x_j \in \mathcal{M}_\ell} \| x_i - x_j \| > 0$. 
+Let $A_n$ be the event such that $\max_k \max_i \|x_{(i+1)}^{(k)} - x_{(i)}^{(k)}\| < \delta$, where $x_{(i)}^{(k)}$ is the $i^{th}$ order statistic on the $k^{th}$ manifold. 
+Then for any $\epsilon > 0$, there exists an $N \in \mathbb{N}$ such that when $n > N$, $P(A_n) > 1 - \epsilon$. 
+\end{theorem}
+
+Sampling a graph from this latent configuration and constructing an ASE
+from the graph introduces error relative to the original latent vectors.
+If the errors are small, which is guaranteed by the consistency property
+of the ASE for sufficiently large \(n\), then sufficient separation
+between embedding points belonging to different manifolds is still
+guaranteed.
+
+\begin{theorem}[Community detection for one-dimensional nonintersecting manifolds with noise]
+\label{nonintersect-with-noise}
+Suppose the same setup as in theorem \ref{nonintersect-no-noise}. 
+Define $y_i = x_i + e_i$ and $B_n$ as the event such that $\|y_{(i+1)}^{(k)} - y_{(i)}^{(k)}\| < \|y_{(i)}^{(k)} - y_{(j)}^{(\ell)}\|$ for any $k \neq \ell$. 
+Here, $y_{(i)}^{(k)}$ is the $i^{th}$ order statistic on the $k^{th}$ manifold plus its corresponding noise. 
+Then if $\max_i \|e_i\| < \delta / 3$, for any $\epsilon > 0$, there exists an $N \in \mathbb{N}$ such that when $n > N$, $P(B_n) > 1 - \epsilon$. 
+\end{theorem}
+
+\begin{theorem}[Community detection for one-dimensional nonintersecting manifold block models]
+\label{nonintersect-mbm}
+Suppose we have the same setup as in theorem \ref{nonintersect-no-noise}. 
+Let $p, q \geq 0$ such that $p + q = d$ where $d$ is the dimension of the range of each $g_k$. 
+Suppose we draw $A \sim \mathrm{GRDPG}_{p,q}(X; \rho_n)$ for some sparsity parameter $\rho_n \in (0, 1]$, and let $\hat{X}$ be the ASE of $A$. 
+Then for any $\epsilon > 0$, there exists an $N \in \mathbb{N}$ such that when $n > N$, single linkage clustering produces zero community detection error with high probability. 
+\end{theorem}
+
 \hypertarget{intersecting-manifolds}{%
@@
 
+\label{section:intersecting}
+
 In this section, we again consider the setting for the RDPG or GRDPG in
 which each community lies on a manifold in the latent space. However,
-this time, we do not assume that the manifolds are nonintersecting.
+this time, we do not assume that the manifolds are nonintersecting. We
+also restrict this case to one-dimensional manifolds which are each
+described by \(g_k : [0, 1] \to \mathcal{X}\). Then we define a mixture
+model as follows:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Draw \(t_1, ..., t_n \stackrel{\mathrm{iid}}{\sim}F\) for probability
+  distribution \(F\) with support \([0, 1]\).
+\item
+  Draw
+  \(z_1, ..., z_n \stackrel{\mathrm{iid}}{\sim}\mathrm{Multinomial}(\alpha_1, ..., \alpha_K)\),
+  the community labels.
+\item
+  Let each \(x_i = g_{z_i}(t_i)\) be the latent vector for vertex
+  \(v_i\), and collect the latent vectors into matrix
+  \(X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top\).
+\item
+  Draw \(A \sim \mathrm{RDPG}(X)\) or
+  \(A \sim \mathrm{GRDPG}_{p,q}(X)\).
+\end{enumerate}
 

### 2022-04-27 draft/draft.Rmd
still working ...

@@
 In this setting, although each manifold is in $d$-dimensional space, since the manifolds are one-dimensional, it is possible to consider an ordering of points sampled on the manifold. 
-In the original latent space in which there is no noise, the condition for which community detection is possible via $\eta$-neighborhood graph clustering or single linkage clustering is each pair of adjacent order statistics on each manifold being less than $\delta$ apart.
+In the original latent space in which there is no noise, the condition for which community detection is possible via $\eta$-neighborhood graph clustering or single linkage clustering is satisfied if each pair of adjacent order statistics on each manifold is less than $\delta$ apart.
 
@@
 Let $x_1, ..., x_n$ be points sampled on $K$ one-dimensional manifolds $\mathcal{M}_1, ..., \mathcal{M}_K$. 
-Suppose $\delta = \min_{k \neq \ell} \min_{x_i \in \mathcal{M}_k, x_j \in \mathcal{M}_\ell} \| x_i - x_j \| > 0$. 
+Suppose $\delta = \min\limits_{k \neq \ell} \min\limits_{x_i \in \mathcal{M}_k, x_j \in \mathcal{M}_\ell} \| x_i - x_j \| > 0$. 
 Let $A_n$ be the event such that $\max_k \max_i \|x_{(i+1)}^{(k)} - x_{(i)}^{(k)}\| < \delta$, where $x_{(i)}^{(k)}$ is the $i^{th}$ order statistic on the $k^{th}$ manifold. 
@@
 Suppose the same setup as in theorem \ref{nonintersect-no-noise}. 
-Define $y_i = x_i + e_i$ and $B_n$ as the event such that $\|y_{(i+1)}^{(k)} - y_{(i)}^{(k)}\| < \|y_{(i)}^{(k)} - y_{(j)}^{(\ell)}\|$ for any $k \neq \ell$. 
-Here, $y_{(i)}^{(k)}$ is the $i^{th}$ order statistic on the $k^{th}$ manifold plus its corresponding noise. 
+Define $y_i = x_i + e_i$ and $B_n$ as the event such that $\|y_{(i+1)}^{(k)} - y_{(i)}^{(k)}\| < \|y_{(i)}^{(k)} - y_{(j)}^{(\ell)}\|$ for any $k \neq \ell$, where, $y_{(i)}^{(k)}$ is the $i^{th}$ order statistic on the $k^{th}$ manifold plus its corresponding noise.
 Then if $\max_i \|e_i\| < \delta / 3$, for any $\epsilon > 0$, there exists an $N \in \mathbb{N}$ such that when $n > N$, $P(B_n) > 1 - \epsilon$. 
@@
 Suppose we draw $A \sim \GRDPG_{p,q}(X; \rho_n)$ for some sparsity parameter $\rho_n \in (0, 1]$, and let $\hat{X}$ be the ASE of $A$. 
-Then for any $\epsilon > 0$, there exists an $N \in \mathbb{N}$ such that when $n > N$, single linkage clustering produces zero community detection error with high probability. 
+Then single linkage clustering produces zero community detection error with high probability as $n \to \infty$. 
 \end{theorem}

### 2022-04-27 draft/draft.tex
still working ...

@@
 is possible via \(\eta\)-neighborhood graph clustering or single linkage
-clustering is each pair of adjacent order statistics on each manifold
-being less than \(\delta\) apart.
+clustering is satisfied if each pair of adjacent order statistics on
+each manifold is less than \(\delta\) apart.
 
@@
 Suppose we draw $A \sim \mathrm{GRDPG}_{p,q}(X; \rho_n)$ for some sparsity parameter $\rho_n \in (0, 1]$, and let $\hat{X}$ be the ASE of $A$. 
-Then for any $\epsilon > 0$, there exists an $N \in \mathbb{N}$ such that when $n > N$, single linkage clustering produces zero community detection error with high probability. 
+Then single linkage clustering produces zero community detection error with high probability as $n \to \infty$. 
 \end{theorem}

### 2022-05-04 draft/draft.Rmd
still working on it

@@
 
-# Latent Structure Block Models
+# Generalized Random Dot Product Graphs with Community Structure
 
-All block models are latent structure models.
+All Bernoulli graphs are generalized random dot product graphs. 
+Whether this is useful for inference depends on the structure of the latent space. 
+In the case of the $\ER$ model, SBM, DCBM, and PABM, the latent structure is linear, and the linearity can be exploited for community detection and parameter estimation. 
+In this section, we discuss general, often nonlinear latent structure models, focusing on those with community structure.
+
+To motivate this, consider a generalization of the $\ER$ model. 
+Recall that when viewed as an RDPG, the latent space of an $\ER$ model consists of one point in Euclidean space. 
+In the following example, instead of fixing the edge probability, it is sampled from a distribution in such a way that when viewed as an RDPG, the latent space consists of a curve.
+
+\begin{example}[Hierarchical $\ER$ model]
+In the $\ER$ model, the edge probability matrix has a fixed value $[P_{ij}] \equiv p \in [0, 1]$. 
+Suppose instead that each $P_{ij}$ is drawn from a distribution with density $f(p) = \frac{1}{\frac{\pi}{2} - 1} \Big(\frac{1}{\sqrt{1 - p^2}} - 1 \Big)$.
+
+Then it can be shown that this model is equivalent to a random dot product graph with a two dimensional latent space in which latent vectors are sampled uniformly along the curve $g(t) = \begin{bmatrix} \cos(\frac{\pi}{2} t) & \sin(\frac{\pi}{2} t) \end{bmatrix}^\top$, $0 \leq t \leq 1$.
+\end{example}
 
@@
 
-Here, we provide two algorithms for MBM community detection. 
+We provide two approaches to community detection for the manifold block model. 
 First, we consider the case in which communities correspond to manifolds in the latent space that do not intersect and are separated by some finite distance.
@@
 In this section, we consider the following scenario: 
-Suppose that each community is represented by a manifold $\mathcal{M}_k$, $k \in \{1, ..., K\}$ in the latent space of a RDPG or GDRPG, and the manifolds do not intersect each other. 
+Suppose that each community is represented by a closed manifold $\mathcal{M}_k$, $k \in \{1, ..., K\}$ in the latent space of a RDPG or GDRPG. 
 Define $\delta = \min\limits_{k \neq \ell} \min\limits_{x \in \mathcal{M}_k, y \in \mathcal{M}_\ell} \|x - y\|$, the minimum distance between two manifolds. 
@@
 Compute $\hat{X}$, the ASE of $A$ using the $p$ most positive and $q$ most negative eigenvalues and their corresponding eigenvectors.\;
-Apply single linkage clustering with $K$ communities on $\hat{X}$\;
+Apply single linkage clustering with $K$ communities on $\hat{X}$.\;
 \caption{ASE clustering for nonintersecting communities.}
@@
 
-Consider the specific case in which each $\mathcal{M}_k \subset \mathbb{R}^d$ is a one-dimensional manifold. 
 Let $F_k$ be a probability distribution with support $\mathcal{M}_k$. 
@@
 
-In this setting, although each manifold is in $d$-dimensional space, since the manifolds are one-dimensional, it is possible to consider an ordering of points sampled on the manifold. 
-In the original latent space in which there is no noise, the condition for which community detection is possible via $\eta$-neighborhood graph clustering or single linkage clustering is satisfied if each pair of adjacent order statistics on each manifold is less than $\delta$ apart.
-
-\begin{theorem}[Community detection for one-dimensional nonintersecting manifolds without noise]
+\begin{theorem}[Community detection for nonintersecting manifolds without noise]
 \label{nonintersect-no-noise}
-Let $x_1, ..., x_n$ be points sampled on $K$ one-dimensional manifolds $\mathcal{M}_1, ..., \mathcal{M}_K$. 
+Let $x_1, ..., x_n$ be points sampled on $K$ manifolds $\mathcal{M}_1, ..., \mathcal{M}_K$. 
 Suppose $\delta = \min\limits_{k \neq \ell} \min\limits_{x_i \in \mathcal{M}_k, x_j \in \mathcal{M}_\ell} \| x_i - x_j \| > 0$. 
-Let $A_n$ be the event such that $\max_k \max_i \|x_{(i+1)}^{(k)} - x_{(i)}^{(k)}\| < \delta$, where $x_{(i)}^{(k)}$ is the $i^{th}$ order statistic on the $k^{th}$ manifold. 
-Then for any $\epsilon > 0$, there exists an $N \in \mathbb{N}$ such that when $n > N$, $P(A_n) > 1 - \epsilon$. 
-\end{theorem}
-
-Sampling a graph from this latent configuration and constructing an ASE from the graph introduces error relative to the original latent vectors. 
-If the errors are small, which is guaranteed by the consistency property of the ASE for sufficiently large $n$, then  sufficient separation between embedding points belonging to different manifolds is still guaranteed.
-
-\begin{theorem}[Community detection for one-dimensional nonintersecting manifolds with noise]
-\label{nonintersect-with-noise}
-Suppose the same setup as in theorem \ref{nonintersect-no-noise}. 
-Define $y_i = x_i + e_i$ and $B_n$ as the event such that $\|y_{(i+1)}^{(k)} - y_{(i)}^{(k)}\| < \|y_{(i)}^{(k)} - y_{(j)}^{(\ell)}\|$ for any $k \neq \ell$, where, $y_{(i)}^{(k)}$ is the $i^{th}$ order statistic on the $k^{th}$ manifold plus its corresponding noise.
-Then if $\max_i \|e_i\| < \delta / 3$, for any $\epsilon > 0$, there exists an $N \in \mathbb{N}$ such that when $n > N$, $P(B_n) > 1 - \epsilon$. 
-\end{theorem}
-
-\begin{theorem}[Community detection for one-dimensional nonintersecting manifold block models]
-\label{nonintersect-mbm}
-Suppose we have the same setup as in theorem \ref{nonintersect-no-noise}. 
-Let $p, q \geq 0$ such that $p + q = d$ where $d$ is the dimension of the range of each $g_k$. 
-Suppose we draw $A \sim \GRDPG_{p,q}(X; \rho_n)$ for some sparsity parameter $\rho_n \in (0, 1]$, and let $\hat{X}$ be the ASE of $A$. 
-Then single linkage clustering produces zero community detection error with high probability as $n \to \infty$. 
+Define $A_n$ as the event that for a $\eta$-neighborhood graph constructed from the sample $x_1, ..., x_n$ consists of exactly $K$ disconnected subgraphs that map on exactly to each manifold for some $\eta \in (0, \delta)$.
+Then for any $\epsilon \in (0, 1)$, there exists an $N$ such that when $n > N$, $A_n$ occurs with high probability. 
+\label{nonintersect-no-noise}
 \end{theorem}
@@
 
-To prove theorems \ref{nonintersect-no-noise}, \ref{nonintersect-with-noise}, and \ref{nonintersect-mbm}, we first need to introduce some theory for order statistics. 
-Let $D_i^{(k)} = X_{(i+1)}^{(k)} - X_{(i)}^{(k)}$, where $X_{(i)}^{(k)}$ is the $i^{th}$ order statistic of univariate sample $X_1^{(k)}, ..., X_n^{(k)} \iid F^{(k)}$ and $F^{(k)}$ is a continuous distribution on manifold $\mathcal{M}_k$.
-If $\max_i D_i^{(k)} < \delta$, then there is sufficient separation of points between each manifold. 
-Then it is sufficient to quantify $P(\max_i D_i^{(k)} > \delta)$ for each $k$ as a function of $n$ and $\delta$ and show that this converges to zero as $n$ grows to $\infty$.
+In order to prove theorem \ref{nonintersect-no-noise}, we first establish that the density of points within one manifold is sufficiently dense. 
+The following lemma is based on lemma 2 of \citet{trosset2020rehabilitating}.
 
-For ease of notation, we drop the superscript denoting community/manifold membership until lemma 6. 
-Denote $f(x)$ as the density of each $F$, $g_i(x)$ as the density of $X_{(i)}$, 
-$g_{ij}(x, y)$ as the joint density of $X_{(i)}, X_{(j)}$, and $h_i(d)$ as the density of $D_i$ (with corresponding capital letters for the cumulative distribution functions).
-
-The following are taken as given[^wikipedia]:
-
-1. $g_i(x) = \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x))^{n-i} f(x)$.
-2. $g_{ij}(x, y) = \frac{n!}{(i-1)! (j-i-1)! (n-j)!} (F(x))^{i-1} (F(y) - F(x))^{j-i-1} (1 - F(y))^{n-j} f(x) f(y)$.
-3. By convolution, $h_i(d) = \int_0^{1} g_{i, i+1} (x, x + d) \dd x$.
-
-[^wikipedia]: https://en.wikipedia.org/wiki/Order_statistic
-
-\begin{lemma}[The probability density function of $D_i$]
-Suppose $X_1, ..., X_n \iid F$ and the support of $F$ is the unit interval. 
-Let $D_i = X_{(i+1)} - X_{(i)}$. 
-Then the density of each $D_i$ is:
-
-\begin{equation}
-\label{eq:pdf}
-h_i(d) = \int_0^{1-d} \frac{n!}{(i-1)! (n-i-1)!} (F(x))^{i-1} (1 - F(x+d))^{n-i-1} f(x) f(x+d) \dd x 
-\end{equation}
+\begin{lemma}
+\label{lem:one-hypercube}
+Let $x_1, ..., x_n \iid F$ with support $[0, 1]^r$, and $f(x) \geq a > 0$ everywhere on the support. 
+Define $H_n$ as the event that an $\eta$-neighborhood graph constructed from the sample is connected for any $\eta > 0$. 
+Then for any $\epsilon > 0$, there exists $N = O \bigg( \frac{\log \epsilon + r \log \eta - \frac{r}{2} \log r}{\log (1 - a \eta^r r^{-r / 2})} \bigg)$ such that $P(E_n) > 1 - \epsilon$ when $n \geq N$.
 \end{lemma}
@@
 \begin{proof}
-This is a direct consequence of 2 and 3. 
-We also note that because the support of $X_i$ is $[0, 1]$, 
-the integral only needs to be evaluated from $0$ to $1 - d$ 
-because of the $f(x+d)$ and $1 - F(x+d)$ terms.
-\end{proof}
-
-\begin{lemma}[The cumulative distribution function of $D_i$]
-Suppose $X_1, ..., X_n \iid F$ and the support of $F$ is the unit interval. 
-Let $D_i = X_{(i+1)} - X_{(i)}$. 
-Then the distribution function of each $D_i$ is:
-
-\begin{equation}
-\label{eq:cdf}
-P(D_i < \delta) = H_i(\delta) = 1 - \int_0^{1-\delta} \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \dd x
-\end{equation}
-\end{lemma}
-
-\begin{proof}
-$$
-\begin{aligned}
-H_i(\delta) & = \int_x^{x+\delta} h_i(d) \dd d \\
-& = \int_x^{x+\delta} \int_0^{1} \frac{n!}{(i-1)! (n-i-1)!} ((F(x))^{i-1} (1 - F(x+d))^{n-i-1} f(x) f(x+d) \dd x \dd d \\
-& = \int_0^{1} \frac{n!}{(i-1)! (n-i-1)!} (F(x))^{i-1} f(x) \int_x^{x+\delta} (1 - F(x+d))^{n-i-1} f(x+d) \dd d \dd x \\
-& = \int_0^{1} \frac{n!}{(i-1)! (n-i-1)!} (F(x))^{i-1} f(x) \int_{F(x)}^{F(x+\delta)} (1 - u)^{n-i-1} \dd u \dd x \\
-& = \int_0^{1} \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} f(x) ((1 - F(x))^{n-i} - (1 - F(x + \delta))^{n-i}) \dd x \\ 
-& = \int_0^1 g_i(x) \dd x - \int_0^1 \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \dd x \\
-& = 1 - \int_0^1 \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \dd x \\
-\end{aligned}
-$$
-
-Because of the $x + \delta$ term and the fact that the support is $[0, 1]$, the integrand is zero above $1 - \delta$, so we are left with
-
-$$
-\begin{aligned}
-& = 1 - \int_0^{1 - \delta} \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \dd x.
-\end{aligned}
-$$
-\end{proof}
-
-We now focus on the case where points are sampled uniformly on the unit interval to build up to more general distributions.
-
-\begin{lemma}[Differences between order statistics of a uniform distribution]
-If $X_1, ..., X_n \iid \Uniform(0, 1)$, then each $D_i \sim \Betadist(1, n)$.
-\end{lemma}
-
-\begin{proof}
-We begin with Eq. (\ref{eq:pdf}), plugging in $f(x) = 1$ and $F(x) = x$:
-
-$$
-h_i(d) = \int_0^{1-d} \frac{n!}{(i-1)! (n-i-1)!} x^{i-1} (1-x-d)^{n-i-1} \dd x.
-$$
-
-Then we proceed with integration by parts, setting 
-$u = x^{i-1} \implies du = (i-1) x^{i-2}$ and 
-$dv = (1-x-d)^{n-i-1} dx \implies v = -\frac{1}{n-i} (1-x-d)^{n-i-1}$. 
-Note that $u v |_0^{1-d} = 0$ in this case. This yields
-
-$$
-= \frac{n!}{(i-1)! (n-i-1)!} \int \frac{i-1}{n-i} x^{i-2} (1-x-d)^{n-i} \dd x.
-$$
-
-Then applying integration by parts again until the $x^p$ term disappears, we get:
-
-$$
-\begin{aligned}
-& = \frac{n!}{(i-1)! (n-i-1)!} \frac{(i-1)!}{(n-i) \cdots (n-2)} \int_0^{1-d} (1-x-d)^{n-2} \dd x \\
-& = -\frac{n (n-1)}{n-1} (1-x-d)^{n-1} \Big|_0^{1-d} \\
-& = n (1 - d)^{n-1}.
-\end{aligned}
-$$
-
-This the density function for $\Betadist(1, n)$, completing the proof.
-\end{proof}
-
-\begin{lemma}
-Let $X_1, ..., X_n \iid \Uniform(0, 1)$. 
-Then for any $\epsilon$ and $\delta > 0$, 
-there exists an $N = O \big(\frac{-\log \epsilon}{\delta} \big)$ such that 
-$P(\max_i X_{(i+1)} - X_{(i)} < \delta) \geq 1 - \epsilon$ when $n > N$.
-\end{lemma}
-
-\begin{proof}
-Since $X_{(i+1)} - X_{(i)} = D_i \sim \Betadist(1, n)$, 
-$P(X_{(i+1)} - X_{(i)} < \delta) = 1 - (1 - \delta)^n $. This yields
-
-$$
-\begin{aligned}
-P(\max_i D_i < \delta) & \geq (P(D_i < \delta))^{n-1} \\
-& = (1 - (1 - \delta)^n)^{n-1} \\
-& \approx e^{-n \exp(-n \delta)}.
-\end{aligned}
-$$
-
-In the limit $n \to \infty$, this goes to 1.
-\end{proof}
-
-Now we extend this lemma to general distributions on the unit interval.
-
-\begin{lemma}
-\label{lem:lim-d}
-Let $X_1, ..., X_n \iid F$ with support $[0, 1]$, and suppose $f(x)$ is continuous and $f(x) \geq a > 0$ everywhere on the support. 
-Let $D_i = X_{(i+1)} - X_{(i)}$. 
-Then for any $\epsilon > 0$, there exists $N > 0$ such that $P(\max_i D_i < \delta) \geq 1 - \epsilon$ when $n > N$.
-\end{lemma}
-
-\begin{proof}
-We start with Eq. (\ref{eq:cdf}):
-
-$$P(D_i \leq \delta) = 1 - \int_0^{1-\delta} \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \dd x.$$
-
-Making the approximation $F(x+\delta) \approx F(x) + \delta f(x)$ 
-and bounding $f(x) \geq a$, we get:
-
-$$P(D_i \leq \delta) \geq 1 - \int_0^{1-\delta} \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x) - a \delta)^{n-i} f(x) \dd x.$$
-
-Then making the substitution $u = F(x) \implies du = f(x) \dd x$, we obtain
-
-$$1 - \int_0^{F(1-\delta)} \frac{n!}{(n-i)! (i-1)!} u^{i-1} (1 - u - a \delta)^{n-i} \dd u$$
-
-Evaluating the integral yields
-
-$$P(D_i < \delta) = 1 - (1 - a \delta)^n + (1 - F(1-\delta) - a \delta)^n.$$
-
-Then as before,
-
-$$
-\begin{aligned}
-P(\max_i D_i < \delta) & = P(\text{all } D_i < \delta) \\
-& = 1 - P(\text{some } D_i > \delta) \\
-& \geq 1 - \sum_i^{n-1} P(D_i > \delta) \\
-& = 1 - (n - 1) (1 - a \delta)^n + (n - 1) (1 - F(1 - \delta) - a \delta)^n
-\end{aligned}
-$$
-
-Since $(1 - a \delta) < 1$ and $(1 - F(1-\delta) - a \delta) < 1$, this converges to $1$ in the limit $n \to \infty$. 
-Then for any $\epsilon \in (0, 1)$, we can always solve for an $n$ such that $1 - (n-1) (1 - a \delta)^n (n-1) (1 - F(1 - \delta) - a \delta)^n \geq 1 - \epsilon$ holds. 
-\end{proof}
-
-\begin{proof}[Proof of theorem \ref{nonintersect-no-noise}]
-This theorem requires two additional conditions: 
-First, each $n_k \to \infty$ as $n \to \infty$, and second, the number of manifolds $K$ remains fixed. 
-These are satisfied by the model assumptions. 
-Then this is a direct consequence of Lemma \ref{lem:lim-d}. 
-Denoting $m = \min_k n_k$ as the sample size of the manifold with the fewest points,
-
-$$
-\begin{aligned}
-P(\max_{k,i} \|X_{(i+1)}^{(k)} - X_{(i)}^{(k)}\| < \delta) & = P(\forall k, i ~ D_i^{(k)} < \delta) \\
-& \geq \prod_{k=1}^K P(D_i^{(k)} < \delta) \\
-& \geq (1 - (m-1) (1 - a \delta)^{m} + (m-1) (1 - F(1-\delta) - a \delta)^{m})^K
-\end{aligned}
-$$
-Since as $n \to \infty$, $m \to \infty$ as well, and $K$ is fixed, this approaches $(1)^K = 1$. Then setting this $\geq 1 - \epsilon$, it is always possible to solve this for $m$. 
-\end{proof}
-
-\begin{proof}[Proof of theorem \ref{nonintersect-with-noise}]
-Let $\max_i \|e_i\| \leq \xi$. 
-\end{proof}
-
-\begin{proof}[Proof of theorem \ref{nonintersect-mbm}]
-\end{proof}
-
-Here, we extend the results on the line to the unit hypercube. The following lemma is based on lemma 2 of \citet{trosset2020rehabilitating}.
-
-\begin{lemma}
-\label{thm:multidim}
-Let $X_1, ..., X_n \iid F$ with support $[0, 1]^r$, and $f(x) \geq a > 0$ everywhere on the support. 
-Define $E_n$ as the event that an $\eta$-neighborhood graph constructed from the sample is connected. 
-Then for any $\epsilon > 0$, there exists $N = O \bigg( \frac{\log \epsilon \eta^r / r^{r/2}}{\log (1 - \frac{a \eta^r}{r^{r / 2}})} \bigg)$ such that $P(E_n) > 1 - \epsilon$ when $n \geq N$.
-\end{lemma}
-
-\begin{proof}[Proof (sketch)]
 Divide the hypercube $[0, 1]^r$ into a grid of sub-hypercubes of side length at most $\eta / \sqrt{r}$. 
@@
 \begin{aligned}
-P(E_n) & = 1 - P(\text{some cells don't contain } X_i) \\
-& \geq 1 - \sum_k^{\lceil \sqrt{r} / \eta \rceil^r} \prod_i^n P(X_i \text{ is not in the } k^{th} \text{ hypercube}) \\
-& \geq 1 - \lceil \sqrt{r} / \eta \rceil^r (1 - a \eta^r / r^{r/2})^n
+P(H_n) & = 1 - P(\text{some cells don't contain } X_i) \\
+& \geq 1 - \sum_m^{\lceil \sqrt{r} / \eta \rceil^r} \prod_i^n P(X_i \text{ is not in the } k^{th} \text{ hypercube}) \\
+& \geq 1 - \lceil \sqrt{r} / \eta \rceil^r (1 - a \eta^r / r^{r/2})^n,
 \end{aligned}
 $$
+which approaches $1$ as $n \to \infty$. 
+Setting this quantity as  $\geq 1 - \epsilon$ and solving for $n$ yields the desired rate. 
+\end{proof}
 
-Setting this $\geq 1 - \epsilon$ and solving for $n$ yields the desired rate. 
+\begin{lemma}
+\label{lem:K-hypercubes}
+Let there be $K \geq 2$ hypercubes $\mathcal{C}_1, ..., \mathcal{C}_K$ of dimension $r_1, ..., r_K$ in $\mathbb{R}^d$ such that $\delta = \min\limits_{k \neq \ell} \min\limits_{x_i \in \mathcal{C}_k, x_j \in \mathcal{C}_\ell} \|x_i - x_j\| > 0$.
+Let $F_k$ be a distribution with support $C_k$ such that its density $f_k$ is nonzero on the support. Let $a = \min_k \min_t f_k(t) > 0$. 
+Define a mixture model as follows:
+
+\begin{enumerate}
+  \item Draw labels $z_1, ..., z_n \iid \Multinomial(\alpha_1, ..., \alpha_K)$. 
+  \item Draw latent vectors each as $x_i \indep F_{z_i}$.
+\end{enumerate}
+
+Define $H'_n$ as the event that an $\eta$-neighborhood graph constructed from this sample consists of exactly $K$ disconnected subgraphs for any $\eta \in (0, \delta)$. 
+Then for any $\epsilon \in (0, 1)$, there exists an $N = O \Big( \frac{\log(1 - (1 - \epsilon)^{1/K}) + d \log \eta - \frac{d}{2} \log d}{\alpha_{min} \log(1 - a \eta^d d^{-d/2})} \Big)$ such that when $n > N$, $P(E_n) > 1 - \epsilon$.
+\end{lemma}
+
+\begin{proof}
+Let $H^{(k)}$ be the event that lemma \ref{lem:one-hypercube} holds for $\mathcal{C}_k$. Then $H^{(k)} = H_{n_k}$ where $H_n$ is defined as in \ref{lem:one-hypercube}. 
+Then $P(H'_n) = P(H_{n_1} \text{ and } ... \text{ and } H_{n_K})$.
+
+$$
+\begin{aligned}
+P(H'_n) & = \prod_k P(H_{n_k}) \\
+& \geq \prod_k 1 - \lceil \sqrt{r_k} / \eta \rceil^{r_k} (1 - a \eta^{r_k} r_k^{-r_k/2})^{n_k} \\
+& \geq \prod_k 1 - \lceil \sqrt{d} / \eta \rceil^{d} (1 - a \eta^{d} d^{-d/2})^{\alpha_{\min} n} \\
+& \geq \big(1 - \lceil \sqrt{d} / \eta \rceil^{d} (1 - a \eta^{d} d^{-d/2})^{\alpha_{\min} n} \big)^K, \\
+\end{aligned}
+$$
+which approaches 1 as $n \to \infty$. 
+Setting this quantity to $\geq 1 - \epsilon$ and solving for $n$ yields the desired rate. 
+\end{proof}
+
+\begin{proof}[Proof of theorem \ref{nonintersect-no-noise}]
 \end{proof}

### 2022-05-04 draft/draft.tex
still working on it

@@
 \usepackage{hyperref}
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 
@@
 
-% Pandoc citation processing
+% tightlist command for lists without linebreak
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 
@@
 
-\hypertarget{latent-structure-block-models}{%
-\section{Latent Structure Block
-Models}\label{latent-structure-block-models}}
+\hypertarget{generalized-random-dot-product-graphs-with-community-structure}{%
+\section{Generalized Random Dot Product Graphs with Community
+Structure}\label{generalized-random-dot-product-graphs-with-community-structure}}
 
-All block models are latent structure models.
+All Bernoulli graphs are generalized random dot product graphs. Whether
+this is useful for inference depends on the structure of the latent
+space. In the case of the \(\text{Erd\"{o}s-R\'{e}nyi}\) model, SBM,
+DCBM, and PABM, the latent structure is linear, and the linearity can be
+exploited for community detection and parameter estimation. In this
+section, we discuss general, often nonlinear latent structure models,
+focusing on those with community structure.
+
+To motivate this, consider a generalization of the
+\(\text{Erd\"{o}s-R\'{e}nyi}\) model. Recall that when viewed as an
+RDPG, the latent space of an \(\text{Erd\"{o}s-R\'{e}nyi}\) model
+consists of one point in Euclidean space. In the following example,
+instead of fixing the edge probability, it is sampled from a
+distribution in such a way that when viewed as an RDPG, the latent space
+consists of a curve.
+
+\begin{example}[Hierarchical $\text{Erd\"{o}s-R\'{e}nyi}$ model]
+In the $\text{Erd\"{o}s-R\'{e}nyi}$ model, the edge probability matrix has a fixed value $[P_{ij}] \equiv p \in [0, 1]$. 
+Suppose instead that each $P_{ij}$ is drawn from a distribution with density $f(p) = \frac{1}{\frac{\pi}{2} - 1} \Big(\frac{1}{\sqrt{1 - p^2}} - 1 \Big)$.
+
+Then it can be shown that this model is equivalent to a random dot product graph with a two dimensional latent space in which latent vectors are sampled uniformly along the curve $g(t) = \begin{bmatrix} \cos(\frac{\pi}{2} t) & \sin(\frac{\pi}{2} t) \end{bmatrix}^\top$, $0 \leq t \leq 1$.
+\end{example}
 
@@
 
-Here, we provide two algorithms for MBM community detection. First, we
-consider the case in which communities correspond to manifolds in the
-latent space that do not intersect and are separated by some finite
-distance. In this scenario, we use the convergence of the ASE to show
-that single linkage clustering on the latent space produces a clustering
-such that the total number of misclustered vertices goes to zero, with
-high probability.
+We provide two approaches to community detection for the manifold block
+model. First, we consider the case in which communities correspond to
+manifolds in the latent space that do not intersect and are separated by
+some finite distance. In this scenario, we use the convergence of the
+ASE to show that single linkage clustering on the latent space produces
+a clustering such that the total number of misclustered vertices goes to
+zero, with high probability.
 
@@
 In this section, we consider the following scenario: Suppose that each
-community is represented by a manifold \(\mathcal{M}_k\),
-\(k \in \{1, ..., K\}\) in the latent space of a RDPG or GDRPG, and the
-manifolds do not intersect each other. Define
+community is represented by a closed manifold \(\mathcal{M}_k\),
+\(k \in \{1, ..., K\}\) in the latent space of a RDPG or GDRPG. Define
 \(\delta = \min\limits_{k \neq \ell} \min\limits_{x \in \mathcal{M}_k, y \in \mathcal{M}_\ell} \|x - y\|\),
@@
 
-Consider the specific case in which each
-\(\mathcal{M}_k \subset \mathbb{R}^d\) is a one-dimensional manifold.
 Let \(F_k\) be a probability distribution with support
@@
 
-In this setting, although each manifold is in \(d\)-dimensional space,
-since the manifolds are one-dimensional, it is possible to consider an
-ordering of points sampled on the manifold. In the original latent space
-in which there is no noise, the condition for which community detection
-is possible via \(\eta\)-neighborhood graph clustering or single linkage
-clustering is satisfied if each pair of adjacent order statistics on
-each manifold is less than \(\delta\) apart.
-
-\begin{theorem}[Community detection for one-dimensional nonintersecting manifolds without noise]
+\begin{theorem}[Community detection for nonintersecting manifolds without noise]
 \label{nonintersect-no-noise}
-Let $x_1, ..., x_n$ be points sampled on $K$ one-dimensional manifolds $\mathcal{M}_1, ..., \mathcal{M}_K$. 
+Let $x_1, ..., x_n$ be points sampled on $K$ manifolds $\mathcal{M}_1, ..., \mathcal{M}_K$. 
 Suppose $\delta = \min\limits_{k \neq \ell} \min\limits_{x_i \in \mathcal{M}_k, x_j \in \mathcal{M}_\ell} \| x_i - x_j \| > 0$. 
-Let $A_n$ be the event such that $\max_k \max_i \|x_{(i+1)}^{(k)} - x_{(i)}^{(k)}\| < \delta$, where $x_{(i)}^{(k)}$ is the $i^{th}$ order statistic on the $k^{th}$ manifold. 
-Then for any $\epsilon > 0$, there exists an $N \in \mathbb{N}$ such that when $n > N$, $P(A_n) > 1 - \epsilon$. 
-\end{theorem}
-
-Sampling a graph from this latent configuration and constructing an ASE
-from the graph introduces error relative to the original latent vectors.
-If the errors are small, which is guaranteed by the consistency property
-of the ASE for sufficiently large \(n\), then sufficient separation
-between embedding points belonging to different manifolds is still
-guaranteed.
-
-\begin{theorem}[Community detection for one-dimensional nonintersecting manifolds with noise]
-\label{nonintersect-with-noise}
-Suppose the same setup as in theorem \ref{nonintersect-no-noise}. 
-Define $y_i = x_i + e_i$ and $B_n$ as the event such that $\|y_{(i+1)}^{(k)} - y_{(i)}^{(k)}\| < \|y_{(i)}^{(k)} - y_{(j)}^{(\ell)}\|$ for any $k \neq \ell$, where, $y_{(i)}^{(k)}$ is the $i^{th}$ order statistic on the $k^{th}$ manifold plus its corresponding noise.
-Then if $\max_i \|e_i\| < \delta / 3$, for any $\epsilon > 0$, there exists an $N \in \mathbb{N}$ such that when $n > N$, $P(B_n) > 1 - \epsilon$. 
-\end{theorem}
-
-\begin{theorem}[Community detection for one-dimensional nonintersecting manifold block models]
-\label{nonintersect-mbm}
-Suppose we have the same setup as in theorem \ref{nonintersect-no-noise}. 
-Let $p, q \geq 0$ such that $p + q = d$ where $d$ is the dimension of the range of each $g_k$. 
-Suppose we draw $A \sim \mathrm{GRDPG}_{p,q}(X; \rho_n)$ for some sparsity parameter $\rho_n \in (0, 1]$, and let $\hat{X}$ be the ASE of $A$. 
-Then single linkage clustering produces zero community detection error with high probability as $n \to \infty$. 
+Define $A_n$ as the event that for a $\eta$-neighborhood graph constructed from the sample $x_1, ..., x_n$ consists of exactly $K$ disconnected subgraphs that map on exactly to each manifold for some $\eta \in (0, \delta)$.
+Then for any $\epsilon \in (0, 1)$, there exists an $N$ such that when $n > N$, $A_n$ occurs with high probability. 
+\label{nonintersect-no-noise}
 \end{theorem}
@@
 
-To prove theorems \ref{nonintersect-no-noise},
-\ref{nonintersect-with-noise}, and \ref{nonintersect-mbm}, we first need
-to introduce some theory for order statistics. Let
-\(D_i^{(k)} = X_{(i+1)}^{(k)} - X_{(i)}^{(k)}\), where \(X_{(i)}^{(k)}\)
-is the \(i^{th}\) order statistic of univariate sample
-\(X_1^{(k)}, ..., X_n^{(k)} \stackrel{\mathrm{iid}}{\sim}F^{(k)}\) and
-\(F^{(k)}\) is a continuous distribution on manifold \(\mathcal{M}_k\).
-If \(\max_i D_i^{(k)} < \delta\), then there is sufficient separation of
-points between each manifold. Then it is sufficient to quantify
-\(P(\max_i D_i^{(k)} > \delta)\) for each \(k\) as a function of \(n\)
-and \(\delta\) and show that this converges to zero as \(n\) grows to
-\(\infty\).
+In order to prove theorem \ref{nonintersect-no-noise}, we first
+establish that the density of points within one manifold is sufficiently
+dense. The following lemma is based on lemma 2 of
+\citet{trosset2020rehabilitating}.
 
-For ease of notation, we drop the superscript denoting
-community/manifold membership until lemma 6. Denote \(f(x)\) as the
-density of each \(F\), \(g_i(x)\) as the density of \(X_{(i)}\),
-\(g_{ij}(x, y)\) as the joint density of \(X_{(i)}, X_{(j)}\), and
-\(h_i(d)\) as the density of \(D_i\) (with corresponding capital letters
-for the cumulative distribution functions).
-
-The following are taken as given\footnote{\url{https://en.wikipedia.org/wiki/Order_statistic}}:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  \(g_i(x) = \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x))^{n-i} f(x)\).
-\item
-  \(g_{ij}(x, y) = \frac{n!}{(i-1)! (j-i-1)! (n-j)!} (F(x))^{i-1} (F(y) - F(x))^{j-i-1} (1 - F(y))^{n-j} f(x) f(y)\).
-\item
-  By convolution,
-  \(h_i(d) = \int_0^{1} g_{i, i+1} (x, x + d) \mathrm{d}x\).
-\end{enumerate}
-
-\begin{lemma}[The probability density function of $D_i$]
-Suppose $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}F$ and the support of $F$ is the unit interval. 
-Let $D_i = X_{(i+1)} - X_{(i)}$. 
-Then the density of each $D_i$ is:
-
-\begin{equation}
-\label{eq:pdf}
-h_i(d) = \int_0^{1-d} \frac{n!}{(i-1)! (n-i-1)!} (F(x))^{i-1} (1 - F(x+d))^{n-i-1} f(x) f(x+d) \mathrm{d}x 
-\end{equation}
+\begin{lemma}
+\label{lem:one-hypercube}
+Let $x_1, ..., x_n \stackrel{\mathrm{iid}}{\sim}F$ with support $[0, 1]^r$, and $f(x) \geq a > 0$ everywhere on the support. 
+Define $H_n$ as the event that an $\eta$-neighborhood graph constructed from the sample is connected for any $\eta > 0$. 
+Then for any $\epsilon > 0$, there exists $N = O \bigg( \frac{\log \epsilon + r \log \eta - \frac{r}{2} \log r}{\log (1 - a \eta^r r^{-r / 2})} \bigg)$ such that $P(E_n) > 1 - \epsilon$ when $n \geq N$.
 \end{lemma}
@@
 \begin{proof}
-This is a direct consequence of 2 and 3. 
-We also note that because the support of $X_i$ is $[0, 1]$, 
-the integral only needs to be evaluated from $0$ to $1 - d$ 
-because of the $f(x+d)$ and $1 - F(x+d)$ terms.
-\end{proof}
-
-\begin{lemma}[The cumulative distribution function of $D_i$]
-Suppose $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}F$ and the support of $F$ is the unit interval. 
-Let $D_i = X_{(i+1)} - X_{(i)}$. 
-Then the distribution function of each $D_i$ is:
-
-\begin{equation}
-\label{eq:cdf}
-P(D_i < \delta) = H_i(\delta) = 1 - \int_0^{1-\delta} \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \mathrm{d}x
-\end{equation}
-\end{lemma}
-
-\begin{proof}
-$$
-\begin{aligned}
-H_i(\delta) & = \int_x^{x+\delta} h_i(d) \mathrm{d}d \\
-& = \int_x^{x+\delta} \int_0^{1} \frac{n!}{(i-1)! (n-i-1)!} ((F(x))^{i-1} (1 - F(x+d))^{n-i-1} f(x) f(x+d) \mathrm{d}x \mathrm{d}d \\
-& = \int_0^{1} \frac{n!}{(i-1)! (n-i-1)!} (F(x))^{i-1} f(x) \int_x^{x+\delta} (1 - F(x+d))^{n-i-1} f(x+d) \mathrm{d}d \mathrm{d}x \\
-& = \int_0^{1} \frac{n!}{(i-1)! (n-i-1)!} (F(x))^{i-1} f(x) \int_{F(x)}^{F(x+\delta)} (1 - u)^{n-i-1} \mathrm{d}u \mathrm{d}x \\
-& = \int_0^{1} \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} f(x) ((1 - F(x))^{n-i} - (1 - F(x + \delta))^{n-i}) \mathrm{d}x \\ 
-& = \int_0^1 g_i(x) \mathrm{d}x - \int_0^1 \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \mathrm{d}x \\
-& = 1 - \int_0^1 \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \mathrm{d}x \\
-\end{aligned}
-$$
-
-Because of the $x + \delta$ term and the fact that the support is $[0, 1]$, the integrand is zero above $1 - \delta$, so we are left with
-
-$$
-\begin{aligned}
-& = 1 - \int_0^{1 - \delta} \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \mathrm{d}x.
-\end{aligned}
-$$
-\end{proof}
-
-We now focus on the case where points are sampled uniformly on the unit
-interval to build up to more general distributions.
-
-\begin{lemma}[Differences between order statistics of a uniform distribution]
-If $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}\mathrm{Uniform}(0, 1)$, then each $D_i \sim \mathrm{Beta}(1, n)$.
-\end{lemma}
-
-\begin{proof}
-We begin with Eq. (\ref{eq:pdf}), plugging in $f(x) = 1$ and $F(x) = x$:
-
-$$
-h_i(d) = \int_0^{1-d} \frac{n!}{(i-1)! (n-i-1)!} x^{i-1} (1-x-d)^{n-i-1} \mathrm{d}x.
-$$
-
-Then we proceed with integration by parts, setting 
-$u = x^{i-1} \implies du = (i-1) x^{i-2}$ and 
-$dv = (1-x-d)^{n-i-1} dx \implies v = -\frac{1}{n-i} (1-x-d)^{n-i-1}$. 
-Note that $u v |_0^{1-d} = 0$ in this case. This yields
-
-$$
-= \frac{n!}{(i-1)! (n-i-1)!} \int \frac{i-1}{n-i} x^{i-2} (1-x-d)^{n-i} \mathrm{d}x.
-$$
-
-Then applying integration by parts again until the $x^p$ term disappears, we get:
-
-$$
-\begin{aligned}
-& = \frac{n!}{(i-1)! (n-i-1)!} \frac{(i-1)!}{(n-i) \cdots (n-2)} \int_0^{1-d} (1-x-d)^{n-2} \mathrm{d}x \\
-& = -\frac{n (n-1)}{n-1} (1-x-d)^{n-1} \Big|_0^{1-d} \\
-& = n (1 - d)^{n-1}.
-\end{aligned}
-$$
-
-This the density function for $\mathrm{Beta}(1, n)$, completing the proof.
-\end{proof}
-
-\begin{lemma}
-Let $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}\mathrm{Uniform}(0, 1)$. 
-Then for any $\epsilon$ and $\delta > 0$, 
-there exists an $N = O \big(\frac{-\log \epsilon}{\delta} \big)$ such that 
-$P(\max_i X_{(i+1)} - X_{(i)} < \delta) \geq 1 - \epsilon$ when $n > N$.
-\end{lemma}
-
-\begin{proof}
-Since $X_{(i+1)} - X_{(i)} = D_i \sim \mathrm{Beta}(1, n)$, 
-$P(X_{(i+1)} - X_{(i)} < \delta) = 1 - (1 - \delta)^n $. This yields
-
-$$
-\begin{aligned}
-P(\max_i D_i < \delta) & \geq (P(D_i < \delta))^{n-1} \\
-& = (1 - (1 - \delta)^n)^{n-1} \\
-& \approx e^{-n \exp(-n \delta)}.
-\end{aligned}
-$$
-
-In the limit $n \to \infty$, this goes to 1.
-\end{proof}
-
-Now we extend this lemma to general distributions on the unit interval.
-
-\begin{lemma}
-\label{lem:lim-d}
-Let $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}F$ with support $[0, 1]$, and suppose $f(x)$ is continuous and $f(x) \geq a > 0$ everywhere on the support. 
-Let $D_i = X_{(i+1)} - X_{(i)}$. 
-Then for any $\epsilon > 0$, there exists $N > 0$ such that $P(\max_i D_i < \delta) \geq 1 - \epsilon$ when $n > N$.
-\end{lemma}
-
-\begin{proof}
-We start with Eq. (\ref{eq:cdf}):
-
-$$P(D_i \leq \delta) = 1 - \int_0^{1-\delta} \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \mathrm{d}x.$$
-
-Making the approximation $F(x+\delta) \approx F(x) + \delta f(x)$ 
-and bounding $f(x) \geq a$, we get:
-
-$$P(D_i \leq \delta) \geq 1 - \int_0^{1-\delta} \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x) - a \delta)^{n-i} f(x) \mathrm{d}x.$$
-
-Then making the substitution $u = F(x) \implies du = f(x) \mathrm{d}x$, we obtain
-
-$$1 - \int_0^{F(1-\delta)} \frac{n!}{(n-i)! (i-1)!} u^{i-1} (1 - u - a \delta)^{n-i} \mathrm{d}u$$
-
-Evaluating the integral yields
-
-$$P(D_i < \delta) = 1 - (1 - a \delta)^n + (1 - F(1-\delta) - a \delta)^n.$$
-
-Then as before,
-
-$$
-\begin{aligned}
-P(\max_i D_i < \delta) & = P(\text{all } D_i < \delta) \\
-& = 1 - P(\text{some } D_i > \delta) \\
-& \geq 1 - \sum_i^{n-1} P(D_i > \delta) \\
-& = 1 - (n - 1) (1 - a \delta)^n + (n - 1) (1 - F(1 - \delta) - a \delta)^n
-\end{aligned}
-$$
-
-Since $(1 - a \delta) < 1$ and $(1 - F(1-\delta) - a \delta) < 1$, this converges to $1$ in the limit $n \to \infty$. 
-Then for any $\epsilon \in (0, 1)$, we can always solve for an $n$ such that $1 - (n-1) (1 - a \delta)^n (n-1) (1 - F(1 - \delta) - a \delta)^n \geq 1 - \epsilon$ holds. 
-\end{proof}
-
-\begin{proof}[Proof of theorem \ref{nonintersect-no-noise}]
-This theorem requires two additional conditions: 
-First, each $n_k \to \infty$ as $n \to \infty$, and second, the number of manifolds $K$ remains fixed. 
-These are satisfied by the model assumptions. 
-Then this is a direct consequence of Lemma \ref{lem:lim-d}. 
-Denoting $m = \min_k n_k$ as the sample size of the manifold with the fewest points,
-
-$$
-\begin{aligned}
-P(\max_{k,i} \|X_{(i+1)}^{(k)} - X_{(i)}^{(k)}\| < \delta) & = P(\forall k, i ~ D_i^{(k)} < \delta) \\
-& \geq \prod_{k=1}^K P(D_i^{(k)} < \delta) \\
-& \geq (1 - (m-1) (1 - a \delta)^{m} + (m-1) (1 - F(1-\delta) - a \delta)^{m})^K
-\end{aligned}
-$$
-Since as $n \to \infty$, $m \to \infty$ as well, and $K$ is fixed, this approaches $(1)^K = 1$. Then setting this $\geq 1 - \epsilon$, it is always possible to solve this for $m$. 
-\end{proof}
-
-\begin{proof}[Proof of theorem \ref{nonintersect-with-noise}]
-Let $\max_i \|e_i\| \leq \xi$. 
-\end{proof}
-
-\begin{proof}[Proof of theorem \ref{nonintersect-mbm}]
-\end{proof}
-
-Here, we extend the results on the line to the unit hypercube. The
-following lemma is based on lemma 2 of
-\citet{trosset2020rehabilitating}.
-
-\begin{lemma}
-\label{thm:multidim}
-Let $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}F$ with support $[0, 1]^r$, and $f(x) \geq a > 0$ everywhere on the support. 
-Define $E_n$ as the event that an $\eta$-neighborhood graph constructed from the sample is connected. 
-Then for any $\epsilon > 0$, there exists $N = O \bigg( \frac{\log \epsilon \eta^r / r^{r/2}}{\log (1 - \frac{a \eta^r}{r^{r / 2}})} \bigg)$ such that $P(E_n) > 1 - \epsilon$ when $n \geq N$.
-\end{lemma}
-
-\begin{proof}[Proof (sketch)]
 Divide the hypercube $[0, 1]^r$ into a grid of sub-hypercubes of side length at most $\eta / \sqrt{r}$. 
@@
 \begin{aligned}
-P(E_n) & = 1 - P(\text{some cells don't contain } X_i) \\
-& \geq 1 - \sum_k^{\lceil \sqrt{r} / \eta \rceil^r} \prod_i^n P(X_i \text{ is not in the } k^{th} \text{ hypercube}) \\
-& \geq 1 - \lceil \sqrt{r} / \eta \rceil^r (1 - a \eta^r / r^{r/2})^n
+P(H_n) & = 1 - P(\text{some cells don't contain } X_i) \\
+& \geq 1 - \sum_m^{\lceil \sqrt{r} / \eta \rceil^r} \prod_i^n P(X_i \text{ is not in the } k^{th} \text{ hypercube}) \\
+& \geq 1 - \lceil \sqrt{r} / \eta \rceil^r (1 - a \eta^r / r^{r/2})^n,
 \end{aligned}
 $$
+which approaches $1$ as $n \to \infty$. 
+Setting this quantity as  $\geq 1 - \epsilon$ and solving for $n$ yields the desired rate. 
+\end{proof}
 
-Setting this $\geq 1 - \epsilon$ and solving for $n$ yields the desired rate. 
+\begin{lemma}
+\label{lem:K-hypercubes}
+Let there be $K \geq 2$ hypercubes $\mathcal{C}_1, ..., \mathcal{C}_K$ of dimension $r_1, ..., r_K$ in $\mathbb{R}^d$ such that $\delta = \min\limits_{k \neq \ell} \min\limits_{x_i \in \mathcal{C}_k, x_j \in \mathcal{C}_\ell} \|x_i - x_j\| > 0$.
+Let $F_k$ be a distribution with support $C_k$ such that its density $f_k$ is nonzero on the support. Let $a = \min_k \min_t f_k(t) > 0$. 
+Define a mixture model as follows:
+
+\begin{enumerate}
+  \item Draw labels $z_1, ..., z_n \stackrel{\mathrm{iid}}{\sim}\mathrm{Multinomial}(\alpha_1, ..., \alpha_K)$. 
+  \item Draw latent vectors each as $x_i \stackrel{\mathrm{ind}}{\sim}F_{z_i}$.
+\end{enumerate}
+
+Define $H'_n$ as the event that an $\eta$-neighborhood graph constructed from this sample consists of exactly $K$ disconnected subgraphs for any $\eta \in (0, \delta)$. 
+Then for any $\epsilon \in (0, 1)$, there exists an $N = O \Big( \frac{\log(1 - (1 - \epsilon)^{1/K}) + d \log \eta - \frac{d}{2} \log d}{\alpha_{min} \log(1 - a \eta^d d^{-d/2})} \Big)$ such that when $n > N$, $P(E_n) > 1 - \epsilon$.
+\end{lemma}
+
+\begin{proof}
+Let $H^{(k)}$ be the event that lemma \ref{lem:one-hypercube} holds for $\mathcal{C}_k$. Then $H^{(k)} = H_{n_k}$ where $H_n$ is defined as in \ref{lem:one-hypercube}. 
+Then $P(H'_n) = P(H_{n_1} \text{ and } ... \text{ and } H_{n_K})$.
+
+$$
+\begin{aligned}
+P(H'_n) & = \prod_k P(H_{n_k}) \\
+& \geq \prod_k 1 - \lceil \sqrt{r_k} / \eta \rceil^{r_k} (1 - a \eta^{r_k} r_k^{-r_k/2})^{n_k} \\
+& \geq \prod_k 1 - \lceil \sqrt{d} / \eta \rceil^{d} (1 - a \eta^{d} d^{-d/2})^{\alpha_{\min} n} \\
+& \geq \big(1 - \lceil \sqrt{d} / \eta \rceil^{d} (1 - a \eta^{d} d^{-d/2})^{\alpha_{\min} n} \big)^K, \\
+\end{aligned}
+$$
+which approaches 1 as $n \to \infty$. 
+Setting this quantity to $\geq 1 - \epsilon$ and solving for $n$ yields the desired rate. 
+\end{proof}
+
+\begin{proof}[Proof of theorem \ref{nonintersect-no-noise}]
 \end{proof}

### 2022-05-04 draft/draft.Rmd
still working on it

@@
 ```{r example1, fig.cap = 'Manifold block model described in Example 1.
-The latent configuration is on the left, and a random dot product graph drawn from the latent configuration is on the middle, and the ASE is on the right.', fig.height = 2.5, fig.width = 7}
+The latent configuration is on the left, a random dot product graph drawn from the latent configuration is on the middle, and the ASE is on the right.', fig.height = 2.5, fig.width = 7}
 set.seed(123456)
@@
 Define $H_n$ as the event that an $\eta$-neighborhood graph constructed from the sample is connected for any $\eta > 0$. 
-Then for any $\epsilon > 0$, there exists $N = O \bigg( \frac{\log \epsilon + r \log \eta - \frac{r}{2} \log r}{\log (1 - a \eta^r r^{-r / 2})} \bigg)$ such that $P(E_n) > 1 - \epsilon$ when $n \geq N$.
+Then for any $\epsilon > 0$, there exists $N = O \bigg( \frac{\log \epsilon + r \log \eta - \frac{r}{2} \log r}{\log (1 - a \eta^r r^{-r / 2})} \bigg)$ such that $P(H_n) > 1 - \epsilon$ when $n \geq N$.
 \end{lemma}
@@
 Setting this quantity as  $\geq 1 - \epsilon$ and solving for $n$ yields the desired rate. 
+\end{proof}
+
+Lemma \ref{lem:one-hypercube} extends to the case in which vectors are drawn from hypercubes with noise, given that the noise is bounded.
+
+\begin{lemma}
+\label{lem:one-hypercube-noise}
+Suppose the setup from lemma \ref{lem:one-hypercube}, but instead of observing each $x_i$, we observe each $y_i = x_i + e_i$ for some $e_1, ..., e_n \in \mathbb{R}^d$. 
+Let $\nu = \max_i \|e_i\| < \infty$ be bounded. 
+Define $\tilde{H}_n$ as the event in which an $(\eta + 2 \nu)$-neighborhood graph constructed from sample $y_1, ..., y_n$ is connected. 
+Then for any $\epsilon > 0$, there exists $N = O \bigg( \frac{\log \epsilon + r \log \eta - \frac{r}{2} \log r}{\log (1 - a \eta^r r^{-r / 2})} \bigg)$ such that $P(\tilde{H}_n) > 1 - \epsilon$ when $n \geq N$.
+\end{lemma}
+
+\begin{proof}
 \end{proof}
@@
 
+It can similarly be shown that the scenario in lemma \ref{lem:K-hypercubes} with bounded noise can be achieved by replacing the $\eta$-neighborhood graph with an $(\eta + 2 \nu)$-neighborhood graph, where $\nu = \max_i \|e_i\|$.
+This time, it is necessary to bound the noise as $\nu < \delta / 3$ to maintain separation between the hypercubes.
+
 \section{Details on Fitting Bezier Curves with Noise}

### 2022-05-04 draft/draft.tex
still working on it

@@
 \caption{Manifold block model described in Example 1.
-The latent configuration is on the left, and a random dot product graph drawn from the latent configuration is on the middle, and the ASE is on the right.}\label{fig:example1}
+The latent configuration is on the left, a random dot product graph drawn from the latent configuration is on the middle, and the ASE is on the right.}\label{fig:example1}
 \end{figure}
@@
 Setting this quantity as  $\geq 1 - \epsilon$ and solving for $n$ yields the desired rate. 
+\end{proof}
+
+Lemma \ref{lem:one-hypercube} extends to the case in which vectors are
+drawn from hypercubes with noise, given that the noise is bounded.
+
+\begin{lemma}
+\label{lem:one-hypercube-noise}
+Suppose the setup from lemma \ref{lem:one-hypercube}, but instead of observing each $x_i$, we observe each $y_i = x_i + e_i$ for some $e_1, ..., e_n \in \mathbb{R}^d$. 
+Let $\nu = \max_i \|e_i\| < \infty$ be bounded. 
+Define $\tilde{H}_n$ as the event in which an $(\eta + 2 \nu)$-neighborhood graph constructed from sample $y_1, ..., y_n$ is connected. 
+Then for any $\epsilon > 0$, there exists $N = O \bigg( \frac{\log \epsilon + r \log \eta - \frac{r}{2} \log r}{\log (1 - a \eta^r r^{-r / 2})} \bigg)$ such that $P(\tilde{H}_n) > 1 - \epsilon$ when $n \geq N$.
+\end{lemma}
+
+\begin{proof}
 \end{proof}
@@
 
+It can similarly be shown that the scenario in lemma
+\ref{lem:K-hypercubes} with bounded noise can be achieved by replacing
+the \(\eta\)-neighborhood graph with an \((\eta + 2 \nu)\)-neighborhood
+graph, where \(\nu = \max_i \|e_i\|\). This time, it is necessary to
+bound the noise as \(\nu < \delta / 3\) to maintain separation between
+the hypercubes.
+
 \section{Details on Fitting Bezier Curves with Noise}

### 2022-05-05 draft/draft.Rmd
changes based on meeting earlier today

@@
 In the $\ER$ model, the edge probability matrix has a fixed value $[P_{ij}] \equiv p \in [0, 1]$.
-Suppose instead that each $P_{ij}$ is drawn from a distribution with density $f(p) = \frac{1}{\frac{\pi}{2} - 1} \Big(\frac{1}{\sqrt{1 - p^2}} - 1 \Big)$.
 
-Then it can be shown that this model is equivalent to a random dot product graph with a two dimensional latent space in which latent vectors are sampled uniformly along the curve $g(t) = \begin{bmatrix} \cos(\frac{\pi}{2} t) & \sin(\frac{\pi}{2} t) \end{bmatrix}^\top$, $0 \leq t \leq 1$.
+Suppose that we have a random dot product graph in which the latent space is $\mathbb{R}^2$ and latent vectors are drawn uniformly from the quarter circle defined by $g(t) = \begin{bmatrix} \cos(\frac{\pi}{2} t) & \sin(\frac{\pi}{2} t) \end{bmatrix}^\top$, $0 \leq t \leq 1$.
+Then it can be shown that in this model, instead of a fixed $P_{ij} = p$, the edge probabilities are distributed with density $f(p) = \frac{2}{\pi - 2} \Big(\frac{1}{\sqrt{1 - p^2}} - 1 \Big)$. 
 \end{example}
 
-\begin{definition}[Manifold block model]
-Let $p, q \geq 0$, $d = p + q \geq 1$, $K \geq 1$, and $n \geq 1$ be integers.
-If there are manifolds $\mathcal{M}_1, ..., \mathcal{M}_K \subset \mathcal{X}$ for $\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p,q} y \in [0, 1] \}$, and 
-$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$ such that each $x_i \in \mathcal{M}_k$ for some $k \in [K]$, then $A \sim \GRDPG_{p,q}(X; \rho_n)$ is a \emph{manifold block model}.
-\end{definition}
-
-In practice, we often define each manifold $\mathcal{M}_k$ by a continuous function $g_k : [0, 1]^r \to \mathcal{X}$, where $1 \leq r < d$ is the dimensionality of manifold $\mathcal{M}_k$. 
-To complete the probability model, we define a probability distribution $F$ with support $[0, 1]^r$ and draw the community memberships from a multinomial distribution.
-
-The full mixture model can be described as follows:
-
-1. Draw labels $z_1, ..., z_n \iid \Multinomial(\alpha_1, ..., \alpha_K)$. 
-2. Draw latent vectors by first drawing each $t_1, ..., t_n \iid F$ and then computing each $x_i = g_{z_i}(t_i)$. 
-3. Let $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$, 
-and draw $A \sim \RDPG(X)$ or $A \sim \GRDPG_{p,q}(X)$.
+By changing the latent structure from a point mass to a curve, we are able to come up with more flexible Bernoulli graph models in which edge probabilities follow more general probability distributions.
+Community structure then can be added by sampling latent vectors from multiple curves. 
+Then the adjacency spectral embedding of the resulting graph allows us to recover that community structure. 
+This is illustrated in the following example.
 
 \begin{example}
-Let $p = 2$, $q = 0$, $K = 2$, and $r = 1$. 
-Define two one-dimensional manifolds by $f_1(t) = \begin{bmatrix} \cos(\frac{\pi}{3} t) & \sin(\frac{\pi}{3} t) \end{bmatrix}^\top$ and $f_2(t) = \begin{bmatrix} 1 - \cos(\frac{\pi}{3} t) & 1 - \sin(\frac{\pi}{3} t) \end{bmatrix}^\top$.
+Define two one-dimensional manifolds in $\mathbb{R}^2$ by $f_1(t) = \begin{bmatrix} \cos(\frac{\pi}{3} t) & \sin(\frac{\pi}{3} t) \end{bmatrix}^\top$ and $f_2(t) = \begin{bmatrix} 1 - \cos(\frac{\pi}{3} t) & 1 - \sin(\frac{\pi}{3} t) \end{bmatrix}^\top$.
 Draw $t_1, ..., t_n \iid \Uniform(0, 1)$ and $z_1, ..., z_n \iid \Multinomial(\frac{1}{2}, \frac{1}{2})$, and compute latent vectors $x_i = f_{z_i}(t_i)$, which are collected in data matrix $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$.
 Finally, let $A \sim \RDPG(X)$.
-Fig \ref{fig:example1} shows the latent configuration drawn from this latent distribution, a random dot product graph drawn from the latent configuration, and the adjacency spectral embedding of the graph.
+Fig.
+\ref{fig:example1} shows the latent configuration drawn from this latent distribution, a random dot product graph drawn from the latent configuration, and the adjacency spectral embedding of the graph.
+Although the community structure is not obvious from the graph, the embedding shows a clear separation between the two communities.
 
@@
 \end{example}
+
+We now formally define the manifold block model.
+
+\begin{definition}[Manifold block model]
+Let $p, q \geq 0$, $d = p + q \geq 1$, $1 \leq r < d$, $K \geq 2$, and $n \geq 1$ be integers.
+Define manifolds $\mathcal{M}_1, ..., \mathcal{M}_K \subset \mathcal{X}$ for $\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p,q} y \in [0, 1] \}$ each by continuous function $g_k : [0, 1] \to \mathcal{X}$,
+and probability distributions $F_1, ..., F_K$ each with support $[0, 1]^r$. 
+Then the following mixture model is a manifold block model:
+
+\begin{enumerate}
+  \item Draw labels $z_1, ..., z_n \iid \Multinomial(\alpha_1, ..., \alpha_K)$. 
+  \item Draw latent vectors by first drawing each $t_i \indep F_{z_i}$ and then computing each $x_i = g_{z_i}(t_i)$. 
+  \item Let $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$, and draw $A \sim \RDPG(X)$ or $A \sim \GRDPG_{p,q}(X)$. 
+\end{enumerate}
+\end{definition}
 
@@
 \begin{proof}
-Let $H^{(k)}$ be the event that lemma \ref{lem:one-hypercube} holds for $\mathcal{C}_k$. Then $H^{(k)} = H_{n_k}$ where $H_n$ is defined as in \ref{lem:one-hypercube}. 
+Let $H^{(k)}$ be the event that lemma \ref{lem:one-hypercube} holds for $\mathcal{C}_k$. Then $H^{(k)} = H_{n_k}$ where $H_n$ is defined as in lemma \ref{lem:one-hypercube}. 
 Then $P(H'_n) = P(H_{n_1} \text{ and } ... \text{ and } H_{n_K})$.

### 2022-05-05 draft/draft.tex
changes based on meeting earlier today

@@
 In the $\text{Erd\"{o}s-R\'{e}nyi}$ model, the edge probability matrix has a fixed value $[P_{ij}] \equiv p \in [0, 1]$.
-Suppose instead that each $P_{ij}$ is drawn from a distribution with density $f(p) = \frac{1}{\frac{\pi}{2} - 1} \Big(\frac{1}{\sqrt{1 - p^2}} - 1 \Big)$.
 
-Then it can be shown that this model is equivalent to a random dot product graph with a two dimensional latent space in which latent vectors are sampled uniformly along the curve $g(t) = \begin{bmatrix} \cos(\frac{\pi}{2} t) & \sin(\frac{\pi}{2} t) \end{bmatrix}^\top$, $0 \leq t \leq 1$.
+Suppose that we have a random dot product graph in which the latent space is $\mathbb{R}^2$ and latent vectors are drawn uniformly from the quarter circle defined by $g(t) = \begin{bmatrix} \cos(\frac{\pi}{2} t) & \sin(\frac{\pi}{2} t) \end{bmatrix}^\top$, $0 \leq t \leq 1$.
+Then it can be shown that in this model, instead of a fixed $P_{ij} = p$, the edge probabilities are distributed with density $f(p) = \frac{2}{\pi - 2} \Big(\frac{1}{\sqrt{1 - p^2}} - 1 \Big)$. 
 \end{example}
 
-\begin{definition}[Manifold block model]
-Let $p, q \geq 0$, $d = p + q \geq 1$, $K \geq 1$, and $n \geq 1$ be integers.
-If there are manifolds $\mathcal{M}_1, ..., \mathcal{M}_K \subset \mathcal{X}$ for $\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p,q} y \in [0, 1] \}$, and 
-$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$ such that each $x_i \in \mathcal{M}_k$ for some $k \in [K]$, then $A \sim \mathrm{GRDPG}_{p,q}(X; \rho_n)$ is a \emph{manifold block model}.
-\end{definition}
-
-In practice, we often define each manifold \(\mathcal{M}_k\) by a
-continuous function \(g_k : [0, 1]^r \to \mathcal{X}\), where
-\(1 \leq r < d\) is the dimensionality of manifold \(\mathcal{M}_k\). To
-complete the probability model, we define a probability distribution
-\(F\) with support \([0, 1]^r\) and draw the community memberships from
-a multinomial distribution.
-
-The full mixture model can be described as follows:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Draw labels
-  \(z_1, ..., z_n \stackrel{\mathrm{iid}}{\sim}\mathrm{Multinomial}(\alpha_1, ..., \alpha_K)\).
-\item
-  Draw latent vectors by first drawing each
-  \(t_1, ..., t_n \stackrel{\mathrm{iid}}{\sim}F\) and then computing
-  each \(x_i = g_{z_i}(t_i)\).
-\item
-  Let \(X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top\), and
-  draw \(A \sim \mathrm{RDPG}(X)\) or
-  \(A \sim \mathrm{GRDPG}_{p,q}(X)\).
-\end{enumerate}
+By changing the latent structure from a point mass to a curve, we are
+able to come up with more flexible Bernoulli graph models in which edge
+probabilities follow more general probability distributions. Community
+structure then can be added by sampling latent vectors from multiple
+curves. Then the adjacency spectral embedding of the resulting graph
+allows us to recover that community structure. This is illustrated in
+the following example.
 
 \begin{example}
-Let $p = 2$, $q = 0$, $K = 2$, and $r = 1$. 
-Define two one-dimensional manifolds by $f_1(t) = \begin{bmatrix} \cos(\frac{\pi}{3} t) & \sin(\frac{\pi}{3} t) \end{bmatrix}^\top$ and $f_2(t) = \begin{bmatrix} 1 - \cos(\frac{\pi}{3} t) & 1 - \sin(\frac{\pi}{3} t) \end{bmatrix}^\top$.
+Define two one-dimensional manifolds in $\mathbb{R}^2$ by $f_1(t) = \begin{bmatrix} \cos(\frac{\pi}{3} t) & \sin(\frac{\pi}{3} t) \end{bmatrix}^\top$ and $f_2(t) = \begin{bmatrix} 1 - \cos(\frac{\pi}{3} t) & 1 - \sin(\frac{\pi}{3} t) \end{bmatrix}^\top$.
 Draw $t_1, ..., t_n \stackrel{\mathrm{iid}}{\sim}\mathrm{Uniform}(0, 1)$ and $z_1, ..., z_n \stackrel{\mathrm{iid}}{\sim}\mathrm{Multinomial}(\frac{1}{2}, \frac{1}{2})$, and compute latent vectors $x_i = f_{z_i}(t_i)$, which are collected in data matrix $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$.
 Finally, let $A \sim \mathrm{RDPG}(X)$.
-Fig \ref{fig:example1} shows the latent configuration drawn from this latent distribution, a random dot product graph drawn from the latent configuration, and the adjacency spectral embedding of the graph.
+Fig.
+\ref{fig:example1} shows the latent configuration drawn from this latent distribution, a random dot product graph drawn from the latent configuration, and the adjacency spectral embedding of the graph.
+Although the community structure is not obvious from the graph, the embedding shows a clear separation between the two communities.
 
@@
 \end{example}
+
+We now formally define the manifold block model.
+
+\begin{definition}[Manifold block model]
+Let $p, q \geq 0$, $d = p + q \geq 1$, $1 \leq r < d$, $K \geq 2$, and $n \geq 1$ be integers.
+Define manifolds $\mathcal{M}_1, ..., \mathcal{M}_K \subset \mathcal{X}$ for $\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p,q} y \in [0, 1] \}$ each by continuous function $g_k : [0, 1] \to \mathcal{X}$,
+and probability distributions $F_1, ..., F_K$ each with support $[0, 1]^r$. 
+Then the following mixture model is a manifold block model:
+
+\begin{enumerate}
+  \item Draw labels $z_1, ..., z_n \stackrel{\mathrm{iid}}{\sim}\mathrm{Multinomial}(\alpha_1, ..., \alpha_K)$. 
+  \item Draw latent vectors by first drawing each $t_i \stackrel{\mathrm{ind}}{\sim}F_{z_i}$ and then computing each $x_i = g_{z_i}(t_i)$. 
+  \item Let $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$, and draw $A \sim \mathrm{RDPG}(X)$ or $A \sim \mathrm{GRDPG}_{p,q}(X)$. 
+\end{enumerate}
+\end{definition}
 

### 2022-05-18 draft/draft.Rmd
still working on proofs

@@
 Suppose $\delta = \min\limits_{k \neq \ell} \min\limits_{x_i \in \mathcal{M}_k, x_j \in \mathcal{M}_\ell} \| x_i - x_j \| > 0$. 
-Define $A_n$ as the event that for a $\eta$-neighborhood graph constructed from the sample $x_1, ..., x_n$ consists of exactly $K$ disconnected subgraphs that map on exactly to each manifold for some $\eta \in (0, \delta)$.
-Then for any $\epsilon \in (0, 1)$, there exists an $N$ such that when $n > N$, $A_n$ occurs with high probability. 
-\label{nonintersect-no-noise}
+Define $A_n$ as the event that a $\eta$-neighborhood graph constructed from the sample $x_1, ..., x_n$ consists of exactly $K$ disconnected subgraphs that map exactly to each manifold for some $\eta \in (0, \delta)$.
+Then for any $\epsilon \in (0, 1)$, there exists an $N$ such that when $n > N$, $P(A_n) > 1 - \epsilon$. 
+\end{theorem}
+
+\begin{theorem}{Community detection for RDPG for which the communities come from nonintersecting manifolds}
+\label{nonintersect-rdpg}
+Let $x_1, ..., x_n$ be points sampled on $K$ manifolds $\mathcal{M}_1, ..., \mathcal{M}_K$. 
+Suppose $\delta = \min\limits_{k \neq \ell} \min\limits_{x_i \in \mathcal{M}_k, x_j \in \mathcal{M}_\ell} \| x_i - x_j \| > 0$. 
+Let $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$ and $A \sim \RDPG(X)$. 
+Define $B_n$ as the event that an $\eta$-neighborhood graph constructed from the ASE of $A$ consists of exactly $K$ disconnected subgraphs that map exactly to each manifold for some $\eta \in (0, \delta)$.
+Then for any $\epsilon \in (0, 1)$, there exists an $N$ such that when $n > N$, $P(B_n) > 1 - \epsilon$. 
 \end{theorem}

### 2022-05-18 draft/draft.tex
still working on proofs

@@
 Suppose $\delta = \min\limits_{k \neq \ell} \min\limits_{x_i \in \mathcal{M}_k, x_j \in \mathcal{M}_\ell} \| x_i - x_j \| > 0$. 
-Define $A_n$ as the event that for a $\eta$-neighborhood graph constructed from the sample $x_1, ..., x_n$ consists of exactly $K$ disconnected subgraphs that map on exactly to each manifold for some $\eta \in (0, \delta)$.
-Then for any $\epsilon \in (0, 1)$, there exists an $N$ such that when $n > N$, $A_n$ occurs with high probability. 
-\label{nonintersect-no-noise}
+Define $A_n$ as the event that a $\eta$-neighborhood graph constructed from the sample $x_1, ..., x_n$ consists of exactly $K$ disconnected subgraphs that map exactly to each manifold for some $\eta \in (0, \delta)$.
+Then for any $\epsilon \in (0, 1)$, there exists an $N$ such that when $n > N$, $P(A_n) > 1 - \epsilon$. 
+\end{theorem}
+
+\begin{theorem}{Community detection for RDPG for which the communities come from nonintersecting manifolds}
+\label{nonintersect-rdpg}
+Let $x_1, ..., x_n$ be points sampled on $K$ manifolds $\mathcal{M}_1, ..., \mathcal{M}_K$. 
+Suppose $\delta = \min\limits_{k \neq \ell} \min\limits_{x_i \in \mathcal{M}_k, x_j \in \mathcal{M}_\ell} \| x_i - x_j \| > 0$. 
+Let $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$ and $A \sim \mathrm{RDPG}(X)$. 
+Define $B_n$ as the event that an $\eta$-neighborhood graph constructed from the ASE of $A$ consists of exactly $K$ disconnected subgraphs that map exactly to each manifold for some $\eta \in (0, \delta)$.
+Then for any $\epsilon \in (0, 1)$, there exists an $N$ such that when $n > N$, $P(B_n) > 1 - \epsilon$. 
 \end{theorem}

### 2022-06-20 draft/draft.Rmd
syncing stuff

@@
 One common analysis task for graph and network data is community detection, which assumes that each vertex of a graph has a hidden community label.
-The goal of the analysis is then to retrieve these labels.
-In order to perform this analysis as a statistical inference task is to define a probability model with inherent community structure.
-We call such models *block models*:
+The goal of the analysis is then to estimate these labels upon observing a graph.
+In order to perform this analysis as a statistical inference task is to define a probability model with inherent community structure. We call such models *block models*: 
 First, each vertex is assigned a label $z_1, ..., z_n \in \{1, 2, ..., K\}$ where $K \ll n$. 
@@
   \item Draw labels $z_1, ..., z_n \iid \Multinomial(\alpha_1, ..., \alpha_K)$. 
-  \item Draw latent vectors by first drawing each $t_i \indep F_{z_i}$ and then computing each $x_i = g_{z_i}(t_i)$. 
+  \item Draw latent vectors by first drawing each $t_i \indep F_{z_i}$ and then compute each $x_i = g_{z_i}(t_i)$. 
   \item Let $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$, and draw $A \sim \RDPG(X)$ or $A \sim \GRDPG_{p,q}(X)$. 
@@
                       animation.dir = animation.dir,
-                      animation.title = paste('random', i, ifelse(intercept,
+                      animation.title = paste('random', 
+                                              i, 
+                                              ifelse(intercept,
                                                      'w-intercept',
@@
 Xhat <- embedding(A, 2, 0)
+n <- nrow(Xhat)
 plot(Xhat, col = z, asp = 1)

### 2022-06-20 draft/draft.tex
syncing stuff

@@
 detection, which assumes that each vertex of a graph has a hidden
-community label. The goal of the analysis is then to retrieve these
-labels. In order to perform this analysis as a statistical inference
-task is to define a probability model with inherent community structure.
-We call such models \emph{block models}: First, each vertex is assigned
-a label \(z_1, ..., z_n \in \{1, 2, ..., K\}\) where \(K \ll n\). Then
-each edge probability \(P_{ij}\) is said to depend on the labels \(z_i\)
-and \(z_j\), possibly along with some other parameters. For example, the
+community label. The goal of the analysis is then to estimate these
+labels upon observing a graph. In order to perform this analysis as a
+statistical inference task is to define a probability model with
+inherent community structure. We call such models \emph{block models}:
+First, each vertex is assigned a label
+\(z_1, ..., z_n \in \{1, 2, ..., K\}\) where \(K \ll n\). Then each edge
+probability \(P_{ij}\) is said to depend on the labels \(z_i\) and
+\(z_j\), possibly along with some other parameters. For example, the
 stochastic block model (SBM) sets a fixed edge probability for each pair
@@
   \item Draw labels $z_1, ..., z_n \stackrel{\mathrm{iid}}{\sim}\mathrm{Multinomial}(\alpha_1, ..., \alpha_K)$. 
-  \item Draw latent vectors by first drawing each $t_i \stackrel{\mathrm{ind}}{\sim}F_{z_i}$ and then computing each $x_i = g_{z_i}(t_i)$. 
+  \item Draw latent vectors by first drawing each $t_i \stackrel{\mathrm{ind}}{\sim}F_{z_i}$ and then compute each $x_i = g_{z_i}(t_i)$. 
   \item Let $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$, and draw $A \sim \mathrm{RDPG}(X)$ or $A \sim \mathrm{GRDPG}_{p,q}(X)$. 
@@
 
+rves clustering.\} \textbackslash end\{algorithm\}
+
 \begin{theorem}
@@
 \end{algorithm}
+
+upervised \(K\)-curves clustering.\} \textbackslash end\{algorithm\}
 

### 2022-06-20 functions.R

@@ -395,29 +395,39 @@ manifold.clustering <- function(X, K = 2,
                                 animation.title = 'plot') {
   n <- nrow(X)
   
+  partial.init <- FALSE
   if (length(initialization) == n) {
     if (all(!is.na(initialization))) {
       z.hat <- initialization
     } else {
+      partial.init <- TRUE
       z.hat.partial <- initialization[!is.na(initialization)]
       X.partial <- X[!is.na(initialization), ]
-      clustering.partial <- manifold.clustering(
-        X = X.partial, 
-        K = K,
-        degree = degree,
-        initialization = z.hat.partial,
-        intercept = intercept,
-        maxit = maxit,
-        k.isomap = k.isomap * length(z.hat.partial) / n,
-        curve.init = curve.init,
-        min.t = min.t, 
-        max.t = max.t, 
-        normalize = normalize,
-        eps = eps,
-        parallel = paralle,
-        verbose = verbose, 
-        animate = FALSE)
-      z.hat <- clustering.partial$z
+      curves <- lapply(seq(K), function(k) {
+        estimate.bezier.curve.2(
+          X.partial[z.hat.partial == k, ],
+          degree = degree, 
+          k.isomap = as.integer(k.isomap * length(z.hat.partial) / n + 1),
+          intercept = intercept,
+          initialization = curve.init,
+          min.t = min.t, 
+          max.t = max.t, 
+          normalize = normalize,
+          parallel = parallel, 
+          eps = eps / K)
+      })
+      distances <- do.call('cbind', lapply(seq(K), function(k) {
+        compute.distances.bezier(X, 
+                                 clustering.partial$p[[k]],
+                                 min.t = min.t,
+                                 max.t = max.t, 
+                                 parallel = parallel,
+                                 intercept = intercept)
+      }))
+      z.hat <- apply(distances, 1, which.min)
+      for (k in seq(K)) {
+        z.hat[initialization == k] <- k
+      }
     }
   } else if (initialization == 'spectral') {
     if (missing(A)) {
@@ -539,7 +549,7 @@ manifold.clustering <- function(X, K = 2,
       
       update.curves.df <- plyr::ldply(seq(K), function(k) {
         X.hat <- 
-          construct.bezier.model.matrix(t.seq, r, intercept) %*% curves[[k]]$p
[…]

### 2022-07-07 draft/draft.Rmd
updated draft

--- 2022-05-05 draft/draft.Rmd
@@
 \begin{definition}[Manifold block model]
+\label{def:manifold-block-model}
 Let $p, q \geq 0$, $d = p + q \geq 1$, $1 \leq r < d$, $K \geq 2$, and $n \geq 1$ be integers.
@@
   \item Draw labels $z_1, ..., z_n \iid \Multinomial(\alpha_1, ..., \alpha_K)$. 
-  \item Draw latent vectors by first drawing each $t_i \indep F_{z_i}$ and then computing each $x_i = g_{z_i}(t_i)$. 
-  \item Let $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$, and draw $A \sim \RDPG(X)$ or $A \sim \GRDPG_{p,q}(X)$. 
+  \item Draw latent vectors by first drawing each $t_i \indep F_{z_i}$ and then compute each $x_i = g_{z_i}(t_i)$. 
+  \item Let $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$, and draw $A \sim \RDPG(X; \rho_n)$ or $A \sim \GRDPG_{p,q}(X; \rho_n)$. 
 \end{enumerate}
@@
 3. Let $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$, 
-and draw $A \sim \RDPG(X)$ or $A \sim \GRDPG_{p,q}(X)$.
+and draw $A \sim \RDPG(X; \rho_n)$ or $A \sim \GRDPG_{p,q}(X; \rho_n)$.
 
@@
 
-\begin{theorem}[Community detection for nonintersecting manifolds without noise]
-\label{nonintersect-no-noise}
-Let $x_1, ..., x_n$ be points sampled on $K$ manifolds $\mathcal{M}_1, ..., \mathcal{M}_K$. 
-Suppose $\delta = \min\limits_{k \neq \ell} \min\limits_{x_i \in \mathcal{M}_k, x_j \in \mathcal{M}_\ell} \| x_i - x_j \| > 0$. 
-Define $A_n$ as the event that for a $\eta$-neighborhood graph constructed from the sample $x_1, ..., x_n$ consists of exactly $K$ disconnected subgraphs that map on exactly to each manifold for some $\eta \in (0, \delta)$.
-Then for any $\epsilon \in (0, 1)$, there exists an $N$ such that when $n > N$, $A_n$ occurs with high probability. 
-\label{nonintersect-no-noise}
+\begin{theorem}[Community detection for the GRDPG for which the communities come from nonintersecting manifolds]
+\label{theorem:nonintersect-grdpg}
+Let $x_1, ..., x_n$ be points sampled on $K$ compact, connected manifolds $\mathcal{M}_1, ..., \mathcal{M}_K \subset \mathbb{R}^d$ each with probability measures $F_1, ..., F_K$, and the manifolds are separated by distance at least $\delta = \min\limits_{k \neq \ell} \min\limits_{x_i \in \mathcal{M}_k, x_j \in \mathcal{M}_\ell} \| x_i - x_j \| > 0$.
+Let $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$ and $A \sim \GRDPG_{p,q}(X; \rho_n)$ for some $p, q \in \mathbb{N}_0$ such that $p + q = d$ and sparsity parameter $\rho_n$ that satisfies $n \rho_n = \omega(\log^c n)$ for some $c > 1$.
+Define $A_n(\eta)$ as the event that an $\eta$-neighborhood graph constructed from the ASE of $A$ consists of exactly $K$ disconnected subgraphs that map exactly to each manifold. 
+Then for any $\eta \in (0, \delta)$, $\lim\limits_{n \to \infty} P(A_n(\eta)) = 1$. 
 \end{theorem}
+
+If the manifolds are one-dimensional, then a more precise rate of convergence can be derived.
 
@@
 \end{algorithm}
-
-\begin{theorem}
-\label{k-curves-clustering}
-Let each $g_k$ be smooth. 
-Then $K$-curves clustering converges to a stationary point of the objective, 
-$\sum_k \sum_{i \in C_k} \|x_i - g_k(t_i)\|^2$.
-\end{theorem}
-
-\begin{proof}
-$K$-curves clustering is a batch coordinate descent algorithm. 
-Thus, in order to show that it converges to a stationary point, it is sufficient to show that each descent step decreases the objective function. 
-\end{proof}
 
@@
 
-In order to prove theorem \ref{nonintersect-no-noise}, we first establish that the density of points within one manifold is sufficiently dense. 
-The following lemma is based on lemma 2 of \citet{trosset2020rehabilitating}.
-
 \begin{lemma}
-\label{lem:one-hypercube}
-Let $x_1, ..., x_n \iid F$ with support $[0, 1]^r$, and $f(x) \geq a > 0$ everywhere on the support. 
-Define $H_n$ as the event that an $\eta$-neighborhood graph constructed from the sample is connected for any $\eta > 0$. 
-Then for any $\epsilon > 0$, there exists $N = O \bigg( \frac{\log \epsilon + r \log \eta - \frac{r}{2} \log r}{\log (1 - a \eta^r r^{-r / 2})} \bigg)$ such that $P(H_n) > 1 - \epsilon$ when $n \geq N$.
+\label{lemma:no-noise}
+Let $x_1, ..., x_n$ be drawn from $K$ compact, connected manifolds $\mathcal{M}_1$, ..., $\mathcal{M}_K$ each with probability measures $F_1$, ..., $F_K$, and the manifolds are separated by distance at least $\delta > 0$.
+Suppose that for any $\epsilon > 0$ and $x$ drawn from each $F_k$, on $\mathcal{M}_k$, $F(B(x, \epsilon)) > 0$ where $B(x, \epsilon)$ is the open ball of radius $\epsilon$ centered at $x$. 
+Let $E_n(\eta)$ denote the event that an $\eta$-neighborhood graph constructed from $x_1, ..., x_n$ is comprised of exactly $K$ disjoint subgraphs that map to each of the $K$ manifolds. 
+Then if each $n_k \to \infty$ as $n \to \infty$, $\lim\limits_{n \to \infty} P(E_n(\eta)) = 1$ for each $\eta \in (0, \delta)$. 
 \end{lemma}
@@
 \begin{proof}
-Divide the hypercube $[0, 1]^r$ into a grid of sub-hypercubes of side length at most $\eta / \sqrt{r}$. 
-$E_n$ is satisfied if each sub-hypercube contains at least one $X_i$ from the sample.
+It is clear that if $\eta \in (0, \delta)$, an $\eta$-neighborhood graph constructed from the sample will always consist of at least $K$ disjoint subgraphs for which no subgraph contains vertices belonging to points from two different manifolds.
+Then it is sufficient to show that for a sufficiently large $n$, any $\eta$-neighborhood graph (where $\eta \in (0, \delta)$) will achieve $E_n$.
+
+Define each $E_{n_k}^{(k)}(\eta)$ as the event that if a sub-sample of size $n_k$ drawn from manifold $\mathcal{M}_k$, every $x \in \mathcal{M}_k$ is within distance $\eta$ of some $x_j$ of the sub-sample.
+Then if $E_{n_k}^{(k)}(\eta)$ is true, the $\eta$-neighborhood graph results in a connected subgraph for points within the $k^{th}$ manifold. 
+By lemma 2 of \citet{trosset2020rehabilitating}, $P((E_{n_k}^{(k)}(\eta))^c) \leq \ell_k (1 - b_k)^{n_k}$ for some $\ell_k \in \mathbb{N}$ and $b_k \in (0, 1]$. 
+If each $E_{n_k}^{(k)}(\eta)$ is true, then $E_n$ is achieved, so $E_n(\eta) = \bigcap_k E_{n_k}^{(k)}(\eta))$. 
+$\bigcap_k E_{n_k}^{(k)}(\eta) = \Big( \bigcup_k (E_{n_k}^{(k)}(\eta))^c \Big)^c$, so it is sufficient to show $\lim\limits_{n \to \infty} P \big( \bigcup_k (E_{n_k}^{(k)}(\eta))^c \big) \to 0$.
 
@@
 \begin{aligned}
-P(H_n) & = 1 - P(\text{some cells don't contain } X_i) \\
-& \geq 1 - \sum_m^{\lceil \sqrt{r} / \eta \rceil^r} \prod_i^n P(X_i \text{ is not in the } k^{th} \text{ hypercube}) \\
-& \geq 1 - \lceil \sqrt{r} / \eta \rceil^r (1 - a \eta^r / r^{r/2})^n,
+P \big( \bigcup_k(E_{n_k}^{(k)})^c \big) & \leq \sum_k P \big((E_{n_k}^{(k)})^c \big) \\
+& \leq \sum_k \ell_k (1 - b_k)^{n_k} \\
+& \leq K \ell_{\max} (1 - b_{\min})^{n_{\min}},
 \end{aligned}
 $$
-which approaches $1$ as $n \to \infty$. 
-Setting this quantity as  $\geq 1 - \epsilon$ and solving for $n$ yields the desired rate. 
+
+which tends to $0$ as $n \to \infty$. 
 \end{proof}
 
-Lemma \ref{lem:one-hypercube} extends to the case in which vectors are drawn from hypercubes with noise, given that the noise is bounded.
+\begin{proof}[Proof of theorem \ref{theorem:nonintersect-grdpg}]
+Define $E_n(\eta)$ as in lemma \ref{lemma:no-noise} and each $\epsilon_i = \|\hat{x}_i - Q_n x_i\|$ where $\hat{x}_i$ is the $i^{th}$ embedding vector and $Q_n$ is some indefinite orthogonal transformation as in \citet{https://doi.org/10.1111/rssb.12509}.
+Let $\epsilon = \max_i \epsilon_i$. 
+Then, by theorem 3 of \citet{https://doi.org/10.1111/rssb.12509}, for some finite $M > 0$, $P \big(\epsilon < M \frac{\log^c n}{\sqrt{n}} \big) \to 1$ as $n \to \infty$.
+Define $D_n$ as the event that $\epsilon < \delta / 2$. 
+For any $\delta > 0$, there is a finite $N$ such that when $n \geq N$, $M \frac{\log^c n}{\sqrt{n}} < \delta / 4$. 
+Then for the same $n$, $P(D_n) = 1$, so $\lim\limits_{n \to \infty} P(D_n) = 1$.
 
-\begin{lemma}
-\label{lem:one-hypercube-noise}
-Suppose the setup from lemma \ref{lem:one-hypercube}, but instead of observing each $x_i$, we observe each $y_i = x_i + e_i$ for some $e_1, ..., e_n \in \mathbb{R}^d$. 
-Let $\nu = \max_i \|e_i\| < \infty$ be bounded. 
-Define $\tilde{H}_n$ as the event in which an $(\eta + 2 \nu)$-neighborhood graph constructed from sample $y_1, ..., y_n$ is connected. 
-Then for any $\epsilon > 0$, there exists $N = O \bigg( \frac{\log \epsilon + r \log \eta - \frac{r}{2} \log r}{\log (1 - a \eta^r r^{-r / 2})} \bigg)$ such that $P(\tilde{H}_n) > 1 - \epsilon$ when $n \geq N$.
-\end{lemma}
+$A_n(\eta)$ is true if both $E_n(\eta)$ and $D_n$ are true, since $E_n(\eta)$ ensures that the $\eta$-neighborhood graph for each manifold is connected and $D_n$ ensures that the $\eta$-neighborhood graph has no edges between embedding points from different manifolds.
 
-\begin{proof}
-\end{proof}
-
-\begin{lemma}
-\label{lem:K-hypercubes}
-Let there be $K \geq 2$ hypercubes $\mathcal{C}_1, ..., \mathcal{C}_K$ of dimension $r_1, ..., r_K$ in $\mathbb{R}^d$ such that $\delta = \min\limits_{k \neq \ell} \min\limits_{x_i \in \mathcal{C}_k, x_j \in \mathcal{C}_\ell} \|x_i - x_j\| > 0$.
-Let $F_k$ be a distribution with support $C_k$ such that its density $f_k$ is nonzero on the support. Let $a = \min_k \min_t f_k(t) > 0$. 
-Define a mixture model as follows:
-
-\begin{enumerate}
-  \item Draw labels $z_1, ..., z_n \iid \Multinomial(\alpha_1, ..., \alpha_K)$. 
-  \item Draw latent vectors each as $x_i \indep F_{z_i}$.
-\end{enumerate}
-
-Define $H'_n$ as the event that an $\eta$-neighborhood graph constructed from this sample consists of exactly $K$ disconnected subgraphs for any $\eta \in (0, \delta)$. 
-Then for any $\epsilon \in (0, 1)$, there exists an $N = O \Big( \frac{\log(1 - (1 - \epsilon)^{1/K}) + d \log \eta - \frac{d}{2} \log d}{\alpha_{min} \log(1 - a \eta^d d^{-d/2})} \Big)$ such that when $n > N$, $P(E_n) > 1 - \epsilon$.
-\end{lemma}
-
-\begin{proof}
-Let $H^{(k)}$ be the event that lemma \ref{lem:one-hypercube} holds for $\mathcal{C}_k$. Then $H^{(k)} = H_{n_k}$ where $H_n$ is defined as in lemma \ref{lem:one-hypercube}. 
-Then $P(H'_n) = P(H_{n_1} \text{ and } ... \text{ and } H_{n_K})$.
+Then to show that $P(A_n) \to 1$:
 
@@
 \begin{aligned}
-P(H'_n) & = \prod_k P(H_{n_k}) \\
-& \geq \prod_k 1 - \lceil \sqrt{r_k} / \eta \rceil^{r_k} (1 - a \eta^{r_k} r_k^{-r_k/2})^{n_k} \\
-& \geq \prod_k 1 - \lceil \sqrt{d} / \eta \rceil^{d} (1 - a \eta^{d} d^{-d/2})^{\alpha_{\min} n} \\
-& \geq \big(1 - \lceil \sqrt{d} / \eta \rceil^{d} (1 - a \eta^{d} d^{-d/2})^{\alpha_{\min} n} \big)^K, \\
+P(A_n) & = P(E_n(\eta) \cap D_n) \\
+& = P((E_n^c(\eta) \cup D_n^c)^c) \\
+& = 1 - P(E_n^c(\eta) \cup D_n^c) \\
+& \geq 1 - P(E_n^c(\eta)) - P(D_n^c) \\ 
+& = P(E_n(\eta)) + P(D_n) - 1,
 \end{aligned}
 $$
-which approaches 1 as $n \to \infty$. 
-Setting this quantity to $\geq 1 - \epsilon$ and solving for $n$ yields the desired rate. 
+which tends toward $1$ as $n \to \infty$ since both $P(E_n(\eta))$ and $P(D_n)$ tend toward $1$ as $n \to \infty$. 
 \end{proof}
 
-\begin{proof}[Proof of theorem \ref{nonintersect-no-noise}]
-\end{proof}
-
-It can similarly be shown that the scenario in lemma \ref{lem:K-hypercubes} with bounded noise can be achieved by replacing the $\eta$-neighborhood graph with an $(\eta + 2 \nu)$-neighborhood graph, where $\nu = \max_i \|e_i\|$.
-This time, it is necessary to bound the noise as $\nu < \delta / 3$ to maintain separation between the hypercubes.
-
 \section{Details on Fitting Bezier Curves with Noise}

### 2022-07-07 draft/draft.tex
updated draft

@@
   \item Draw latent vectors by first drawing each $t_i \stackrel{\mathrm{ind}}{\sim}F_{z_i}$ and then compute each $x_i = g_{z_i}(t_i)$. 
-  \item Let $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$, and draw $A \sim \mathrm{RDPG}(X)$ or $A \sim \mathrm{GRDPG}_{p,q}(X)$. 
+  \item Let $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$, and draw $A \sim \mathrm{RDPG}(X; \rho_n)$ or $A \sim \mathrm{GRDPG}_{p,q}(X; \rho_n)$. 
 \end{enumerate}
@@
   Let \(X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top\), and
-  draw \(A \sim \mathrm{RDPG}(X)\) or
-  \(A \sim \mathrm{GRDPG}_{p,q}(X)\).
+  draw \(A \sim \mathrm{RDPG}(X; \rho_n)\) or
+  \(A \sim \mathrm{GRDPG}_{p,q}(X; \rho_n)\).
 \end{enumerate}
@@
 
-\begin{theorem}[Community detection for nonintersecting manifolds without noise]
-\label{nonintersect-no-noise}
-Let $x_1, ..., x_n$ be points sampled on $K$ manifolds $\mathcal{M}_1, ..., \mathcal{M}_K$. 
-Suppose $\delta = \min\limits_{k \neq \ell} \min\limits_{x_i \in \mathcal{M}_k, x_j \in \mathcal{M}_\ell} \| x_i - x_j \| > 0$. 
-Define $A_n$ as the event that a $\eta$-neighborhood graph constructed from the sample $x_1, ..., x_n$ consists of exactly $K$ disconnected subgraphs that map exactly to each manifold for some $\eta \in (0, \delta)$.
-Then for any $\epsilon \in (0, 1)$, there exists an $N$ such that when $n > N$, $P(A_n) > 1 - \epsilon$. 
+\begin{theorem}[Community detection for the GRDPG for which the communities come from nonintersecting manifolds]
+\label{theorem:nonintersect-grdpg}
+Let $x_1, ..., x_n$ be points sampled on $K$ compact, connected manifolds $\mathcal{M}_1, ..., \mathcal{M}_K \subset \mathbb{R}^d$ each with probability measures $F_1, ..., F_K$, and the manifolds are separated by distance at least $\delta = \min\limits_{k \neq \ell} \min\limits_{x_i \in \mathcal{M}_k, x_j \in \mathcal{M}_\ell} \| x_i - x_j \| > 0$.
+Let $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$ and $A \sim \mathrm{GRDPG}_{p,q}(X; \rho_n)$ for some $p, q \in \mathbb{N}_0$ such that $p + q = d$ and sparsity parameter $\rho_n$ that satisfies $n \rho_n = \omega(\log^c n)$ for some $c > 1$.
+Define $A_n(\eta)$ as the event that an $\eta$-neighborhood graph constructed from the ASE of $A$ consists of exactly $K$ disconnected subgraphs that map exactly to each manifold. 
+Then for any $\eta \in (0, \delta)$, $\lim\limits_{n \to \infty} P(A_n(\eta)) = 1$. 
 \end{theorem}
 
-\begin{theorem}{Community detection for RDPG for which the communities come from nonintersecting manifolds}
-\label{nonintersect-rdpg}
-Let $x_1, ..., x_n$ be points sampled on $K$ manifolds $\mathcal{M}_1, ..., \mathcal{M}_K$. 
-Suppose $\delta = \min\limits_{k \neq \ell} \min\limits_{x_i \in \mathcal{M}_k, x_j \in \mathcal{M}_\ell} \| x_i - x_j \| > 0$. 
-Let $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$ and $A \sim \mathrm{RDPG}(X)$. 
-Define $B_n$ as the event that an $\eta$-neighborhood graph constructed from the ASE of $A$ consists of exactly $K$ disconnected subgraphs that map exactly to each manifold for some $\eta \in (0, \delta)$.
-Then for any $\epsilon \in (0, 1)$, there exists an $N$ such that when $n > N$, $P(B_n) > 1 - \epsilon$. 
-\end{theorem}
+If the manifolds are one-dimensional, then a more precise rate of
+convergence can be derived.
 
@@
 rves clustering.\} \textbackslash end\{algorithm\}
-
-\begin{theorem}
-\label{k-curves-clustering}
-Let each $g_k$ be smooth. 
-Then $K$-curves clustering converges to a stationary point of the objective, 
-$\sum_k \sum_{i \in C_k} \|x_i - g_k(t_i)\|^2$.
-\end{theorem}
-
-\begin{proof}
-$K$-curves clustering is a batch coordinate descent algorithm. 
-Thus, in order to show that it converges to a stationary point, it is sufficient to show that each descent step decreases the objective function. 
-\end{proof}
 
@@
 
-In order to prove theorem \ref{nonintersect-no-noise}, we first
-establish that the density of points within one manifold is sufficiently
-dense. The following lemma is based on lemma 2 of
-\citet{trosset2020rehabilitating}.
-
 \begin{lemma}
-\label{lem:one-hypercube}
-Let $x_1, ..., x_n \stackrel{\mathrm{iid}}{\sim}F$ with support $[0, 1]^r$, and $f(x) \geq a > 0$ everywhere on the support. 
-Define $H_n$ as the event that an $\eta$-neighborhood graph constructed from the sample is connected for any $\eta > 0$. 
-Then for any $\epsilon > 0$, there exists $N = O \bigg( \frac{\log \epsilon + r \log \eta - \frac{r}{2} \log r}{\log (1 - a \eta^r r^{-r / 2})} \bigg)$ such that $P(H_n) > 1 - \epsilon$ when $n \geq N$.
+\label{lemma:no-noise}
+Let $x_1, ..., x_n$ be drawn from $K$ compact, connected manifolds $\mathcal{M}_1$, ..., $\mathcal{M}_K$ each with probability measures $F_1$, ..., $F_K$, and the manifolds are separated by distance at least $\delta > 0$.
+Suppose that for any $\epsilon > 0$ and $x$ drawn from each $F_k$, on $\mathcal{M}_k$, $F(B(x, \epsilon)) > 0$ where $B(x, \epsilon)$ is the open ball of radius $\epsilon$ centered at $x$. 
+Let $E_n(\eta)$ denote the event that an $\eta$-neighborhood graph constructed from $x_1, ..., x_n$ is comprised of exactly $K$ disjoint subgraphs that map to each of the $K$ manifolds. 
+Then if each $n_k \to \infty$ as $n \to \infty$, $\lim\limits_{n \to \infty} P(E_n(\eta)) = 1$ for each $\eta \in (0, \delta)$. 
 \end{lemma}
@@
 \begin{aligned}
-P(H_n) & = 1 - P(\text{some cells don't contain } X_i) \\
-& \geq 1 - \sum_m^{\lceil \sqrt{r} / \eta \rceil^r} \prod_i^n P(X_i \text{ is not in the } k^{th} \text{ hypercube}) \\
-& \geq 1 - \lceil \sqrt{r} / \eta \rceil^r (1 - a \eta^r / r^{r/2})^n,
+P \big( \bigcup_k(E_{n_k}^{(k)})^c \big) & \leq \sum_k P \big((E_{n_k}^{(k)})^c \big) \\
+& \leq \sum_k \ell_k (1 - b_k)^{n_k} \\
+& \leq K \ell_{\max} (1 - b_{\min})^{n_{\min}},
 \end{aligned}
 $$
-which approaches $1$ as $n \to \infty$. 
-Setting this quantity as  $\geq 1 - \epsilon$ and solving for $n$ yields the desired rate. 
+
+which tends to $0$ as $n \to \infty$. 
 \end{proof}
 
-Lemma \ref{lem:one-hypercube} extends to the case in which vectors are
-drawn from hypercubes with noise, given that the noise is bounded.
+\begin{proof}[Proof of theorem \ref{theorem:nonintersect-grdpg}]
+Define $E_n(\eta)$ as in lemma \ref{lemma:no-noise} and each $\epsilon_i = \|\hat{x}_i - Q_n x_i\|$ where $\hat{x}_i$ is the $i^{th}$ embedding vector and $Q_n$ is some indefinite orthogonal transformation as in \citet{https://doi.org/10.1111/rssb.12509}.
+Let $\epsilon = \max_i \epsilon_i$. 
+Then, by theorem 3 of \citet{https://doi.org/10.1111/rssb.12509}, for some finite $M > 0$, $P \big(\epsilon < M \frac{\log^c n}{\sqrt{n}} \big) \to 1$ as $n \to \infty$.
+Define $D_n$ as the event that $\epsilon < \delta / 2$. 
+For any $\delta > 0$, there is a finite $N$ such that when $n \geq N$, $M \frac{\log^c n}{\sqrt{n}} < \delta / 4$. 
+Then for the same $n$, $P(D_n) = 1$, so $\lim\limits_{n \to \infty} P(D_n) = 1$.
 
-\begin{lemma}
-\label{lem:one-hypercube-noise}
-Suppose the setup from lemma \ref{lem:one-hypercube}, but instead of observing each $x_i$, we observe each $y_i = x_i + e_i$ for some $e_1, ..., e_n \in \mathbb{R}^d$. 
-Let $\nu = \max_i \|e_i\| < \infty$ be bounded. 
-Define $\tilde{H}_n$ as the event in which an $(\eta + 2 \nu)$-neighborhood graph constructed from sample $y_1, ..., y_n$ is connected. 
-Then for any $\epsilon > 0$, there exists $N = O \bigg( \frac{\log \epsilon + r \log \eta - \frac{r}{2} \log r}{\log (1 - a \eta^r r^{-r / 2})} \bigg)$ such that $P(\tilde{H}_n) > 1 - \epsilon$ when $n \geq N$.
-\end{lemma}
+$A_n(\eta)$ is true if both $E_n(\eta)$ and $D_n$ are true, since $E_n(\eta)$ ensures that the $\eta$-neighborhood graph for each manifold is connected and $D_n$ ensures that the $\eta$-neighborhood graph has no edges between embedding points from different manifolds.
 
-\begin{proof}
-\end{proof}
-
-\begin{lemma}
-\label{lem:K-hypercubes}
-Let there be $K \geq 2$ hypercubes $\mathcal{C}_1, ..., \mathcal{C}_K$ of dimension $r_1, ..., r_K$ in $\mathbb{R}^d$ such that $\delta = \min\limits_{k \neq \ell} \min\limits_{x_i \in \mathcal{C}_k, x_j \in \mathcal{C}_\ell} \|x_i - x_j\| > 0$.
-Let $F_k$ be a distribution with support $C_k$ such that its density $f_k$ is nonzero on the support. Let $a = \min_k \min_t f_k(t) > 0$. 
-Define a mixture model as follows:
-
-\begin{enumerate}
-  \item Draw labels $z_1, ..., z_n \stackrel{\mathrm{iid}}{\sim}\mathrm{Multinomial}(\alpha_1, ..., \alpha_K)$. 
-  \item Draw latent vectors each as $x_i \stackrel{\mathrm{ind}}{\sim}F_{z_i}$.
-\end{enumerate}
-
-Define $H'_n$ as the event that an $\eta$-neighborhood graph constructed from this sample consists of exactly $K$ disconnected subgraphs for any $\eta \in (0, \delta)$. 
-Then for any $\epsilon \in (0, 1)$, there exists an $N = O \Big( \frac{\log(1 - (1 - \epsilon)^{1/K}) + d \log \eta - \frac{d}{2} \log d}{\alpha_{min} \log(1 - a \eta^d d^{-d/2})} \Big)$ such that when $n > N$, $P(E_n) > 1 - \epsilon$.
-\end{lemma}
-
-\begin{proof}
-Let $H^{(k)}$ be the event that lemma \ref{lem:one-hypercube} holds for $\mathcal{C}_k$. Then $H^{(k)} = H_{n_k}$ where $H_n$ is defined as in lemma \ref{lem:one-hypercube}. 
-Then $P(H'_n) = P(H_{n_1} \text{ and } ... \text{ and } H_{n_K})$.
+Then to show that $P(A_n) \to 1$:
 
@@
 \begin{aligned}
-P(H'_n) & = \prod_k P(H_{n_k}) \\
-& \geq \prod_k 1 - \lceil \sqrt{r_k} / \eta \rceil^{r_k} (1 - a \eta^{r_k} r_k^{-r_k/2})^{n_k} \\
-& \geq \prod_k 1 - \lceil \sqrt{d} / \eta \rceil^{d} (1 - a \eta^{d} d^{-d/2})^{\alpha_{\min} n} \\
-& \geq \big(1 - \lceil \sqrt{d} / \eta \rceil^{d} (1 - a \eta^{d} d^{-d/2})^{\alpha_{\min} n} \big)^K, \\
+P(A_n) & = P(E_n(\eta) \cap D_n) \\
+& = P((E_n^c(\eta) \cup D_n^c)^c) \\
+& = 1 - P(E_n^c(\eta) \cup D_n^c) \\
+& \geq 1 - P(E_n^c(\eta)) - P(D_n^c) \\ 
+& = P(E_n(\eta)) + P(D_n) - 1,
 \end{aligned}
 $$
-which approaches 1 as $n \to \infty$. 
-Setting this quantity to $\geq 1 - \epsilon$ and solving for $n$ yields the desired rate. 
+which tends toward $1$ as $n \to \infty$ since both $P(E_n(\eta))$ and $P(D_n)$ tend toward $1$ as $n \to \infty$. 
 \end{proof}
-
-\begin{proof}[Proof of theorem \ref{nonintersect-no-noise}]
-\end{proof}
-
-It can similarly be shown that the scenario in lemma
-\ref{lem:K-hypercubes} with bounded noise can be achieved by replacing
-the \(\eta\)-neighborhood graph with an \((\eta + 2 \nu)\)-neighborhood
-graph, where \(\nu = \max_i \|e_i\|\). This time, it is necessary to
-bound the noise as \(\nu < \delta / 3\) to maintain separation between
-the hypercubes.
 

### 2022-07-07 draft/draft.Rmd
updated draft

@@
 
+```{r}
+clustering.df <- readr::read_csv('../simulations/balanced.csv')
+```
+
+```{r}
+clust.summary.df <- clustering.df %>% 
+  dplyr::group_by(n, nsamp) %>% 
+  dplyr::summarise(
+    med.count = median(error.count),
+    first.q.count = quantile(error.count, .25),
+    third.q.count = quantile(error.count, .75),
+    med.rate = median(error.rate),
+    first.q.rate = quantile(error.rate, .25),
+    third.q.rate = quantile(error.rate, .75)
+  ) %>% 
+  dplyr::ungroup()
+
+n.vec <- unique(clustering.df$n)
+
+ggplot(clust.summary.df) + 
+  theme_bw() + 
+  theme(text = element_text(size = 10)) + 
+  scale_y_log10() +
+  scale_x_log10(breaks = n.vec) +
+  labs(y = 'community detection error rate',
+       colour = NULL, shape = NULL) + 
+  geom_line(aes(x = n, y = med.rate, colour = factor(nsamp))) + 
+  geom_point(aes(x = n, y = med.rate, colour = factor(nsamp))) + 
+  geom_errorbar(aes(x = n, 
+                    ymin = first.q.rate, ymax = third.q.rate, 
+                    colour = factor(nsamp))) + 
+  scale_colour_brewer(palette = 'Set1')
+```
+
 # Discussion

### 2022-07-19 draft/draft.Rmd
trying to rework proof of theorem 1

@@
 \newtheorem{lemma}{Lemma}
+\newtheorem{corollary}{Corollary}
 \newtheorem{proposition}{Proposition}
@@
 source('~/dev/pabm-grdpg/functions.R')
-source('https://mtrosset.pages.iu.edu/Courses/675/graph.r')
+# source('https://mtrosset.pages.iu.edu/Courses/675/graph.r')
 
@@
 Define $A_n(\eta)$ as the event that an $\eta$-neighborhood graph constructed from the ASE of $A$ consists of exactly $K$ disconnected subgraphs that map exactly to each manifold. 
-Then for any $\eta \in (0, \delta)$, $\lim\limits_{n \to \infty} P(A_n(\eta)) = 1$. 
+Then for some $C > 0$ and any $\eta \in (0, C \delta)$, $\lim\limits_{n \to \infty} P(A_n(\eta)) = 1$. 
 \end{theorem}
@@
 
+\begin{corollary}
+Let $x_1, ..., x_n$ be drawn from $K$ compact, connected, one-dimensional manifolds $\mathcal{M}_1, ..., \mathcal{M}_K$, each with probability measures $F_1, ..., F_K$, and the manifolds are spearated by distance at least $\delta > 0$.
+
+\end{corollary}
+
 \begin{proof}[Proof of theorem \ref{theorem:nonintersect-grdpg}]
-Define $E_n(\eta)$ as in lemma \ref{lemma:no-noise} and each $\epsilon_i = \|\hat{x}_i - Q_n x_i\|$ where $\hat{x}_i$ is the $i^{th}$ embedding vector and $Q_n$ is some indefinite orthogonal transformation as in \citet{https://doi.org/10.1111/rssb.12509}.
-Let $\epsilon = \max_i \epsilon_i$. 
-Then, by theorem 3 of \citet{https://doi.org/10.1111/rssb.12509}, for some finite $M > 0$, $P \big(\epsilon < M \frac{\log^c n}{\sqrt{n}} \big) \to 1$ as $n \to \infty$.
-Define $D_n$ as the event that $\epsilon < \delta / 2$. 
-For any $\delta > 0$, there is a finite $N$ such that when $n \geq N$, $M \frac{\log^c n}{\sqrt{n}} < \delta / 4$. 
-Then for the same $n$, $P(D_n) = 1$, so $\lim\limits_{n \to \infty} P(D_n) = 1$.
+Define $E_n(\eta)$ as in lemma \ref{lemma:no-noise} for manifolds $Q_n(\mathcal{M}_1), ..., Q_n(\mathcal{M}_K)$ and each $e_i = \hat{x}_i - Q_n x_i$ where $\hat{x}_i$ is the $i^{th}$ embedding vector and $Q_n$ is some indefinite orthogonal transformation as in \citet{https://doi.org/10.1111/rssb.12509}.
+Since $Q_n$ is a linear map, for any $\eta \in (0, \|Q_n\| \delta)$, $P(E_n(\eta)) \to 1$ as $n \to \infty$. 
+Let $\epsilon_i = \|e_i\|$, $\epsilon = \max_i \epsilon_i$, and $C_n = \|Q_n\|$.
 
-$A_n(\eta)$ is true if both $E_n(\eta)$ and $D_n$ are true, since $E_n(\eta)$ ensures that the $\eta$-neighborhood graph for each manifold is connected and $D_n$ ensures that the $\eta$-neighborhood graph has no edges between embedding points from different manifolds.
+$A_n(\eta)$ is true if $\eta < \min\limits_{k, \ell} \min\limits_{x_i \in \mathcal{M}_k, x_j \in \mathcal{M}_\ell} \|\hat{x}_i - \hat{x}_j\|$, which is defined as event $D_n(\eta)$, and $\eta \geq \max\limits_k \max\limits_{x_i, x_j \in \mathcal{M}_k} \|\hat{x}_i - \hat{x}_j\|$, which is defined as event $\hat{E}_n(\eta)$.
+
+For any $x_i, x_j$ from different manifolds, $\|\hat{x}_i - \hat{x}_j\| \geq C_n \delta - 2 \epsilon$ if $2 \epsilon \leq C_n \delta$. 
+By theorem 3 of \citet{https://doi.org/10.1111/rssb.12509}, for some finite $M > 0$, $P \big(\epsilon < M \frac{\log^c n}{\sqrt{n}} \big) \to 1$ as $n \to \infty$, so $P(C_n \delta < 2 \epsilon) \leq P(C_n \delta < 2 M n^{1/2} \log^c n) \to 0$ since $C_n \delta > 0$.
+Then since $P(C_n \delta - 2 \epsilon > 0) \to 1$, there is an $\epsilon \in (0, C_n \delta - 2 \epsilon)$ with probability 1. 
+Thus, $P(D_n(\eta)) \to 1$.
 
@@
 \begin{aligned}
-P(A_n) & = P(E_n(\eta) \cap D_n) \\
-& = P((E_n^c(\eta) \cup D_n^c)^c) \\
-& = 1 - P(E_n^c(\eta) \cup D_n^c) \\
-& \geq 1 - P(E_n^c(\eta)) - P(D_n^c) \\ 
-& = P(E_n(\eta)) + P(D_n) - 1,
+P(A_n) & = P(\hat{E}_n(\eta) \cap D_n) \\
+& = P((\hat{E}_n^c(\eta) \cup D_n^c)^c) \\
+& = 1 - P(\hat{E}_n^c(\eta) \cup D_n^c) \\
+& \geq 1 - P(\hat{E}_n^c(\eta)) - P(D_n^c) \\ 
+& = P(\hat{E}_n(\eta)) + P(D_n) - 1,
 \end{aligned}

### 2022-07-19 draft/draft.tex
updated draft
trying to rework proof of theorem 1

@@
 
+\begin{center}\includegraphics{draft_files/figure-latex/unnamed-chunk-17-1} \end{center}
+
 \hypertarget{discussion}{%

### 2023-05-12 sdss-2023/poster.Rmd
sdss slides and poster

---
title: Manifold Clustering in the Setting of Generalized Random Dot Product Graphs
author:
  - name: John Koo
    affil: 1
  - name: Minh Tang
    affil: 2
  - name: Michael W. Trosset
    affil: 1
affiliation:
  - num: 1
    address: Indiana University
  - num: 2
    address: North Carolina State University
title_textsize: "56pt"
author_textsize: "32pt"
affiliation_textsize: "8pt"
sectitle_textsize: "36pt"
sectitle2_textsize: "36pt"
body_textsize: "32pt"
column_numbers: 4
poster_height: "20.5in"
poster_width: "36.5in"
column_margins: "0.1in"
primary_colour: "#990000"
secondary_colour: "#000000"
accent_colour: "#990000"
titlebox_borderwidth: "0.1cm"
titlebox_bordercol: "#000000"
title_textcol: "#ffffff"
author_textcol: "#ffffff"
sectitle2_textcol: "#990000"
sectitle_borderwidth: "0mm"
font_family: "Times New Roman"
logoleft_name: poster_files/trident_medium.png
output: 
  posterdown::posterdown_html:
    self_contained: true
    keep_tex: true
header-includes:
- \usepackage{setspace}
- \usepackage{float}
- \usepackage{mathtools}
- \usepackage{natbib}
- \usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
- \setcitestyle{numbers,square,comma}
- \usepackage{verbatim}
- \usepackage{amsthm}
- \usepackage{comment}
- \usepackage{graphicx}
- \setbeamertemplate{itemize items}[circle]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      comment = NA, 
                      warning = FALSE, 
                      message = FALSE,
                      fig.align = 'center',
                      fig.lp = '')
import::from(magrittr, `%>%`)
library(ggplot2)
library(ggraph)
source('~/dev/pabm-grdpg/functions.R')
source('~/dev/manifold-block-models/functions.R')
```

\newcommand{\diag}{\text{diag}}
\newcommand{\tr}{\text{Tr}}
\newcommand{\blockdiag}{\text{blockdiag}}
\newcommand{\indep}{\stackrel{\text{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\text{iid}}{\sim}}
\newcommand{\Bernoulli}{\text{Bernoulli}}
\newcommand{\Betadist}{\text{Beta}}
\newcommand{\BG}{\text{BernoulliGraph}}
\newcommand{\Categorical}{\text{Cat}}
\newcommand{\Uniform}{\text{Uniform}}
\newcommand{\RDPG}{\text{RDPG}}
\newcommand{\GRDPG}{\text{GRDPG}}
\newcommand{\PABM}{\text{PABM}}
\newtheorem{theorem}{Theorem}

# Generalized Random Dot Product Graphs

The *generalized random dot product graph* is a random graph model in which each vertex $v_i$ has a corresponding hidden vector $x_i \in \mathbb{R}^{p+q}$ and each edge probability is the indefinite inner product of the corresponding pair of hidden vectors, i.e., $P_{ij} = x_i^\top I_{p, q} x_j$, $I_{p,q} = \Bigl[\begin{smallmatrix} I_{p} & 0 \\ 0 & - I_{q} \end{smallmatrix} \Bigr]$

## Adjacency Spectral Embedding
Approximate $A$ by spectral decomposition $A \approx V_{p, q} \Lambda_{p, q} V_{p, q}^\top$. 
The subscript $p, q$ denotes the $p$ most positive and $q$ most negative eigenvalues and corresponding eigenvectors. 
Each $\hat{x}_i$, the $i^{th}$ row of $\hat{X} = V_{p, q} |\Lambda_{p, q}|^{1/2},$ estimates the relative position of its corresponding latent vector $x_i$, up to an indefinite orthogonal transformation.

**Theorem** (Rubin-Delanchy et al. 2022): $\max_i \| \hat{x}_i - Q_n x_i \| = O_P \Big( \frac{\log^c n}{n^{1/2}} \Big)$ for some $Q_n \in \mathbb{O}(p, q)$.

```{r, cache = TRUE}
n1 <- 2 ** 5
n2 <- 2 ** 5
n <- n1 + n2
z <- c(rep(1, n1), rep(2, n2))
u1 <- runif(n1)
u2 <- runif(n2)
x1 <- cos(pi / 3 * u1)
y1 <- sin(pi / 3 * u1)
x2 <- 1 + cos(pi / 3 * u2 + pi)
y2 <- 1 + sin(pi / 3 * u2 + pi)

data.matrix <- cbind(c(x1, x2), c(y1, y2))
P <- data.matrix %*% diag(c(1, 1)) %*% t(data.matrix)
A <- draw.graph(P)
A.graph <- tidygraph::as_tbl_graph(A, directed = FALSE)

Xhat <- embedding(A, 2, 0)
Xhat[, 1] <- -Xhat[, 1]
```

```{r, fig.height = 2.25, fig.width = 6, out.width = '100%'}
latent.config.plot <- ggplot() + 
  geom_point(aes(x = data.matrix[, 1], y = data.matrix[, 2], 
                 colour = factor(z))) + 
  # coord_fixed() + 
  theme_bw() + 
  theme(legend.position = 'none') + 
  labs(x = NULL, y = NULL, colour = NULL, title = 'Latent Vectors') + 
  scale_colour_brewer(palette = 'Set1') +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

graph.plot <- ggraph(A, layout = 'stress') + 
  geom_edge_link(alpha = 1 / sqrt(n) / 2) + 
  geom_node_point(aes(colour = factor(z)),
                  show.legend = FALSE) + 
  labs(colour = NULL, 
       title = 'GRDPG') + 
  # theme_graph() + 
  theme_void() + 
  scale_colour_brewer(palette = 'Set1')

ase.plot <- ggplot() + 
  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2],
                 colour = factor(z))) + 
  # coord_fixed() + 
  theme_bw() + 
  theme(legend.position = 'none') + 
  labs(x = NULL, y = NULL, colour = NULL, title = 'ASE') + 
  scale_colour_brewer(palette = 'Set1') +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

gridExtra::grid.arrange(latent.config.plot, graph.plot, ase.plot, nrow = 1)
```

# Block Models as GRDPGs

It has been previously shown that the SBM, DCBM, and PABM are GRDPGs in which the communities lie on point masses, line segments, and subspaces, respectively.

```{r, cache = TRUE, fig.width = 5, fig.height = 2.3, out.width = '100%'}
n <- 64
n1 <- n / 2
n2 <- n / 2
z <- c(rep(1, n1), rep(2, n2))
p <- 1/2
q <- 1/2
r <- 1/8
P <- matrix(r, nrow = n, ncol = n)
P[seq(n1), seq(n1)] <- p
P[seq(n1 + 1, n), seq(n1 + 1, n)] <- q
X <- embedding(P, 2, 0)
A <- draw.graph(P)
Xhat <- embedding(A, 2, 0)
A.graph <- tidygraph::as_tbl_graph(A, directed = FALSE)

sbm.graph.plot <- ggraph(A, layout = 'stress') + 
  geom_edge_link(alpha = 1 / sqrt(n) / 2) + 
  geom_node_point(aes(colour = factor(z)),
                  show.legend = FALSE) + 
  labs(colour = NULL, 
       title = 'SBM') + 
  # theme_graph() + 
  theme_void() + 
  scale_colour_brewer(palette = 'Set1')

sbm.ase.plot <- ggplot() + 
  geom_point(aes(x = Xhat[, 2], y = Xhat[, 1],
                 colour = factor(z)),
             shape = 4) + 
  geom_point(aes(x = X[, 2], y = X[, 1], 
                 colour = factor(z))) + 
  # coord_fixed() +
  theme_bw() + 
  labs(x = NULL, y = NULL, colour = NULL, 
       title = 'Point Masses') + 
  scale_colour_brewer(palette = 'Set1') +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = 'none')

gridExtra::grid.arrange(sbm.graph.plot, sbm.ase.plot, nrow = 1)
```

```{r, cache = TRUE, fig.width = 5, fig.height = 2.3, out.width = '100%'}
n <- 128
n1 <- n / 2
n2 <- n / 2
z <- c(rep(1, n1), rep(2, n2))
p <- 1/2
q <- 1/2
r <- 1/8
P <- matrix(r, nrow = n, ncol = n)
P[seq(n1), seq(n1)] <- p
P[seq(n1 + 1, n), seq(n1 + 1, n)] <- q
omega <- rbeta(n, 2, 1)
P <- P * (omega %*% t(omega))
X <- embedding(P, 2, 0)
A <- draw.graph(P)
Xhat <- embedding(A, 2, 0)
Xhat <- Xhat
A.graph <- tidygraph::as_tbl_graph(A, directed = FALSE)

dcbm.graph.plot <- ggraph(A, layout = 'stress') + 
  geom_edge_link(alpha = 1 / sqrt(n) / 2) + 
  geom_node_point(aes(colour = factor(z)),
                  show.legend = FALSE) + 
  labs(colour = NULL, 
       title = 'DCBM') + 
  # theme_graph() + 
  theme_void() + 
  scale_colour_brewer(palette = 'Set1')

dcbm.ase.plot <- ggplot() + 
  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2],
                 colour = factor(z)),
             shape = 4) + 
  geom_point(aes(x = X[, 1], y = X[, 2], 
                 colour = factor(z))) + 
  # coord_fixed() +
  theme_bw() + 
  labs(x = NULL, y = NULL, colour = NULL, 
       title = 'Rays') + 
  scale_colour_brewer(palette = 'Set1') +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = 'none')

gridExtra::grid.arrange(dcbm.graph.plot, dcbm.ase.plot, nrow = 1)
```

```{r, cache = TRUE, fig.width = 5, fig.height = 2.3, out.width = '100%'}
n <- 64
Pz <- generate.P.beta(n)
P <- Pz$P
z <- Pz$clustering
X <- embedding(P, 3, 1)
A <- draw.graph(P)
Xhat <- embedding(A, 3, 1)
# Xhat[, 2] <- -Xhat[, 2]
A.graph <- tidygraph::as_tbl_graph(A, directed = FALSE)

pabm.graph.plot <- ggraph(A, layout = 'stress') + 
  geom_edge_link(alpha = 1 / sqrt(n) / 2) + 
  geom_node_point(aes(colour = factor(z)),
                  show.legend = FALSE) + 
  labs(colour = NULL, 
       title = 'PABM') + 
  # theme_graph() + 
  theme_void() + 
  scale_colour_brewer(palette = 'Set1')

pabm.ase.plot <- ggplot() + 
  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2],
                 colour = factor(z)),
             shape = 4) + 
  geom_point(aes(x = X[, 1], y = X[, 2], 
                 colour = factor(z))) + 
  # coord_fixed() +
  theme_bw() + 
  theme(legend.position = 'none',
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) + 
  labs(x = NULL, y = NULL, colour = NULL, 
       title = 'Subspaces (Projected)') + 
  scale_colour_brewer(palette = 'Set1')

gridExtra::grid.arrange(pabm.graph.plot, pabm.ase.plot, nrow = 1)
```

# GRDPGs with Nonlinear Latent Structure

```{r mbconnectome-graph, fig.height = 3, figh.width = 6, cache = TRUE, out.width = '67%'}
library(mbstructure)
data(MBconnectome)

graph.r <- generate.graph(newrdat, vdf.right)
A.r <- as.matrix(igraph::as_adjacency_matrix(graph.r$g, type = 'both'))
z.r <- as.numeric(graph.r$vdf$type)
n.r <- length(z.r)
A.r <- sign(A.r + t(A.r))
diag(A.r) <- 0
graph.r <- ggraph(A.r, layout = 'stress') + 
  geom_edge_link(alpha = 1 / sqrt(n.r) / 2) + 
  geom_node_point(aes(colour = factor(z.r)), 
                  show.legend = FALSE) + 
  labs(colour = NULL) + 
  theme_void() + 
  scale_colour_brewer(palette = 'Set1')

graph.l <- generate.graph(newldat, vdf.left)
A.l <- as.matrix(igraph::as_adjacency_matrix(graph.l$g, type = 'both'))
z.l <- as.numeric(graph.l$vdf$type)
n.l <- length(z.l)
A.l <- sign(A.l + t(A.l))
diag(A.l) <- 0
graph.l <- ggraph(A.l, layout = 'stress') + 
  geom_edge_link(alpha = 1 / sqrt(n.l) / 2) + 
  geom_node_point(aes(colour = factor(z.l)), 
                  show.legend = FALSE) + 
  labs(colour = NULL) + 
  theme_void() + 
  scale_colour_brewer(palette = 'Set1')

# gridExtra::grid.arrange(graph.l, graph.r, nrow = 1)
```

```{r mbconnectome-ase, fig.height = 3.75, fig.width = 7, out.width = '100%', cache = TRUE}
Xhat.l <- embedding(A.l, 2, 1)
ase.l <- ggplot() + 
  geom_point(aes(x = Xhat.l[, 1], y = Xhat.l[, 2],
                 colour = factor(z.l))) + 
  coord_fixed(.5) +
  theme_bw() + 
  theme(legend.position = 'none') +
  scale_colour_brewer(palette = 'Set1') + 
  labs(x = NULL, y = NULL, colour = NULL) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

Xhat.r <- embedding(A.r, 2, 1)
ase.r <- ggplot() + 
  geom_point(aes(x = Xhat.r[, 1], y = Xhat.r[, 2],
                 colour = factor(z.r))) + 
  coord_fixed(.5) +
  theme_bw() + 
  theme(legend.position = 'none') +
  scale_colour_brewer(palette = 'Set1') + 
  labs(x = NULL, y = NULL, colour = NULL) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

gridExtra::grid.arrange(graph.l, graph.r, ase.l, ase.r, nrow = 2)
```

## Manifold Block Model

Let $p, q \geq 0$, $d = p + q \geq 1$, $1 \leq r < d$, $K \geq 2$, and $n > K$ be integers. 
Define manifolds $\mathcal{M}_1, ..., \mathcal{M}_K \in \mathcal{X}$ for $\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p,q} y \in [0, 1]\}$ each by continuous function $g_k : [0, 1]^r \to \mathcal{X}$. 
Define probability distribution $F$ with support $[0, 1]^r$. 
Then the following mixture model is a *manifold block model*:

1. Draw labels $z_1, ..., z_n \iid \Categorical(\alpha_1, ..., \alpha_K)$.
2. Draw latent vectors by first taking $t_1,..., t_n \iid F$ and then computing each $x_i = g_{z_i}(t_i)$. 
3. Compile the latent vectors into data matrix $X = [ x_1 \mid \cdots \mid x_n ]^\top$ and define the adjacency matrix as $A \sim \GRDPG_{p,q}(X)$.

# $K$-Curves Clustering

1. Compute $X$, the ASE of $A$ using the $p$ most positive and $q$ most negative eigenvalues and their corresponding eigenvectors.
2. Initialize community labels $z_1, ..., z_n$.
3. While change in $\sum_k \sum_{i \in C_k} \|x_i - g_k(t_i)\|^2$ is less than $\epsilon$: 
    i. For $k = 1, ..., K$:
        a. Define $X_k$ as the rows of $X$ for which $z_i = k$.
        b. Fit curve $g_k$ and positions $t_{k_i}$ to $X_k$ by minimizing $\sum_{k_i} \|x_{k_i} - g_k(t_{k_i})\|^2$.
    ii. For $k = 1, ..., K$: 
        a. Assign $z_i \leftarrow \arg\min_\ell \|x_i - g_\ell(t_i)\|^2$.

```{r mbconnectome-kcurves, fig.height = 2, fig.width = 8, cache = TRUE, out.width = '100%'}
set.seed(123456)

z.init.r <- sample.points(z.r, 6)
z.init.l <- sample.points(z.l, 6)

mbconnectome.out.r <- manifold.clustering(Xhat.r, 
                                          length(unique(z.r)), 
                                          initialization = z.init.r,
                                          intercept = FALSE, 
                                          parallel = TRUE, 
                                          verbose = FALSE, 
                                          animate = FALSE, 
                                          curve.init = 'x')
fitted.r <- plot.estimated.curves(Xhat.r, mbconnectome.out.r) + 
  scale_colour_brewer(palette = 'Set1') + 
  coord_fixed(.5) +
  labs(x = NULL, y = NULL) + 
  ylim(min(Xhat.r[, 2]), max(Xhat.r[, 2])) + 
  xlim(min(Xhat.r[, 1]), max(Xhat.r[, 1]))

mbconnectome.out.l <- manifold.clustering(Xhat.l, 
                                          length(unique(z.l)), 
                                          initialization = z.init.l,
                                          intercept = FALSE, 
                                          parallel = TRUE, 
                                          verbose = FALSE, 
                                          animate = FALSE, 
                                          curve.init = 'x')
fitted.l <- plot.estimated.curves(Xhat.l, mbconnectome.out.l) + 
  scale_colour_brewer(palette = 'Set1') + 
  coord_fixed(.5) +
  labs(x = NULL, y = NULL) + 
  ylim(min(Xhat.l[, 2]), max(Xhat.l[, 2])) + 
  xlim(min(Xhat.l[, 1]), max(Xhat.l[, 1]))

gridExtra::grid.arrange(fitted.l, fitted.r, nrow = 1)
```

# Conclusion

Block models can be expressed as GRDPGs in which the communities are linear structures in the latent space. 
We propose the manifold block model to characterize nonlinear latent structures and the $K$-curves clustering algorithm to estimate these structures for community detection.

### 2023-05-12 sdss-2023/slides.Rmd
sdss slides and poster

---
output: 
  beamer_presentation:
    keep_tex: true
    fig_crop: no
    theme: 'default'
    colortheme: 'beaver'
    includes:
      in_header: 
        - page_headers.tex
header-includes:
- \usepackage{setspace}
- \usepackage{float}
- \usepackage{mathtools}
- \usepackage{natbib}
- \usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
- \setcitestyle{numbers,square,comma}
- \usepackage{verbatim}
- \usepackage{amsthm}
- \usepackage{comment}
- \usepackage{graphicx}
- \setbeamertemplate{itemize items}[circle]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      comment = NA, 
                      warning = FALSE, 
                      message = FALSE,
                      fig.align = 'center',
                      fig.lp = '')
options(xtable.comment = FALSE, 
        xtable.table.placement = 'H')
library(ggplot2)
library(ggraph)
import::from(magrittr, `%>%`)
theme_set(theme_bw())

source('~/dev/pabm-grdpg/functions.R')
source('~/dev/manifold-block-models/functions.R')
set.seed(314159)
```

## {.plain}

\center

\LARGE

\textcolor{darkred}{Manifold Clustering in the Setting of Generalized Random Dot Product Graphs}

\normalsize

SDSS Lightning Presentation

May 2023

::: columns

:::: {.column width='33%'}

```{r, out.width = '67px'}
knitr::include_graphics('john-koo.jpeg')
```

John Koo,  
Postdoctoral Fellow,  
Indiana University

::::

:::: {.column width='33%'}

```{r, out.width = '67px'}
knitr::include_graphics('minh-tang.jpg')
```

Minh Tang,  
Assistant Professor,  
NC State University

::::

:::: {.column width='33%'}

```{r, out.width = '67px'}
knitr::include_graphics('michael-trosset.jpg')
```

Michael W. Trosset,  
Professor of Statistics,  
Indiana University

::::

:::

## Community Detection for Networks

\newcommand{\diag}{\text{diag}}
\newcommand{\tr}{\text{Tr}}
\newcommand{\blockdiag}{\text{blockdiag}}
\newcommand{\indep}{\stackrel{\text{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\text{iid}}{\sim}}
\newcommand{\Bernoulli}{\text{Bernoulli}}
\newcommand{\Betadist}{\text{Beta}}
\newcommand{\BG}{\text{BernoulliGraph}}
\newcommand{\Categorical}{\text{Categorical}}
\newcommand{\Uniform}{\text{Uniform}}
\newcommand{\RDPG}{\text{RDPG}}
\newcommand{\GRDPG}{\text{GRDPG}}
\newcommand{\PABM}{\text{PABM}}

```{r out.width = '50%', fig.height = 3, fig.width = 4}
n <- 64
Pz <- generate.P.beta(n, 4, 4, 1, 1, 4)
P <- Pz$P
z <- Pz$clustering
A <- draw.graph(P)
qgraph::qgraph(A, groups = factor(z), legend = FALSE)
```

\begin{center}

How might we cluster the nodes of a network?

\end{center}

## Connecting Block Models to the GRDPG

::: columns

:::: {.column width='67%'}

```{r, cache = TRUE, fig.width = 6, fig.height = 2, out.width = '100%'}
n <- 64
n1 <- n / 2
n2 <- n / 2
z <- c(rep(1, n1), rep(2, n2))
p <- 1/2
q <- 1/2
r <- 1/8
P <- matrix(r, nrow = n, ncol = n)
P[seq(n1), seq(n1)] <- p
P[seq(n1 + 1, n), seq(n1 + 1, n)] <- q
X <- embedding(P, 2, 0)
A <- draw.graph(P)
Xhat <- embedding(A, 2, 0)
A.graph <- tidygraph::as_tbl_graph(A, directed = FALSE)

sbm.graph.plot <- ggraph(A, layout = 'stress') + 
  geom_edge_link(alpha = 1 / sqrt(n) / 2) + 
  geom_node_point(aes(colour = factor(z)),
                  show.legend = FALSE) + 
  labs(colour = NULL, 
       title = 'SBM') + 
  # theme_graph() + 
  theme_void() + 
  scale_colour_brewer(palette = 'Set1')

sbm.ase.plot <- ggplot() + 
  geom_point(aes(x = Xhat[, 2], y = Xhat[, 1],
                 colour = factor(z)),
             shape = 4) + 
  geom_point(aes(x = X[, 2], y = X[, 1], 
                 colour = factor(z))) + 
  # coord_fixed() +
  theme_bw() + 
  theme(legend.position = 'none') + 
  labs(x = NULL, y = NULL, colour = NULL, 
       title = 'Point Masses') + 
  scale_colour_brewer(palette = 'Set1')

gridExtra::grid.arrange(sbm.graph.plot, sbm.ase.plot, nrow = 1)
```

```{r, cache = TRUE, fig.width = 6, fig.height = 2, out.width = '100%'}
n <- 128
n1 <- n / 2
n2 <- n / 2
z <- c(rep(1, n1), rep(2, n2))
p <- 1/2
q <- 1/2
r <- 1/8
P <- matrix(r, nrow = n, ncol = n)
P[seq(n1), seq(n1)] <- p
P[seq(n1 + 1, n), seq(n1 + 1, n)] <- q
omega <- rbeta(n, 2, 1)
P <- P * (omega %*% t(omega))
X <- embedding(P, 2, 0)
A <- draw.graph(P)
Xhat <- embedding(A, 2, 0)
# Xhat <- -Xhat
A.graph <- tidygraph::as_tbl_graph(A, directed = FALSE)

dcbm.graph.plot <- ggraph(A, layout = 'stress') + 
  geom_edge_link(alpha = 1 / sqrt(n) / 2) + 
  geom_node_point(aes(colour = factor(z)),
                  show.legend = FALSE) + 
  labs(colour = NULL, 
       title = 'DCBM') + 
  # theme_graph() + 
  theme_void() + 
  scale_colour_brewer(palette = 'Set1')

dcbm.ase.plot <- ggplot() + 
  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2],
                 colour = factor(z)),
             shape = 4) + 
  geom_point(aes(x = X[, 1], y = X[, 2], 
                 colour = factor(z))) + 
  # coord_fixed() +
  theme_bw() + 
  theme(legend.position = 'none') + 
  labs(x = NULL, y = NULL, colour = NULL, 
       title = 'Rays') + 
  scale_colour_brewer(palette = 'Set1')

gridExtra::grid.arrange(dcbm.graph.plot, dcbm.ase.plot, nrow = 1)
```

```{r, cache = TRUE, fig.width = 6, fig.height = 2, out.width = '100%'}
n <- 64
Pz <- generate.P.beta(n)
P <- Pz$P
z <- Pz$clustering
X <- embedding(P, 3, 1)
A <- draw.graph(P)
Xhat <- embedding(A, 3, 1)
Xhat[, 1] <- -Xhat[, 1]
A.graph <- tidygraph::as_tbl_graph(A, directed = FALSE)

pabm.graph.plot <- ggraph(A, layout = 'stress') + 
  geom_edge_link(alpha = 1 / sqrt(n) / 2) + 
  geom_node_point(aes(colour = factor(z)),
                  show.legend = FALSE) + 
  labs(colour = NULL, 
       title = 'PABM') + 
  # theme_graph() + 
  theme_void() + 
  scale_colour_brewer(palette = 'Set1')

pabm.ase.plot <- ggplot() + 
  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2],
                 colour = factor(z)),
             shape = 4) + 
  geom_point(aes(x = X[, 1], y = X[, 2], 
                 colour = factor(z))) + 
  # coord_fixed() +
  theme_bw() + 
  theme(legend.position = 'none') + 
  labs(x = NULL, y = NULL, colour = NULL, 
       title = 'Subspaces (Projected)') + 
  scale_colour_brewer(palette = 'Set1')

gridExtra::grid.arrange(pabm.graph.plot, pabm.ase.plot, nrow = 1)
```

::::

:::: {.column width='33%'}

&nbsp;

&nbsp;

* K-means clustering
* Gaussian mixture models

&nbsp;

* K-means with cosine similarity
* GMM on angles

&nbsp;

&nbsp;

&nbsp;

* Orthogonal Spectral Clustering
* Sparse Subspace Clustering

::::

:::

## Nonlinear Communty Structure

```{r mbconnectome-graph, fig.height = 3, figh.width = 6, cache = TRUE, out.width = '67%'}
library(mbstructure)
data(MBconnectome)

graph.r <- generate.graph(newrdat, vdf.right)
A.r <- as.matrix(igraph::as_adjacency_matrix(graph.r$g, type = 'both'))
z.r <- as.numeric(graph.r$vdf$type)
n.r <- length(z.r)
A.r <- sign(A.r + t(A.r))
diag(A.r) <- 0
graph.r <- ggraph(A.r, layout = 'stress') + 
  geom_edge_link(alpha = 1 / sqrt(n.r) / 2) + 
  geom_node_point(aes(colour = factor(z.r)), 
                  show.legend = FALSE) + 
  labs(colour = NULL) + 
  theme_void() + 
  scale_colour_brewer(palette = 'Set1')

graph.l <- generate.graph(newldat, vdf.left)
A.l <- as.matrix(igraph::as_adjacency_matrix(graph.l$g, type = 'both'))
z.l <- as.numeric(graph.l$vdf$type)
n.l <- length(z.l)
A.l <- sign(A.l + t(A.l))
diag(A.l) <- 0
graph.l <- ggraph(A.l, layout = 'stress') + 
  geom_edge_link(alpha = 1 / sqrt(n.l) / 2) + 
  geom_node_point(aes(colour = factor(z.l)), 
                  show.legend = FALSE) + 
  labs(colour = NULL) + 
  theme_void() + 
  scale_colour_brewer(palette = 'Set1')

# gridExtra::grid.arrange(graph.l, graph.r, nrow = 1)
```

```{r mbconnectome-ase, fig.height = 6, fig.width = 8, out.width = '100%', cache = TRUE}
Xhat.l <- embedding(A.l, 2, 1)
ase.l <- ggplot() + 
  geom_point(aes(x = Xhat.l[, 1], y = Xhat.l[, 2],
                 colour = factor(z.l))) + 
  coord_fixed(.5) +
  theme_bw() + 
  theme(legend.position = 'none') +
  scale_colour_brewer(palette = 'Set1') + 
  labs(x = NULL, y = NULL, colour = NULL)

Xhat.r <- embedding(A.r, 2, 1)
ase.r <- ggplot() + 
  geom_point(aes(x = Xhat.r[, 1], y = Xhat.r[, 2],
                 colour = factor(z.r))) + 
  coord_fixed(.5) +
  theme_bw() + 
  theme(legend.position = 'none') +
  scale_colour_brewer(palette = 'Set1') + 
  labs(x = NULL, y = NULL, colour = NULL)

gridExtra::grid.arrange(graph.l, graph.r, ase.l, ase.r, nrow = 2)
```

## Manifold Block Model

Let $p, q \geq 0$, $d = p + q \geq 1$, $1 \leq r < d$, $K \geq 2$, and $n > K$ be integers. 
Define manifolds $\mathcal{M}_1, ..., \mathcal{M}_K \in \mathcal{X}$ for $\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p,q} y \in [0, 1]\}$ each by continuous function $g_k : [0, 1]^r \to \mathcal{X}$. 
Define probability distribution $F$ with support $[0, 1]^r$. 
Then the following mixture model is a *manifold block model*:

\begin{enumerate}
\item Draw labels $z_1, ..., z_n \iid \Categorical(\alpha_1, ..., \alpha_K)$.
\item Draw latent vectors by first taking $t_1,..., t_n \iid F$ and then computing each $x_i = g_{z_i}(t_i)$. 
\item Compile the latent vectors into data matrix $X = [ x_1 \mid \cdots \mid x_n ]^\top$ and define the adjacency matrix as $A \sim \GRDPG_{p,q}(X)$. 
\end{enumerate}

## Manifold Block Model

1. $z_1, ..., z_n \iid \Categorical(1/2, 1/2)$
2. $t_1, ..., t_n \iid \Uniform(0, 1)$
3. $x_i = g_{z_i}(t_i)$ 
    * $g_1(t) = [t^2, 2 t (1-t)]^\top$
    * $g_2(t) = [2 t (1-t), (1-t)^2]^\top$
4. $A \sim \GRDPG_{2, 0}(X)$

```{r intersect-curves-example, fig.height = 2, fig.width = 8, fig.cap = 'Latent vectors on intersecting curves (left), along with an RDPG drawn from this configuration (center) and its ASE (right).', cache = TRUE, out.width='100%'}
f1 <- function(t) {
  x1 <- t ^ 2
  x2 <- 2 * t - 2 * t ^ 2
  return(cbind(x1, x2))
}

f2 <- function(t) {
  x1 <- 2 * t - 2 * t ^ 2
  x2 <- 1 - 2 * t + t ^ 2
  return(cbind(x1, x2))
}

set.seed(314159)

n1 <- 2 ** 7
n2 <- n1
n <- n1 + n2
z <- c(rep(1, n1), rep(2, n2))

a <- 1
b <- 1
t1 <- rbeta(n1, a, b)
t2 <- rbeta(n2, a, b)
T <- rbind(t1, t2)

X1 <- f1(t1)
X2 <- f2(t2)
X <- rbind(X1, X2)

latent.vecs <- ggplot() + 
  geom_point(aes(x = X[, 1], y = X[, 2], colour = factor(z)), size = .5) + 
  # coord_fixed() + 
  labs(x = expression(x[1]), y = expression(x[2]), colour = NULL) + 
  theme_bw() + 
  theme(legend.position = 'none') + 
  scale_colour_brewer(palette = 'Set1')

P <- X %*% t(X)
diag(P) <- 0
A <- draw.graph(P)

rdpg <- ggraph(A, layout = 'stress') + 
  geom_edge_link(alpha = 1 / sqrt(n) / 2) + 
  geom_node_point(aes(colour = factor(z)), 
                  show.legend = FALSE) + 
  labs(colour = NULL) + 
  theme_void() + 
  scale_colour_brewer(palette = 'Set1')

Xhat <- embedding(A, 2, 0)

ase <- ggplot() + 
  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2], colour = factor(z)), size = .5) + 
  # coord_fixed() + 
  labs(x = expression(x[1]), y = expression(x[2]), colour = NULL) + 
  theme_bw() + 
  theme(legend.position = 'none') + 
  scale_colour_brewer(palette = 'Set1')

gridExtra::grid.arrange(latent.vecs, rdpg, ase, nrow = 1)
```

## $K$-Curves Clustering

\begin{algorithm}[H]
\label{alg:kcurves}
\scriptsize
\DontPrintSemicolon
\SetAlgoLined
\KwData{Adjacency matrix $A$, number of communities $K$, embedding dimensions $p$, $q$, stopping criterion $\epsilon$}
\KwResult{Community assignments $1, ..., K$, curves $g_1, ..., g_K$}
Compute $X$, the ASE of $A$ using the $p$ most positive and $q$ most negative eigenvalues and their corresponding eigenvectors.\;
Initialize community labels $z_1, ..., z_n$.\;
\Repeat {the change in $\sum_k \sum_{i \in C_k} \|x_i - g_k(t_i)\|^2$ is less than $\epsilon$} {
\For {$k = 1, ..., K$} {
Define $X_k$ as the rows of $X$ for which $z_i = k$.\;
Fit curve $g_k$ and positions $t_{k_i}$ to $X_k$ by minimizing $\sum_{k_i} \|x_{k_i} - g_k(t_{k_i})\|^2$.\;
}
\For {$k = 1, ..., K$} {
Assign $z_i \leftarrow \arg\min_\ell \|x_i - g_\ell(t_i)\|^2$.\;
}
}
\caption{$K$-curves clustering.}
\end{algorithm}

## Example

```{r mbconnectome-ase-2, fig.height = 2, fig.width = 8, out.width = '100%', cache = TRUE}
Xhat.l <- embedding(A.l, 2, 1)
ase.l <- ggplot() + 
  geom_point(aes(x = Xhat.l[, 1], y = Xhat.l[, 2],
                 colour = factor(z.l))) + 
  coord_fixed(.5) +
  theme_bw() + 
  theme(legend.position = 'none') +
  scale_colour_brewer(palette = 'Set1') + 
  labs(x = NULL, y = NULL, colour = NULL)

Xhat.r <- embedding(A.r, 2, 1)
ase.r <- ggplot() + 
  geom_point(aes(x = Xhat.r[, 1], y = Xhat.r[, 2],
                 colour = factor(z.r))) + 
  coord_fixed(.5) +
  theme_bw() + 
  theme(legend.position = 'none') +
  scale_colour_brewer(palette = 'Set1') + 
  labs(x = NULL, y = NULL, colour = NULL)

gridExtra::grid.arrange(ase.l, ase.r, nrow = 1)
```

```{r mbconnectome-kcurves, fig.height = 2, fig.width = 8, cache = TRUE, out.width = '100%'}
set.seed(123456)

z.init.r <- sample.points(z.r, 6)
z.init.l <- sample.points(z.l, 6)

mbconnectome.out.r <- manifold.clustering(Xhat.r, 
                                          length(unique(z.r)), 
                                          initialization = z.init.r,
                                          intercept = FALSE, 
                                          parallel = TRUE, 
                                          verbose = FALSE, 
                                          animate = FALSE, 
                                          curve.init = 'x')
fitted.r <- plot.estimated.curves(Xhat.r, mbconnectome.out.r) + 
  scale_colour_brewer(palette = 'Set1') + 
  coord_fixed(.5) +
  labs(x = NULL, y = NULL) + 
  ylim(min(Xhat.r[, 2]), max(Xhat.r[, 2])) + 
  xlim(min(Xhat.r[, 1]), max(Xhat.r[, 1]))

mbconnectome.out.l <- manifold.clustering(Xhat.l, 
                                          length(unique(z.l)), 
                                          initialization = z.init.l,
                                          intercept = FALSE, 
                                          parallel = TRUE, 
                                          verbose = FALSE, 
                                          animate = FALSE, 
                                          curve.init = 'x')
fitted.l <- plot.estimated.curves(Xhat.l, mbconnectome.out.l) + 
  scale_colour_brewer(palette = 'Set1') + 
  coord_fixed(.5) +
  labs(x = NULL, y = NULL) + 
  ylim(min(Xhat.l[, 2]), max(Xhat.l[, 2])) + 
  xlim(min(Xhat.l[, 1]), max(Xhat.l[, 1]))

gridExtra::grid.arrange(fitted.l, fitted.r, nrow = 1)
```

### 2023-05-12 sdss-2023/slides.tex
sdss slides and poster

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  ignorenonframetext,
]{beamer}
\usepackage{pgfpages}
\setbeamertemplate{caption}[numbered]
\setbeamertemplate{caption label separator}{: }
\setbeamercolor{caption name}{fg=normal text.fg}
\beamertemplatenavigationsymbolsempty
% Prevent slide breaks in the middle of a paragraph
\widowpenalties 1 10000
\raggedbottom
\setbeamertemplate{part page}{
  \centering
  \begin{beamercolorbox}[sep=16pt,center]{part title}
    \usebeamerfont{part title}\insertpart\par
  \end{beamercolorbox}
}
\setbeamertemplate{section page}{
  \centering
  \begin{beamercolorbox}[sep=12pt,center]{part title}
    \usebeamerfont{section title}\insertsection\par
  \end{beamercolorbox}
}
\setbeamertemplate{subsection page}{
  \centering
  \begin{beamercolorbox}[sep=8pt,center]{part title}
    \usebeamerfont{subsection title}\insertsubsection\par
  \end{beamercolorbox}
}
\AtBeginPart{
  \frame{\partpage}
}
\AtBeginSection{
  \ifbibliography
  \else
    \frame{\sectionpage}
  \fi
}
\AtBeginSubsection{
  \frame{\subsectionpage}
}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usecolortheme{beaver}
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\newif\ifbibliography
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\iffalse
https://tex.stackexchange.com/questions/392324/how-to-exclude-total-slide-number-from-beamer-slide
\fi
\setbeamertemplate{navigation symbols}{} 
\setbeamertemplate{footline}{\quad\hfill\insertframenumber\strut\quad}
\usefonttheme[onlymath]{serif}
\setbeamertemplate{itemize items}[circle]
\author[John Koo]{John Koo}
\usepackage{setspace}
\usepackage{float}
\usepackage{mathtools}
\usepackage{natbib}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\setcitestyle{numbers,square,comma}
\usepackage{verbatim}
\usepackage{amsthm}
\usepackage{comment}
\usepackage{graphicx}
\setbeamertemplate{itemize items}[circle]
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\begin{frame}[plain]{}
\protect\hypertarget{section}{}
\center

\LARGE

\textcolor{darkred}{Manifold Clustering in the Setting of Generalized Random Dot Product Graphs}

\normalsize

SDSS Lightning Presentation

May 2023

\begin{columns}[T]
\begin{column}{0.33\textwidth}
\begin{center}\includegraphics[width=67px]{john-koo} \end{center}

John Koo,\\
Postdoctoral Fellow,\\
Indiana University
\end{column}

\begin{column}{0.33\textwidth}
\begin{center}\includegraphics[width=67px]{minh-tang} \end{center}

Minh Tang,\\
Assistant Professor,\\
NC State University
\end{column}

\begin{column}{0.33\textwidth}
\begin{center}\includegraphics[width=67px]{michael-trosset} \end{center}

Michael W. Trosset,\\
Professor of Statistics,\\
Indiana University
\end{column}
\end{columns}
\end{frame}

\begin{frame}{Community Detection for Networks}
\protect\hypertarget{community-detection-for-networks}{}
\newcommand{\diag}{\text{diag}}
\newcommand{\tr}{\text{Tr}}
\newcommand{\blockdiag}{\text{blockdiag}}
\newcommand{\indep}{\stackrel{\text{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\text{iid}}{\sim}}
\newcommand{\Bernoulli}{\text{Bernoulli}}
\newcommand{\Betadist}{\text{Beta}}
\newcommand{\BG}{\text{BernoulliGraph}}
\newcommand{\Categorical}{\text{Categorical}}
\newcommand{\Uniform}{\text{Uniform}}
\newcommand{\RDPG}{\text{RDPG}}
\newcommand{\GRDPG}{\text{GRDPG}}
\newcommand{\PABM}{\text{PABM}}

\begin{center}\includegraphics[width=0.5\linewidth]{slides_files/figure-beamer/unnamed-chunk-4-1} \end{center}

\begin{center}

How might we cluster the nodes of a network?

\end{center}
\end{frame}

\begin{frame}{Connecting Block Models to the GRDPG}
\protect\hypertarget{connecting-block-models-to-the-grdpg}{}
\begin{columns}[T]
\begin{column}{0.67\textwidth}
\begin{center}\includegraphics[width=1\linewidth]{slides_files/figure-beamer/unnamed-chunk-5-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{slides_files/figure-beamer/unnamed-chunk-6-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{slides_files/figure-beamer/unnamed-chunk-7-1} \end{center}
\end{column}

\begin{column}{0.33\textwidth}
~

~

\begin{itemize}
\tightlist
\item
  K-means clustering
\item
  Gaussian mixture models
\end{itemize}

~

\begin{itemize}
\tightlist
\item
  K-means with cosine similarity
\item
  GMM on angles
\end{itemize}

~

~

~

\begin{itemize}
\tightlist
\item
  Orthogonal Spectral Clustering
\item
  Sparse Subspace Clustering
\end{itemize}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{Nonlinear Communty Structure}
\protect\hypertarget{nonlinear-communty-structure}{}
\begin{center}\includegraphics[width=1\linewidth]{slides_files/figure-beamer/mbconnectome-ase-1} \end{center}
\end{frame}

\begin{frame}{Manifold Block Model}
\protect\hypertarget{manifold-block-model}{}
Let \(p, q \geq 0\), \(d = p + q \geq 1\), \(1 \leq r < d\),
\(K \geq 2\), and \(n > K\) be integers. Define manifolds
\(\mathcal{M}_1, ..., \mathcal{M}_K \in \mathcal{X}\) for
\(\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p,q} y \in [0, 1]\}\)
each by continuous function \(g_k : [0, 1]^r \to \mathcal{X}\). Define
probability distribution \(F\) with support \([0, 1]^r\). Then the
following mixture model is a \emph{manifold block model}:

\begin{enumerate}
\item Draw labels $z_1, ..., z_n \stackrel{\text{iid}}{\sim}\text{Categorical}(\alpha_1, ..., \alpha_K)$.
\item Draw latent vectors by first taking $t_1,..., t_n \stackrel{\text{iid}}{\sim}F$ and then computing each $x_i = g_{z_i}(t_i)$. 
\item Compile the latent vectors into data matrix $X = [ x_1 \mid \cdots \mid x_n ]^\top$ and define the adjacency matrix as $A \sim \text{GRDPG}_{p,q}(X)$. 
\end{enumerate}
\end{frame}

\begin{frame}{Manifold Block Model}
\protect\hypertarget{manifold-block-model-1}{}
\begin{enumerate}
\tightlist
\item
  \(z_1, ..., z_n \stackrel{\text{iid}}{\sim}\text{Categorical}(1/2, 1/2)\)
\item
  \(t_1, ..., t_n \stackrel{\text{iid}}{\sim}\text{Uniform}(0, 1)\)
\item
  \(x_i = g_{z_i}(t_i)\)

  \begin{itemize}
  \tightlist
  \item
    \(g_1(t) = [t^2, 2 t (1-t)]^\top\)
  \item
    \(g_2(t) = [2 t (1-t), (1-t)^2]^\top\)
  \end{itemize}
\item
  \(A \sim \text{GRDPG}_{2, 0}(X)\)
\end{enumerate}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{slides_files/figure-beamer/intersect-curves-example-1}

}

\caption{Latent vectors on intersecting curves (left), along with an RDPG drawn from this configuration (center) and its ASE (right).}\label{intersect-curves-example}
\end{figure}
\end{frame}

\begin{frame}{\(K\)-Curves Clustering}
\protect\hypertarget{k-curves-clustering}{}
\begin{algorithm}[H]
\label{alg:kcurves}
\scriptsize
\DontPrintSemicolon
\SetAlgoLined
\KwData{Adjacency matrix $A$, number of communities $K$, embedding dimensions $p$, $q$, stopping criterion $\epsilon$}
\KwResult{Community assignments $1, ..., K$, curves $g_1, ..., g_K$}
Compute $X$, the ASE of $A$ using the $p$ most positive and $q$ most negative eigenvalues and their corresponding eigenvectors.\;
Initialize community labels $z_1, ..., z_n$.\;
\Repeat {the change in $\sum_k \sum_{i \in C_k} \|x_i - g_k(t_i)\|^2$ is less than $\epsilon$} {
\For {$k = 1, ..., K$} {
Define $X_k$ as the rows of $X$ for which $z_i = k$.\;
Fit curve $g_k$ and positions $t_{k_i}$ to $X_k$ by minimizing $\sum_{k_i} \|x_{k_i} - g_k(t_{k_i})\|^2$.\;
}
\For {$k = 1, ..., K$} {
Assign $z_i \leftarrow \arg\min_\ell \|x_i - g_\ell(t_i)\|^2$.\;
}
}
\caption{$K$-curves clustering.}
\end{algorithm}
\end{frame}

\begin{frame}{Example}
\protect\hypertarget{example}{}
\begin{center}\includegraphics[width=1\linewidth]{slides_files/figure-beamer/mbconnectome-ase-2-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{slides_files/figure-beamer/mbconnectome-kcurves-1} \end{center}
\end{frame}

\end{document}

### 2023-05-12 sdss-2023/poster.Rmd
minor adjustments to poster

@@
 
-# Block Models as GRDPGs
+# Block Models as Linear GRDPGs
 
-It has been previously shown that the SBM, DCBM, and PABM are GRDPGs in which the communities lie on point masses, line segments, and subspaces, respectively.
+It has been previously shown that the SBM, DCBM, and PABM are GRDPGs in which the communities correspond to linear structures in the latent space.
 
@@
                  colour = factor(z.l))) + 
-  coord_fixed(.5) +
+  # coord_fixed(.5) +
   theme_bw() + 
@@
                  colour = factor(z.r))) + 
-  coord_fixed(.5) +
+  # coord_fixed(.5) +
   theme_bw() + 
@@
   scale_colour_brewer(palette = 'Set1') + 
-  coord_fixed(.5) +
+  coord_fixed(.4) +
   labs(x = NULL, y = NULL) + 
   ylim(min(Xhat.r[, 2]), max(Xhat.r[, 2])) + 
-  xlim(min(Xhat.r[, 1]), max(Xhat.r[, 1]))
+  xlim(min(Xhat.r[, 1]), max(Xhat.r[, 1])) +
+  theme(axis.title.x = element_blank(),
+        axis.text.x = element_blank(),
+        axis.ticks.x = element_blank(),
+        axis.title.y = element_blank(),
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank())
 
@@
   scale_colour_brewer(palette = 'Set1') + 
-  coord_fixed(.5) +
+  coord_fixed(.4) +
   labs(x = NULL, y = NULL) + 
   ylim(min(Xhat.l[, 2]), max(Xhat.l[, 2])) + 
-  xlim(min(Xhat.l[, 1]), max(Xhat.l[, 1]))
+  xlim(min(Xhat.l[, 1]), max(Xhat.l[, 1])) +
+  theme(axis.title.x = element_blank(),
+        axis.text.x = element_blank(),
+        axis.ticks.x = element_blank(),
+        axis.title.y = element_blank(),
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank())
 
@@
 Block models can be expressed as GRDPGs in which the communities are linear structures in the latent space. 
-We propose the manifold block model to characterize nonlinear latent structures and the $K$-curves clustering algorithm to estimate these structures for community detection.
+We propose the manifold block model to extend this to nonlinear latent structures and the $K$-curves clustering algorithm to estimate these structures for community detection.

### 2023-05-12 sdss-2023/poster.Rmd
minor adjustments to poster

@@
 
-It has been previously shown that the SBM, DCBM, and PABM are GRDPGs in which the communities correspond to linear structures in the latent space.
+SBM, DCBM, and PABM are GRDPGs in which the communities correspond to linear structures in the latent space.
 
-```{r, cache = TRUE, fig.width = 5, fig.height = 2.3, out.width = '100%'}
+```{r, cache = TRUE, fig.width = 5, fig.height = 2.4, out.width = '100%'}
 n <- 64
@@
 
-```{r, cache = TRUE, fig.width = 5, fig.height = 2.3, out.width = '100%'}
+```{r, cache = TRUE, fig.width = 5, fig.height = 2.4, out.width = '100%'}
 n <- 128
@@
 
-```{r, cache = TRUE, fig.width = 5, fig.height = 2.3, out.width = '100%'}
+```{r, cache = TRUE, fig.width = 5, fig.height = 2.4, out.width = '100%'}
 n <- 64

### 2023-05-22 sdss-2023/poster-2.Rmd
alternative poster

--- 2023-05-12 sdss-2023/poster.Rmd
@@
     address: North Carolina State University
-title_textsize: "56pt"
+title_textsize: "48pt"
 author_textsize: "32pt"
 affiliation_textsize: "8pt"
-sectitle_textsize: "36pt"
-sectitle2_textsize: "36pt"
+sectitle_textsize: "34pt"
+sectitle2_textsize: "34pt"
 body_textsize: "32pt"
@@
 poster_width: "36.5in"
-column_margins: "0.1in"
+column_margins: "0.01in"
 primary_colour: "#990000"
@@
 accent_colour: "#990000"
-titlebox_borderwidth: "0.1cm"
+titlebox_borderwidth: "0.01cm"
 titlebox_bordercol: "#000000"
@@
 sectitle_borderwidth: "0mm"
-font_family: "Times New Roman"
+font_family: "Latin Modern Math"
 logoleft_name: poster_files/trident_medium.png
@@
 source('~/dev/manifold-block-models/functions.R')
+
+doMC::registerDoMC(8)
 ```
@@
 
-# Generalized Random Dot Product Graphs
-
-The *generalized random dot product graph* is a random graph model in which each vertex $v_i$ has a corresponding hidden vector $x_i \in \mathbb{R}^{p+q}$ and each edge probability is the indefinite inner product of the corresponding pair of hidden vectors, i.e., $P_{ij} = x_i^\top I_{p, q} x_j$, $I_{p,q} = \Bigl[\begin{smallmatrix} I_{p} & 0 \\ 0 & - I_{q} \end{smallmatrix} \Bigr]$
-
-## Adjacency Spectral Embedding
-Approximate $A$ by spectral decomposition $A \approx V_{p, q} \Lambda_{p, q} V_{p, q}^\top$. 
-The subscript $p, q$ denotes the $p$ most positive and $q$ most negative eigenvalues and corresponding eigenvectors. 
-Each $\hat{x}_i$, the $i^{th}$ row of $\hat{X} = V_{p, q} |\Lambda_{p, q}|^{1/2},$ estimates the relative position of its corresponding latent vector $x_i$, up to an indefinite orthogonal transformation.
-
-**Theorem** (Rubin-Delanchy et al. 2022): $\max_i \| \hat{x}_i - Q_n x_i \| = O_P \Big( \frac{\log^c n}{n^{1/2}} \Big)$ for some $Q_n \in \mathbb{O}(p, q)$.
-
-```{r, cache = TRUE}
-n1 <- 2 ** 5
-n2 <- 2 ** 5
-n <- n1 + n2
-z <- c(rep(1, n1), rep(2, n2))
-u1 <- runif(n1)
-u2 <- runif(n2)
-x1 <- cos(pi / 3 * u1)
-y1 <- sin(pi / 3 * u1)
-x2 <- 1 + cos(pi / 3 * u2 + pi)
-y2 <- 1 + sin(pi / 3 * u2 + pi)
-
-data.matrix <- cbind(c(x1, x2), c(y1, y2))
-P <- data.matrix %*% diag(c(1, 1)) %*% t(data.matrix)
-A <- draw.graph(P)
-A.graph <- tidygraph::as_tbl_graph(A, directed = FALSE)
-
-Xhat <- embedding(A, 2, 0)
-Xhat[, 1] <- -Xhat[, 1]
-```
-
-```{r, fig.height = 2.25, fig.width = 6, out.width = '100%'}
-latent.config.plot <- ggplot() + 
-  geom_point(aes(x = data.matrix[, 1], y = data.matrix[, 2], 
-                 colour = factor(z))) + 
-  # coord_fixed() + 
-  theme_bw() + 
-  theme(legend.position = 'none') + 
-  labs(x = NULL, y = NULL, colour = NULL, title = 'Latent Vectors') + 
-  scale_colour_brewer(palette = 'Set1') +
-  theme(axis.title.x = element_blank(),
-        axis.text.x = element_blank(),
-        axis.ticks.x = element_blank(),
-        axis.title.y = element_blank(),
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank())
-
-graph.plot <- ggraph(A, layout = 'stress') + 
-  geom_edge_link(alpha = 1 / sqrt(n) / 2) + 
-  geom_node_point(aes(colour = factor(z)),
-                  show.legend = FALSE) + 
-  labs(colour = NULL, 
-       title = 'GRDPG') + 
-  # theme_graph() + 
-  theme_void() + 
-  scale_colour_brewer(palette = 'Set1')
-
-ase.plot <- ggplot() + 
-  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2],
-                 colour = factor(z))) + 
-  # coord_fixed() + 
-  theme_bw() + 
-  theme(legend.position = 'none') + 
-  labs(x = NULL, y = NULL, colour = NULL, title = 'ASE') + 
-  scale_colour_brewer(palette = 'Set1') +
-  theme(axis.title.x = element_blank(),
-        axis.text.x = element_blank(),
-        axis.ticks.x = element_blank(),
-        axis.title.y = element_blank(),
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank())
-
-gridExtra::grid.arrange(latent.config.plot, graph.plot, ase.plot, nrow = 1)
-```
-
 # Block Models as Linear GRDPGs
 
-SBM, DCBM, and PABM are GRDPGs in which the communities correspond to linear structures in the latent space.
+SBM, DCBM, and PABM are generalized random dot product graphs in which the communities correspond to linear structures in the latent space.
 
-```{r, cache = TRUE, fig.width = 5, fig.height = 2.4, out.width = '100%'}
+```{r, cache = TRUE, fig.width = 5, fig.height = 2.25, out.width = '100%'}
+set.seed(314159)
+
 n <- 64
@@
 Xhat <- embedding(A, 2, 0)
+Xhat <- -Xhat
 A.graph <- tidygraph::as_tbl_graph(A, directed = FALSE)
@@
 
-```{r, cache = TRUE, fig.width = 5, fig.height = 2.4, out.width = '100%'}
+```{r, cache = TRUE, fig.width = 5, fig.height = 2.25, out.width = '100%'}
+set.seed(314159)
+
 n <- 128
@@
 
-```{r, cache = TRUE, fig.width = 5, fig.height = 2.4, out.width = '100%'}
+```{r, cache = TRUE, fig.width = 5, fig.height = 2.25, out.width = '100%'}
+set.seed(314159)
+
 n <- 64
@@
 diag(A.r) <- 0
-graph.r <- ggraph(A.r, layout = 'stress') + 
+A.r.graph <- tidygraph::as_tbl_graph(unname(A.r), directed = FALSE)
+graph.r <- ggraph(A.r.graph, layout = 'stress') + 
   geom_edge_link(alpha = 1 / sqrt(n.r) / 2) + 
@@
 diag(A.l) <- 0
-graph.l <- ggraph(A.l, layout = 'stress') + 
+A.l.graph <- tidygraph::as_tbl_graph(unname(A.l), directed = FALSE)
+graph.l <- ggraph(A.l.graph, layout = 'stress') + 
   geom_edge_link(alpha = 1 / sqrt(n.l) / 2) + 
@@
 
-```{r mbconnectome-ase, fig.height = 3.75, fig.width = 7, out.width = '100%', cache = TRUE}
+```{r mbconnectome-ase, fig.height = 5, fig.width = 7, out.width = '75%', cache = TRUE}
 Xhat.l <- embedding(A.l, 2, 1)
@@
 
-Let $p, q \geq 0$, $d = p + q \geq 1$, $1 \leq r < d$, $K \geq 2$, and $n > K$ be integers. 
-Define manifolds $\mathcal{M}_1, ..., \mathcal{M}_K \in \mathcal{X}$ for $\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p,q} y \in [0, 1]\}$ each by continuous function $g_k : [0, 1]^r \to \mathcal{X}$.
+Let $p, q \geq 0$, $d = p + q \geq 1$, $1 \geq r < d$, $K \geq 2$, and $n > K$ be integers. 
+Let $\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p,q} y \in [0, 1]\}$ and define manifolds $\mathcal{M}_1, ..., \mathcal{M}_K \subset \mathcal{X}$ each by continuous function $g_k : [0, 1]^r \to \mathcal{X}$.
 Define probability distribution $F$ with support $[0, 1]^r$. 
@@
 2. Initialize community labels $z_1, ..., z_n$.
-3. While change in $\sum_k \sum_{i \in C_k} \|x_i - g_k(t_i)\|^2$ is less than $\epsilon$: 
+3. While change in the loss function $L = \sum_k \sum_{i \in C_k} \|x_i - g_k(t_i)\|^2$ is less than $\epsilon$: 
     i. For $k = 1, ..., K$:
         a. Define $X_k$ as the rows of $X$ for which $z_i = k$.
-        b. Fit curve $g_k$ and positions $t_{k_i}$ to $X_k$ by minimizing $\sum_{k_i} \|x_{k_i} - g_k(t_{k_i})\|^2$.
+        b. Fit curve $g_k$ and positions $t_{k_i}$ by minimizing $\sum_{k_i} \|x_{k_i} - g_k(t_{k_i})\|^2$.
     ii. For $k = 1, ..., K$: 
@@
 
-```{r mbconnectome-kcurves, fig.height = 2, fig.width = 8, cache = TRUE, out.width = '100%'}
+```{r mbconnectome-kcurves, fig.height = 2, fig.width = 6, cache = TRUE, out.width = '67%'}
 set.seed(123456)
@@
 
+**Theorem**. 
+Let an MBM be such that the manifolds are described as polynomial curves of order $R$. 
+Suppose that for each community $k$, we have labels for at least $R + 1$ vertices. 
+Then $K$-curves clustering outputs estimates such that  
+$$L(\hat{z}_1, ..., \hat{z}_n, \hat{g}_1, ..., \hat{g}_K; X) \stackrel{p}{\to} 0.$$
+
+# Simulation
+
+Latent vectors were drawn uniformly on three intersecting quadratic curves in $\mathbb{R}^3$ (left) to construct a GRDPG (middle). 
+Curves were then fitted to the ASE (right) and embedding vectors were assigned labels based on proximity to the curves.
+
+```{r three-curves, cache = TRUE, fig.height = 2, fig.width = 6, out.width = '100%'}
+set.seed(314159)
+par(mfrow = c(1, 3))
+
+n <- 512
+
+p.list <- list(matrix(c(0, 1, 0, 
+                        0, 0, 1,
+                        0, 0, 0),
+                      nrow = 3, ncol = 3),
+               matrix(c(0, 0, 0, 
+                        0, 0, 1,
+                        0, 1, 0),
+                      nrow = 3, ncol = 3),
+               matrix(c(0, 1, 0, 
+                        0, 0, 1,
+                        0, 1, 0),
+                      nrow = 3, ncol = 3))
+
+K <- length(p.list)
+
+z <- sample(seq(K), n, replace = TRUE)
+z <- sort(z)
+
+t <- runif(n)
+
+X <- lapply(seq(K), function(k) {
+  bezier.curve(t[z == k], 
+               p.list[[k]], 
+               intercept = TRUE)
+}) %>% 
+  do.call('rbind', .)
+
+p <- ncol(X)
+q <- 0
+
+P <- grdpg.edge.prob.matrix(X, p, q)
+A <- draw.graph(P)
+Xhat <- embedding(A, p, q)
+
+scatterplot3d::scatterplot3d(X[, 1], X[, 2], X[, 3],
+                             color = z + 1, 
+                             mar = c(0, 0, 0, 0),
+                             xlab = NA, ylab = NA, zlab = NA,
+                             # pch = '.', 
+                             asp = 1,
+                             tick.marks = FALSE)
+
+qgraph::qgraph(A, 
+               groups = factor(z), 
+               legend = FALSE,
+               mar = c(0, 0, 0, 0))
+
+scatterplot3d::scatterplot3d(Xhat[, 1], Xhat[, 2], Xhat[, 3], 
+                             color = z + 1, 
+                             mar = c(0, 0, 0, 0),
+                             xlab = NA, ylab = NA, zlab = NA, 
+                             # pch = '.', 
+                             asp = 1,
+                             # angle = 40,
+                             tick.marks = FALSE)
+```
+
+```{r sim-curves-3, fig.height = 3, fig.width = 8, out.width = '100%', cache = TRUE}
+n.vec <- 2 ^ c(7, 8, 9, 10, 11)
+nsamp.vec <- c(0, 4, 8)
+
+clustering.df <- readr::read_csv('~/dev/manifold-block-models/simulations/balanced-3/balanced-3.csv')
+
+clust.summary.df <- clustering.df %>% 
+  dplyr::group_by(n, nsamp) %>% 
+  dplyr::summarise(
+    med.count = median(error.count),
+    first.q.count = quantile(error.count, .25),
+    third.q.count = quantile(error.count, .75),
+    med.rate = median(error.rate),
+    first.q.rate = quantile(error.rate, .25),
+    third.q.rate = quantile(error.rate, .75)
+  ) %>% 
+  dplyr::ungroup()
+
+ggplot(clust.summary.df) + 
+  theme_bw() + 
+  theme(text = element_text(size = 10)) + 
+  scale_y_log10() +
+  scale_x_log10(breaks = n.vec) +
+  labs(y = 'community detection error rate',
+       x = 'n',
+       colour = 'known labels per community', 
+       shape = NULL) + 
+  geom_line(aes(x = n, y = med.rate, colour = factor(nsamp))) + 
+  geom_point(aes(x = n, y = med.rate, colour = factor(nsamp))) + 
+  geom_errorbar(aes(x = n, 
+                    ymin = first.q.rate, ymax = third.q.rate, 
+                    colour = factor(nsamp)),
+                width = .125) + 
+  theme(legend.position = 'bottom') + 
+  scale_colour_brewer(palette = 'Set1')
+```
+
 # Conclusion

### 2023-05-22 sdss-2023/poster.Rmd
alternative poster

@@
     address: North Carolina State University
-title_textsize: "56pt"
+title_textsize: "52pt"
 author_textsize: "32pt"
 affiliation_textsize: "8pt"
-sectitle_textsize: "36pt"
-sectitle2_textsize: "36pt"
+sectitle_textsize: "34pt"
+sectitle2_textsize: "34pt"
 body_textsize: "32pt"
@@
 poster_width: "36.5in"
-column_margins: "0.1in"
+column_margins: "0.05in"
 primary_colour: "#990000"
@@
 source('~/dev/manifold-block-models/functions.R')
+doMC::registerDoMC(8)
 ```
 
+\singlespacing
 \newcommand{\diag}{\text{diag}}

### 2023-05-24 sdss-2023/slides.Rmd
updated slides

@@
 
-gridExtra::grid.arrange(latent.vecs, rdpg, ase, nrow = 1)
+# gridExtra::grid.arrange(latent.vecs, rdpg, ase, nrow = 1)
+gridExtra::grid.arrange(latent.vecs, rdpg, nrow = 1)
 ```
@@
 ```
+
+## Asymptotic Results
+
+**Theorem**. 
+Let an MBM be such that the manifolds are described by functions $g_1(t), ..., g_K(t)$ which are polynomial curves of order $R$. 
+Define the loss of $K$-curves clustering as:
+
+$$L(\hat{z}_1, ..., \hat{z}_n, \hat{g}_1, ..., \hat{g}_K; A) = \sum_k \sum_{i : \hat{z}_i = k} \|\hat{x}_{i} - \hat{g}_k(t_{i})\|^2,$$
+
+where $\hat{x}_i$ are the embedding vectors of $A$. 
+Suppose that for each community $k$, we have labels for at least $R + 1$ vertices. 
+Then as $n \to \infty$, $K$-curves clustering outputs estimates such that
+
+$$L(\hat{z}_1, ..., \hat{z}_n, \hat{g}_1, ..., \hat{g}_K; A) \stackrel{p}{\to} 0.$$
+
+## Simulation
+
+```{r three-curves, cache = TRUE, fig.height = 2, fig.width = 6, out.width = '100%'}
+set.seed(314159)
+par(mfrow = c(1, 3))
+
+n <- 256
+
+p.list <- list(matrix(c(0, 1, 0, 
+                        0, 0, 1,
+                        0, 0, 0),
+                      nrow = 3, ncol = 3),
+               matrix(c(0, 0, 0, 
+                        0, 0, 1,
+                        0, 1, 0),
+                      nrow = 3, ncol = 3),
+               matrix(c(0, 1, 0, 
+                        0, 0, 1,
+                        0, 1, 0),
+                      nrow = 3, ncol = 3))
+
+K <- length(p.list)
+
+z <- sample(seq(K), n, replace = TRUE)
+z <- sort(z)
+
+t <- runif(n)
+
+X <- lapply(seq(K), function(k) {
+  bezier.curve(t[z == k], 
+               p.list[[k]], 
+               intercept = TRUE)
+}) %>% 
+  do.call('rbind', .)
+
+p <- ncol(X)
+q <- 0
+
+P <- grdpg.edge.prob.matrix(X, p, q)
+A <- draw.graph(P)
+Xhat <- embedding(A, p, q)
+
+degs <- colSums(A)
+A <- A[degs > 0, degs > 0]
+z <- z[degs > 0]
+X <- X[degs > 0, ]
+Xhat <- Xhat[degs > 0, ]
+
+scatterplot3d::scatterplot3d(X[, 1], X[, 2], X[, 3],
+                             color = z + 1, 
+                             mar = c(0, 0, 0, 0),
+                             xlab = NA, ylab = NA, zlab = NA,
+                             # pch = '.', 
+                             asp = 1,
+                             tick.marks = FALSE)
+
+qgraph::qgraph(A, 
+               groups = factor(z), 
+               borders = FALSE,
+               legend = FALSE,
+               esize = .001,
+               mar = c(0, 0, 0, 0))
+
+scatterplot3d::scatterplot3d(Xhat[, 1], Xhat[, 2], Xhat[, 3], 
+                             color = z + 1, 
+                             mar = c(0, 0, 0, 0),
+                             xlab = NA, ylab = NA, zlab = NA, 
+                             # pch = '.', 
+                             asp = 1,
+                             # angle = 40,
+                             tick.marks = FALSE)
+```
+
+```{r sim-curves-3, fig.height = 2.5, fig.width = 6, out.width = '100%', cache = TRUE}
+n.vec <- 2 ^ c(7, 8, 9, 10, 11)
+nsamp.vec <- c(0, 4, 8)
+
+clustering.df <- readr::read_csv('~/dev/manifold-block-models/simulations/balanced-3/balanced-3.csv')
+
+clust.summary.df <- clustering.df %>% 
+  dplyr::group_by(n, nsamp) %>% 
+  dplyr::summarise(
+    med.count = median(error.count),
+    first.q.count = quantile(error.count, .25),
+    third.q.count = quantile(error.count, .75),
+    med.rate = median(error.rate),
+    first.q.rate = quantile(error.rate, .25),
+    third.q.rate = quantile(error.rate, .75)
+  ) %>% 
+  dplyr::ungroup()
+
+ggplot(clust.summary.df) + 
+  theme_bw() + 
+  theme(text = element_text(size = 10)) + 
+  scale_y_log10() +
+  scale_x_log10(breaks = n.vec) +
+  labs(y = 'error rate',
+       x = 'n',
+       colour = 'known labels per community', 
+       shape = NULL) + 
+  geom_line(aes(x = n, y = med.rate, colour = factor(nsamp))) + 
+  geom_point(aes(x = n, y = med.rate, colour = factor(nsamp))) + 
+  geom_errorbar(aes(x = n, 
+                    ymin = first.q.rate, ymax = third.q.rate, 
+                    colour = factor(nsamp)),
+                width = .125) + 
+  theme(legend.position = 'bottom') + 
+  scale_colour_brewer(palette = 'Set1')
+```
+
+## Thank you!
+
+Code and drafts available at https://github.com/johneverettkoo/manifold-block-models

### 2023-05-24 sdss-2023/slides.tex
updated slides

@@
 
+\begin{frame}{Asymptotic Results}
+\protect\hypertarget{asymptotic-results}{}
+\textbf{Theorem}. Let an MBM be such that the manifolds are described by
+functions \(g_1(t), ..., g_K(t)\) which are polynomial curves of order
+\(R\). Define the loss of \(K\)-curves clustering as:
+
+\[L(\hat{z}_1, ..., \hat{z}_n, \hat{g}_1, ..., \hat{g}_K; A) = \sum_k \sum_{i : \hat{z}_i = k} \|\hat{x}_{i} - \hat{g}_k(t_{i})\|^2,\]
+
+where \(\hat{x}_i\) are the embedding vectors of \(A\). Suppose that for
+each community \(k\), we have labels for at least \(R + 1\) vertices.
+Then as \(n \to \infty\), \(K\)-curves clustering outputs estimates such
+that
+
+\[L(\hat{z}_1, ..., \hat{z}_n, \hat{g}_1, ..., \hat{g}_K; A) \stackrel{p}{\to} 0.\]
+\end{frame}
+
+\begin{frame}{Simulation}
+\protect\hypertarget{simulation}{}
+\begin{center}\includegraphics[width=1\linewidth]{slides_files/figure-beamer/three-curves-1} \end{center}
+
+\begin{center}\includegraphics[width=1\linewidth]{slides_files/figure-beamer/sim-curves-3-1} \end{center}
+\end{frame}
+
+\begin{frame}{Thank you!}
+\protect\hypertarget{thank-you}{}
+Code and drafts available at
+\url{https://github.com/johneverettkoo/manifold-block-models}
+\end{frame}
+
 \end{document}

### 2023-05-24 sdss-2023/slides.Rmd
minor changes
minor updates

@@
 
-# gridExtra::grid.arrange(latent.vecs, rdpg, ase, nrow = 1)
-gridExtra::grid.arrange(latent.vecs, rdpg, nrow = 1)
+gridExtra::grid.arrange(latent.vecs, rdpg, ase, nrow = 1)
+# gridExtra::grid.arrange(latent.vecs, rdpg, nrow = 1)
 ```
@@
 Suppose that for each community $k$, we have labels for at least $R + 1$ vertices. 
-Then as $n \to \infty$, $K$-curves clustering outputs estimates such that
+Then as $n \to \infty$, $K$-curves clustering outputs estimators such that
 

### 2023-05-24 sdss-2023/slides.Rmd
minor updates

@@
 
-```{r three-curves, cache = TRUE, fig.height = 2, fig.width = 6, out.width = '100%'}
+```{r three-curves, cache = TRUE, fig.height = 4, fig.width = 12, out.width = '100%'}
 set.seed(314159)
@@
 
-n <- 256
+n <- 512
 
@@
 degs <- colSums(A)
-A <- A[degs > 0, degs > 0]
-z <- z[degs > 0]
-X <- X[degs > 0, ]
-Xhat <- Xhat[degs > 0, ]
+A.trunc <- A[degs > 0, degs > 0]
+z.trunc <- z[degs > 0]
+# X <- X[degs > 0, ]
+# Xhat <- Xhat[degs > 0, ]
 
@@
 
-qgraph::qgraph(A, 
-               groups = factor(z), 
+qgraph::qgraph(A.trunc, 
+               groups = factor(z.trunc), 
                borders = FALSE,
@@
                     colour = factor(nsamp)),
-                width = .125) + 
+                width = .05) + 
   theme(legend.position = 'bottom') + 

### 2023-05-24 sdss-2023/poster-2.Rmd
minor fix to poster
minor changes
minor updates

@@
     ii. For $k = 1, ..., K$: 
-        a. Assign $z_i \leftarrow \arg\min_\ell \|x_i - g_\ell(t_i)\|^2$.
+        a. Assign $z_i \leftarrow \arg\min_k \|x_i - g_k(t_i)\|^2$.
 
@@
 Suppose that for each community $k$, we have labels for at least $R + 1$ vertices. 
-Then $K$-curves clustering outputs estimates such that  
-$$L(\hat{z}_1, ..., \hat{z}_n, \hat{g}_1, ..., \hat{g}_K; X) \stackrel{p}{\to} 0.$$
+Then $K$-curves clustering outputs estimators such that  
+$$L(\hat{z}_1, ..., \hat{z}_n, \hat{g}_1, ..., \hat{g}_K; A) \stackrel{p}{\to} 0.$$
 
@@
 
+degs <- colSums(A)
+A.trunc <- A[degs > 0, degs > 0]
+z.trunc <- z[degs > 0]
+# X <- X[degs > 0, ]
+# Xhat <- Xhat[degs > 0, ]
+
 scatterplot3d::scatterplot3d(X[, 1], X[, 2], X[, 3],
@@
 
-qgraph::qgraph(A, 
-               groups = factor(z), 
+qgraph::qgraph(A.trunc, 
+               groups = factor(z.trunc), 
                legend = FALSE,
@@
   scale_x_log10(breaks = n.vec) +
-  labs(y = 'community detection error rate',
+  labs(y = 'error rate',
        x = 'n',
@@
                     colour = factor(nsamp)),
-                width = .125) + 
+                width = .0625) + 
   theme(legend.position = 'bottom') + 

### 2023-05-24 sdss-2023/slides.Rmd
minor updates

@@
 Let $p, q \geq 0$, $d = p + q \geq 1$, $1 \leq r < d$, $K \geq 2$, and $n > K$ be integers. 
-Define manifolds $\mathcal{M}_1, ..., \mathcal{M}_K \in \mathcal{X}$ for $\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p,q} y \in [0, 1]\}$ each by continuous function $g_k : [0, 1]^r \to \mathcal{X}$.
+Define manifolds $\mathcal{M}_1, ..., \mathcal{M}_K \subset \mathcal{X}$ for $\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p,q} y \in [0, 1]\}$ each by continuous function $g_k : [0, 1]^r \to \mathcal{X}$.
 Define probability distribution $F$ with support $[0, 1]^r$. 

### 2023-05-24 sdss-2023/slides.Rmd
minor revisions

@@
 Define $X_k$ as the rows of $X$ for which $z_i = k$.\;
-Fit curve $g_k$ and positions $t_{k_i}$ to $X_k$ by minimizing $\sum_{k_i} \|x_{k_i} - g_k(t_{k_i})\|^2$.\;
+Fit curve $g_k$ and positions $t_{i}$ to $X_k$ by minimizing $\sum_{i \in C_k} \|x_{i} - g_k(t_{i})\|^2$.\;
 }
-\For {$k = 1, ..., K$} {
-Assign $z_i \leftarrow \arg\min_\ell \|x_i - g_\ell(t_i)\|^2$.\;
+\For {$i = 1, ..., n$} {
+Assign $z_i \leftarrow \arg\min_k \|x_i - g_k(t_i)\|^2$.\;
 }

### 2023-05-24 sdss-2023/slides.tex
minor changes
minor updates
minor revisions

@@
 \(K \geq 2\), and \(n > K\) be integers. Define manifolds
-\(\mathcal{M}_1, ..., \mathcal{M}_K \in \mathcal{X}\) for
+\(\mathcal{M}_1, ..., \mathcal{M}_K \subset \mathcal{X}\) for
 \(\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p,q} y \in [0, 1]\}\)
@@
 each community \(k\), we have labels for at least \(R + 1\) vertices.
-Then as \(n \to \infty\), \(K\)-curves clustering outputs estimates such
-that
+Then as \(n \to \infty\), \(K\)-curves clustering outputs estimators
+such that
 

### 2023-05-25 sdss-2023/poster-2.Rmd
minor revisions
minor fix
minor adjustments

@@
 
-Let $p, q \geq 0$, $d = p + q \geq 1$, $1 \geq r < d$, $K \geq 2$, and $n > K$ be integers. 
+Let $p, q \geq 0$, $d = p + q \geq 1$, $1 \leq r < d$, $K \geq 2$, and $n > K$ be integers. 
 Let $\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p,q} y \in [0, 1]\}$ and define manifolds $\mathcal{M}_1, ..., \mathcal{M}_K \subset \mathcal{X}$ each by continuous function $g_k : [0, 1]^r \to \mathcal{X}$.
@@
         a. Define $X_k$ as the rows of $X$ for which $z_i = k$.
-        b. Fit curve $g_k$ and positions $t_{k_i}$ by minimizing $\sum_{k_i} \|x_{k_i} - g_k(t_{k_i})\|^2$.
-    ii. For $k = 1, ..., K$: 
+        b. Fit curve $g_k$ and positions $t_{i}$ by minimizing $\sum_{i \in C_k} \|x_{i} - g_k(t_{i})\|^2$.
+    ii. For $i = 1, ..., n$: 
         a. Assign $z_i \leftarrow \arg\min_k \|x_i - g_k(t_i)\|^2$.
@@
 
-```{r sim-curves-3, fig.height = 3, fig.width = 8, out.width = '100%', cache = TRUE}
+```{r sim-curves-3, fig.height = 2, fig.width = 5, out.width = '100%', cache = TRUE}
 n.vec <- 2 ^ c(7, 8, 9, 10, 11)

### 2024-02-21 functions.R
hopefully fixed bug where code hangs

@@ -52,9 +52,9 @@ estimate.one.t.bezier <- function(x, p,
                     intercept = intercept, 
                     method = 'SANN')$par
   } else if (method == 'polyroot') {
-    if (!intercept) {
-      p <- rbind(0, p)
-    }
+    # if (!intercept) {
+    #   p <- rbind(0, p)
+    # }
     
     # p0 <- p[1, ]
     # p1 <- p[2, ]
@@ -65,20 +65,22 @@ estimate.one.t.bezier <- function(x, p,
     # b3 <- -sum((p0 - 2 * p1 + p2) ^ 2)
     # b <- c(b0, b1, b2, b3)
     
-    b <- bezier.coefs.to.d.polynomial.coefs(p, x, TRUE)
+    b <- bezier.coefs.to.d.polynomial.coefs(p, x, intercept)
+    b <- round(b, -log10(tol))
+    print(b)
     roots <- polyroot(b)
     roots <- Re(roots[abs(Im(roots)) < tol])
     roots <- roots[roots >= min.t]
     roots <- roots[roots <= max.t]
     roots <- c(roots, min.t, max.t)
     if (length(roots) > 1) {
-      minima <- roots[which.min(sapply(roots, l, intercept = TRUE))]
+      minima <- roots[which.min(sapply(roots, l, intercept = intercept))]
     } else if (length(roots) == 1) {
       minima <- roots
     } else if (length(roots) == 0) {
       warning('failed to find a real root')
       minima <- c(min.t, max.t)
-      minima <- minima[which.min(sapply(minima, l, intercept = TRUE))]
+      minima <- minima[which.min(sapply(minima, l, intercept = intercept))]
     }
   } else {
     stop('invalid method for finding t')

### 2024-02-21 functions.R
removed unnecessary print call

@@ -52,9 +52,9 @@ estimate.one.t.bezier <- function(x, p,
                     intercept = intercept, 
                     method = 'SANN')$par
   } else if (method == 'polyroot') {
-    # if (!intercept) {
-    #   p <- rbind(0, p)
-    # }
+    if (!intercept) {
+      p <- rbind(0, p)
+    }
     
     # p0 <- p[1, ]
     # p1 <- p[2, ]
@@ -65,22 +65,22 @@ estimate.one.t.bezier <- function(x, p,
     # b3 <- -sum((p0 - 2 * p1 + p2) ^ 2)
     # b <- c(b0, b1, b2, b3)
     
-    b <- bezier.coefs.to.d.polynomial.coefs(p, x, intercept)
+    b <- bezier.coefs.to.d.polynomial.coefs(p, x, TRUE)
     b <- round(b, -log10(tol))
-    print(b)
+    # print(b)
     roots <- polyroot(b)
     roots <- Re(roots[abs(Im(roots)) < tol])
     roots <- roots[roots >= min.t]
     roots <- roots[roots <= max.t]
     roots <- c(roots, min.t, max.t)
     if (length(roots) > 1) {
-      minima <- roots[which.min(sapply(roots, l, intercept = intercept))]
+      minima <- roots[which.min(sapply(roots, l, intercept = TRUE))]
     } else if (length(roots) == 1) {
       minima <- roots
     } else if (length(roots) == 0) {
       warning('failed to find a real root')
       minima <- c(min.t, max.t)
-      minima <- minima[which.min(sapply(minima, l, intercept = intercept))]
+      minima <- minima[which.min(sapply(minima, l, intercept = TRUE))]
     }
   } else {
     stop('invalid method for finding t')

### 2024-02-22 draft/draft.Rmd
minor changes

@@
 ---
-title: Manifold Clustering in the Generalized Random Dot Product Graph
+title: The Latent Structure Block Model
 
@@
 output: rticles::asa_article
-urlcolor: blue
+# urlcolor: blue
 header-includes:
@@
 - \usepackage{amsfonts}
+- \usepackage{pdfpages}
+- \usepackage{xcolor}
 ---
@@
 The goal of the analysis is then to estimate these labels upon observing a graph. 
-In order to perform this analysis as a statistical inference task is to define a probability model with inherent community structure. We call such models *block models*: 
+One approach is to define a random graph model that includes community labels. 
 First, each vertex is assigned a label $z_1, ..., z_n \in \{1, 2, ..., K\}$ where $K \ll n$. 
-Then each edge probability $P_{ij}$ is said to depend on the labels $z_i$ and $z_j$, possibly along with some other parameters. 
-For example, the stochastic block model (SBM) sets a fixed edge probability for each pair of communities, i.e., $P_{ij} = \omega_{z_i, z_j}$. 
-The degree-corrected block model (DCBM) assigns an additional parameter $\theta_i$ to each vertex by which edge probabilities are scaled, i.e., $P_{ij} = \theta_i \theta_j \omega_{z_i, z_j}$. 
+Then the edge probabilities are restricted such that they depend on the labels of the vertices, and possibly some other parameters specific to each vertex, i.e., $P_{ij} = g(z_i, z_j, \phi_i, \phi_j)$, where $\phi_i$ is a set of parameters corresponding to vertex $i$.
+We call such models *block models*. 
+The simplest and best known block model is the stochastic block model (SBM), which sets a fixed edge probability for each pair of communities, i.e., $P_{ij} = \theta_{z_i, z_j}$. 
+The degree-corrected block model (DCBM) assigns an additional parameter $\omega_i$ to each vertex by which edge probabilities are scaled, i.e., $P_{ij} = \omega_i \omega_j \theta_{z_i, z_j}$. 
 The popularity adjusted block model (PABM) assigns $K$ parameters to each vertex $\lambda_{i1}, \lambda_{i2}, ..., \lambda_{iK}$ that describe that vertex's affinity toward each community; 
 the edge probability between vertices $i$ and $j$ is then defined as the product of vertex $i$'s affinity toward vertex $j$'s community and vertex $j$'s affinity toward vertex $i$'s community, i.e., $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$.
-
 The three block model types, as well as the $\ER$ model, impose structure on $P$, including on the rank of $P$. 
@@
 This includes the SBM, in which communities correspond to point masses, DCBM, in which communities correspond to line segments, and PABM, in which communities correspond to orthogonal subspaces. 
-In this work, we extend this idea to communities organized into more general latent structures.
-In particular, we assume that each community corresponds to a manifold in the latent space.
+In this work, we extend this idea to communities organized into more general latent structures. In particular, we assume that each community corresponds to a manifold in the latent space.
 
@@
 All Bernoulli graphs are generalized random dot product graphs. 
-Whether this is useful for inference depends on the structure of the latent space. 
-In the case of the $\ER$ model, SBM, DCBM, and PABM, the latent structure is linear, and the linearity can be exploited for community detection and parameter estimation. 
-In this section, we discuss general, often nonlinear latent structure models, focusing on those with community structure.
+For a specific example, consider the two-community DCBM and its GRDPG latent space representation. 
+In the DCBM, the edge probability between each pair of vertices is the product of the baseline community-wise edge probability and the degree correction factors of the vertices, or $P_{ij} = \theta_{z_i, z_j} \omega_i \omega_j$.
 
@@
 
-\begin{algorithm}[h]
+\begin{algorithm}[H]
 \DontPrintSemicolon
@@
 ```{r}
-clustering.df <- readr::read_csv('../simulations/balanced.csv')
+clustering.df <- readr::read_csv('../simulations/balanced-2/balanced.csv')
 ```
@@
 \section{Details on Fitting Bezier Curves with Noise}
+
+# References

### 2024-02-22 draft/draft.tex
minor changes

@@
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
 % !TeX program = pdfLaTeX
@@
 \usepackage{enumerate}
-\usepackage{natbib}
+\usepackage[]{natbib}
 \usepackage{textcomp}
-\usepackage[hyphens]{url} % not crucial - just used below for the URL
-\usepackage{hyperref}
 
@@
 \addtolength{\oddsidemargin}{-.5in}%
-\addtolength{\evensidemargin}{-.5in}%
+\addtolength{\evensidemargin}{-1in}%
 \addtolength{\textwidth}{1in}%
-\addtolength{\textheight}{1.3in}%
-\addtolength{\topmargin}{-.8in}%
+\addtolength{\textheight}{1.7in}%
+\addtolength{\topmargin}{-1in}%
 
@@
 \usepackage{amsfonts}
+\usepackage{pdfpages}
+\usepackage{xcolor}
+
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\hypersetup{
+  pdftitle={The Latent Structure Block Model},
+  pdfkeywords={block models, community detection, coordinate descent,
+latent structure models, manifold clustering, random dot product graph},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
 
@@
 {
-  \title{\bf Manifold Clustering in the Generalized Random Dot Product
-Graph}
+  \title{\bf The Latent Structure Block Model}
 
@@
   \begin{center}
-    {\LARGE\bf Manifold Clustering in the Generalized Random Dot Product
-Graph}
+    {\LARGE\bf The Latent Structure Block Model}
   \end{center}
@@
 latent structure models, manifold clustering, random dot product graph
+
 \vfill
@@
 \newpage
-\spacingset{1.45} % DON'T change the spacing!
+\spacingset{1.9} % DON'T change the spacing!
 
@@
 community label. The goal of the analysis is then to estimate these
-labels upon observing a graph. In order to perform this analysis as a
-statistical inference task is to define a probability model with
-inherent community structure. We call such models \emph{block models}:
-First, each vertex is assigned a label
-\(z_1, ..., z_n \in \{1, 2, ..., K\}\) where \(K \ll n\). Then each edge
-probability \(P_{ij}\) is said to depend on the labels \(z_i\) and
-\(z_j\), possibly along with some other parameters. For example, the
-stochastic block model (SBM) sets a fixed edge probability for each pair
-of communities, i.e., \(P_{ij} = \omega_{z_i, z_j}\). The
+labels upon observing a graph. One approach is to define a random graph
+model that includes community labels. First, each vertex is assigned a
+label \(z_1, ..., z_n \in \{1, 2, ..., K\}\) where \(K \ll n\). Then the
+edge probabilities are restricted such that they depend on the labels of
+the vertices, and possibly some other parameters specific to each
+vertex, i.e., \(P_{ij} = g(z_i, z_j, \phi_i, \phi_j)\), where \(\phi_i\)
+is a set of parameters corresponding to vertex \(i\). We call such
+models \emph{block models}. The simplest and best known block model is
+the stochastic block model (SBM), which sets a fixed edge probability
+for each pair of communities, i.e., \(P_{ij} = \theta_{z_i, z_j}\). The
 degree-corrected block model (DCBM) assigns an additional parameter
-\(\theta_i\) to each vertex by which edge probabilities are scaled,
-i.e., \(P_{ij} = \theta_i \theta_j \omega_{z_i, z_j}\). The popularity
+\(\omega_i\) to each vertex by which edge probabilities are scaled,
+i.e., \(P_{ij} = \omega_i \omega_j \theta_{z_i, z_j}\). The popularity
 adjusted block model (PABM) assigns \(K\) parameters to each vertex
@@
 affinity toward vertex \(i\)'s community, i.e.,
-\(P_{ij} = \lambda_{i z_j} \lambda_{j z_i}\).
-
-The three block model types, as well as the
-\(\text{Erd\"{o}s-R\'{e}nyi}\) model, impose structure on \(P\),
-including on the rank of \(P\). \(P\) has rank 1 for the
-\(\text{Erd\"{o}s-R\'{e}nyi}\) model, rank \(K\) for the SBM and DCBM,
-and rank \(K^2\) for the PABM. This provides the intuition behind
+\(P_{ij} = \lambda_{i z_j} \lambda_{j z_i}\). The three block model
+types, as well as the \(\text{Erd\"{o}s-R\'{e}nyi}\) model, impose
+structure on \(P\), including on the rank of \(P\). \(P\) has rank 1 for
+the \(\text{Erd\"{o}s-R\'{e}nyi}\) model, rank \(K\) for the SBM and
+DCBM, and rank \(K^2\) for the PABM. This provides the intuition behind
 another family of Bernoulli graphs called the \emph{random dot product
@@
 
-All Bernoulli graphs are generalized random dot product graphs. Whether
-this is useful for inference depends on the structure of the latent
-space. In the case of the \(\text{Erd\"{o}s-R\'{e}nyi}\) model, SBM,
-DCBM, and PABM, the latent structure is linear, and the linearity can be
-exploited for community detection and parameter estimation. In this
-section, we discuss general, often nonlinear latent structure models,
-focusing on those with community structure.
+All Bernoulli graphs are generalized random dot product graphs. For a
+specific example, consider the two-community DCBM and its GRDPG latent
+space representation. In the DCBM, the edge probability between each
+pair of vertices is the product of the baseline community-wise edge
+probability and the degree correction factors of the vertices, or
+\(P_{ij} = \theta_{z_i, z_j} \omega_i \omega_j\).
 
@@
 
-\bibliographystyle{agsm}
+\bibliographystyle{plain}
+\renewcommand\refname{References}
 \bibliography{bibliography.bib}

### 2024-02-23 functions.R
adding another initialization option

@@ -66,7 +66,7 @@ estimate.one.t.bezier <- function(x, p,
     # b <- c(b0, b1, b2, b3)
     
     b <- bezier.coefs.to.d.polynomial.coefs(p, x, TRUE)
-    b <- round(b, -log10(tol))
+    b <- round(b, -log10(tol) * 2)
     # print(b)
     roots <- polyroot(b)
     roots <- Re(roots[abs(Im(roots)) < tol])
@@ -225,7 +225,7 @@ estimate.bezier.curve.2 <- function(X,
   W <- diag(weights)
   
   if (missing(init.params)) {
-    if (initialization == 'random') {
+    if (initialization[1] == 'random') {
       if (intercept) {
         p <- matrix(rnorm(d * (degree + 1)), nrow = degree + 1, ncol = d)
       } else {
@@ -242,7 +242,7 @@ estimate.bezier.curve.2 <- function(X,
         }
       }
       
-    } else if (initialization == 'isomap') {
+    } else if (initialization[1] == 'isomap') {
       isomap.out <- estimate.bezier.curve.isomap(X, 
                                                  degree = degree, 
                                                  k = k.isomap, 
@@ -250,10 +250,14 @@ estimate.bezier.curve.2 <- function(X,
                                                  intercept = intercept)
       p <- isomap.out$p
       t.hat <- isomap.out$t
-    } else  if (initialization == 'x') {
+    } else if (initialization[1] == 'x') {
       t.hat <- X[, 1]
       if (!intercept) t.hat <- abs(t.hat)
       t.hat <- normalize.umvue(t.hat)
+      T <- construct.bezier.model.matrix(t.hat, degree, intercept = intercept)
+      p <- solve(t(T) %*% W %*% T, t(T) %*% W %*% X)
+    } else if (length(initialization) == nrow(X)) {
+      t.hat <- initialization
       T <- construct.bezier.model.matrix(t.hat, degree, intercept = intercept)
       p <- solve(t(T) %*% W %*% T, t(T) %*% W %*% X)
     } else {
